### 2020-07-02 __latexindent_temp.tex
Template added

\documentclass{article}

\usepackage{epcc}
\usepackage{graphics}
\usepackage{booktabs}
\usepackage{mathrsfs}

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
% this will make nested numbering all arabic numbers

% This example file shows how a report can be laid out using Latex. It
% does not use any special local features so should be portable to other
% places.
% 
% When producing draft copies of a thesis you may want to print only
% selected pages of the thesis. To do this use the command
% 
% dvips -f -p 4 -n 3 myfile.dvi | lpr
% 
% where -p 4 means start printing at page 4 (ie the page that will be
% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
% This example will print pages 4, 5 and 6.
% 
% If you want to print the report and also save paper you can print more
% than one page on each sheet of paper. Use the command
% 
% dvips -f myfile.dvi | psnup -2 | lpr
% 
% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
%
% To produce a PDF version you can create a PostScript copy first
%
% dvips -f myfile.dvi > myfile.pdf
%
% and then convert it
%
% distill myfile.ps
%
% or you can go straight to PDF
%
% pdflatex myfile
%
% Note that pdflatex expects all included figures to be in PDF too. See
% the includegraphics command below.

% This document contains many cross-references and forward references,
% eg in constructing a table of contents, so Latex may need to be run
% twice to get all the references correct. If you need to run Latex twice
% you may get the warning:
% 
% LaTeX Warning: Label(s) may have changed. Rerun to get cross-references right.

\begin{document}

\pagenumbering{roman}

\title{Project Preparation Report\\Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
\author{Chaolin Han}
\date{\today}

\makeEPCCtitle

\newpage

\tableofcontents

% \newpage

% \section*{Acknowledgements}

% This template is a slightly modified version of the one developed by
% Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
% acknowledgement follows:

% {\em This template has been produced with help from many former students who
% have shown different ways of doing things. Please make suggestions for
% further improvements.}

% \section*{Note}

% This template was originally created for a full MSc dissertation.
% Although I have modified the format slightly (eg there are no separate
% chapters), I have not substantially changed the text or the advice
% appearing in the \LaTeX comments (source lines beginning with \%). You
% should therefore be aware that for a shorter report you may
% not need to go into as much detail as indicated here. For example, you
% might not need to include any acknowledgements and any reference list is
% likely to be very short.

\newpage
\pagenumbering{arabic}

\section{Introduction}

This report focuses on the work in the preparation phase of 
MSc High-performance Computing with Data Analysis Dissertation Project. 
The project is built around 
a biological dynamical systems model with parameter estimation and model comparison 
being the
primary tasks. The complex interactions between several cells and molecules
in zebrafish spinal cord regeneration is the modelling target.

Previous dissertation review section gives comments on the content 
and presentation of a dissertation report. The background of this project is 
introduced in detail regarding several aspects of the project with the motivation
and related literature discussed. 
The preliminary work includes background 
research, literature review, tools and techniques research and their results 
are illustrated. The final project 
proposal and workplan breakdown are given with regards to the conclusion from 
background research and discussion with supervisors, together with a 
estimated project timeline. Risk analysis and possible 
outline of dissertation report are also addressed.

The project aims at the parameter estimation and model comparison. Approximate 
Bayesian computation (ABC) approaches will be used to estimate parameters of mathmatical

\section{Previous Dissertation Review}

The reviewed previous dissertation is {\em Modelling Room Acoustics with a 
GPU} by Jacob Josiah Webber, finished in August 2017.
\subsection*{Summary}

Finite-difference time-domain (FDTD) techniques are introduced to numerically solve an acoustics model. The sound will be affected by the room it is in. The solution of wave equations of the sound can be computationally approximated using FDTD. It is a traditional simulation job for GPU. To improve the performance, the author proposed a ‘bounding box’ method, whose accompanied data structure significantly reduced memory usage and resultant simulation computation time is saved.

\subsection*{Does the dissertation adequately explain the context in which the problem is set?}

The motivation is addressed in Chapter 1 (Introductions) and the first part of Chapter 2. It briefly introduced together with its performance concerns. More details could be added to describe the meaning of this modelling mission and performance optimisation mission with relevant application or implementation examples given to clarify it.

The background and literature review is represented adequately in details. Section 2.1 derived formulas about acoustics wave, applied finite-difference on them and obtained an iterative propagation computation equation (Equation 2.15) which is the operation o the numerical modelling. Related boundary conditions and stencilling operations are also properly described. Later sections (2.3 and 2.4) addressed the GPU-based parallelisation and FDTD performance issues in the acoustics modelling context.

The background and literature review section is well structured and addressed and explains the problem background in a way that readers without much-related background knowledge (e.g. acoustics, GPU structures) could understand it.

\subsection*{Has the student adequately explained the scope of the problem, the approach that they would take, how success would be measured and reflected on the outcomes?}

The aim of the project is to propose a new method and improve the modelling performance. To address the benefits of the proposed method, the author gave both theoretical  (section 3.3)and experimental (section 5.2) analysis to prove its effectiveness. The software development process (section 4.2) and theoretical derivation and implementation of the ‘bounding box’ method (2.4.1 and 3.2) are reasonable and adequate.

The reproducibility is achieved by giving the specific hardware platforms (NVIDIA K20 GPU server) and its environment (Appendix B), as well as source code and building instructions. As the author addressed in section 5.3, the replication of the program should output a consistent timing results to ones in Table 5.1 and 5.2.

In regard of the limitations, as stated in the conclusion section, ‘particular effort was made to avoid choosing rooms that were overly suited to the new method’; however more test case should be given to make the results more convincing. Another limitation has been stated in the introductory paragraphs of Chapter 4, where experimentation and results analysis are not fully explored due to lack of time.

The overall outcome is satisfying as it matches the theoretical prediction and the improvement is significant compared to previous method, regardless of the above limitations.

\subsection*{Quality of presentation}
The language is accurate and descriptions of the theories, designs and implementations are clear to follow.

The listed figures are informative. E.g., Figure 2.1 gives clear representation of the stencilling operations, and Figure 3.1 is a suitable example to illustrate the theories.

The dissertation is well structured such that the process of the project and the works in each stage of the project are included. The tables and algorithm blocks are all represented in clear academical forms, so do their placement and alignment. The formulas and equations, algorithm blocks and figures are all consistently numbered and referred properly.

The references are managed in a consistent manner and are correctly referred when needed. The appendices include original unsummarised data, code and GPU device information which provide meaning supplement when readers want to check some details not mentioned in the main body. However some other content should also be presented here, e.g. implementation environment, experiment input (boundary settings) and benchmarking tools/methods.

\section{Background and Literature Review}

Mathematical models are helpful in resolving complex interactions in real-life 
dynamical systems. The project will use a mathematical model to represent 
zebrafish spinal cord regeneration process, then determine its parameter 
estimation and compare it with alternative models.

Zebrafish is one kind of freshwater fish that is widely used in scientific 
research for its regenerative abilities. Zebrafish can regenerate their spinal 
cord after injury. Tsarouchas et al.\cite{ref:Tsarouchas} revealed how the immune system 
responds to 
allow axons of nerve cells growth across the lesion site. The experiments showed 
evidence that several kinds of immune cells and cytokines play determinative 
roles in regulating inflammation and repair. Cytokines are small proteins 
secreted by certain cells to pass signals across cells. The regeneration 
process is controlled by two cytokines, Il-1$\beta$ and Tnf-$\alpha$. Generally, 
Il-1$\beta$ 
inhibits the regeneration while Tnf-$\alpha$ promotes it, both of which are 
dynamically controlled by mainly two kinds of immune cells (i.e. macrophage, 
neutrophil). Based on these experimental results the initial 
hypothesis is proposed that the repair process are dynamically controlled by 
interactions between these cells and cytokines. Figure \ref{fig:1} illustrates the 
basic interactions.

\begin{figure}

\begin{center}
\resizebox{0.60\hsize}{!}{\includegraphics{fig/fig1}}
\end{center}

\caption{Interactions in zebrafish spinal cord regeneration, proposed by Tsarouchas 
et al\cite{ref:Tsarouchas}. Lines ended with an arrow indicate promotion and lines 
ended with a short segment indicate inhibition}
\label{fig:1}

\end{figure}

Mathematical models used to describe complex interactions of dynamical systems 
are usually in the form of differential equations. Analytically solving differential 
equations is not always feasible and the solution can be extremely complicated 
especially for complex systems, so often dynamical system models are solved 
numerically to generate data, which can be compared with real data and determine 
model error. According to our hypothesis an initial ordinary differential equation 
system can be built by describing the dynamics of number or concentration of immune 
cells and cytokines. The initial equations shown in (\ref{eq:1}) are preliminary equations 
that represent the dynamical model, which might be adjusted or calibrated according 
to modelling results and real-world data.

In (\ref{eq:1}), $N$ and $\Phi$ represent the
number of neutrophil and macrophage respectively, $\alpha$ and $\beta$ denotes the
relative concentration of cytokines Tnf-$\alpha$ and Il-1$\beta$. 
$\lambda$, $\kappa$, $\mu$, $s$, $i$ and $\nu$ with different subscripts are 
all model parameters i.e. coefficient in the interactions.

\begin{equation} \label{eq:1}
\begin{array}{l}
\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\\\
\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\\\
\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}}-\mu_\beta\beta\\\\
\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
\end{array}
\end{equation}

A mathematical model would introduce parameters (e.g. $\kappa$, $\mu$ ,etc in (1)), which are to be determined to 
give a specific model using observed data. A traditional method to find the best desired parameter sets 
is to apply an optimisation-based fitting, where the discrepancy of the model’s 
predictions and experimental data is minimised. Alternatively, by adopting Bayesian 
inference, we can focus on the experimental data and try to infer the parameter sets. 
Many approaches have been proposed and used in the area of systems biology\cite{ref:Stumpf}. Compared 
to optimisation-based fitting methods, Bayesian inference approaches can relieve the 
concerns of over-fitting and local optima. It focuses on posterior distribution which 
represents parameters’ ‘uncertainty’ given the observation data. As Bayes rule 
stated, the posterior distribution $p(\theta|D)$ is proportional to product of likelihood and 
prior
\begin{equation} \label{eq:2}
p(\theta|D)\propto l(\theta|D)p(\theta)
\end{equation}
where $\theta$ denotes parameter vector, $l$ is the likelihood function and $D$ is the observed data.

The prior distribution $p(\theta)$ is given to specify the initial estimates on 
parameters before observing the data; the likelihood is given as conditional 
probability $l(\theta|D)$ which indicates how well the observed data is explained by 
the parameter $\theta$.

% It can be hard to find analytical expression for likelihood, especially in large 
% complex models; instead computational approaches such as Markov Chain Monte Carlo 
% (MCMC)\cite{ref:Gilks} and sequential Monte Carlo (SMC)\cite{ref:Toni} are proposed to 
% solve this problem. 
% These approaches try to approximate the 
% posterior distribution by simulating the data and they are known as approximate Bayesian 
% computation (ABC). S.A. Sisson\cite{ref:Sisson} illustrated that sequential Monte Carlo sampler can 
% be more powerful and efficient than MCMC and traditional ABC rejection methods.

It can be hard to find an analytical expression for likelihood, especially in large 
complex models. Computational approaches such as Markov Chain Monte Carlo\cite{ref:Gilks} techniques
could be used 
to solve this, where sampling methods are provided to computationally evaluate the likelihood.
Approximate Bayesian computation (ABC) methods are alternative solutions. They are likelihood-free 
methods that simulate data from model and approximate posterior distribution 
instead of evaluating the likelihood. There are several ABC approaches e.g. ABC-rejection, ABC-MCMC
and ABC-SMC (sequential Monte Carlo)\cite{ref:Toni}. S.A. Sisson\cite{ref:Sisson} suggested that sequential Monte Carlo sampler in ABC can 
be more powerful and efficient than ABC-MCMC and traditional ABC rejection methods.

ABC approaches have been widely used in biology applications due to their 
‘likelihood-free’ feature. In many systems biology models, the likelihood is 
intractable, meaning that it cannot be defined as a function of the 
parameters $\theta$ analytically. Liepe et al.\cite{ref:Liepe} showed a 
successful application of 
ABC-SMC in systems biology 
with a software package ABC-SysBio, where ABC-SMC is used both in parameter estimation and model 
selection.

Parameters of the zebrafish spinal cord regeneration model can be estimated 
using ABC-SMC as it is hard to find exact likelihood here and an efficient 
approximation approaches might help us in exploring the uncertainty in parameter 
sets and comparing our model with alternative ones.

% see the man page for dvips for details of the special command which is
% much more powerful than is shown here. It allows offsets in the
% horizontal and vertical and scaling in x and y.

% choosing suitable values for offset and scale can be a tiring matter
% of trial and error.

% note that labels do not need to include a description of the object
% they are labelling but it can be helpful, eg \label{fig:figurename}.

\section{Preliminary Investigations}

Preliminary works include background material study and tools research, which aims to build a concrete start of the project and explore the uncertainty and suitability of tools.

\subsection{Background Research}

The two key background papers\cite{ref:Tsarouchas,ref:Liepe} were studied firstly. 
Tsarouchas et al.\cite{ref:Tsarouchas} illustrated 
the biology process (zebrafish spinal cord regeneration after injury) that we want to 
model and introduced the dynamic controls within the process in details. As 
shown in the paper the process requires related biology 
background knowledge, especially in immunology, thus related necessary background
materials were studied to form a more comprehensive understanding and could help in 
future analyse of results. The other background paper focuses on the algorithms that 
are expected to be implemented in the project. More materials related to Bayesian 
inference were studied, including tutorials, books and articles. Basic theories of 
Bayesian inference and practical examples were studied with several ABC tutorial examples 
practised to improve understanding. Other background materials include papers on 
ABC-SMC algorithm, including the theory, implementation steps and comparison with 
other ABC algorithms.

The mathematical models were also preliminarily studied, with concerns over model 
building, data processing and solving. Ordinary differential equation computational 
solution and its solver were reviewed in preparation for the future actual 
implementation.

The background research is the most important work in the preparation phase, 
as it converted the first impression and interests of the project into concrete 
knowledge in the biology and statistical inference basis such that the mission, 
risks and challenges of the project could be discussed and analysed; it also 
ensured an easier technically start of the project.

\subsection{Tools Research}
Besides the background research, tools that might be used in the project, mainly 
the packages and libraries, were studied, compared and evaluated.

There are currently varies of tools and softwares that implement ABC-SMC. A desirable tool should take many technical aspects into account, e.g. working environment, mathematical models solvers, parallelisation support and reliability.

ABC-SMC is expected to be applied in Python or R languages, as most of the published packages are implemented on these two languages; further studies may use other languages if a more suitable software package for that is found or we need to build our own code for optimising.

The studied packages are listed in Figure \ref{fig:2}. Their user guidance, 
documentations (if exist) are viewed, from which their features are concluded 
in a note (taskPackages.md under tasks directory in project repository);
their main features are listed below.

\begin{figure}

	\begin{center}
	\resizebox{1.00\hsize}{!}{\includegraphics{fig/fig2}}
	\end{center}

	\caption{Studied packages}
	\label{fig:2}

	\end{figure}

1. pyABC

\quad $\bullet$ Built-in support for multiple types of  parallelisation

\quad $\bullet$ Built-in visualisation

2. astroABC

\quad $\bullet$ Support parallel sampling using MPI or multiprocessing

\quad $\bullet$ Each run can be paused and resumed later

3. ELFI

\quad $\bullet$ Connect to ipyparallel for more parallelisation options

\quad $\bullet$ Using directed acyclic graph as model

4. ABC-SysBio

\quad $\bullet$ Experimental CUDA accelerated parallelisation

5. EasyABC

\quad $\bullet$ Multicore parallelisation

\quad $\bullet$ Offer several different sequential algorithms

6. smfsb

\quad $\bullet$ Provide simple and basic output of parameter estimations.

Some examples given in their documentations have been practised to explore the details and features of ABC-SMC algorithm in these packages and determine which of them meet our requirements. The three preferred packages are pyABC, EasyABC and ELFI. All of them are well documented and support different kinds of parallelisation; other features include support for built-in visualisation, pause-and-restart and easy to use. The choice is limited in the preliminary practise and may be changed accordingly in the future practical implementation as more packages will be studied.

Tools research provides grounded comprehension and conclusion on packages 
choosing in this project and finds the initial tools to start with. More 
tools will be studied and possibly used in the future.

\section{Final Proposal}

The project research consists of six parts. The first two parts are preparation work that would result in a mathematical model to implement our algorithms. Part 3, 4 and 5 are the practical session including several experiments and implementations to explore and analyse the model. The last part is to organise the works into report and dissertation.

\subsection{Background Research and Project Preparation}

This part focus on the preparation of the project, including background reading, tools research and regular discussion with supervisors. It will build a concrete ground to start the later tasks. Most of the work have already been done.

\subsection{Dynamical System Modelling}

Non-linear differential equations will be adapted to represent the dynamic system of zebrafish spinal cord regeneration process. Under the hypothesis1, there are complex interactions between two types of cytokines and three types of cells at the wounded site and the expected mathematical model representing by ODE systems will be built to describe these interactions.

\subsection{Parameter Estimating}

The parameters in the ODE system will be estimated using experimental data1. ABC-SMC will be applied to approximate the posterior distribution of parameters and produce reliable parameter estimates of the model. A distance function will be built to measure the discrepancy between the model and simulated data before applying the algorithm, together with coping with missing data in certain time points.

As the SMC sampling process is computation-intensive and thus time-consuming, a parallel sampling technique will be of our interest.

\subsection{Results Interpretation and Analysing}

This part focuses on the evaluation of the parameterised model and analysis of the SMC results. Inference results will be tested using synthetic data and used to evaluate how well the model fits the data and how much confidence we have on the parameter estimates.

\subsection{Model Comparison}

ABC-SMC can also be implemented to compare our model with alternative ones. Similar procedures will be applied to model candidates and result in an approximation of each model’s marginal distribution, which could be the main criteria for our model selection.

\subsection{Performance Analysis}

The performance data of the sampling process will be analysed to reveal performance of the parallelisation implementation, including parallel speedup and efficiency. Trade-offs between cost, runtime and energy will also be discussed.

\subsection{Dissertation and Report Writing}

A project study report and final dissertation will be written to address all works and results in the project research process.

\subsection{Estimated Computational Resources Usage}

The computational resources usage estimation is a provisional rough estimation, 
which can be changed as the implementation and performance test proceeds.

\textbf{Cirrus}

Most of the computational work in this project is expected to be deployed on Cirrus.
\begin{itemize}
	\item CPU hrs needed (standard compute nodes): 7200 CPUh (2 hrs per run $\times$ 
	36 cores$\times$ 100 runs)
	\item Disk storage needed: 10 GB, including data and softwares
	\item Access to GPU compute nodes is also needed for possible parallelisation using CUDA
\end{itemize}

\section{Workplan}

Each part of the project research is planned as followed. The first stage is to prepare for the project in Semester 2 and possibly in the Easter holidays; the second stage will start at 18 May when the last examination ends.

\begin{enumerate}
	\item Background Research and Project Preparation, Dynamical System Modelling

	$\bullet$ Start time: 21/01/2020

	$\bullet$ Duration: 8 weeks

	$\bullet$ The later Easter holidays could be used to do more research

	\item Parameterising

	$\bullet$ Start time: 18/05/2020

	$\bullet$ Estimated duration: 7 weeks

	\item Results Interpretation and Analysing

	$\bullet$ Start date: 22/06/2020

	$\bullet$ Estimated duration: 5 weeks

	\item Model comparison

	$\bullet$ Start date: 15/06/2020

	$\bullet$ Estimated duration: 5 weeks

	\item Performance Analysis

	$\bullet$ Start date: 13/07/2020

	$\bullet$ Estimated duration: 2 weeks

	\item Dissertation and report writing

	$\bullet$ Start date: 20/07/2020

	$\bullet$ Estimated duration: 5 weeks

\end{enumerate}

The Gantt chart is shown in Figure \ref{fig:3}. The two stages are separately listed.

\begin{figure}

	\begin{center}
	\resizebox{0.80\hsize}{!}{\includegraphics{fig/fig3}}
	\end{center}

	\caption{Workplan Gantt chart}
	\label{fig:3}

	\end{figure}

\section{Risk Analysis}

Possible risks and challenges are addressed in this section. These risks are concluded from the preliminary researches. Though we tried to predict and evaluate all risks in the project, new risks may appear as the project proceeds.

\subsection{Technical Problems}

\textbf{Tools/software problems}

\begin{itemize}
	\item There could be unforeseen bugs in the code of the tools or software that are used in the project; softwares may crash, leading to possible data damaging. It is usually unpreventable.

	Probability: moderate

	Mitigation: in this case, we may need to find and fix the bug, switch tool/softwares and/or make a data protection plan to reduce the impact.

	\item There could be problems in the implementation environment where the compatibility is our main concern. The implementation may require certain environment conditions (e.g. operating system, compiler version, dependencies, etc) to be performed, or they could be incompatible with the hardware (e.g. CUDA acceleration and multithreading).

	Probability: high

	Mitigation: we have studied a variety of packages (section 4.2) to ensure that we have alternatives when it is necessary to switch tools, and will take more tools into the backup plan. Another solution is to change or built the required environment if applicable, i.e. switching OS, compile environment or hardware platform.

	\item The worst case could be that we may need to write our code to implement the ABC-SMC algorithm and complete the rest study, which would heavily increase the workload.

	Probability: low

	Mitigation: The workplan should be adjusted once it happens to ensure more time devoted to algorithm code and debug process. The rest work may be simplified.

\end{itemize}

\textbf{Hardware problems}

\begin{itemize}
	\item Hardware failures and hardware unavailability could intrude considerable impact on the project process. Hardware failure may cause data damaging; hardware unavailability may break the estimated workplan.

	Probability: low

	Mitigation: a regular data backup should be planned where experiments records, results and configurations should be well protected. In this case to proceed the project we may need to seek for hardware alternatives.
\end{itemize}

\subsection{Parallelisation Concerns}

\begin{itemize}
	\item The project seeks to implement a parallel sampling. The first risk could be that different tools use different parallel techniques targeting different hardwares, where the parallel performance comparison across hardware is a problem.

	Probability: high

	Mitigation: more study on the cross-platform benchmark should be conducted to give a reasonable comparison of performance.

	\item Another risk can be the tricky parallel implementations, as there are different parallel systems with different configuration; it may cost time to find which kinds of implementation would work before we proceed to parallel performance analyse.

	Probability: moderate

	Mitigation: more investigation of the parallel implementations should be performed, which can reduce the effort in finding working schemes along with investigations in section 4.2.

	\item The execution time of parameterising using ABC-SMC is mainly determined by problem size and sampling parameters (number of generations, tolerance sequence and population size); it may cost time to obtain a reasonable parameter input that avoids long execution time and in the meantime give precise results. Another trade-off we may face could be the common scalability trade-offs where a suitable scale-up is to be discussed.

	Probability: high

	Mitigation: It is usually unavoidable. To minimise the cost and effort a plan on the implementation experiment should be made and discussed where only necessary experiment should be conducted and redundant parameter space enumeration should be avoided. 
\end{itemize}

\subsection{Modelling}

\begin{itemize}
	\item The main risk in the modelling process is model misspecification.

	Probability: moderate

	Mitigation: we may need to calibrate our model or adopt another model. These changes could consequently affect all later works in parameterising and analysis and the workplace should be adjusted accordingly.

	\item There is also risks in experiment data processing: variables are measured in different time steps and data on gene expression uses different scales from the differential equations. These problems will introduce some difficulties in the model fitting.

	Probability: high

	Mitigation: the processing methods should be studied and backup plan should be made to enable a successful fitting.
\end{itemize}

\subsection{Other Risks}

As far as this project concerns, certain cases of force majeure could affect the project progress which are low in probability. It should be noted that the current Coronavirus epidemic could affect some part of the project process, e.g. supervision meetings and access to certain physical resources.

The project may face failure, e.g. undesirables results, critical technical problems and running out of time. These failures should be noticed as early as possible through the risk analysis and management throughout the project time and be avoided as much as we can.

\section{Outline of the Dissertation Report}

A possible outline of the Dissertation Report is structured as below.

\begin{enumerate}
	\item Introduction
	\item Background and Literature Review
	\begin{enumerate}
		\item Zebrafish spinal cord regeneration
		\item Mathematical modelling
		\item Bayesian Inference
	\end{enumerate}
	\item Modelling
	\item Design and Implementation
	\begin{enumerate}
		\item Parameter estimation
		\item Model comparison
		\item High-performance implementations
	\end{enumerate}
	\item Results and Analysing
	\begin{enumerate}
		\item ABC-SMC results
		\item Performance analysis
		\item Discussions
	\end{enumerate}
	\item Future Works
	\item Conclusions
\end{enumerate}

\pagebreak
\begin{thebibliography}{100}

\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N. and Feng, Y., 
2018. Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration. Nature communications, 9(1), pp.1-17.

\bibitem{ref:Stumpf}Stumpf, M., Balding, D.J. and Girolami, M., 
2011. Handbook of statistical systems biology. John Wiley \& Sons.

\bibitem{ref:Gilks}Gilks, W.R., Richardson, S. and Spiegelhalter, D., 
1995. Markov chain Monte Carlo in practice. Chapman and Hall/CRC.

\bibitem{ref:Toni}Toni, T., Welch, D., Strelkowa, N., Ipsen, A. and Stumpf, M.P., 
2009. Approximate Bayesian computation scheme for parameter inference and model selection in dynamical systems. Journal of the Royal Society Interface, 6(31), pp.187-202.

\bibitem{ref:Sisson}Sisson, S.A., Fan, Y. and Tanaka, M.M., 
2007. Sequential monte carlo without likelihoods. Proceedings of the National Academy of Sciences, 104(6), pp.1760-1765.

\bibitem{ref:Liepe}Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P. and Stumpf, M.P., 
2014. A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation. Nature protocols, 9(2), p.439.

\end{thebibliography}

\end{document}

### 2020-07-02 MSc_dissertation.tex
Ready for work

--- 2020-07-02 __latexindent_temp.tex
@@
 
-The project aims at the parameter estimation and model comparison. Approximate 
-Bayesian computation (ABC) approaches will be used to estimate parameters of mathmatical
+The project aims at the parameter estimation and model comparison. As a kind of 
+computational Bayesian inference methods, Approximate 
+Bayesian computation (ABC) approaches will be used to estimate parameters of mathematical models and
+compare our model with alternative ones. A parallel implementation is of our interests 
+and performance data and inference outputs from ABC will be analysed then to evaluate the result.
 
@@
 \frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\\\
-\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}}-\mu_\beta\beta\\\\
+\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\\\
 \frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
@@
 
-A mathematical model would introduce parameters (e.g. $\kappa$, $\mu$ ,etc in (1)), which are to be determined to 
+A mathematical model will introduce parameters (e.g. $\kappa$, $\mu$ ,etc in (\ref{eq:1})), which are to be determined to 
 give a specific model using observed data. A traditional method to find the best desired parameter sets 
@@
 
-1. pyABC
+\begin{enumerate}
+	\item pyABC
+	\begin{itemize}
+		\item It has built-in support for multiple types of parallelisation, e.g., 
+		multicore processing and distributed cluster
 
-\quad $\bullet$ Built-in support for multiple types of  parallelisation
+		\item Stored runs can be resumed
 
-\quad $\bullet$ Built-in visualisation
+		\item Several built-in visualisation for the results are provided
+	\end{itemize}
 
-2. astroABC
+	\item astroABC
+	\begin{itemize}
+		\item It supports parallel sampling using MPI or multiprocessing
 
-\quad $\bullet$ Support parallel sampling using MPI or multiprocessing
+		\item Every run of the sampling can be paused and resumed later
+	\end{itemize}
 
-\quad $\bullet$ Each run can be paused and resumed later
+	\item ELFI
+	\begin{itemize}
+		\item It allows user to connect to \texttt{ipyparallel} for more parallelisation options
 
-3. ELFI
+		\item It uses directed acyclic graph as input model; a conversion is needed before implementation
+	\end{itemize}
 
-\quad $\bullet$ Connect to ipyparallel for more parallelisation options
+	\item ABC-SysBio
+	\begin{itemize}
+		\item It has experimental CUDA accelerated parallelisation support
+		\item It is specially designed for systems biology applications
 
-\quad $\bullet$ Using directed acyclic graph as model
+	\end{itemize}
 
-4. ABC-SysBio
+	\item EasyABC
+	\begin{itemize}
+		\item It supports multicore parallelisation
 
-\quad $\bullet$ Experimental CUDA accelerated parallelisation
+		\item Several different sequential algorithms are offered
+	\end{itemize}
 
-5. EasyABC
+	\item smfsb
+	\begin{itemize}
+		\item It provides simple and basic output of parameter estimations
+	\end{itemize}
 
-\quad $\bullet$ Multicore parallelisation
-
-\quad $\bullet$ Offer several different sequential algorithms
-
-6. smfsb
-
-\quad $\bullet$ Provide simple and basic output of parameter estimations.
+\end{enumerate}
 

### 2020-07-04 MSc-dissertation.tex
New

\documentclass[12pt,a4paper]{report}

\usepackage{epcc}
\usepackage{graphics}

% This example file shows how a thesis can be laid out using Latex. It
% does not use any special local features so should be portable to other
% places.
% 
% When producing draft copies of a thesis you may want to print only
% selected pages of the thesis. To do this use the command
% 
% dvips -f -p 4 -n 3 myfile.dvi | lpr
% 
% where -p 4 means start printing at page 4 (ie the page that will be
% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
% This example will print pages 4, 5 and 6.
% 
% If you want to print the thesis and also save paper you can print more
% than one page on each sheet of paper. Use the command
% 
% dvips -f myfile.dvi | psnup -2 | lpr
% 
% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
%
% To produce a PDF version you can create a PostScript copy first
%
% dvips -f myfile.dvi > myfile.pdf
%
% and then convert it
%
% distill myfile.ps
%
% or you can go straight to PDF
%
% pdflatex myfile
%
% Note that pdflatex expects all included figures to be in PDF too. See
% the includegraphics command below.

% This document contains many cross-references and forward references,
% eg in constructing a table of contents, so Latex may need to be run
% twice to get all the references correct. If you need to run Latex twice
% you may get the warning:
% 
% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.

% the following 4 lines are the content of the smallmargins.sty file
% but including them explicitly makes this more portable.
%AC%\oddsidemargin=0.1in
%AC%\topmargin=-0.5in
%AC%\textheight=9in
%AC%\textwidth=6.25in

%AC%\parskip 10pt
%AC%\parindent 0in

\begin{document}

%AC%\pagestyle{myheadings}
%AC%\markright{D.~S.~Henty}

%\title{A Latex thesis example}
%\author{D.~S.~Henty}
%\date{\today}

%\maketitle

\pagenumbering{roman}

\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
\author{Chaolin Han}
\date{\today}

\makeEPCCtitle

\thispagestyle{empty}

\vspace{11cm}

\begin{center}

\large{MSc in High Performance Computing}

\large{The University of Edinburgh}

\large{Year of Presentation: 2020}

\end{center}

\newpage

\begin{abstract}
This is the bit where you summarise what is in your thesis.
\end{abstract}

\pagenumbering{roman}

\tableofcontents
\listoftables
\listoffigures

\begin{titlepage}
\vspace*{2in}
% an acknowledgements section is completely optional but if you decide
% not to include it you should still include an empty {titlepage}
% environment as this initialises things like section and page numbering.
\section*{Acknowledgements}

This template is a slightly modified version of the one developed by
Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
acknowledgement follows:

{\em This template has been produced with help from many former students who
have shown different ways of doing things. Please make suggestions for
further improvements.}

\end{titlepage}

\pagenumbering{arabic}

\chapter{Introduction}
This should contain a description of your project and the problem you
are trying to solve. Where appropriate you should also include
references to work which has already been done on your topic and
anything else which lets you set your work in context.

One of the things you will need to do is to ensure that you have a
suitable list of references.  To do this you should see \cite{ref:lam}
or some other suitable reference.  Note the format of the citation used
here is the style favoured in this department.  Here is another
reference \cite{ref:bloggs} for good measure.

You will also want to make sure you have no spelling or grammatical
mistakes. To help idwentify spelling mistukes you caan use the commands
{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
that spelling mistakes are not the only errors which can occur. Spelling
checkers will not find errors which are, in fact, valid words such as
{\em there} for {\em their}, nor will they find repeated words which
sometimes occur if your concentration is broken when typing. {\bf There
is no substitute for thorough proof reading!}

\chapter{Background}

\section{Zebrafish spinal cord regeneration}
This is just to show how to break things into sections.

Some paragraphs in this demonstration document are here to provide some
padding so that sections last for more than one page to illustrate what
happens on subsequent pages. Note that the page numbering style is usually
different on the first page of a new chapter than on subsequent pages.

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.

\section{Mathematical modelling}
Some bits are hard.

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.

\subsection{Hard bits}
You might want to include an equation here:

\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
\label{equation:delsplit}
% note that the label is optional but allows you to refer to this
% equation later.
\end{equation}

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.

Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.

\subsection{Even harder bits}
This might be one of the places where you might want to refer to
equation \ref{equation:delsplit}. You will usually need to use the
Latex command twice to make cross-references like this work properly.
The cross-reference information is stored in the {\em .aux} file so don't
delete it.

\subsubsection{Numbering}
You can keep subdividing but eventually you get to a level where
numbering stops. This text is in a subsubsection which is not numbered
by default.

\paragraph{More on numbering:}

This text is in a paragraph which is also not numbered by default and
the ``title'' of the paragraph is not on a separate line.
If you want to increase the depth to which sections are numbered you
should see the section on setting the secnumdepth counter in the manual.

\section{Bayesian inference}

\chapter{Mathematical modelling}

You might sometimes want to include equations without numbering them.

\begin{displaymath}
E=mc^{2}
\end{displaymath}

You might also want to include diagrams.  The example shows the use of
the special command which allows existing postscript files to be
included.  You would normally keep your figures separate from the text. 
These pictures might be satellite images or postscript output from some
program such as IDL, PV-WAVE, Uniras or xpaint.

Below I create a figure which is centred and stretched to 30\% of the
width of the page \verb+{0.30\hsize}+ and with the height stretched by
the same amount \verb+{!}+ to preserve the aspect ratio. If you omit the
extension (ie .eps, .ps or .pdf) on the file name then latex will pick
up the postscript copy whereas pdflatex will automatically pick up the
PDF version.

\begin{figure}

\begin{center}
\resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
\end{center}

\caption{The University Crest}
\label{fig:eucrest}

\end{figure}

% see the man page for dvips for details of the special command which is
% much more powerful than is shown here. It allows offsets in the
% horizontal and vertical and scaling in x and y.

% choosing suitable values for offset and scale can be a tiring matter
% of trial and error.

% note that labels do not need to include a description of the object
% they are labelling but it can be helpful, eg \label{fig:figurename}.

You can use a label on a figure to refer to it later. The university
crest is in \ref{fig:eucrest}. Note that you should not use phrases like
``the figure above'' or ``the following figure'' since Latex may move
the figure relative to the text if it cannot be fitted onto the current page.

\chapter{Implementations and Experiments}

\section{Parameter estimation}
Here are some results.

\section{Model comparison}
Here are some results.

\section{Performance experiments}

The performance experiments are designed to explore the parallel performance of ABC SMC implementations. Usually ABC SMC is a time-consuming and computation intensive task and usually is executed on large clusters. The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.

First experiments are designed to illustrate the scaling-up performance. The program used here is a

% in practice you would probably keep this in a separate file and use
% the \include{filename} command to insert it here.

\chapter{Results Analysis}

\section{ABC SMC results}

\section{Performance experiments}

\section{Discussions}
This is the place to put your conclusions about your work. You can
split it into different sections if appropriate. You may want to include
a section of future work which could be carried out to continue your
research.

\chapter{Future works}

\chapter{Conclusions}

\appendix
% the appendix command just changes heading styles for appendices.

\chapter{Stuff which is too detailed}

Appendices should contain all the material which is considered too
detailed to be included in the main bod but which is, nevertheless,
important enough to be included in the thesis.

\chapter{Stuff which no-one will read}

Some people include in their thesis a lot of detail, particularly
computer code, which no-one will ever read. You should be careful that
anything like this you include should contain some element of uniqueness
which justifies its inclusion.

\begin{thebibliography}{100}

\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
and Reference Manual.} Addison Wesley. pp242.

\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
in Environments} Int. Journal of Silly Findings. pp 23-29.

\end{thebibliography}

\end{document}

### 2020-07-08 MSc-dissertation.tex
New

@@
 
-First experiments are designed to illustrate the scaling-up performance. The program used here is a
+First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
 
-% in practice you would probably keep this in a separate file and use
-% the \include{filename} command to insert it here.
+\begin{itemize}
+    \item Prior distribution: default to log-uniform distribution [$1\times 10^{-6}$, 50] for all the 12 parameters
+    \item threshold schedule: median epsilon
+    \item No factors, no adaptive distance or adaptive population applied
+    \item Population size is 2000, with 20 generations
+\end{itemize}
+
+HOW PYABC parallelise the sampling
+
+For HPC systems like Cirrus, \verb|pyabc| uses \verb|multiprocessing| for multi-core parallel sampling.
+By default if the number of cores is not specified, it will automatically read the number of available cores and use them all.
+Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
+
+The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54, 72 cores respectively.
+Each run is repeated 5 times.
+The average execution time, required sampling numbers are recorded.
+Hyperthreading is enabled when using 54 and 72 cores.
+The access to the node that contains computation cores is exclusive, such that the execution would not be affected by other programs of operations.
+
+The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the must time-consuming part.
+The rest part of the program is mostly not parallelised, e.g.
+database I/O and reductions operations.
+The sampling process involves sampling, perturbation and test of the acceptance criteria, all of which are computation-intensive.
+
+In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources is limited[REF].
+The performance experiment result could provide a reference that illustrate that how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
+
+The performance could also be analysed given a profiling report.
+The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck.
+according to which we could
 

### 2020-07-10 MSc-dissertation.tex
New para

@@
 \section{Zebrafish spinal cord regeneration}
-This is just to show how to break things into sections.
-
-Some paragraphs in this demonstration document are here to provide some
-padding so that sections last for more than one page to illustrate what
-happens on subsequent pages. Note that the page numbering style is usually
-different on the first page of a new chapter than on subsequent pages.
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
-rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.
 
 \section{Mathematical modelling}
-Some bits are hard.
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
-rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.
 
@@
 \end{equation}
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
-rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.
-
-Here is a padding paragraph.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb. 
-Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet
-more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.  Rhubarb. 
-More rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.   Rhubarb.  More
-rhubarb.  Yet more rhubarb.   Rhubarb.  More rhubarb.  Yet more
-rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.    Rhubarb.  More
-rhubarb.  Yet more rhubarb.  Rhubarb.  More rhubarb.  Yet more rhubarb.
 
@@
 
+\section{Software tools}
+
 \chapter{Mathematical modelling}
 
-You might sometimes want to include equations without numbering them.
+Mathematical models can describe different kinds of dynamic system, and can be used as a guide to prediction and analysis.
+An ideal model in this case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  trend against time.
 
-\begin{displaymath}
-E=mc^{2}
-\end{displaymath}
+Models for the regeneration process is in the form of ordinary differential equations, specifically the time differential form.
+Terms in the ODE are mostly explainable and corresponds interaction paths.
 
-You might also want to include diagrams.  The example shows the use of
-the special command which allows existing postscript files to be
-included.  You would normally keep your figures separate from the text. 
-These pictures might be satellite images or postscript output from some
-program such as IDL, PV-WAVE, Uniras or xpaint.
+\paragraph{Observed data}
 
-Below I create a figure which is centred and stretched to 30\% of the
-width of the page \verb+{0.30\hsize}+ and with the height stretched by
-the same amount \verb+{!}+ to preserve the aspect ratio. If you omit the
-extension (ie .eps, .ps or .pdf) on the file name then latex will pick
-up the postscript copy whereas pdflatex will automatically pick up the
-PDF version.
+Our models is built regarding the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
+The measurement includes the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
+As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
+According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
+$N$ and $\Phi$ is of the unit 'number of cells', $\beta$ and $\alpha$ is of the unit 'relative concentration'.
+
+It is noted that the variance of the measured data is relatively high.
+
+our hypothesis
+
+\section{Basic model}
+
+\section{Alternative models}
 
@@
 
+\subsection{Scaling-up}
+
 First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
@@
 
+\subsection{Profiling}
+
 The performance could also be analysed given a profiling report.
-The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck.
-according to which we could
+The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck, according to which we could find the hot-spot of program and given possible suggestions on improving the performance.
+
+In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
 
@@
 
+\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N.
+and Feng, Y., 2018.
+Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
+Nature communications, 9(1), pp.1-17.
+
 \end{thebibliography}

### 2020-07-10 MSc-dissertation.tex
New para

@@
 
-\paragraph{Observed data}
+\section{Observed data}
 
@@
 According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit 'number of cells', $\beta$ and $\alpha$ is of the unit 'relative concentration'.
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration'.
 
 It is noted that the variance of the measured data is relatively high.
+The summary statistic used for the parameter estimations is mean of measurement, assuming that measured data is Gaussian-like distributed.
+The distribution of the measured data points is plotted and examined to see if the measurement mean could represent the distribution.
+The result is that at most time points the measurement values are Gaussian-distributed, some distributions are skewed.
+One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
+Mean value can summarise most data measures and thus is still used as the target summarised observed data.
 
-our hypothesis
+\section{Hypothesis and Models}
 
-\section{Basic model}
+\subsection{Basic model}
 
-\section{Alternative models}
+\subsection{Alternative models}
 
-\begin{figure}
+\subsection{Model evaluation and comparison}
 
-\begin{center}
-\resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
-\end{center}
+% \begin{figure}
 
-\caption{The University Crest}
-\label{fig:eucrest}
+% \begin{center}
+% \resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
+% \end{center}
 
-\end{figure}
+% \caption{The University Crest}
+% \label{fig:eucrest}
+
+% \end{figure}
 
@@
 
-You can use a label on a figure to refer to it later. The university
-crest is in \ref{fig:eucrest}. Note that you should not use phrases like
-``the figure above'' or ``the following figure'' since Latex may move
-the figure relative to the text if it cannot be fitted onto the current page.
+% You can use a label on a figure to refer to it later. The university
+% crest is in \ref{fig:eucrest}. Note that you should not use phrases like
+% ``the figure above'' or ``the following figure'' since Latex may move
+% the figure relative to the text if it cannot be fitted onto the current page.
 

### 2020-07-11 __latexindent_temp.tex
Model 1

@@
-\documentclass{article}
+\documentclass[12pt,a4paper]{report}
 
@@
 \usepackage{graphics}
-\usepackage{booktabs}
-\usepackage{mathrsfs}
+\usepackage{mathtools}
 
-\renewcommand{\labelenumii}{\theenumii}
-\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
-% this will make nested numbering all arabic numbers
-
-% This example file shows how a report can be laid out using Latex. It
+% This example file shows how a thesis can be laid out using Latex. It
 % does not use any special local features so should be portable to other
@@
 % 
-% If you want to print the report and also save paper you can print more
+% If you want to print the thesis and also save paper you can print more
 % than one page on each sheet of paper. Use the command
@@
 % 
-% LaTeX Warning: Label(s) may have changed. Rerun to get cross-references right.
+% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.
+
+% the following 4 lines are the content of the smallmargins.sty file
+% but including them explicitly makes this more portable.
+%AC%\oddsidemargin=0.1in
+%AC%\topmargin=-0.5in
+%AC%\textheight=9in
+%AC%\textwidth=6.25in
+
+%AC%\parskip 10pt
+%AC%\parindent 0in
 
@@
 
+%AC%\pagestyle{myheadings}
+%AC%\markright{D.~S.~Henty}
+
+%\title{A Latex thesis example}
+%\author{D.~S.~Henty}
+%\date{\today}
+
+%\maketitle
+
 \pagenumbering{roman}
 
-\title{Project Preparation Report\\Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
+\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
 \author{Chaolin Han}
@@
 
+\thispagestyle{empty}
+
+\vspace{11cm}
+
+\begin{center}
+
+\large{MSc in High Performance Computing}
+
+\large{The University of Edinburgh}
+
+\large{Year of Presentation: 2020}
+
+\end{center}
+
 \newpage
 
+\begin{abstract}
+This is the bit where you summarise what is in your thesis.
+\end{abstract}
+
+\pagenumbering{roman}
+
 \tableofcontents
+\listoftables
+\listoffigures
 
-% \newpage
+\begin{titlepage}
+\vspace*{2in}
+% an acknowledgements section is completely optional but if you decide
+% not to include it you should still include an empty {titlepage}
+% environment as this initialises things like section and page numbering.
+\section*{Acknowledgements}
 
-% \section*{Acknowledgements}
+This template is a slightly modified version of the one developed by
+Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
+acknowledgement follows:
 
-% This template is a slightly modified version of the one developed by
-% Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
-% acknowledgement follows:
+{\em This template has been produced with help from many former students who
+have shown different ways of doing things. Please make suggestions for
+further improvements.}
 
-% {\em This template has been produced with help from many former students who
-% have shown different ways of doing things. Please make suggestions for
-% further improvements.}
+\end{titlepage}
 
-% \section*{Note}
-
-% This template was originally created for a full MSc dissertation.
-% Although I have modified the format slightly (eg there are no separate
-% chapters), I have not substantially changed the text or the advice
-% appearing in the \LaTeX comments (source lines beginning with \%). You
-% should therefore be aware that for a shorter report you may
-% not need to go into as much detail as indicated here. For example, you
-% might not need to include any acknowledgements and any reference list is
-% likely to be very short.
-
-\newpage
 \pagenumbering{arabic}
 
-\section{Introduction}
+\chapter{Introduction}
+This should contain a description of your project and the problem you
+are trying to solve. Where appropriate you should also include
+references to work which has already been done on your topic and
+anything else which lets you set your work in context.
 
-This report focuses on the work in the preparation phase of 
-MSc High-performance Computing with Data Analysis Dissertation Project. 
-The project is built around 
-a biological dynamical systems model with parameter estimation and model comparison 
-being the
-primary tasks. The complex interactions between several cells and molecules
-in zebrafish spinal cord regeneration is the modelling target.
+One of the things you will need to do is to ensure that you have a
+suitable list of references.  To do this you should see \cite{ref:lam}
+or some other suitable reference.  Note the format of the citation used
+here is the style favoured in this department.  Here is another
+reference \cite{ref:bloggs} for good measure.
 
-Previous dissertation review section gives comments on the content 
-and presentation of a dissertation report. The background of this project is 
-introduced in detail regarding several aspects of the project with the motivation
-and related literature discussed. 
-The preliminary work includes background 
-research, literature review, tools and techniques research and their results 
-are illustrated. The final project 
-proposal and workplan breakdown are given with regards to the conclusion from 
-background research and discussion with supervisors, together with a 
-estimated project timeline. Risk analysis and possible 
-outline of dissertation report are also addressed.
+You will also want to make sure you have no spelling or grammatical
+mistakes. To help idwentify spelling mistukes you caan use the commands
+{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
+that spelling mistakes are not the only errors which can occur. Spelling
+checkers will not find errors which are, in fact, valid words such as
+{\em there} for {\em their}, nor will they find repeated words which
+sometimes occur if your concentration is broken when typing. {\bf There
+is no substitute for thorough proof reading!}
 
-The project aims at the parameter estimation and model comparison. Approximate 
-Bayesian computation (ABC) approaches will be used to estimate parameters of mathmatical
+\chapter{Background}
 
-\section{Previous Dissertation Review}
+\section{Zebrafish spinal cord regeneration}
 
-The reviewed previous dissertation is {\em Modelling Room Acoustics with a 
-GPU} by Jacob Josiah Webber, finished in August 2017.
-\subsection*{Summary}
+\section{Mathematical modelling}
 
-Finite-difference time-domain (FDTD) techniques are introduced to numerically solve an acoustics model.
-The sound will be affected by the room it is in.
-The solution of wave equations of the sound can be computationally approximated using FDTD.
-It is a traditional simulation job for GPU.
-To improve the performance, the author proposed a ‘bounding box’ method, whose accompanied data structure significantly reduced memory usage and resultant simulation computation time is saved.
+\subsection{Hard bits}
+You might want to include an equation here:
 
-\subsection*{Does the dissertation adequately explain the context in which the problem is set?}
+\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
+\label{equation:delsplit}
+% note that the label is optional but allows you to refer to this
+% equation later.
+\end{equation}
 
-The motivation is addressed in Chapter 1 (Introductions) and the first part of Chapter 2.
-It briefly introduced together with its performance concerns.
-More details could be added to describe the meaning of this modelling mission and performance optimisation mission with relevant application or implementation examples given to clarify it.
+\subsection{Even harder bits}
+This might be one of the places where you might want to refer to
+equation \ref{equation:delsplit}. You will usually need to use the
+Latex command twice to make cross-references like this work properly.
+The cross-reference information is stored in the {\em .aux} file so don't
+delete it.
 
-The background and literature review is represented adequately in details.
-Section 2.1 derived formulas about acoustics wave, applied finite-difference on them and obtained an iterative propagation computation equation (Equation 2.15) which is the operation o the numerical modelling.
-Related boundary conditions and stencilling operations are also properly described.
-Later sections (2.3 and 2.4) addressed the GPU-based parallelisation and FDTD performance issues in the acoustics modelling context.
+\subsubsection{Numbering}
+You can keep subdividing but eventually you get to a level where
+numbering stops. This text is in a subsubsection which is not numbered
+by default.
 
-The background and literature review section is well structured and addressed and explains the problem background in a way that readers without much-related background knowledge (e.g.
-acoustics, GPU structures) could understand it.
+\paragraph{More on numbering:}
 
-\subsection*{Has the student adequately explained the scope of the problem, the approach that they would take, how success would be measured and reflected on the outcomes?}
+This text is in a paragraph which is also not numbered by default and
+the ``title'' of the paragraph is not on a separate line.
+If you want to increase the depth to which sections are numbered you
+should see the section on setting the secnumdepth counter in the manual.
 
-The aim of the project is to propose a new method and improve the modelling performance.
-To address the benefits of the proposed method, the author gave both theoretical  (section 3.3)and experimental (section 5.2) analysis to prove its effectiveness.
-The software development process (section 4.2) and theoretical derivation and implementation of the ‘bounding box’ method (2.4.1 and 3.2) are reasonable and adequate.
+\section{Bayesian inference}
 
-The reproducibility is achieved by giving the specific hardware platforms (NVIDIA K20 GPU server) and its environment (Appendix B), as well as source code and building instructions.
-As the author addressed in section 5.3, the replication of the program should output a consistent timing results to ones in Table 5.1 and 5.2.
+\section{Software tools}
 
-In regard of the limitations, as stated in the conclusion section, ‘particular effort was made to avoid choosing rooms that were overly suited to the new method’; however more test case should be given to make the results more convincing.
-Another limitation has been stated in the introductory paragraphs of Chapter 4, where experimentation and results analysis are not fully explored due to lack of time.
+\chapter{Mathematical modelling}
 
-The overall outcome is satisfying as it matches the theoretical prediction and the improvement is significant compared to previous method, regardless of the above limitations.
+Mathematical models can describe different kinds of dynamic system, and can be used as a guide to prediction and analysis.
+An ideal model in this case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  trend against time.
 
-\subsection*{Quality of presentation}
-The language is accurate and descriptions of the theories, designs and implementations are clear to follow.
+Models for the regeneration process is in the form of ordinary differential equations, specifically the time differential form.
+Terms in the ODE are mostly explainable and corresponds interaction paths.
 
-The listed figures are informative. E.g., Figure 2.1 gives clear representation of the stencilling operations, and Figure 3.1 is a suitable example to illustrate the theories.
+\section{Observed data}
 
-The dissertation is well structured such that the process of the project and the works in each stage of the project are included.
-The tables and algorithm blocks are all represented in clear academical forms, so do their placement and alignment.
-The formulas and equations, algorithm blocks and figures are all consistently numbered and referred properly.
+Our models is built regarding the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
+The measurement includes the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
+As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
+According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration'.
 
-The references are managed in a consistent manner and are correctly referred when needed.
-The appendices include original unsummarised data, code and GPU device information which provide meaning supplement when readers want to check some details not mentioned in the main body.
-However some other content should also be presented here, e.g.
-implementation environment, experiment input (boundary settings) and benchmarking tools/methods.
-
-\section{Background and Literature Review}
-
-Mathematical models are helpful in resolving complex interactions in real-life 
-dynamical systems. The project will use a mathematical model to represent 
-zebrafish spinal cord regeneration process, then determine its parameter 
-estimation and compare it with alternative models.
-
-Zebrafish is one kind of freshwater fish that is widely used in scientific 
-research for its regenerative abilities. Zebrafish can regenerate their spinal 
-cord after injury. Tsarouchas et al.\cite{ref:Tsarouchas} revealed how the immune system 
-responds to 
-allow axons of nerve cells growth across the lesion site. The experiments showed 
-evidence that several kinds of immune cells and cytokines play determinative 
-roles in regulating inflammation and repair. Cytokines are small proteins 
-secreted by certain cells to pass signals across cells. The regeneration 
-process is controlled by two cytokines, Il-1$\beta$ and Tnf-$\alpha$. Generally, 
-Il-1$\beta$ 
-inhibits the regeneration while Tnf-$\alpha$ promotes it, both of which are 
-dynamically controlled by mainly two kinds of immune cells (i.e. macrophage, 
-neutrophil). Based on these experimental results the initial 
-hypothesis is proposed that the repair process are dynamically controlled by 
-interactions between these cells and cytokines. Figure \ref{fig:1} illustrates the 
-basic interactions.
+It is noted that the variance of the measured data is relatively high.
+The summary statistic used for the parameter estimations is mean of measurement, assuming that measured data is Gaussian-like distributed.
+The distribution of the measured data points is plotted and examined to see if the measurement mean could represent the distribution.
+The result is that at most time points the measurement values are Gaussian-distributed, some distributions are skewed.
+One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
+Mean value can summarise most data measures and thus is still used as the target summarised observed data.
+A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}
 
@@
 \begin{center}
-\resizebox{0.60\hsize}{!}{\includegraphics{fig/fig1}}
+\resizebox{1.0\hsize}{!}{\includegraphics{fig/obs_curve.png}}
 \end{center}
 
-\caption{Interactions in zebrafish spinal cord regeneration, proposed by Tsarouchas 
-et al\cite{ref:Tsarouchas}. Lines ended with an arrow indicate promotion and lines 
-ended with a short segment indicate inhibition}
-\label{fig:1}
+\caption{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
+\label{fig:obs_data}
 
@@
 
-Mathematical models used to describe complex interactions of dynamical systems 
-are usually in the form of differential equations. Analytically solving differential 
-equations is not always feasible and the solution can be extremely complicated 
-especially for complex systems, so often dynamical system models are solved 
-numerically to generate data, which can be compared with real data and determine 
-model error. According to our hypothesis an initial ordinary differential equation 
-system can be built by describing the dynamics of number or concentration of immune 
-cells and cytokines. The initial equations shown in (\ref{eq:1}) are preliminary equations 
-that represent the dynamical model, which might be adjusted or calibrated according 
-to modelling results and real-world data.
+\section{Hypothesis and Models}
 
-In (\ref{eq:1}), $N$ and $\Phi$ represent the
-number of neutrophil and macrophage respectively, $\alpha$ and $\beta$ denotes the
-relative concentration of cytokines Tnf-$\alpha$ and Il-1$\beta$. 
-$\lambda$, $\kappa$, $\mu$, $s$, $i$ and $\nu$ with different subscripts are 
-all model parameters i.e. coefficient in the interactions.
+5 models in total are proposed according to different hypothesis.
+At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
+After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can be more close to the true process, recover more features of the observed data or test our hypothesis.
 
-\begin{equation} \label{eq:1}
-\begin{array}{l}
-\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\\\
-\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\\\
-\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}}-\mu_\beta\beta\\\\
+All these models assumes the interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for parameter estimation.
+Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data.
+
+\subsection{Basic model}
+
+A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
+
+\begin{equation*}\label{eq:pareto mle2}
+    \begin{multlined}
+        \frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
 \frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-\end{array}
-\end{equation}
+   \end{multlined}
+  \end{equation*}
 
-A mathematical model would introduce parameters (e.g. $\kappa$, $\mu$ ,etc in (1)), which are to be determined to 
-give a specific model using observed data. A traditional method to find the best desired parameter sets 
-is to apply an optimisation-based fitting, where the discrepancy of the model’s 
-predictions and experimental data is minimised. Alternatively, by adopting Bayesian 
-inference, we can focus on the experimental data and try to infer the parameter sets. 
-Many approaches have been proposed and used in the area of systems biology\cite{ref:Stumpf}. Compared 
-to optimisation-based fitting methods, Bayesian inference approaches can relieve the 
-concerns of over-fitting and local optima. It focuses on posterior distribution which 
-represents parameters’ ‘uncertainty’ given the observation data. As Bayes rule 
-stated, the posterior distribution $p(\theta|D)$ is proportional to product of likelihood and 
-prior
-\begin{equation} \label{eq:2}
-p(\theta|D)\propto l(\theta|D)p(\theta)
-\end{equation}
-where $\theta$ denotes parameter vector, $l$ is the likelihood function and $D$ is the observed data.
+\subsection{Alternative models}
 
-The prior distribution $p(\theta)$ is given to specify the initial estimates on 
-parameters before observing the data; the likelihood is given as conditional 
-probability $l(\theta|D)$ which indicates how well the observed data is explained by 
-the parameter $\theta$.
+\subsection{Model evaluation and comparison}
 
-% It can be hard to find analytical expression for likelihood, especially in large 
-% complex models; instead computational approaches such as Markov Chain Monte Carlo 
-% (MCMC)\cite{ref:Gilks} and sequential Monte Carlo (SMC)\cite{ref:Toni} are proposed to 
-% solve this problem. 
-% These approaches try to approximate the 
-% posterior distribution by simulating the data and they are known as approximate Bayesian 
-% computation (ABC). S.A. Sisson\cite{ref:Sisson} illustrated that sequential Monte Carlo sampler can 
-% be more powerful and efficient than MCMC and traditional ABC rejection methods.
+% \begin{figure}
 
-It can be hard to find an analytical expression for likelihood, especially in large 
-complex models. Computational approaches such as Markov Chain Monte Carlo\cite{ref:Gilks} techniques
-could be used 
-to solve this, where sampling methods are provided to computationally evaluate the likelihood.
-Approximate Bayesian computation (ABC) methods are alternative solutions. They are likelihood-free 
-methods that simulate data from model and approximate posterior distribution 
-instead of evaluating the likelihood. There are several ABC approaches e.g. ABC-rejection, ABC-MCMC
-and ABC-SMC (sequential Monte Carlo)\cite{ref:Toni}. S.A. Sisson\cite{ref:Sisson} suggested that sequential Monte Carlo sampler in ABC can 
-be more powerful and efficient than ABC-MCMC and traditional ABC rejection methods.
+% \begin{center}
+% \resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
+% \end{center}
 
-ABC approaches have been widely used in biology applications due to their 
-‘likelihood-free’ feature. In many systems biology models, the likelihood is 
-intractable, meaning that it cannot be defined as a function of the 
-parameters $\theta$ analytically. Liepe et al.\cite{ref:Liepe} showed a 
-successful application of 
-ABC-SMC in systems biology 
-with a software package ABC-SysBio, where ABC-SMC is used both in parameter estimation and model 
-selection.
+% \caption{The University Crest}
+% \label{fig:eucrest}
 
-Parameters of the zebrafish spinal cord regeneration model can be estimated 
-using ABC-SMC as it is hard to find exact likelihood here and an efficient 
-approximation approaches might help us in exploring the uncertainty in parameter 
-sets and comparing our model with alternative ones.
+% \end{figure}
 
@@
 
-\section{Preliminary Investigations}
+% You can use a label on a figure to refer to it later. The university
+% crest is in \ref{fig:eucrest}. Note that you should not use phrases like
+% ``the figure above'' or ``the following figure'' since Latex may move
+% the figure relative to the text if it cannot be fitted onto the current page.
 
-Preliminary works include background material study and tools research, which aims to build a concrete start of the project and explore the uncertainty and suitability of tools.
+\chapter{Implementations and Experiments}
 
-\subsection{Background Research}
+\section{Parameter estimation}
+Here are some results.
 
-The two key background papers\cite{ref:Tsarouchas,ref:Liepe} were studied firstly. 
-Tsarouchas et al.\cite{ref:Tsarouchas} illustrated 
-the biology process (zebrafish spinal cord regeneration after injury) that we want to 
-model and introduced the dynamic controls within the process in details. As 
-shown in the paper the process requires related biology 
-background knowledge, especially in immunology, thus related necessary background
-materials were studied to form a more comprehensive understanding and could help in 
-future analyse of results. The other background paper focuses on the algorithms that 
-are expected to be implemented in the project. More materials related to Bayesian 
-inference were studied, including tutorials, books and articles. Basic theories of 
-Bayesian inference and practical examples were studied with several ABC tutorial examples 
-practised to improve understanding. Other background materials include papers on 
-ABC-SMC algorithm, including the theory, implementation steps and comparison with 
-other ABC algorithms.
+\section{Model comparison}
+Here are some results.
 
-The mathematical models were also preliminarily studied, with concerns over model 
-building, data processing and solving. Ordinary differential equation computational 
-solution and its solver were reviewed in preparation for the future actual 
-implementation.
+\section{Performance experiments}
 
-The background research is the most important work in the preparation phase, 
-as it converted the first impression and interests of the project into concrete 
-knowledge in the biology and statistical inference basis such that the mission, 
-risks and challenges of the project could be discussed and analysed; it also 
-ensured an easier technically start of the project.
+The performance experiments are designed to explore the parallel performance of ABC SMC implementations.
+Usually ABC SMC is a time-consuming and computation intensive task and usually is executed on large clusters.
+The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
 
-\subsection{Tools Research}
-Besides the background research, tools that might be used in the project, mainly 
-the packages and libraries, were studied, compared and evaluated.
+\subsection{Scaling-up}
 
-There are currently varies of tools and softwares that implement ABC-SMC.
-A desirable tool should take many technical aspects into account, e.g.
-working environment, mathematical models solvers, parallelisation support and reliability.
+First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
 
-ABC-SMC is expected to be applied in Python or R languages, as most of the published packages are implemented on these two languages; further studies may use other languages if a more suitable software package for that is found or we need to build our own code for optimising.
-
-The studied packages are listed in Figure \ref{fig:2}. Their user guidance, 
-documentations (if exist) are viewed, from which their features are concluded 
-in a note (taskPackages.md under tasks directory in project repository);
-their main features are listed below.
-
-\begin{figure}
-
-	\begin{center}
-	\resizebox{1.00\hsize}{!}{\includegraphics{fig/fig2}}
-	\end{center}
-
-	\caption{Studied packages}
-	\label{fig:2}
-
-	\end{figure}
-
-1. pyABC
-
-\quad $\bullet$ Built-in support for multiple types of  parallelisation
-
-\quad $\bullet$ Built-in visualisation
-
-2. astroABC
-
-\quad $\bullet$ Support parallel sampling using MPI or multiprocessing
-
-\quad $\bullet$ Each run can be paused and resumed later
-
-3. ELFI
-
-\quad $\bullet$ Connect to ipyparallel for more parallelisation options
-
-\quad $\bullet$ Using directed acyclic graph as model
-
-4. ABC-SysBio
-
-\quad $\bullet$ Experimental CUDA accelerated parallelisation
-
-5. EasyABC
-
-\quad $\bullet$ Multicore parallelisation
-
-\quad $\bullet$ Offer several different sequential algorithms
-
-6. smfsb
-
-\quad $\bullet$ Provide simple and basic output of parameter estimations.
-
-Some examples given in their documentations have been practised to explore the details and features of ABC-SMC algorithm in these packages and determine which of them meet our requirements.
-The three preferred packages are pyABC, EasyABC and ELFI.
-All of them are well documented and support different kinds of parallelisation; other features include support for built-in visualisation, pause-and-restart and easy to use.
-The choice is limited in the preliminary practise and may be changed accordingly in the future practical implementation as more packages will be studied.
-
-Tools research provides grounded comprehension and conclusion on packages 
-choosing in this project and finds the initial tools to start with. More 
-tools will be studied and possibly used in the future.
-
-\section{Final Proposal}
-
-The project research consists of six parts.
-The first two parts are preparation work that would result in a mathematical model to implement our algorithms.
-Part 3, 4 and 5 are the practical session including several experiments and implementations to explore and analyse the model.
-The last part is to organise the works into report and dissertation.
-
-\subsection{Background Research and Project Preparation}
-
-This part focus on the preparation of the project, including background reading, tools research and regular discussion with supervisors.
-It will build a concrete ground to start the later tasks.
-Most of the work have already been done.
-
-\subsection{Dynamical System Modelling}
-
-Non-linear differential equations will be adapted to represent the dynamic system of zebrafish spinal cord regeneration process.
-Under the hypothesis1, there are complex interactions between two types of cytokines and three types of cells at the wounded site and the expected mathematical model representing by ODE systems will be built to describe these interactions.
-
-\subsection{Parameter Estimating}
-
-The parameters in the ODE system will be estimated using experimental data1.
-ABC-SMC will be applied to approximate the posterior distribution of parameters and produce reliable parameter estimates of the model.
-A distance function will be built to measure the discrepancy between the model and simulated data before applying the algorithm, together with coping with missing data in certain time points.
-
-As the SMC sampling process is computation-intensive and thus time-consuming, a parallel sampling technique will be of our interest.
-
-\subsection{Results Interpretation and Analysing}
-
-This part focuses on the evaluation of the parameterised model and analysis of the SMC results.
-Inference results will be tested using synthetic data and used to evaluate how well the model fits the data and how much confidence we have on the parameter estimates.
-
-\subsection{Model Comparison}
-
-ABC-SMC can also be implemented to compare our model with alternative ones.
-Similar procedures will be applied to model candidates and result in an approximation of each model’s marginal distribution, which could be the main criteria for our model selection.
-
-\subsection{Performance Analysis}
-
-The performance data of the sampling process will be analysed to reveal performance of the parallelisation implementation, including parallel speedup and efficiency.
-Trade-offs between cost, runtime and energy will also be discussed.
-
-\subsection{Dissertation and Report Writing}
-
-A project study report and final dissertation will be written to address all works and results in the project research process.
-
-\subsection{Estimated Computational Resources Usage}
-
-The computational resources usage estimation is a provisional rough estimation, 
-which can be changed as the implementation and performance test proceeds.
-
-\textbf{Cirrus}
-
-Most of the computational work in this project is expected to be deployed on Cirrus.
 \begin{itemize}
-	\item CPU hrs needed (standard compute nodes): 7200 CPUh (2 hrs per run $\times$ 
-	36 cores$\times$ 100 runs)
-	\item Disk storage needed: 10 GB, including data and softwares
-	\item Access to GPU compute nodes is also needed for possible parallelisation using CUDA
+    \item Prior distribution: default to log-uniform distribution [$1\times 10^{-6}$, 50] for all the 12 parameters
+    \item threshold schedule: median epsilon
+    \item No factors, no adaptive distance or adaptive population applied
+    \item Population size is 2000, with 20 generations
 \end{itemize}
 
-\section{Workplan}
+HOW PYABC parallelise the sampling
 
-Each part of the project research is planned as followed.
-The first stage is to prepare for the project in Semester 2 and possibly in the Easter holidays; the second stage will start at 18 May when the last examination ends.
+For HPC systems like Cirrus, \verb|pyabc| uses \verb|multiprocessing| for multi-core parallel sampling.
+By default if the number of cores is not specified, it will automatically read the number of available cores and use them all.
+Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
 
-\begin{enumerate}
-	\item Background Research and Project Preparation, Dynamical System Modelling
+The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54, 72 cores respectively.
+Each run is repeated 5 times.
+The average execution time, required sampling numbers are recorded.
+Hyperthreading is enabled when using 54 and 72 cores.
+The access to the node that contains computation cores is exclusive, such that the execution would not be affected by other programs of operations.
 
-	$\bullet$ Start time: 21/01/2020
+The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the must time-consuming part.
+The rest part of the program is mostly not parallelised, e.g.
+database I/O and reductions operations.
+The sampling process involves sampling, perturbation and test of the acceptance criteria, all of which are computation-intensive.
 
-	$\bullet$ Duration: 8 weeks
+In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources is limited[REF].
+The performance experiment result could provide a reference that illustrate that how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
 
-	$\bullet$ The later Easter holidays could be used to do more research
+\subsection{Profiling}
 
-	\item Parameterising
+The performance could also be analysed given a profiling report.
+The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck, according to which we could find the hot-spot of program and given possible suggestions on improving the performance.
 
-	$\bullet$ Start time: 18/05/2020
+In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
 
-	$\bullet$ Estimated duration: 7 weeks
+\chapter{Results Analysis}
 
-	\item Results Interpretation and Analysing
+\section{ABC SMC results}
 
-	$\bullet$ Start date: 22/06/2020
+\section{Performance experiments}
 
-	$\bullet$ Estimated duration: 5 weeks
+\section{Discussions}
+This is the place to put your conclusions about your work. You can
+split it into different sections if appropriate. You may want to include
+a section of future work which could be carried out to continue your
+research.
 
-	\item Model comparison
+\chapter{Future works}
 
-	$\bullet$ Start date: 15/06/2020
+\chapter{Conclusions}
 
-	$\bullet$ Estimated duration: 5 weeks
+\appendix
+% the appendix command just changes heading styles for appendices.
 
-	\item Performance Analysis
+\chapter{Stuff which is too detailed}
 
-	$\bullet$ Start date: 13/07/2020
+Appendices should contain all the material which is considered too
+detailed to be included in the main bod but which is, nevertheless,
+important enough to be included in the thesis.
 
-	$\bullet$ Estimated duration: 2 weeks
+\chapter{Stuff which no-one will read}
 
-	\item Dissertation and report writing
+Some people include in their thesis a lot of detail, particularly
+computer code, which no-one will ever read. You should be careful that
+anything like this you include should contain some element of uniqueness
+which justifies its inclusion.
 
-	$\bullet$ Start date: 20/07/2020
-
-	$\bullet$ Estimated duration: 5 weeks
-
-\end{enumerate}
-
-The Gantt chart is shown in Figure \ref{fig:3}. The two stages are separately listed.
-
-\begin{figure}
-
-	\begin{center}
-	\resizebox{0.80\hsize}{!}{\includegraphics{fig/fig3}}
-	\end{center}
-
-	\caption{Workplan Gantt chart}
-	\label{fig:3}
-
-	\end{figure}
-
-\section{Risk Analysis}
-
-Possible risks and challenges are addressed in this section.
-These risks are concluded from the preliminary researches.
-Though we tried to predict and evaluate all risks in the project, new risks may appear as the project proceeds.
-
-\subsection{Technical Problems}
-
-\textbf{Tools/software problems}
-
-\begin{itemize}
-	\item There could be unforeseen bugs in the code of the tools or software that are used in the project; softwares may crash, leading to possible data damaging. It is usually unpreventable.
-
-	Probability: moderate
-
-	Mitigation: in this case, we may need to find and fix the bug, switch tool/softwares and/or make a data protection plan to reduce the impact.
-
-	\item There could be problems in the implementation environment where the compatibility is our main concern.
-The implementation may require certain environment conditions (e.g.
-operating system, compiler version, dependencies, etc) to be performed, or they could be incompatible with the hardware (e.g.
-CUDA acceleration and multithreading).
-
-	Probability: high
-
-	Mitigation: we have studied a variety of packages (section 4.2) to ensure that we have alternatives when it is necessary to switch tools, and will take more tools into the backup plan.
-Another solution is to change or built the required environment if applicable, i.e.
-switching OS, compile environment or hardware platform.
-
-	\item The worst case could be that we may need to write our code to implement the ABC-SMC algorithm and complete the rest study, which would heavily increase the workload.
-
-	Probability: low
-
-	Mitigation: The workplan should be adjusted once it happens to ensure more time devoted to algorithm code and debug process. The rest work may be simplified.
-
-\end{itemize}
-
-\textbf{Hardware problems}
-
-\begin{itemize}
-	\item Hardware failures and hardware unavailability could intrude considerable impact on the project process.
-Hardware failure may cause data damaging; hardware unavailability may break the estimated workplan.
-
-	Probability: low
-
-	Mitigation: a regular data backup should be planned where experiments records, results and configurations should be well protected.
-In this case to proceed the project we may need to seek for hardware alternatives.
-\end{itemize}
-
-\subsection{Parallelisation Concerns}
-
-\begin{itemize}
-	\item The project seeks to implement a parallel sampling.
-The first risk could be that different tools use different parallel techniques targeting different hardwares, where the parallel performance comparison across hardware is a problem.
-
-	Probability: high
-
-	Mitigation: more study on the cross-platform benchmark should be conducted to give a reasonable comparison of performance.
-
-	\item Another risk can be the tricky parallel implementations, as there are different parallel systems with different configuration; it may cost time to find which kinds of implementation would work before we proceed to parallel performance analyse.
-
-	Probability: moderate
-
-	Mitigation: more investigation of the parallel implementations should be performed, which can reduce the effort in finding working schemes along with investigations in section 4.2.
-
-	\item The execution time of parameterising using ABC-SMC is mainly determined by problem size and sampling parameters (number of generations, tolerance sequence and population size); it may cost time to obtain a reasonable parameter input that avoids long execution time and in the meantime give precise results.
-Another trade-off we may face could be the common scalability trade-offs where a suitable scale-up is to be discussed.
-
-	Probability: high
-
-	Mitigation: It is usually unavoidable.
-To minimise the cost and effort a plan on the implementation experiment should be made and discussed where only necessary experiment should be conducted and redundant parameter space enumeration should be avoided.
-\end{itemize}
-
-\subsection{Modelling}
-
-\begin{itemize}
-	\item The main risk in the modelling process is model misspecification.
-
-	Probability: moderate
-
-	Mitigation: we may need to calibrate our model or adopt another model.
-These changes could consequently affect all later works in parameterising and analysis and the workplace should be adjusted accordingly.
-
-	\item There is also risks in experiment data processing: variables are measured in different time steps and data on gene expression uses different scales from the differential equations.
-These problems will introduce some difficulties in the model fitting.
-
-	Probability: high
-
-	Mitigation: the processing methods should be studied and backup plan should be made to enable a successful fitting.
-\end{itemize}
-
-\subsection{Other Risks}
-
-As far as this project concerns, certain cases of force majeure could affect the project progress which are low in probability.
-It should be noted that the current Coronavirus epidemic could affect some part of the project process, e.g.
-supervision meetings and access to certain physical resources.
-
-The project may face failure, e.g.
-undesirables results, critical technical problems and running out of time.
-These failures should be noticed as early as possible through the risk analysis and management throughout the project time and be avoided as much as we can.
-
-\section{Outline of the Dissertation Report}
-
-A possible outline of the Dissertation Report is structured as below.
-
-\begin{enumerate}
-	\item Introduction
-	\item Background and Literature Review
-	\begin{enumerate}
-		\item Zebrafish spinal cord regeneration
-		\item Mathematical modelling
-		\item Bayesian Inference
-	\end{enumerate}
-	\item Modelling
-	\item Design and Implementation
-	\begin{enumerate}
-		\item Parameter estimation
-		\item Model comparison
-		\item High-performance implementations
-	\end{enumerate}
-	\item Results and Analysing
-	\begin{enumerate}
-		\item ABC-SMC results
-		\item Performance analysis
-		\item Discussions
-	\end{enumerate}
-	\item Future Works
-	\item Conclusions
-\end{enumerate}
-
-\pagebreak
 \begin{thebibliography}{100}
 
-\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N. and Feng, Y., 
-2018. Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration. Nature communications, 9(1), pp.1-17.
+\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
+and Reference Manual.} Addison Wesley. pp242.
 
-\bibitem{ref:Stumpf}Stumpf, M., Balding, D.J. and Girolami, M., 
-2011. Handbook of statistical systems biology. John Wiley \& Sons.
+\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
+in Environments} Int. Journal of Silly Findings. pp 23-29.
 
-\bibitem{ref:Gilks}Gilks, W.R., Richardson, S. and Spiegelhalter, D., 
-1995. Markov chain Monte Carlo in practice. Chapman and Hall/CRC.
-
-\bibitem{ref:Toni}Toni, T., Welch, D., Strelkowa, N., Ipsen, A. and Stumpf, M.P., 
-2009. Approximate Bayesian computation scheme for parameter inference and model selection in dynamical systems. Journal of the Royal Society Interface, 6(31), pp.187-202.
-
-\bibitem{ref:Sisson}Sisson, S.A., Fan, Y. and Tanaka, M.M., 
-2007. Sequential monte carlo without likelihoods. Proceedings of the National Academy of Sciences, 104(6), pp.1760-1765.
-
-\bibitem{ref:Liepe}Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P. and Stumpf, M.P., 
-2014. A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation. Nature protocols, 9(2), p.439.
+\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N.
+and Feng, Y., 2018.
+Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
+Nature communications, 9(1), pp.1-17.
 

### 2020-07-12 MSc-dissertation.tex
New
Model 4 and 5

@@
 \usepackage{graphics}
+% \usepackage{mathtools}
+\usepackage{amsmath}
 
@@
 Mean value can summarise most data measures and thus is still used as the target summarised observed data.
+A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}
+
+\begin{figure}
+\begin{center}
+\resizebox{1.0\hsize}{!}{\includegraphics{fig/obs_curve.png}}
+\end{center}
+
+\caption{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
+\label{fig:obs_data}
+
+\end{figure}
 
@@
 
+5 models in total are proposed according to different hypothesis.
+At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
+After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can be more close to the true process, recover more features of the observed data or test our hypothesis.
+
+All these models assumes the interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for parameter estimation.
+Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data.
+
 \subsection{Basic model}
 
+A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
+A interaction map illustrate the model is shown in Figure \ref{fig:m1}.
+This model is a simplification a the process described in \cite{ref:Tsarouchas}.
+To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
+number of cells as `cell', and denote the unit of $\beta$ and $\aleph$ i.e.
+relative mRNA expression as `unit' for simplicity.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{0.4\hsize}{!}{\includegraphics{fig/model1.png}}
+    \end{center}
+
+    \caption{Interactions modelled in the basic model (model 1) based on Tsarouchas et al.\cite{ref:Tsarouchas}.
+Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
+    \label{fig:m1}
+
+\end{figure}
+
+It is assumed that there are negative feedbacks for all the variables. MORE DESCRIBE.
+
+\begin{align}
+    \label{eq:model1}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c c c|} 
+ \hline
+ Parameter & Definition & Units\\ [0.5ex] 
+ \hline\hline
+ $\lambda_N$ & Self-increase rate of neutrophil & $cell/h$  \\ 
+ $\kappa_{N\beta}$ & Promoting effect coefficient by il-1$\beta$ & $cell/(unit\cdotp h)$\\
+ $\mu_N$ & Coefficient of negative feedback of $N$ & $h^{-1}$ \\
+ $\nu_{N\Phi}$ & Coefficient of inhibition of both $N$ and $\Phi$& $cell^{-1}\cdotp h^{-1}$ \\
+ \hline
+ $\lambda_\Phi$ & Self-increase rate of macrophage & $cell/h$ \\
+ $\kappa_{\Phi\beta}$ & Promoting effect coefficient by il-1$\beta$ & $cell/(unit\cdotp h)$ \\
+ $\mu_\Phi$ &  Coefficient of negative feedback of $\Phi$ & $h^{-1}$ \\
+ \hline
+ $s_{\beta N}$ & Production rate from $N$ & $unit/(cell\cdotp h)$ \\
+ $i_{\beta\Phi}$ & Coefficient of inhibition to the production & $cell^{-1}$ \\
+ $\mu_\beta$ & Coefficient of negative feedback of $\beta$ & $h^{-1}$ \\
+ \hline
+ $s_{\alpha\Phi}$ & Production rate from $\Phi$ & $unit/(cell\cdotp h)$ \\
+ $\mu_\alpha$ & Coefficient of negative feedback of $\alpha$ & $h^{-1}$ \\
+[1ex] 
+ \hline
+\end{tabular}
+\caption{Parameters introduced in the basic model (model 1)}
+\label{table:m1}
+\end{table}
+
 \subsection{Alternative models}
+
+\paragraph{Model 2 and model 3}
+
+As the observed data indicates, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
+Regarding this, the self-increase term $\lambda$ cannot be constant, thus a exponentially decay $\lambda$ term is introduced and model 2 is proposed as Eqn.
+\ref{eq:model2}.
+The inhibition of il-1$\beta$ being produced by neutrophil, i.e.
+$i_{\beta\Phi}$ is considered to be ignored, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
+This case corresponds to model 3, written as Eqn.
+\ref{eq:model3}.
+
+Model 2 and model 3 has one extra parameter $a$ which is a coefficient in the exponentially decay determining the decay speed, with the unit $h^{-1}$.
+
+\begin{align}
+    \label{eq:model2}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{align}
+    \label{eq:model3}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+Model 3 can be regarded as a simplification of model 2, as it can be regarded as model 2 with parameter $i_{\beta\Phi}=0$.
+For these three models, model 1 is a naive one that is proposed at very first time and used as a `template' to build and test parameter inference framework.
+As the implementation is successful, model 2 and 3 is proposed.
+After fitting model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points which are close to steady states for neutrophil and macrophage.
+model 3 is a small simplification of model 2 and theoretically less general than model 2.
+
+\paragraph{Model 4 and model 5}
+
+After the first model selection experiment, it is found that some significant features presented in the observed data is not recovered by any of the model.
+According to that, attempts are tried to introduce more interactions within the dynamic system considering the both biological and  mathematical context.
+Extra promoting effect to the expression of tnf-$\alpha$ is considered, by either add a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or add a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
+\ref{eq:model4}) and model 5 (Eqn.
+\ref{eq:model5}).
+
+\begin{align}
+    \label{eq:model4}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha+d_{\beta\alpha}\beta
+    \end{split}
+\end{align}
+
+\begin{align}
+    \label{eq:model5}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=(s_{\alpha\Phi}+f_{\beta\alpha}\beta)\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c c|} 
+     \hline
+     Parameter & Definition & Units\\ [0.5ex] 
+     \hline\hline
+     $d_{\beta\alpha}$ & Coefficient of promoting effect from $\beta$ & $h^{-1}$  \\ 
+     \hline
+     $f_{\beta\alpha}$ & Coefficient of promoting the production of $\alpha$ by $\beta$ & $cell^{-1}\cdotp h^{-1}$  \\ 
+    [1ex] 
+     \hline
+    \end{tabular}
+    \caption{Parameters introduced in model 4 and 5}
+    \label{table:m45}
+\end{table}
 

### 2020-07-12 MSc-dissertation.tex
New
New para

@@
 
+\subsection{Limitations}
+
 % \begin{figure}
@@
 
+This chapter focuses on the implementation of the ABC SMC on the proposed models.
+Estimation of  parameters in the proposed models is the main task of this section.
+ABC SMC-based models selection is another task that can be easily implemented with the parameter estimation framework.
+
+The build and test is under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
+Some other code e.g.
+shell script and \verb|R| code are also used to perform the experiments and analyse the results.
+Tested software and system environment can be found at Appendix A.
+The build and test is performed on local computers and HPC clusters available within EPCC.
+
 \section{Parameter estimation}
-Here are some results.
+
+\subsection{ABC SMC and model hyperparameters}
+
+\subsection{Implementations on models}
 
 \section{Model comparison}
-Here are some results.
 
@@
 
-\chapter{Results Analysis}
+\chapter{Results and Discussions}
 
@@
 \section{Discussions}
-This is the place to put your conclusions about your work. You can
-split it into different sections if appropriate. You may want to include
-a section of future work which could be carried out to continue your
-research.
 
@@
 
-\chapter{Stuff which is too detailed}
+\chapter{System and software environment}
 
-Appendices should contain all the material which is considered too
-detailed to be included in the main bod but which is, nevertheless,
-important enough to be included in the thesis.
+\section{ABC SMC implementation}
+
+\section{Data analysis}
 
@@
 and Feng, Y., 2018.
-Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
+Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
 Nature communications, 9(1), pp.1-17.
+
+\bibitem{ref:pyabc} Klinger, E., Rickert, D. and Hasenauer, J., 2018. pyABC: distributed, likelihood-free inference. Bioinformatics, 34(20), pp.3591-3593.
 

### 2020-07-13 MSc-dissertation.tex
Skeleton added

@@
 
+[background and motivation]
+
+[what is done]
+
 \chapter{Background}
@@
 
+[describe the process and how cells and cytokines are involved]
+
 \section{Mathematical modelling}
 
-\subsection{Hard bits}
-You might want to include an equation here:
+[how this can be modelled as dynamic systems]
 
-\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
-\label{equation:delsplit}
-% note that the label is optional but allows you to refer to this
-% equation later.
-\end{equation}
-
-\subsection{Even harder bits}
-This might be one of the places where you might want to refer to
-equation \ref{equation:delsplit}. You will usually need to use the
-Latex command twice to make cross-references like this work properly.
-The cross-reference information is stored in the {\em .aux} file so don't
-delete it.
-
-\subsubsection{Numbering}
-You can keep subdividing but eventually you get to a level where
-numbering stops. This text is in a subsubsection which is not numbered
-by default.
-
-\paragraph{More on numbering:}
-
-This text is in a paragraph which is also not numbered by default and
-the ``title'' of the paragraph is not on a separate line.
-If you want to increase the depth to which sections are numbered you
-should see the section on setting the secnumdepth counter in the manual.
+[general topics of mathematical models: how to build model according to interactions; how to calibrate/ parameterise the model; how to solve the model; dynamics; how to evaluate the model]
 
@@
 
+[how to parameterise the model]
+
+[how the infer the parameter given data]
+
+[likelihood-free inference and information about ABC SMC]
+
+[model comparison, ABC SMC for model comparison]
+
 \section{Software tools}
+
+[existing tools and software for this task]
+
+[workflow and developing]
 
@@
 
+[feasible or not; goodness of fit; features of these models]
+
+[bayes factor for model selection]
+
 \subsection{Limitations}
+
+[hypothesis]
+
+[available data]
+
+[model misspecification]
 
@@
 
+[implementation details and options/ settings that are affecting the ABC]
+
 \subsection{ABC SMC and model hyperparameters}
+
+[including the options that are tested/ tried/ not adopted]
+
+[how do these options correspond to biological model and how to set them accordingly for proposed model]
+
+[experiments plan]
 
@@
 
+[ABC implementations details, e.g. distance, population, ODE related functions]
+
 \section{Model comparison}
+
+[comparison of model 1, 2 and 3]
+
+[further comparison of model 3, 4 and 5]
 
@@
 The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
+
+[MORE meanings of the performance study]
 
@@
 
-HOW PYABC parallelise the sampling
+[HOW PYABC parallelise the sampling]
 
@@
 
-The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54, 72 cores respectively.
+The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54 and 72 cores respectively.
 Each run is repeated 5 times.
@@
 
+[WHAT results is outputted and WHAT topics are discussed]
+
 \section{ABC SMC results}
+
+[inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]
+
+[model selections result; bayes factor]
 
@@
 
+[scaling-up performance: speed-up and efficiency]
+
+[profiling results: hot-spot and possible improvements]
+
 \section{Discussions}
+
+[goodness of fit (evaluation); preferred models and effects of modification to basic model]
+
+[efficiency and trade-offs in scaling-up]
+
+[POSSIBLE: compare to exact inference; compared to other implementations]
+
+[limitations]
 
@@
 
+[something not done as scheduled]
+
+[interesting topics to look inside]
+
+[moving to generalisations]
+
 \chapter{Conclusions}
+
+[what is studied]
+
+[to what extend]
+
+[how good is the result]
+
+[summary of the findings and advise]
+
+[thanks]
 

### 2020-07-15 MSc-dissertation.tex
New para

@@
 
+As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed model is firstly examined before actual inference using the experimental observed data.
+
+In the first part, synthetic data is used with known true parameter values.
+The algorithm implementation is evaluated by the efficiency and goodness of resultant model under different implementation options and ABC SMC hyperparameters.
+
+Then according to findings in the first part, several options with good performance are tried in the inference using experimental observed data (Figure \ref{fig:obs_data}).
+To obtain a more accurate and general results, these experiments are repeated for 3 or more times.
+
 \subsection{ABC SMC and model hyperparameters}
@@
 
+The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options. Among them the following is studied using synthetic data:
+
+\paragraph{Perturbation kernels}
+
+Perturbation kernels work in the sampling process.
+In each generation $t$, samples are firstly taken from the previous population $\{\theta_{t-1}\}$ with weights $\{w_{t-1}\}$, then perturbed using the perturbation kernel $K(\theta|\theta^*)$.
+We keep sampling until $N$ particles are accepted, where $N$ is the pre-set population size.
+After that the new weights are calculated and normalised.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
+    \end{center}
+
+    \caption{ABC SMC sampling process} 
+    \label{fig:smc}
+
+\end{figure}
+
+The perturbation kernel is called in the sampling of every particles and determines how the new perturbed particle is chosen, thus it is influential to both computation complexity and the resultant posterior distribution.
+Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations, or could generate a population with a higher acceptance rate if the successive epsilon values are close; A kernel with wide variance, or spreading out in a large space could help in resolving the local optimal problem by a more thoroughly exploration of the parameter space, however it can be more computation-intensive and result in a lower acceptance rate.
+A desired optimal kernel should balance the trade-offs; their property and criteria is discussed in \cite{ref:kernel}.
+
+There are several common choice of perturbation kernels. Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.
+
 [experiments plan]
 
-\subsection{Implementations on models}
+\subsection{Implementations and code}
 
 [ABC implementations details, e.g. distance, population, ODE related functions]
+
+[how the code is developed and built]
 
@@
 
+\bibitem{ref:kernel} Filippi, S., Barnes, C.
+P., Cornebise, J., and Stumpf, M..H.
+(2013).
+On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo.
+Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107
+
 \end{thebibliography}

### 2020-07-15 MSc-dissertation.tex
New para

@@
 
-\subsection{ABC SMC and model hyperparameters}
+\subsection{ABC SMC and model hyperparameters experiments}
 
 [including the options that are tested/ tried/ not adopted]
+
+[including experiments results and discussions]
 
@@
 
-\paragraph{Perturbation kernels}
+\subsubsection{Perturbation kernels}
 
@@
 
-There are several common choice of perturbation kernels. Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.
+There are several common choice of perturbation kernels.
+Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.
+A covariance matrix $\Sigma_t$ of accepted particles is calculated form previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$.
+It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in relecting the true posterior structure \cite{ref:kernel}.
+It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks.
+The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `narrower distributed' perturbation result.
+
+Local M-NN kernel provided by \verb|pyABC| is also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.
+
+The kernel experiments is designed to explore the efficiency of SMC on our dynamical systems.
+Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rate would be our preference.
+The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters.
+The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference.
+As the threshold schedule is fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit i.e.
+recover of target data trends/features and errors of the inferred parameters compared to true values is not discussed.
+
+\subsubsection{Kernel experiment results}
+
+The true value of parameters that are used to generate synthetic data is listed in Appendix B.1.
+These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data.
+The target data [FIGURE HERE] (synthetic data) is then generated using these parameter values via ODE solver in \verb|Python|.
+
+\paragraph{Adaptive functiosn and factors}
+
+\paragraph{Data size, prior range, population size and generations} 1
 
@@
 
-\chapter{Stuff which no-one will read}
+\chapter{Data and settings}
+
+\section{Infer-back experiments}
 
@@
 On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo.
-Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107
+Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107.
+
+\bibitem{ref:abcsysbio} Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P., and Stumpf, M.P.H., 2014.
+A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation.
+Nature Protocols.
+9 (2) pp.
+439–456.
+
+\bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C.
+2017.
+Comparing two sequentialMonte Carlo samplers for exact andapproximate Bayesian inference on biologicalmodels.J.
+R.
+Soc.
+Interface14: 20170340.
+
+\bibitem{ref:disease} Minter, A. and Retkute, R., 2019. Approximate Bayesian Computation for infectious disease modelling. Epidemics, 29, p.100368.
 

### 2020-07-16 MSc-dissertation.tex
New para

@@
 
-The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options. Among them the following is studied using synthetic data:
+The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options.
+To firstly test the ability of inference and observe the results of different options, synthetic data with known true parameter values are used as target data.
+The true value of parameters that are used to generate synthetic data is listed in Appendix B.1.
+These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data (Figure \ref{fig:infer_back_data}), the target data is then generated using these parameter values via ODE solver in \verb|Python|.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
+    \end{center}
+
+    \caption{Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot)} 
+    \label{fig:infer_back_data}
+
+\end{figure}
+
+Among them the following topics are studied using synthetic data.
 
@@
 \subsubsection{Kernel experiment results}
-
-The true value of parameters that are used to generate synthetic data is listed in Appendix B.1.
-These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data.
-The target data [FIGURE HERE] (synthetic data) is then generated using these parameter values via ODE solver in \verb|Python|.
 
@@
 
-Some people include in their thesis a lot of detail, particularly
-computer code, which no-one will ever read. You should be careful that
-anything like this you include should contain some element of uniqueness
-which justifies its inclusion.
+\subsection{Parameter values used to generate synthetic data}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c c|} 
+     \hline
+     Parameter & Value & Unit\\ [0.5ex] 
+     \hline\hline
+     $\lambda_N$ & 2.1989 & $cell/h$  \\ 
+     $\kappa_{N\beta}$ & 3.9627 & $cell/(unit\cdotp h)$\\
+     $\mu_N$ & 1.7219 & $h^{-1}$\\
+     $\nu_{N\Phi}$ & 0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
+     \hline
+     $\lambda_\Phi$ & 1.3146 & $cell/h$ \\
+     $\kappa_{\Phi\beta}$ & 0.1235 & $cell/(unit\cdotp h)$ \\
+     $\mu_\Phi$ & 0.1454 & $h^{-1}$ \\
+     \hline
+     $s_{\beta N}$ & 6.5536 & $unit/(cell\cdotp h)$ \\
+     $i_{\beta\Phi}$ & 1.7062 & $cell^{-1}$ \\
+     $\mu_\beta$ & 0.5212 & $h^{-1}$ \\
+     \hline
+     $s_{\alpha\Phi}$ & 10.2416 & $unit/(cell\cdotp h)$ \\
+     $\mu_\alpha$ & 19.6642 & $h^{-1}$ \\
+    [1ex] 
+     \hline
+    \end{tabular}
+    \caption{Parameter values}
+    \label{table:m1}
+\end{table}
 

### 2020-07-16 MSc-dissertation.tex
New pic

@@
 
-\caption{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
+\caption[Mean of the observed data]%
+{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
 \label{fig:obs_data}
@@
 
-    \caption{Interactions modelled in the basic model (model 1) based on Tsarouchas et al.\cite{ref:Tsarouchas}.
+    \caption[Interactions modelled in the basic model]%
+    {Interactions modelled in the basic model (model 1) based on Tsarouchas et al.\cite{ref:Tsarouchas}.
 Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
@@
 
-    \caption{Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot)} 
+    \caption[Synthetic data generated with known parameter values]%
+    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot)} 
     \label{fig:infer_back_data}
@@
 \subsubsection{Kernel experiment results}
+
+To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e. $\epsilon_t$ set to 10.0. From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel1.png}}
+    \end{center}
+
+    \caption[Total required samples of different kernels]%
+        {Total required samples of different kernels. Different color represents different generations (bottom to top: generation 1 to generation 20)}
+    \label{fig:kernel1}
+
+    \vspace*{\floatsep}
+
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
+    \end{center}
+
+    \caption{Acceptance rates of different kernels}
+    \label{fig:acceptance1}
+
+\end{figure}
+
+% \begin{figure}
+%     \begin{center}
+%     \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
+%     \end{center}
+
+%     \caption{Acceptance rates of different kernels}
+%     \label{fig:acceptance1}
+
+% \end{figure}
 

### 2020-07-16 MSc-dissertation.tex
New skeleton

@@
 
-To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e. $\epsilon_t$ set to 10.0. From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared
+To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e.
+$\epsilon_t$ set to 10.0.
+From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared.
+Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
+has higher acceptance rates among almost all generations. Local M-NN with M=750 is the slowest kernel.
+
+For local M-NN kernels, generally a greater M will lead to lower acceptance rate and 
+more required particles in each generation.
+As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve.
+Consequently, if greater M is test, then it will have even lower acceptance rates; the maximum M=2000 (whole population is considered) will have the lowest acceptance rates.
+
+Multivariate normal kernel has a performance between local M-NN M=250 and M=100.
+It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
+`Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
+Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1.
+Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
+
+It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
+This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
+local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold.
+In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
+Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}.
+One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
+
+\paragraph{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
+local M-NN with M$\leq 250$ (for population size of 2000, i.e.
+12.5\% of the population) and multivariate normal with scaling.
+Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.
 
@@
 
-\paragraph{Adaptive functiosn and factors}
+\subsubsection{Adaptive functiosn and factors}
 
-\paragraph{Data size, prior range, population size and generations} 1
+[how adaptive distance work]
+
+[what is factor]
+
+[what is adaptive distance]
+
+\subsubsection{Experiments results}
+
+\paragraph{Conclusion}
+
+\subsubsection{Data size, prior range, population size and generations}
 
 [experiments plan]
+
+\subsubsection{Experiments results}
+
+\paragraph{Conclusion}
 
@@
 
+\section{Kernel experiment: median epsilon schedule}
+
+\begin{figure}
+
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel2.png}}
+    \end{center}
+
+    \caption{Total sampling size} 
+    \label{fig:kernel2}
+
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance2.png}}
+        \end{center}
+
+        \caption{Acceptance rates} 
+        \label{fig:acceptance2}
+
+\end{figure}
+
 \begin{thebibliography}{100}

### 2020-07-17 MSc-dissertation.tex
New para

@@
 
-Multivariate normal kernel has a performance between local M-NN M=250 and M=100.
-It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
-`Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
-Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1.
-Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
-
-It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
-This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
-local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold.
-In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
-Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}.
-One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
-
-\paragraph{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
-local M-NN with M$\leq 250$ (for population size of 2000, i.e.
-12.5\% of the population) and multivariate normal with scaling.
-Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.
-
 \begin{figure}
@@
 
+Multivariate normal kernel has a performance between local M-NN M=250 and M=100.
+It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
+`Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
+Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1.
+Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
+
+It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
+This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
+local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold.
+In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
+Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}.
+One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
+
+\paragraph{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
+local M-NN with M$\leq 250$ (for population size of 2000, i.e.
+12.5\% of the population) and multivariate normal with scaling.
+Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.
+
 % \begin{figure}
@@
 
-\subsubsection{Adaptive functiosn and factors}
+\subsubsection{Adaptive functions and factors}
 
 [how adaptive distance work]
+
+SMC can be more adaptive by introducing adaptive distance function and adaptive population; besides, factors can be applied to manually `normalise' the data.
+
+The trivial distance function is set to Euclidean distance (2-norm distance)
+
+\begin{align}
+    \label{eq:dis}
+    D=\sqrt{\sum_i \Delta x_i^2}
+\end{align}
+
+where $i$ is the index of data points and $\Delta x_i$ is the discrepancy between  observed data and simulated data at data point $i$, i.e.
+$\Delta x_i = x_{i, simulated}-x_{i, observed}$.
+In this case data points are 12 time points of four variables i.e.
+48 data points in total.
+
+A weighted 2-norm distance can be written as
+
+\begin{align}
+    \label{dis_w}
+    D=\sqrt{\sum_i w_i \Delta x_i^2}
+\end{align}
+
+where a weight $w_i$ is assigned to every data point.
+If data point $i$ have a higher weight, then the data value is regeared to be more informative, i.e.
+giving more help in inferring the true posteriors; weights can be either pre-set according to prior knowledge of the problem, or using adaptive method to be dynamically calculated according to , known as adaptive distance.
+
+Adaptive distance is changing the weights of all data points after each generation iteration, trying to assign informative data points higher weights.
+Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}
+
+\begin{align}
+    \label{dis_w}
+    D=\sqrt{\sum_i f_iw_i \Delta x_i^2}
+\end{align}
+
+Factor is helpful as an option for efficiency.
+When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data point, or behave like a normalisation that can balance the scales of different summarised statistics (here is the 48 mean values).
+There are two appliance case studied: (1) factors used as an normalisation option ans (2) factors used to give more focus on some features of the observed data.
+For (2), as we noticed that the main features (e.g.
+rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points i.e.
+0 - 30 hpl, factors are tried in the later parameter estimation of real data where a more accurate fit of the curve is desired.
+
+Adaptive population \cite{ref:adpt_pop} can adapt the population size of each generation according to the mean correlation of variation error of the previous population.
+In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.
 
@@
 \subsubsection{Experiments results}
+
+[picture here]
+
+[what is the result]
 
@@
 
+\bibitem{ref:adpt_dis} Prangle, D., 2017. Adapting the ABC distance function. Bayesian Analysis, 12(1), pp.289-309.
+
+\bibitem{ref:adpt_pop} Klinger, E.
+and Hasenauer, J., 2017, September.
+A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo.
+In International Conference on Computational Methods in Systems Biology (pp.
+128-144).
+Springer, Cham.
+
 \end{thebibliography}

### 2020-07-17 __latexindent_temp.tex
New para

@@
 \usepackage{graphics}
-\usepackage{mathtools}
+% \usepackage{mathtools}
+\usepackage{amsmath}
 
@@
 \begin{figure}
-
 \begin{center}
@@
 A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
+A interaction map illustrate the model is shown in Figure \ref{fig:m1}.
+This model is a simplification a the process described in \cite{ref:Tsarouchas}.
+To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
+number of cells as `cell', and denote the unit of $\beta$ and $\aleph$ i.e.
+relative mRNA expression as `unit' for simplicity.
 
-\begin{equation*}\label{eq:pareto mle2}
-    \begin{multlined}
-        \frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
-\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-   \end{multlined}
-  \end{equation*}
+\begin{figure}
+    \begin{center}
+    \resizebox{0.4\hsize}{!}{\includegraphics{fig/model1.png}}
+    \end{center}
+
+    \caption[Interactions modelled in the basic model]%
+    {Interactions modelled in the basic model (model 1) based on Tsarouchas et al.\cite{ref:Tsarouchas}.
+Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
+    \label{fig:m1}
+
+\end{figure}
+
+It is assumed that there are negative feedbacks for all the variables. MORE DESCRIBE.
+
+\begin{align}
+    \label{eq:model1}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c c c|} 
+ \hline
+ Parameter & Definition & Units\\ [0.5ex] 
+ \hline\hline
+ $\lambda_N$ & Self-increase rate of neutrophil & $cell/h$  \\ 
+ $\kappa_{N\beta}$ & Promoting effect coefficient by il-1$\beta$ & $cell/(unit\cdotp h)$\\
+ $\mu_N$ & Coefficient of negative feedback of $N$ & $h^{-1}$ \\
+ $\nu_{N\Phi}$ & Coefficient of inhibition of both $N$ and $\Phi$& $cell^{-1}\cdotp h^{-1}$ \\
+ \hline
+ $\lambda_\Phi$ & Self-increase rate of macrophage & $cell/h$ \\
+ $\kappa_{\Phi\beta}$ & Promoting effect coefficient by il-1$\beta$ & $cell/(unit\cdotp h)$ \\
+ $\mu_\Phi$ &  Coefficient of negative feedback of $\Phi$ & $h^{-1}$ \\
+ \hline
+ $s_{\beta N}$ & Production rate from $N$ & $unit/(cell\cdotp h)$ \\
+ $i_{\beta\Phi}$ & Coefficient of inhibition to the production & $cell^{-1}$ \\
+ $\mu_\beta$ & Coefficient of negative feedback of $\beta$ & $h^{-1}$ \\
+ \hline
+ $s_{\alpha\Phi}$ & Production rate from $\Phi$ & $unit/(cell\cdotp h)$ \\
+ $\mu_\alpha$ & Coefficient of negative feedback of $\alpha$ & $h^{-1}$ \\
+[1ex] 
+ \hline
+\end{tabular}
+\caption{Parameters introduced in the basic model (model 1)}
+\label{table:m1}
+\end{table}
 
@@
 
+\paragraph{Model 2 and model 3}
+
+As the observed data indicates, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
+Regarding this, the self-increase term $\lambda$ cannot be constant, thus a exponentially decay $\lambda$ term is introduced and model 2 is proposed as Eqn.
+\ref{eq:model2}.
+The inhibition of il-1$\beta$ being produced by neutrophil, i.e.
+$i_{\beta\Phi}$ is considered to be ignored, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
+This case corresponds to model 3, written as Eqn.
+\ref{eq:model3}.
+
+Model 2 and model 3 has one extra parameter $a$ which is a coefficient in the exponentially decay determining the decay speed, with the unit $h^{-1}$.
+
+\begin{align}
+    \label{eq:model2}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{align}
+    \label{eq:model3}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+Model 3 can be regarded as a simplification of model 2, as it can be regarded as model 2 with parameter $i_{\beta\Phi}=0$.
+For these three models, model 1 is a naive one that is proposed at very first time and used as a `template' to build and test parameter inference framework.
+As the implementation is successful, model 2 and 3 is proposed.
+After fitting model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points which are close to steady states for neutrophil and macrophage.
+model 3 is a small simplification of model 2 and theoretically less general than model 2.
+
+\paragraph{Model 4 and model 5}
+
+After the first model selection experiment, it is found that some significant features presented in the observed data is not recovered by any of the model.
+According to that, attempts are tried to introduce more interactions within the dynamic system considering the both biological and  mathematical context.
+Extra promoting effect to the expression of tnf-$\alpha$ is considered, by either add a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or add a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
+\ref{eq:model4}) and model 5 (Eqn.
+\ref{eq:model5}).
+
+\begin{align}
+    \label{eq:model4}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha+d_{\beta\alpha}\beta
+    \end{split}
+\end{align}
+
+\begin{align}
+    \label{eq:model5}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=(s_{\alpha\Phi}+f_{\beta\alpha}\beta)\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c c|} 
+     \hline
+     Parameter & Definition & Units\\ [0.5ex] 
+     \hline\hline
+     $d_{\beta\alpha}$ & Coefficient of promoting effect from $\beta$ & $h^{-1}$  \\ 
+     \hline
+     $f_{\beta\alpha}$ & Coefficient of promoting the production of $\alpha$ by $\beta$ & $cell^{-1}\cdotp h^{-1}$  \\ 
+    [1ex] 
+     \hline
+    \end{tabular}
+    \caption{Parameters introduced in model 4 and 5}
+    \label{table:m45}
+\end{table}
+
 \subsection{Model evaluation and comparison}
+
+[feasible or not; goodness of fit; features of these models]
+
+[bayes factor for model selection]
+
+\subsection{Limitations}
+
+[hypothesis]
+
+[available data]
+
+[model misspecification]
 
@@
 
+This chapter focuses on the implementation of the ABC SMC on the proposed models.
+Estimation of  parameters in the proposed models is the main task of this section.
+ABC SMC-based models selection is another task that can be easily implemented with the parameter estimation framework.
+
+The build and test is under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
+Some other code e.g.
+shell script and \verb|R| code are also used to perform the experiments and analyse the results.
+Tested software and system environment can be found at Appendix A.
+The build and test is performed on local computers and HPC clusters available within EPCC.
+
 \section{Parameter estimation}
-Here are some results.
+
+[implementation details and options/ settings that are affecting the ABC]
+
+As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed model is firstly examined before actual inference using the experimental observed data.
+
+In the first part, synthetic data is used with known true parameter values.
+The algorithm implementation is evaluated by the efficiency and goodness of resultant model under different implementation options and ABC SMC hyperparameters.
+
+Then according to findings in the first part, several options with good performance are tried in the inference using experimental observed data (Figure \ref{fig:obs_data}).
+To obtain a more accurate and general results, these experiments are repeated for 3 or more times.
+
+\subsection{ABC SMC and model hyperparameters experiments}
+
+[including the options that are tested/ tried/ not adopted]
+
+[including experiments results and discussions]
+
+[how do these options correspond to biological model and how to set them accordingly for proposed model]
+
+The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options.
+To firstly test the ability of inference and observe the results of different options, synthetic data with known true parameter values are used as target data.
+The true value of parameters that are used to generate synthetic data is listed in Appendix B.1.
+These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data (Figure \ref{fig:infer_back_data}), the target data is then generated using these parameter values via ODE solver in \verb|Python|.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
+    \end{center}
+
+    \caption[Synthetic data generated with known parameter values]%
+    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot)} 
+    \label{fig:infer_back_data}
+
+\end{figure}
+
+Among them the following topics are studied using synthetic data.
+
+\subsubsection{Perturbation kernels}
+
+Perturbation kernels work in the sampling process.
+In each generation $t$, samples are firstly taken from the previous population $\{\theta_{t-1}\}$ with weights $\{w_{t-1}\}$, then perturbed using the perturbation kernel $K(\theta|\theta^*)$.
+We keep sampling until $N$ particles are accepted, where $N$ is the pre-set population size.
+After that the new weights are calculated and normalised.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
+    \end{center}
+
+    \caption{ABC SMC sampling process} 
+    \label{fig:smc}
+
+\end{figure}
+
+The perturbation kernel is called in the sampling of every particles and determines how the new perturbed particle is chosen, thus it is influential to both computation complexity and the resultant posterior distribution.
+Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations, or could generate a population with a higher acceptance rate if the successive epsilon values are close; A kernel with wide variance, or spreading out in a large space could help in resolving the local optimal problem by a more thoroughly exploration of the parameter space, however it can be more computation-intensive and result in a lower acceptance rate.
+A desired optimal kernel should balance the trade-offs; their property and criteria is discussed in \cite{ref:kernel}.
+
+There are several common choice of perturbation kernels.
+Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.
+A covariance matrix $\Sigma_t$ of accepted particles is calculated form previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$.
+It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in relecting the true posterior structure \cite{ref:kernel}.
+It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks.
+The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `narrower distributed' perturbation result.
+
+Local M-NN kernel provided by \verb|pyABC| is also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.
+
+The kernel experiments is designed to explore the efficiency of SMC on our dynamical systems.
+Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rate would be our preference.
+The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters.
+The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference.
+As the threshold schedule is fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit i.e.
+recover of target data trends/features and errors of the inferred parameters compared to true values is not discussed.
+
+\subsubsection{Kernel experiment results}
+
+To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e.
+$\epsilon_t$ set to 10.0.
+From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared.
+Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
+has higher acceptance rates among almost all generations. Local M-NN with M=750 is the slowest kernel.
+
+For local M-NN kernels, generally a greater M will lead to lower acceptance rate and 
+more required particles in each generation.
+As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve.
+Consequently, if greater M is test, then it will have even lower acceptance rates; the maximum M=2000 (whole population is considered) will have the lowest acceptance rates.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel1.png}}
+    \end{center}
+
+    \caption[Total required samples of different kernels]%
+        {Total required samples of different kernels. Different color represents different generations (bottom to top: generation 1 to generation 20)}
+    \label{fig:kernel1}
+
+    \vspace*{\floatsep}
+
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
+    \end{center}
+
+    \caption{Acceptance rates of different kernels}
+    \label{fig:acceptance1}
+
+\end{figure}
+
+Multivariate normal kernel has a performance between local M-NN M=250 and M=100.
+It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
+`Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
+Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1.
+Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
+
+It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
+This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
+local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold.
+In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
+Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}.
+One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
+
+\paragraph{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
+local M-NN with M$\leq 250$ (for population size of 2000, i.e.
+12.5\% of the population) and multivariate normal with scaling.
+Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.
+
+% \begin{figure}
+%     \begin{center}
+%     \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
+%     \end{center}
+
+%     \caption{Acceptance rates of different kernels}
+%     \label{fig:acceptance1}
+
+% \end{figure}
+
+\subsubsection{Adaptive functions and factors}
+
+[how adaptive distance work]
+
+SMC can be more adaptive by introducing adaptive distance function and adaptive population; besides, factors can be applied to manually `normalise' the data.
+
+The trivial distance function is set to Euclidean distance (2-norm distance)
+
+\begin{align}
+    \label{eq:dis}
+    D=\sqrt{\sum_i \Delta x_i^2}
+\end{align}
+
+where $i$ is the index of data points and $\Delta x_i$ is the discrepancy between  observed data and simulated data at data point $i$, i.e.
+$\Delta x_i = x_{i, simulated}-x_{i, observed}$.
+In this case data points are 12 time points of four variables i.e.
+48 data points in total.
+
+A weighted 2-norm distance can be written as
+
+\begin{align}
+    \label{dis_w}
+    D=\sqrt{\sum_i w_i \Delta x_i^2}
+\end{align}
+
+where a weight $w_i$ is assigned to every data point.
+If data point $i$ have a higher weight, then the data value is regeared to be more informative, i.e.
+giving more help in inferring the true posteriors; weights can be either pre-set according to prior knowledge of the problem, or using adaptive method to be dynamically calculated according to , known as adaptive distance.
+
+Adaptive distance is changing the weights of all data points after each generation iteration, trying to assign informative data points higher weights.
+Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}
+
+\begin{align}
+    \label{dis_w}
+    D=\sqrt{\sum_i f_iw_i \Delta x_i^2}
+\end{align}
+
+Factor is helpful as an option for efficiency.
+When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data point, or behave like a normalisation that can balance the scales of different summarised statistics (here is the 48 mean values).
+There are two appliance case studied: (1) factors used as an normalisation option ans (2) factors used to give more focus on some features of the observed data.
+For (2), as we noticed that the main features (e.g.
+rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points i.e.
+0 - 30 hpl, factors are tried in the later parameter estimation of real data where a more accurate fit of the curve is desired.
+
+Adaptive population \cite{ref:adpt_pop} can adapt the population size of each generation according to the
+
+[what is factor]
+
+[what is adaptive distance]
+
+\subsubsection{Experiments results}
+
+\paragraph{Conclusion}
+
+\subsubsection{Data size, prior range, population size and generations}
+
+[experiments plan]
+
+\subsubsection{Experiments results}
+
+\paragraph{Conclusion}
+
+\subsection{Implementations and code}
+
+[ABC implementations details, e.g. distance, population, ODE related functions]
+
+[how the code is developed and built]
 
 \section{Model comparison}
-Here are some results.
+
+[comparison of model 1, 2 and 3]
+
+[further comparison of model 3, 4 and 5]
 
@@
 
-\chapter{Results Analysis}
+\chapter{Results and Discussions}
+
+[WHAT results is outputted and WHAT topics are discussed]
 
@@
 
+[inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]
+
+[model selections result; bayes factor]
+
 \section{Performance experiments}
 
+[scaling-up performance: speed-up and efficiency]
+
+[profiling results: hot-spot and possible improvements]
+
 \section{Discussions}
-This is the place to put your conclusions about your work. You can
-split it into different sections if appropriate. You may want to include
-a section of future work which could be carried out to continue your
-research.
+
+[goodness of fit (evaluation); preferred models and effects of modification to basic model]
+
+[efficiency and trade-offs in scaling-up]
+
+[POSSIBLE: compare to exact inference; compared to other implementations]
+
+[limitations]
 
@@
 
-\chapter{Stuff which is too detailed}
+\chapter{System and software environment}
 
-Appendices should contain all the material which is considered too
-detailed to be included in the main bod but which is, nevertheless,
-important enough to be included in the thesis.
+\section{ABC SMC implementation}
 
-\chapter{Stuff which no-one will read}
+\section{Data analysis}
 
-Some people include in their thesis a lot of detail, particularly
-computer code, which no-one will ever read. You should be careful that
-anything like this you include should contain some element of uniqueness
-which justifies its inclusion.
+\chapter{Data and settings}
+
+\section{Infer-back experiments}
+
+\subsection{Parameter values used to generate synthetic data}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c c|} 
+     \hline
+     Parameter & Value & Unit\\ [0.5ex] 
+     \hline\hline
+     $\lambda_N$ & 2.1989 & $cell/h$  \\ 
+     $\kappa_{N\beta}$ & 3.9627 & $cell/(unit\cdotp h)$\\
+     $\mu_N$ & 1.7219 & $h^{-1}$\\
+     $\nu_{N\Phi}$ & 0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
+     \hline
+     $\lambda_\Phi$ & 1.3146 & $cell/h$ \\
+     $\kappa_{\Phi\beta}$ & 0.1235 & $cell/(unit\cdotp h)$ \\
+     $\mu_\Phi$ & 0.1454 & $h^{-1}$ \\
+     \hline
+     $s_{\beta N}$ & 6.5536 & $unit/(cell\cdotp h)$ \\
+     $i_{\beta\Phi}$ & 1.7062 & $cell^{-1}$ \\
+     $\mu_\beta$ & 0.5212 & $h^{-1}$ \\
+     \hline
+     $s_{\alpha\Phi}$ & 10.2416 & $unit/(cell\cdotp h)$ \\
+     $\mu_\alpha$ & 19.6642 & $h^{-1}$ \\
+    [1ex] 
+     \hline
+    \end{tabular}
+    \caption{Parameter values}
+    \label{table:m1}
+\end{table}
+
+\section{Kernel experiment: median epsilon schedule}
+
+\begin{figure}
+
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel2.png}}
+    \end{center}
+
+    \caption{Total sampling size} 
+    \label{fig:kernel2}
+
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance2.png}}
+        \end{center}
+
+        \caption{Acceptance rates} 
+        \label{fig:acceptance2}
+
+\end{figure}
 
@@
 and Feng, Y., 2018.
-Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
+Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
 Nature communications, 9(1), pp.1-17.
+
+\bibitem{ref:pyabc} Klinger, E., Rickert, D. and Hasenauer, J., 2018. pyABC: distributed, likelihood-free inference. Bioinformatics, 34(20), pp.3591-3593.
+
+\bibitem{ref:kernel} Filippi, S., Barnes, C.
+P., Cornebise, J., and Stumpf, M..H.
+(2013).
+On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo.
+Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107.
+
+\bibitem{ref:abcsysbio} Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P., and Stumpf, M.P.H., 2014.
+A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation.
+Nature Protocols.
+9 (2) pp.
+439–456.
+
+\bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C.
+2017.
+Comparing two sequentialMonte Carlo samplers for exact andapproximate Bayesian inference on biologicalmodels.J.
+R.
+Soc.
+Interface14: 20170340.
+
+\bibitem{ref:disease} Minter, A. and Retkute, R., 2019. Approximate Bayesian Computation for infectious disease modelling. Epidemics, 29, p.100368.
+
+\bibitem{ref:adpt_dis} Prangle, D., 2017. Adapting the ABC distance function. Bayesian Analysis, 12(1), pp.289-309.
+
+\bibitem{ref:adpt_pop} Klinger, E.
+and Hasenauer, J., 2017, September.
+A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo.
+In International Conference on Computational Methods in Systems Biology (pp.
+128-144).
+Springer, Cham.
 

### 2020-07-20 MSc-dissertation.tex
Nee pare

@@
 \chapter{Introduction}
-This should contain a description of your project and the problem you
-are trying to solve. Where appropriate you should also include
-references to work which has already been done on your topic and
-anything else which lets you set your work in context.
 
-One of the things you will need to do is to ensure that you have a
-suitable list of references.  To do this you should see \cite{ref:lam}
-or some other suitable reference.  Note the format of the citation used
-here is the style favoured in this department.  Here is another
-reference \cite{ref:bloggs} for good measure.
+% This should contain a description of your project and the problem you
+% are trying to solve. Where appropriate you should also include
+% references to work which has already been done on your topic and
+% anything else which lets you set your work in context.
 
-You will also want to make sure you have no spelling or grammatical
-mistakes. To help idwentify spelling mistukes you caan use the commands
-{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
-that spelling mistakes are not the only errors which can occur. Spelling
-checkers will not find errors which are, in fact, valid words such as
-{\em there} for {\em their}, nor will they find repeated words which
-sometimes occur if your concentration is broken when typing. {\bf There
-is no substitute for thorough proof reading!}
+% One of the things you will need to do is to ensure that you have a
+% suitable list of references.  To do this you should see \cite{ref:lam}
+% or some other suitable reference.  Note the format of the citation used
+% here is the style favoured in this department.  Here is another
+% reference \cite{ref:bloggs} for good measure.
+
+% You will also want to make sure you have no spelling or grammatical
+% mistakes. To help idwentify spelling mistukes you caan use the commands
+% {\em ispell} or {\em spell}. See the appropriate manual pages. Remember
+% that spelling mistakes are not the only errors which can occur. Spelling
+% checkers will not find errors which are, in fact, valid words such as
+% {\em there} for {\em their}, nor will they find repeated words which
+% sometimes occur if your concentration is broken when typing. {\bf There
+% is no substitute for thorough proof reading!}
 
@@
 
-[picture here]
+[FIGURE here]
 
-[what is the result]
+Two types of factors and adaptive distance function are tested by running the same number of generations (20 generation, 2000 particles per generation).
+From FIGURE, by applying factors the total required samples are less, and the adaptive distance requires much less samples; their acceptance rates are also higher than the standard one.
+It seems that adaptive function and factors could help to converge more quickly.
 
-\paragraph{Conclusion}
+However considering the result after 20 generation (FIGURE), the resultant model gives the contrary preference.
+Applying factors and using adaptive distance can lead the resultant model less accurate (after the same generation iterations): they have wider inter-quantile range and may requires more generations to converge.
 
-\subsubsection{Data size, prior range, population size and generations}
+\paragraph{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases.
+However, as the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, a direct compare of the efficiency under the same threshold schedule is unfeasible; after the same number of generations they can reach a small $\epsilon_t$ but still the approximated posterior is not as accurate as the standard implementation.
+AS a result, our implementation of the parameter inference will only consider standard distance function and factors are only tested for improvements (FIGURE).
+
+\subsubsection{Data size, prior distribution range, population size and generations}
 
@@
 
+Some other hyperparameters, e.g.
+feed-in data size, prior range, number of generations and number of particles in each generation is also of our interest.
+Experiments are designed to explore how these options affect the goodness of fit and efficiency of the implementation.
+
+Regarding the present dynamic systems model, the SMC is implemented with two data size options, two prior distribution (standard and wider prior range).
+The iteration options i.e.
+population size and number of generaitions are also tried with different values.
+These experiments intends to give suggestions on the later SMC implementation on experimental data.
+
 \subsubsection{Experiments results}
 
-\paragraph{Conclusion}
+The results from data size and prior distribution range is shown in FIGURE.
+The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement.
+Although much less (60\%) data is used, the total required sampling numbers is not much less (XX\%).
+Although the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data (FIGURE), cares should be taken when using less data point, or using other more summative statistics, e.g.
+mean, standard deviation, peak values, starting and ending values etc.
+As shown in \cite{ref:disease}, less data can lead to a model with more variance, under-fitting or missing some local features e.g.
+local peak.
+
+The prior range and distribution can largely affect the execution of SMC (FIGURE and FIGURE).
+Samples are taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space.
+It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.
+
+[discuss uniform and log uniform when result ready]
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/size1.png}}
+    \end{center}
+
+    \caption[Data size and prior distribution range experiment]%
+        {Data size and prior distribution range experiment. (a) Total required samples, (b) Epsilon values under median epsilon schedule}
+    \label{fig:size}
+
+\end{figure}
+
+[FIGURE HERE]
+
+\paragraph{Summary} This subsection aims to preliminarily explore ABC SMC settings using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options.
+Conclusion can be drawn to choose efficient and proper options of ABC SMC to be applied on the target models with experimental data.
+It also provides reference for a general inference task using SMC with high dimensional parameter space.
 
@@
     \end{tabular}
-    \caption{Parameter values}
+    \caption{Parameter values used for model 1}
     \label{table:m1}
@@
 
-\section{Kernel experiment: median epsilon schedule}
+\subsection{Kernel experiment: median epsilon schedule}
 
@@
 
+\subsection{Other experiment details}
+
 \begin{thebibliography}{100}
-
-\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
-and Reference Manual.} Addison Wesley. pp242.
-
-\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
-in Environments} Int. Journal of Silly Findings. pp 23-29.
 

### 2020-07-22 MSc-dissertation.tex
New para

@@
 \usepackage{graphics}
+\usepackage{hyperref}
 % \usepackage{mathtools}
@@
 
-\chapter{Implementations and Experiments}
+\chapter{Parameter estimation and model comparison}
 
@@
 The build and test is performed on local computers and HPC clusters available within EPCC.
+
+\section{Implementations and code}
+
+[ABC implementations details, e.g. distance, population, ODE related functions]
+
+According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| is adopted in this project.
+\verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}.
+\verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for inference and analysis tasks.
+Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.
+
+The code for the project is developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version is deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task.
+The scaling-up experiment is performed on Cirrus, the profiling and its analysis are performed on local machine.
+Some development and debug are on the remote machines to ensure experiments run correctly on compute node.
+
+The ODE model-related code and utilities are firstly developed to enable reusable functions, variables and data calls.
+It includes the code for the ODE models, ODE solver, data structures, data format transform functions etc.
+By doing this in the implementation code file we can set options and activate ABC SMC easily, without creating definition and deplicated code for the models and parameter priors.
+
+Some modifications are made to the source code of \verb|pyABC|.
+For example, the laboratorial measured data are not considered `complete' for modeling and analysis purpose: there are different measurement time points for cells and cytokines, e.g.
+cell numbers is not measured at 0.25 hpl and cytokines' expression is not measured at time point 120 hpl.
+These missing values are treated as `ignored' when calculating the distance between observed data and simulated data.
+To cope with this built-in distance function code is modified.
+
+Analysis code is also wrapped up into functions which can be easily called after an ABC SMC run is finished.
+They includes multiple visualisations and summary statistic of the results.
+For different tasks there are also separated analysis code file.
+Other analysis tools used in the project include built-in database visualisation tool, R code and Microsoft Excel.
+
+The project code is managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also make the cross machine synchronisation of code versions convenient.
+
+After several development iterations, the code files for ABC SMC implementation now mainly has two types: code file for infer-back experiments (synthetic data), code for parameter inference (experimental data).
+We can just add more models in these code files for a model comparison task.
+
+[how the code is developed and built]
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
     \begin{center}
@@
 
-\subsection{Implementations and code}
-
-[ABC implementations details, e.g. distance, population, ODE related functions]
-
-[how the code is developed and built]
-
 \section{Model comparison}
@@
 
+\subsection{Local machine}
+
+Local development machine is a Mac laptop, running on macOS 10.15.6. The environment of the development is listed in Table \ref{table:local_macine}.
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c|} 
+     \hline
+     Environment & Version \\ [0.5ex] 
+     \hline\hline
+     $\lambda_N$ & 2.1989  \\ 
+     $\kappa_{N\beta}$ & 3.9627  \\
+     $\mu_N$ & 1.7219 \\
+     $\nu_{N\Phi}$  0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
+     \hline
+     $\lambda_\Phi$ & 1.3146  $cell/h$ \\
+     $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$ \\
+     $\mu_\Phi$  0.1454 & $h^{-1}$ \\
+     \hline
+     $s_{\beta N}$ &6.5536  $unit/(cell\cdotp h)$ \\
+     $i_{\beta\Phi}$  1.7062 & $cell^{-1}$ \\
+     $\mu_\beta$ &0.5212  $h^{-1}$ \\
+     \hline
+     $s_{\alpha\Phi}$  10.2416 & $unit/(cell\cdotp h)$ \\
+     $\mu_\alpha$ 19.6642 & $h^{-1}$ \\
+    [1ex] 
+     \hline
+    \end{tabular}
+    \caption{Environment on local machine}
+    \label{table:local_macine}
+\end{table}
+
+\subsection{Remote machine}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c|} 
+     \hline
+     Environment & Version \\ [0.5ex] 
+     \hline\hline
+     $\lambda_N$ & 2.1989  \\ 
+     $\kappa_{N\beta}$ & 3.9627  \\
+     $\mu_N$ & 1.7219 \\
+     $\nu_{N\Phi}$  0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
+     \hline
+     $\lambda_\Phi$ & 1.3146  $cell/h$ \\
+     $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$ \\
+     $\mu_\Phi$  0.1454 & $h^{-1}$ \\
+     \hline
+     $s_{\beta N}$ &6.5536  $unit/(cell\cdotp h)$ \\
+     $i_{\beta\Phi}$  1.7062 & $cell^{-1}$ \\
+     $\mu_\beta$ &0.5212  $h^{-1}$ \\
+     \hline
+     $s_{\alpha\Phi}$  10.2416 & $unit/(cell\cdotp h)$ \\
+     $\mu_\alpha$ 19.6642 & $h^{-1}$ \\
+    [1ex] 
+     \hline
+    \end{tabular}
+    \caption{Environment on remote machine}
+    \label{table:remote_macine}
+\end{table}
+
 \section{Data analysis}

### 2020-07-22 MSc-dissertation.tex
New structure

@@
 
-\section{Parameter estimation}
+% \section{Parameter estimation}
 
-[implementation details and options/ settings that are affecting the ABC]
+% [implementation details and options/ settings that are affecting the ABC]
+
+\section{ABC SMC and model hyperparameters experiments}
 
@@
 To obtain a more accurate and general results, these experiments are repeated for 3 or more times.
-
-\subsection{ABC SMC and model hyperparameters experiments}
 
@@
 
-\subsubsection{Perturbation kernels}
+\subsection{Perturbation kernels}
 
@@
 
-\paragraph{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
+\subsubsection{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
 local M-NN with M$\leq 250$ (for population size of 2000, i.e.
@@
 
-\subsubsection{Adaptive functions and factors}
+\subsection{Adaptive functions and factors}
 
@@
 
-\paragraph{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases.
+\subsubsection{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases.
 However, as the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, a direct compare of the efficiency under the same threshold schedule is unfeasible; after the same number of generations they can reach a small $\epsilon_t$ but still the approximated posterior is not as accurate as the standard implementation.
@@
 
-\subsubsection{Data size, prior distribution range, population size and generations}
+\subsection{Data size, prior distribution range, population size and generations}
 
@@
 
-\paragraph{Summary} This subsection aims to preliminarily explore ABC SMC settings using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options.
+\subsubsection{Summary} This subsection aims to preliminarily explore ABC SMC settings using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options.
 Conclusion can be drawn to choose efficient and proper options of ABC SMC to be applied on the target models with experimental data.
@@
 
-\section{Model comparison}
+\section{Parameter estimation and model comparison}
 
-[comparison of model 1, 2 and 3]
+[separated ABC run on model 1, 2 and 3]
+
+[model comparison among model 1, 2, 3]
+
+[features that derive model 4 and 5]
 
@@
 
-\section{Performance experiments}
+\chapter{Performance experiments}
 
@@
 
-\subsection{Scaling-up}
+\section{Scaling-up}
 
@@
 
-\subsection{Profiling}
+\subsubsection{Results}
+
+[scaling-up performance: speed-up and efficiency]
+
+\section{Profiling}
 
@@
 In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
+
+\subsubsection{Results}
+
+[profiling results: hot-spot and possible improvements]
 
@@
 [model selections result; bayes factor]
-
-\section{Performance experiments}
-
-[scaling-up performance: speed-up and efficiency]
-
-[profiling results: hot-spot and possible improvements]
 

### 2020-07-23 MSc-dissertation.tex
New fig

@@
 
+Using the suggested options in the previous section, the target data (Figure \ref{fig:obs_data}) is prepared as input for the SMC inference framework.
+Regarding the result, the population size and number of generations are adjusted several times to obtain an general informative posterior which should be stable across repeated runs and avoid local optimal as much as possible.
+
+Parameter estimation of model 1, 2 and 3 was tried with different prior distribution setting: distribution range [0, 25] and [0, 75], amd distribution type uniform and log-uniform.
+All these ABC SMC runs uses 2,000 particles per population and 30 generations.
+
+Log-uniform-shaped prior seems to give a better fit of the model, and wider prior range [0, 75] does not give a more accurate result.
+Then distribution range [0, 25] was tried, in case some parameter have a true posterior laying between [25, 50].
+
+The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve123.png}}
+    \end{center}
+
+    \caption{Simulated data from the last population of model 1, 2 and 3}
+    \label{fig:result123}
+
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \resizebox{0.8\hsize}{!}{\includegraphics{fig/eps_acc123.png}}
+    \end{center}
+
+    \caption{Epsilon trends and acceptance rates of model 1, 2 and 3}
+    \label{fig:result123_2}
+
+\end{figure}
+
 [model comparison among model 1, 2, 3]

### 2020-07-25 MSc-dissertation.tex
New pic
New para

@@
 The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}.
+For model 3, the acceptance rates and epsilon path are presented in Figure \ref{fig:result123_2}.
+For model 3, the estimated parameters' posterior is shown in Figure \ref{fig:para1}.
+Figure \ref{fig:result123} plots the estimated curve of models, using the mean value of each parameter's approximated posterior; also 1000 particles (each particle is a parameter set) were sampled and used to generate simulated trajectories of the four variable, and the mean (blue) and 25th to 75th percentile range (grey) are calculated from the 1000 simulated data.
+
+Before applying model comparison methods, some features of the three model can still be compared.
+From simulated curves, It can be seen that all the inter-quartile ranges are tight around the mean value curve, which indicates the approximated posterior are concentrated to some degree; the peak-shaped posterior distributions (Figure \ref{fig:para1}) agrees with this.
+All epsilon values converges to a steady level, and model 3 has a lower final epsilon value and consequently the simulated trajectory is more close to the observed data (model 3, Figure \ref{fig:result123}).
+
+The acceptance rates is fluctuating and have a gradually increase trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated.
+Compared to model 1, model 2 gives a better fit of the decreasing trend at the second half of the observed data, by using a exponentially decaying
+ self-increase rate $\lambda_N$ instead of a constant.
+The resultant simulated data from model 2 and 3 are in the similar trends and the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'.
+We expect model 3 to be the best model as it reaches a threshold value and intuitively gives a better fit, although the fit could not be considered generally a good representation of the biological process that we want to model, as the simulated data are significantly biased from the observed data for tnf-$\alpha$.
+A sharp peak at 4 hpl is observed from the measurement (Figure \ref{fig:obs_data}) but no similar features are represented by any of the three models.
 
@@
 
-\end{figure}
-
-\begin{figure}
-    \begin{center}
-    \resizebox{0.8\hsize}{!}{\includegraphics{fig/eps_acc123.png}}
+    \vspace*{\floatsep}
+
+    \begin{center}
+        \resizebox{0.6\hsize}{!}{\includegraphics{fig/eps_acc123.png}}
         \end{center}
@@
 
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/para1234.png}}
+    \end{center}
+
+    \caption{Estimated posterior distribution of parameters in model 3}
+    \label{fig:para1}
+
+\end{figure}
+
 [model comparison among model 1, 2, 3]
+
+Next a model selection experiment with different prior was conducted, using the same ABC-SMC framework.
+As in the parameter inference, two distributions (uniform and log-uniform) are tested with three different interval ([0, 25], [0, 50] and [0, 75]).
+The results show that model 3 wins with nearly 100\% model probability at most runs after 30 generations, except log-uniform [0, 75].
+All the model probability results are of great discrepancy (e.g.
+100\% model 3 and 0\% for model 1 and 2), so Bayesian factor is not calculated.
+
+Figure \ref{fig:cmp1} shows the model selection process, under log-uniform [0, 50] prior distribution.
+Model 3 gains total advantage over model 1 and 2 after 7th population.
+Together with conclusions above, model 3 is considered to be the the best model to represent the target data so far.
+
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/cmp1.png}}
+    \end{center}
+
+    \caption{Model comparison of model 1, 2 and 3}
+    \label{fig:cmp1}
+
+\end{figure}
 

### 2020-07-25 MSc-dissertation.tex
New para

@@
 \usepackage{amsmath}
+\usepackage{float}
 
@@
 \begin{align}
-    \label{dis_w}
+    \label{dis_f}
     D=\sqrt{\sum_i f_iw_i \Delta x_i^2}
@@
 
+\subsection{Model 1, 2 and 3}
+
 [separated ABC run on model 1, 2 and 3]
@@
 The acceptance rates is fluctuating and have a gradually increase trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated.
-Compared to model 1, model 2 gives a better fit of the decreasing trend at the second half of the observed data, by using a exponentially decaying
- self-increase rate $\lambda_N$ instead of a constant.
+Compared to model 1, model 2 gives a better fit of the decreasing trend at the second half of the observed data, by using a exponentially decaying self-increase rate $\lambda_N$ instead of a constant.
 The resultant simulated data from model 2 and 3 are in the similar trends and the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'.
@@
 
-    \caption{Simulated data from the last population of model 1, 2 and 3}
+    \caption{Simulated trajectory from the last population of model 1, 2 and 3}
     \label{fig:result123}
@@
     \label{fig:para1}
+
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/cmp1.png}}
+        \end{center}
+
+        \caption{Model comparison of model 1, 2 and 3}
+        \label{fig:cmp1}
 
@@
 
+\begin{table}[t!]
+    \centering
+    \begin{tabular}{|c c c|} 
+     \hline
+     Parameter & Estimated mean  & Estimated mean  \\ 
+      & value (uniform) &value (log-uniform) \\[0.5ex] 
+     \hline\hline
+     $\lambda_N$ & 14.156 & 13.296  \\ 
+     $a$ & 9.670 & 0.017  \\ 
+     $\kappa_{N\beta}$ & 18.001 & 0.259 \\
+     $\mu_N$ & 13.196 & 0.018 \\
+     $\nu_{N\Phi}$ & 0.742 & 0.131 \\
+     \hline
+     $\kappa_{\Phi\beta}$ & 0.239 & 0.156 \\
+     $\mu_\Phi$ & 0.154 & 0.035 \\
+     \hline
+     $s_{\beta N}$ & 6.752 & 1.213 \\
+     $\mu_\beta$ & 6.378 & 1.315 \\
+     \hline
+     $s_{\alpha\Phi}$ & 16.997 & 5.826 \\
+     $\mu_\alpha$ & 11.859 & 2.434\\
+     \hline
+    \end{tabular}
+    \caption{Estimated parameter values of model 3}
+    \label{table:estimated1}
+\end{table}
+
+Compared to uniform distributed prior, we found that log-uniform distributed prior is more suitable in the inference.
+Log-uniform results in an obviously better fit and narrower inter-quantile interval (Figure \ref{fig:result123} and \ref{fig:resultCurve_uni}), and the final epsilon value is also smaller.
+Log-uniform tends to assign values that are close to the left boundary (i.e.
+zero in our case) higher density, i.e.
+they are more likely to be sampled in the first population.
+As a result, the estimated parameter posteriors are expected to have smaller mean and/or skewed to left of the x-axis.
+The estimated values proved this (see \ref{table:estimated1}; also observed in other models and prior interval experiments).
+It indicates that some of the true parameter tend to have small values that are close to zero, and using log-uniform distribution could results in a faster and more satisfying inference.
+Regarding this, further experiments all use log-uniform distribution only.
+
+\subsection{Model 4 and 5}
+
+[features that derive model 4 and 5]
+
+We observed that the existing models cannot well represent the trajectory of relative expression of tnf-$\alpha$: from experimental data, a rapid increase of tnf-$\alpha$ expression is observed in 0--4 hpl, followed by a rapid drop; all the models cannot well recover this feature and most test results see a less-rapid increase followed by a gradually decree and the observed peak value is not reached (e.g.
+Figure \ref{fig:result123}).
+Also in some other case e.g.
+Figure \ref{fig:resultCurve_uni}, tnf-$\alpha$ trajectories is well fitted at the cost of under-fitting $\Phi$.
+Hence efforts had been spent to propose more alternative models that can possibly solve this problem.
+
+The two proposed alternative models is based on model 3 and try to add extra term in the tnf-$\alpha$ equation $\mathrm{d} \alpha/\mathrm{d} t$.
+As macrophage is the main source of tnf-$\alpha$ production, and a similar sharp increase-and-drop trend is observed in the il-1$\beta$ trajectory, two hypothesis and corresponding terms are propose as follows.
+
+\paragraph{Model 4} It is assumed that the expression of il-1$\beta$ would have a promoting effect, representing by a phenomenological term $d_{\beta\alpha}\beta$ and $d_{\beta\alpha}$ is a new model parameter (positive constant).
+The promoting effect here is regarded to have the equivalent effect as directly promoting by il-1$\beta$, but the underlying mechanism is unclear so far.
+In mathematical view, this additional term can accelerate the expression of tnf-$\alpha$ in the first few time points and help to recover the peak-shaped trajectory.
+The proposed model is written as Equation \ref{eq:model4}.
+
+\paragraph{Model 5} Alternatively, we considered promotions to the tnf-$\alpha$ production, i.e.
+$s_{\alpha\Phi}\Phi$.
+It is assumed that il-$\beta$ could accelerate the production process.
+An additional term $f_{\beta\alpha}\beta$ is introduced to production rate, with $f_{\beta\alpha}$ being a new model parameter (positive constant).
+This model meets the biological context considering the source of tnf-$\alpha$.
+The proposed model is written as Equation \ref{eq:model5}.
+
+[further comparison of model 3, 4 and 5]
+
+Model 4 and 5 are base on model 3, so in this phase experiments of these three models are conducted for separately parameter inference and overall model comparison.
+Based on our experience in conducted experiments, we used the following setting for parameter inference:
+
+\begin{itemize}
+    \item population size: 2000 and 5000
+    \item number of populations: 30 generations
+    \item prior: [0, 50] log-uniform
+    \item perturbation kernel: multivariate normal kernel
+\end{itemize}
+
+Factors (Equation \ref{dis_f}) are also tried to find if we can force the inference framework to give more importance on the first few data points.
+All these runs are conducted on remote machines and the output database files are retrieved form analysis.
+
+Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 5000.
+
 \begin{figure}
     \begin{center}
-    \resizebox{1.0\hsize}{!}{\includegraphics{fig/cmp1.png}}
-    \end{center}
-
-    \caption{Model comparison of model 1, 2 and 3}
-    \label{fig:cmp1}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve345.png}}
+    \end{center}
+
+    \caption{Simulated data from the last population of model 3, 4 and 5}
+    \label{fig:resultCurve345}
 
 \end{figure}
-
-[features that derive model 4 and 5]
-
-[further comparison of model 3, 4 and 5]
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
 
@@
 
-\subsection{Other experiment details}
+\section{Parameter inference and model comparison results}
+
+\begin{figure}[H]
+
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve_uni.png}}
+    \end{center}
+
+    \caption{Total sampling size} 
+    \label{fig:resultCurve_uni}
+
+\end{figure}
 

### 2020-07-27 MSc-dissertation.tex
New stage

@@
 
-[how the infer the parameter given data]
+[how to infer the parameter given data]
 
@@
 
-[workflow and developing]
+[workflow and developing process]
 
@@
 
-Mathematical models can describe different kinds of dynamic system, and can be used as a guide to prediction and analysis.
-An ideal model in this case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  trend against time.
-
-Models for the regeneration process is in the form of ordinary differential equations, specifically the time differential form.
+Mathematical models can describe different kinds of dynamic systems, and thus can be used as a guide for prediction and analysis.
+An ideal model in our case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  data features in different time point.
+
+Proposed models for the regeneration process are in the form of ordinary differential equations (ODE), specifically the time differential form.
 Terms in the ODE are mostly explainable and corresponds interaction paths.
@@
 
-Our models is built regarding the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement includes the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
+Our models is built on the basis of the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
+The measurement data include the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
 As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
-According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration'.
+According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$).
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration' (compared to the value at first time point).
 
 It is noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations is mean of measurement, assuming that measured data is Gaussian-like distributed.
-The distribution of the measured data points is plotted and examined to see if the measurement mean could represent the distribution.
-The result is that at most time points the measurement values are Gaussian-distributed, some distributions are skewed.
+The summary statistic used for the parameter estimations is mean of measurement data at each time point, assuming that measured data is Gaussian-like distributed.
+To validate this, the distribution of the measured data points is plotted and examined to see if the mean value can represent the distribution.
+The result is that at most time points the measurement values are Gaussian-distributed, although some distributions are skewed.
 One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
-Mean value can summarise most data measures and thus is still used as the target summarised observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}
+Mean value can summarise most data and thus is still used as the target observed data.
+A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}.
 
@@
 At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
-After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can be more close to the true process, recover more features of the observed data or test our hypothesis.
-
-All these models assumes the interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for parameter estimation.
-Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data.
+After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can represent the observed regeneration process better or can be used to test our hypothesis.
+
+All these models assume the involved interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for the inference task.
+Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data available.
 
@@
 
-A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
-A interaction map illustrate the model is shown in Figure \ref{fig:m1}.
-This model is a simplification a the process described in \cite{ref:Tsarouchas}.
+A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the inference framework code.
+An interaction map of the model is shown in Figure \ref{fig:m1}.
+This model is a simplification a the process described in \cite{ref:Tsarouchas} with some minor interactions ignored.
 To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
-number of cells as `cell', and denote the unit of $\beta$ and $\aleph$ i.e.
+number of cells as `cell', and denote the unit of $\beta$ and $\alpha$ i.e.
 relative mRNA expression as `unit' for simplicity.
@@
 
-As the observed data indicates, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
-Regarding this, the self-increase term $\lambda$ cannot be constant, thus a exponentially decay $\lambda$ term is introduced and model 2 is proposed as Eqn.
-\ref{eq:model2}.
-The inhibition of il-1$\beta$ being produced by neutrophil, i.e.
-$i_{\beta\Phi}$ is considered to be ignored, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
+As the observed data indicate, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
+Regarding this, the self-increase parameter $\lambda$ cannot be constant, thus a exponentially decaying $\lambda$ term is introduced and model 2 is proposed.
+(Eqn.
+\ref{eq:model2}).
+Also, the inhibition of il-1$\beta$ production, i.e.
+term $i_{\beta\Phi}\Phi$ is considered to be ignored for a more simple model, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
 This case corresponds to model 3, written as Eqn.
@@
 
-Model 2 and model 3 has one extra parameter $a$ which is a coefficient in the exponentially decay determining the decay speed, with the unit $h^{-1}$.
+Model 2 and model 3 introduce one extra parameter $a$, which is a coefficient in the exponentially decaying $\lambda$, determining the decay speed, with the unit $h^{-1}$.
 
@@
 
-Model 3 can be regarded as a simplification of model 2, as it can be regarded as model 2 with parameter $i_{\beta\Phi}=0$.
-For these three models, model 1 is a naive one that is proposed at very first time and used as a `template' to build and test parameter inference framework.
-As the implementation is successful, model 2 and 3 is proposed.
-After fitting model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points which are close to steady states for neutrophil and macrophage.
-model 3 is a small simplification of model 2 and theoretically less general than model 2.
+Model 3 can be regarded as a simplification of model 2, as it can be treated as model 2 with parameter $i_{\beta\Phi}=0$.
+Among the proposed three models, model 1 is a naive one that was proposed at very first time and used as an ODE `template' to build and test parameter inference framework.
+As the implementation was successful, model 2 and 3 was proposed as we were trying to calibrate some terms and find a better model.
+After fitting, model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points (which relate to the steady state).
+Model 3 makes a small simplification on model 2 and is theoretically less `general' than model 2.
 
@@
 
-After the first model selection experiment, it is found that some significant features presented in the observed data is not recovered by any of the model.
-According to that, attempts are tried to introduce more interactions within the dynamic system considering the both biological and  mathematical context.
-Extra promoting effect to the expression of tnf-$\alpha$ is considered, by either add a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or add a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
+After the first model selection experiment, it was found that some significant features presented in the observed data were not recovered by any of the existing model.
+To resolve this, attempts were tried to introduce more interactions within the dynamic system considering the biological and  mathematical context.
+Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or adding a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
 \ref{eq:model4}) and model 5 (Eqn.
@@
      $f_{\beta\alpha}$ & Coefficient of promoting the production of $\alpha$ by $\beta$ & $cell^{-1}\cdotp h^{-1}$  \\ 
-    [1ex] 
      \hline
@@
 
-[feasible or not; goodness of fit; features of these models]
+[general topics on evaluating a model]
 
@@
 
-This chapter focuses on the implementation of the ABC SMC on the proposed models.
-Estimation of  parameters in the proposed models is the main task of this section.
-ABC SMC-based models selection is another task that can be easily implemented with the parameter estimation framework.
-
-The build and test is under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
+Estimation of  parameters in the proposed models is the main task of this section, which involves the implementation of ABC SMC and several experiments.
+ABC SMC-based model selection is another task that can be easily conducted with the developed parameter inference framework.
+
+The build and test of code were under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
 Some other code e.g.
 shell script and \verb|R| code are also used to perform the experiments and analyse the results.
-Tested software and system environment can be found at Appendix A.
-The build and test is performed on local computers and HPC clusters available within EPCC.
+Software and system environment used in our implementation can be found at Appendix A.
+The build and test were performed on local computers and HPC facilities available within EPCC.
 
@@
 
-According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| is adopted in this project.
+[how the code is developed and built]
+
+According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
 \verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}.
-\verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for inference and analysis tasks.
+\verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks.
 Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.
 
-The code for the project is developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version is deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task.
-The scaling-up experiment is performed on Cirrus, the profiling and its analysis are performed on local machine.
-Some development and debug are on the remote machines to ensure experiments run correctly on compute node.
-
-The ODE model-related code and utilities are firstly developed to enable reusable functions, variables and data calls.
+The inference framework code for the project was developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task.
+The scaling-up experiment was performed on Cirrus, and the profiling and its analysis were performed on local machine.
+Some development and debug were on the remote machines to ensure experiments run correctly on compute node.
+
+The ODE model-related code and utilities were firstly developed to enable reusable functions, variables and data calls.
 It includes the code for the ODE models, ODE solver, data structures, data format transform functions etc.
-By doing this in the implementation code file we can set options and activate ABC SMC easily, without creating definition and deplicated code for the models and parameter priors.
-
-Some modifications are made to the source code of \verb|pyABC|.
-For example, the laboratorial measured data are not considered `complete' for modeling and analysis purpose: there are different measurement time points for cells and cytokines, e.g.
-cell numbers is not measured at 0.25 hpl and cytokines' expression is not measured at time point 120 hpl.
-These missing values are treated as `ignored' when calculating the distance between observed data and simulated data.
+By doing this in a separate code file, we can change options and activate ABC SMC easily in the implementation code file, without creating definition and duplicated code for the models and parameter priors.
+
+Some modifications were made to the source code of \verb|pyABC|.
+For example, the laboratorial measured data were not considered `complete' for modeling and analysis purpose: the available time points for cells and cytokines are different, e.g.
+cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl.
+These missing values were treated as `ignored' when calculating the distance between observed data and simulated data.
 To cope with this built-in distance function code is modified.
 
-Analysis code is also wrapped up into functions which can be easily called after an ABC SMC run is finished.
-They includes multiple visualisations and summary statistic of the results.
+Analysis code was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results.
 For different tasks there are also separated analysis code file.
@@
 
-The project code is managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also make the cross machine synchronisation of code versions convenient.
-
-After several development iterations, the code files for ABC SMC implementation now mainly has two types: code file for infer-back experiments (synthetic data), code for parameter inference (experimental data).
-We can just add more models in these code files for a model comparison task.
-
-[how the code is developed and built]
+The project code was managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also made the cross-hardware synchronisation of code versions convenient.
+
+After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (udsing synthetic data), and code files for parameter inference (using experimental data).
+Model selection can be easily done by adding more model to the implementation code file.
 
@@
 
-As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed model is firstly examined before actual inference using the experimental observed data.
+As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental observed data.
 
@@
 
-\subsection{Data size, prior distribution range, population size and generations}
+\subsection{Data size, prior distribution and population}
 
@@
 Some other hyperparameters, e.g.
-feed-in data size, prior range, number of generations and number of particles in each generation is also of our interest.
-Experiments are designed to explore how these options affect the goodness of fit and efficiency of the implementation.
-
-Regarding the present dynamic systems model, the SMC is implemented with two data size options, two prior distribution (standard and wider prior range).
-The iteration options i.e.
-population size and number of generaitions are also tried with different values.
-These experiments intends to give suggestions on the later SMC implementation on experimental data.
+feed-in data size, prior distribution, number of populations and number of particles in each population were also of our interest.
+Experiments were designed to explore how these options affect the goodness of fit and efficiency of the implementation.
+
+Regarding the dynamic systems model, SMC was applied with two data size options, three prior distribution (uniform distribution, wider uniform distribution and log-uniform distribution).
+The population options i.e.
+population size and number of populations were also tried with different values.
+These experiments intended to give suggestions on the later SMC implementation on experimental data.
 
@@
 
-Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 5000.
+Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 2000. It can be seen that
 
@@
 
-\chapter{Results and Discussions}
-
-[WHAT results is outputted and WHAT topics are discussed]
-
-\section{ABC SMC results}
-
-[inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]
-
-[model selections result; bayes factor]
-
-\section{Discussions}
-
-[goodness of fit (evaluation); preferred models and effects of modification to basic model]
-
-[efficiency and trade-offs in scaling-up]
-
-[POSSIBLE: compare to exact inference; compared to other implementations]
-
-[limitations]
+% \chapter{Results and Discussions}
+
+% [WHAT results is outputted and WHAT topics are discussed]
+
+% \section{ABC SMC results}
+
+% [inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]
+
+% [model selections result; bayes factor]
+
+% \section{Discussions}
+
+% [goodness of fit (evaluation); preferred models and effects of modification to basic model]
+
+% [efficiency and trade-offs in scaling-up]
+
+% [POSSIBLE: compare to exact inference; compared to other implementations]
+
+% [limitations]
 

### 2020-07-29 MSc-dissertation.tex
New para

@@
 Factors (Equation \ref{dis_f}) are also tried to find if we can force the inference framework to give more importance on the first few data points.
+In this experiment, for each trajectory first 8 data points are assigned higher factors (0.75), and the rest 4 data points are assigned with lower factors (0.25).
 All these runs are conducted on remote machines and the output database files are retrieved form analysis.
 
-Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 2000. It can be seen that
+Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 2000.
+It can be seen that the addressed problem in the trajectory of tnf-$\alpha$ is partially relieved in model 4 and 5: compared to model 3, a peak appears around 4 hpl; it can represent more features in the observed data and consequently the final reached epsilon (after 30 populations) is smaller than that of model 3, which proves that our proposed modifications in model 4 and 5 are helpful in recovering more features in observed data.
+
+By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
+Model 1, 2 and 3 con also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- is not fitted.
+Some features, e.g.
+fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.
+
+After applying factors, the results were not largely different.
+As expected, applying factors made the data points at first half (where exists most of the fluctuations) more `focused', however the inferred models are close to one without factors, which could indicate that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.
 

### 2020-07-29 MSc-dissertation.tex
New para

@@
 
+\begin{figure}
+    \begin{center}
+    \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve345.png}}
+    \end{center}
+
+    \caption{Simulated data from the last population of model 3, 4 and 5}
+    \label{fig:resultCurve345}
+
+\end{figure}
+
 By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
@@
 
-\begin{figure}
-    \begin{center}
-    \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve345.png}}
-    \end{center}
-
-    \caption{Simulated data from the last population of model 3, 4 and 5}
-    \label{fig:resultCurve345}
-
-\end{figure}
+When we are trying more populations, an `over-fitting' behaviour is observed at the trajectory of macrophage (FIGURE).
+In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).
+
+Next, we performed a model selection on model 3, 4 and 5.
+FIGURE shows the model probabilities at different generations.
+From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred.
+Again, as DISCUSSED in the model comparison of model 1, 2 and 3, we considered the the violently changing model probability is related to the inference ability at certain epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
+Given the present result, model 5 is regarded as the best model.
+Some further experiments tried larger population size (10,000 particles per population) and more populations (36 populations), but the resultant model did not have a significantly better fit of the observed data, although much more computation resources are used.
+
+A inferred parameter posterior distribution of model 5 is shown in FIGURE. SOME DESCRIPTION.
+
+\subsection{Sensitivity of parameters}
+
+[why last PCs]
+
+[which parameters are most sensitive]
+
+[what does that mean]
 
@@
 
+[possible reasons]
+
 \section{Profiling}

### 2020-07-31 MSc-dissertation.tex
New para

@@
 
-A inferred parameter posterior distribution of model 5 is shown in FIGURE. SOME DESCRIPTION.
+A inferred parameter posterior distribution of model 5 is shown in FIGURE, the estimated mean values is shown in TABLE. SOME DESCRIPTION.
 
@@
 
+[compared to credible intervals plot]
+
 [what does that mean]
 
+A quick sensitivity analysis of the dynamical systems models can be quantified using principle component analysis (PCA) \cite{Toni}, where the first PCs corresponds to sloppy parameters while the last PCs corresponds to stiff parameters \cite{sensitivity}.
+PCA can only provide rough approximated sensitivity behaviour.
+As models 5 won in the model selection, a further interest is to explore the parameter-to-model sensitivity
+
+FIGURE shows the results of PCA on the last population of inferred model5.
+The last PC (PC12) is mostly consists of the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive.
+If we conclude the last 5 PCs (FIGURE), then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$.
+As a conclusion, model 5 is sensitive to changes in the above parameters, which all come from the equation of neutrophil and macrophage i.e.
+equations of the cells; it may suggests that given the observed data, the inferred model is more sensitive to cells' dynamic rather than cytokines' dynamics.
+
+Also, some visualisations of the last population agreed with conclusions from PCA.
+Approximated posteriors of the last population plotted in FIGURE and credible ranges of parameters across populations in FIGURE shows that the 5 parameters identified by PCA is indeed `stiff' parameters.
+
 \chapter{Performance experiments}
 
-The performance experiments are designed to explore the parallel performance of ABC SMC implementations.
-Usually ABC SMC is a time-consuming and computation intensive task and usually is executed on large clusters.
+The performance experiments were designed to explore the parallel performance of ABC SMC inference framework which were used in the parameter estimation and model selection tasks.
+Usually ABC SMC is a time-consuming and computation intensive task and ideally executed on large clusters.
 The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
@@
 
-The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54 and 72 cores respectively.
-Each run is repeated 5 times.
+The program is executed on Cirrus, using 8, 16, 24, 36, 54 and 72 cores respectively.
+Each run is repeated 10 times.
 The average execution time, required sampling numbers are recorded.
@@
 
+[large variation in required sampling numbers]
+
 [possible reasons]
+
+The recorded execution time using different number of cores is shown in FIGURE.
+According to that, speed-up and efficiency can be calculated.
+Here 8 cores is regarded as the baseline for speed-up and efficiency, as the serial version (using only one core) can take quite a long time to finish 10 repeats.
+When increasing number of cores, the execution time drops fast at first but the decreasing rate (dropping speed, i.e.
+the derivative of the execution time against time) is gradually reduced to zero; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (FIGURE).
+
+Denote the number of cores used in the ABC SMC runs as variable $p$.
+The speedup curve shows a linear trend when $p\leq 36$; when $p\geq 36$ the speedup stays nearly unchanged.
+$p=36$ is an inflection point, where 36 is the maximum numbers available physical cores in a compute node of Cirrus.
+A constant drop of efficiency is observed when increasing $p$; smaller $p$ ($p<36$) gives an efficiency higher than 65\% but greater $p$ e.g.
+$p=72$ results in a low efficiency (34\%).
+
+A high variance of total required samples was observed in the scaling-up experiments
 
@@
 
+\bibitem{Toni} Toni, T., Welch, D., Strelkowa, N., Ipsen, A.
+and Stumpf, M.P., 2009.
+Approximate Bayesian computation scheme for parameter inference and model selection in dynamical systems.
+Journal of the Royal Society Interface, 6(31), pp.187-202.
+
+\bibitem{sensitivity} Gutenkunst, R.N., Waterfall, J.J., Casey, F.P., Brown, K.S., Myers, C.R.
+and Sethna, J.P., 2007.
+Universally sloppy parameter sensitivities in systems biology models.
+PLoS Comput Biol, 3(10), p.e189.
+
 \end{thebibliography}

### 2020-07-31 MSc-dissertation.tex
New para

@@
 
-A high variance of total required samples was observed in the scaling-up experiments
+A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations.
+Moreover, we found that the execution time is not ideally proportional to the total required samples (FIGURE).
+A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs.
+Some runs that are stuck in a local optimum and require smaller target epsilon will take huge amount of samples to move away from the local optimum.
+This have been observed when the distribution of posterior is plotted before and after jumping away from local modes (FIGURE and more explain).
+Also, time spent in each sample using sanme number of cores is not constant in our case, which might be a result of the parallel schedule; ideally a serial run using only one core will see a nearly constant time spent in every particle.
+
+Due to the unstable sampling numbers, our interests switched to the per-particle performance, where the average sampling numbers per second (FIGURE) and average time spent in one particle (FIGURE) under different of cores ($p$) are plotted.
+The log-like average time per particle curve shows that using more cores can improve the sampling efficiency, but when $p$ is at a high level (e.g.
+36 in our case), the improvement will be less significant.
 

### 2020-08-01 MSc-dissertation.tex
Fig fix

@@
 \caption[Mean of the observed data]%
-{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
+{Mean of the observed data for neutrophil ($N$), macrophage ($\Phi$), il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
 \label{fig:obs_data}
@@
     \caption[Synthetic data generated with known parameter values]%
-    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot)} 
+    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot).
+Known parameter values are obtained from a least square fitting of the observed data}
     \label{fig:infer_back_data}
@@
 
-Among them the following topics are studied using synthetic data.
+[TABLE for known parameter values]
+
+[why use synthetic data]
+
+Among them the following topics are studied using the synthetic data.
 

### 2020-08-01 MSc-dissertation.tex
Fig fix

@@
 
-\begin{figure}
+\begin{figure}[t!]
     \begin{center}
@@
 
-    \caption{ABC SMC sampling process} 
+    \caption[ABC SMC sampling process]%
+    {ABC SMC sampling process.
+$\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$.
+$d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data.
+for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
     \label{fig:smc}
@@
     \caption[Total required samples of different kernels]%
-        {Total required samples of different kernels. Different color represents different generations (bottom to top: generation 1 to generation 20)}
+        {Total required samples of different kernels after 20 populations (2000 particles in each population).
+Different color represents different generations (bottom to top: population 1 to population 20)}
     \label{fig:kernel1}
@@
 
-    \caption{Acceptance rates of different kernels}
+    \caption[Acceptance rates of different kernels]%
+    {Acceptance rates of different kernels in 20 populations. Each population has 2000 particles}
     \label{fig:acceptance1}
@@
 When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data point, or behave like a normalisation that can balance the scales of different summarised statistics (here is the 48 mean values).
-There are two appliance case studied: (1) factors used as an normalisation option ans (2) factors used to give more focus on some features of the observed data.
+There are two appliance case studied: (1) factors used as an normalisation option and (2) factors used to give more focus on some features of the observed data.
 For (2), as we noticed that the main features (e.g.

### 2020-08-01 MSc-dissertation.tex
New fig

@@
 
-    \caption{Simulated trajectory from the last population of model 1, 2 and 3}
+    \caption[Simulated trajectory from the 20th population of model 1, 2 and 3]%
+    {Simulated trajectory from the last population of model 1, 2 and 3.
+Error bar indicates SEM.
+500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from}
     \label{fig:result123}

### 2020-08-01 MSc-dissertation.tex
Fix fig

@@
 
-    \caption{Estimated posterior distribution of parameters in model 3}
+    \caption[Estimated posterior distribution of parameters in model 3]%
+    {Estimated posterior distribution of parameters in model 3. Shaded range indicates 25\%--75\% quantile of the population}
     \label{fig:para1}
@@
 
-        \caption{Model comparison of model 1, 2 and 3}
+        \caption[Model comparison of model 1, 2 and 3]%
+        {Model comparison of model 1, 2 and 3. From population 8 to population 30, model3 always wins}
         \label{fig:cmp1}
@@
      \hline\hline
-     $\lambda_N$ & 14.156 & 13.296  \\ 
-     $a$ & 9.670 & 0.017  \\ 
-     $\kappa_{N\beta}$ & 18.001 & 0.259 \\
-     $\mu_N$ & 13.196 & 0.018 \\
+     $\lambda_N$ & 14.1 & 13.3 \\ 
+     $a$ & 9.67 & 0.0170  \\ 
+     $\kappa_{N\beta}$ & 18.0 & 0.259 \\
+     $\mu_N$ & 13.2 & 0.0180 \\
      $\nu_{N\Phi}$ & 0.742 & 0.131 \\
@@
      $\kappa_{\Phi\beta}$ & 0.239 & 0.156 \\
-     $\mu_\Phi$ & 0.154 & 0.035 \\
-     \hline
-     $s_{\beta N}$ & 6.752 & 1.213 \\
-     $\mu_\beta$ & 6.378 & 1.315 \\
-     \hline
-     $s_{\alpha\Phi}$ & 16.997 & 5.826 \\
-     $\mu_\alpha$ & 11.859 & 2.434\\
+     $\mu_\Phi$ & 0.154 & 0.0350 \\
+     \hline
+     $s_{\beta N}$ & 6.75 & 1.21 \\
+     $\mu_\beta$ & 6.38 & 1.32 \\
+     \hline
+     $s_{\alpha\Phi}$ & 17.0 & 5.83 \\
+     $\mu_\alpha$ & 11.9 & 2.43\\
      \hline
     \end{tabular}
-    \caption{Estimated parameter values of model 3}
+    \caption[Estimated parameter values of model 3]
+    {Estimated parameter values of model 3, using uniform prior distribution and log-uniform prior distribution respectively}
     \label{table:estimated1}
@@
      \hline\hline
-     $\lambda_N$ & 2.1989 & $cell/h$  \\ 
+     $\lambda_N$ & 2.20 & $cell/h$  \\ 
      $\kappa_{N\beta}$ & 3.9627 & $cell/(unit\cdotp h)$\\

### 2020-08-01 MSc-dissertation.tex
Fix fig

@@
  $\mu_\alpha$ & Coefficient of negative feedback of $\alpha$ & $h^{-1}$ \\
-[1ex] 
  \hline
 \end{tabular}
-\caption{Parameters introduced in the basic model (model 1)}
+\caption{Parameters introduced in the basic model (model 1) [REMOVE unit]}
 \label{table:m1}
@@
      $\lambda_N$ & 2.20 & $cell/h$  \\ 
-     $\kappa_{N\beta}$ & 3.9627 & $cell/(unit\cdotp h)$\\
-     $\mu_N$ & 1.7219 & $h^{-1}$\\
-     $\nu_{N\Phi}$ & 0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
-     \hline
-     $\lambda_\Phi$ & 1.3146 & $cell/h$ \\
-     $\kappa_{\Phi\beta}$ & 0.1235 & $cell/(unit\cdotp h)$ \\
-     $\mu_\Phi$ & 0.1454 & $h^{-1}$ \\
-     \hline
-     $s_{\beta N}$ & 6.5536 & $unit/(cell\cdotp h)$ \\
-     $i_{\beta\Phi}$ & 1.7062 & $cell^{-1}$ \\
-     $\mu_\beta$ & 0.5212 & $h^{-1}$ \\
-     \hline
-     $s_{\alpha\Phi}$ & 10.2416 & $unit/(cell\cdotp h)$ \\
-     $\mu_\alpha$ & 19.6642 & $h^{-1}$ \\
-    [1ex] 
+     $\kappa_{N\beta}$ & 3.96 & $cell/(unit\cdotp h)$\\
+     $\mu_N$ & 1.72 & $h^{-1}$\\
+     $\nu_{N\Phi}$ & 0.220 & $cell^{-1}\cdotp h^{-1}$ \\
+     \hline
+     $\lambda_\Phi$ & 1.31 & $cell/h$ \\
+     $\kappa_{\Phi\beta}$ & 0.124 & $cell/(unit\cdotp h)$ \\
+     $\mu_\Phi$ & 0.145 & $h^{-1}$ \\
+     \hline
+     $s_{\beta N}$ & 6.554 & $unit/(cell\cdotp h)$ \\
+     $i_{\beta\Phi}$ & 1.71 & $cell^{-1}$ \\
+     $\mu_\beta$ & 0.521 & $h^{-1}$ \\
+     \hline
+     $s_{\alpha\Phi}$ & 10.2 & $unit/(cell\cdotp h)$ \\
+     $\mu_\alpha$ & 19.1 & $h^{-1}$ \\
      \hline
@@
 
-    \caption{Total sampling size} 
+    \caption[Total sampling size of different kernels, using median epsilon strategy]
+    {Total sampling size of different kernels, using median epsilon strategy. Different color represents different generations (bottom to top: population 1 to population 20)}
     \label{fig:kernel2}
@@
 
-        \caption{Acceptance rates} 
+        \caption[Acceptance rates of different kernels, using median epsilon strategy]
+        {Acceptance rates of different kernels, using median epsilon strategy. Each population has 2000 particles} 
         \label{fig:acceptance2}

### 2020-08-04 MSc-dissertation.tex
Inro

@@
 
-% This example file shows how a thesis can be laid out using Latex. It
-% does not use any special local features so should be portable to other
-% places.
-% 
-% When producing draft copies of a thesis you may want to print only
-% selected pages of the thesis. To do this use the command
-% 
-% dvips -f -p 4 -n 3 myfile.dvi | lpr
-% 
-% where -p 4 means start printing at page 4 (ie the page that will be
-% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
-% This example will print pages 4, 5 and 6.
-% 
-% If you want to print the thesis and also save paper you can print more
-% than one page on each sheet of paper. Use the command
-% 
-% dvips -f myfile.dvi | psnup -2 | lpr
-% 
-% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
-%
-% To produce a PDF version you can create a PostScript copy first
-%
-% dvips -f myfile.dvi > myfile.pdf
-%
-% and then convert it
-%
-% distill myfile.ps
-%
-% or you can go straight to PDF
-%
-% pdflatex myfile
-%
-% Note that pdflatex expects all included figures to be in PDF too. See
-% the includegraphics command below.
-
-% This document contains many cross-references and forward references,
-% eg in constructing a table of contents, so Latex may need to be run
-% twice to get all the references correct. If you need to run Latex twice
-% you may get the warning:
-% 
-% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.
-
-% the following 4 lines are the content of the smallmargins.sty file
-% but including them explicitly makes this more portable.
-%AC%\oddsidemargin=0.1in
-%AC%\topmargin=-0.5in
-%AC%\textheight=9in
-%AC%\textwidth=6.25in
-
-%AC%\parskip 10pt
-%AC%\parindent 0in
-
 \begin{document}
 
-%AC%\pagestyle{myheadings}
-%AC%\markright{D.~S.~Henty}
-
-%\title{A Latex thesis example}
-%\author{D.~S.~Henty}
-%\date{\today}
-
-%\maketitle
-
 \pagenumbering{roman}
@@
 \title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
+
 \author{Chaolin Han}
+
 \date{\today}
@@
 
-% You will also want to make sure you have no spelling or grammatical
-% mistakes. To help idwentify spelling mistukes you caan use the commands
-% {\em ispell} or {\em spell}. See the appropriate manual pages. Remember
-% that spelling mistakes are not the only errors which can occur. Spelling
-% checkers will not find errors which are, in fact, valid words such as
-% {\em there} for {\em their}, nor will they find repeated words which
-% sometimes occur if your concentration is broken when typing. {\bf There
-% is no substitute for thorough proof reading!}
-
 [background and motivation]
@@
 [what is done]
+
+While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions of cells and chemicals in a biological system.
+The objective process in this dissertation is an example of complex biology system found in tissue regeneration.
+Zebrafish has the ability to regenerate its spinal cord after injury; at the injury-site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.
+
+Our interest lies in the modeling of the regeneration process that happens in the injury-site, especially in the parameter estimation and comparison of different models.
+According to available researches \cite{ref:Tsarouchas}, a mathematical model is proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments.
+However, we would like to explore the the `best fit' if the model which comes with a set of parameter estimates given the observed data.
+Also, we wish to compare different it with different models and possibly find the best model to describe the data without losing biological and mathematical properties in the regeneration process.
+
+Fo the parameter estimation tasks, Bayesian inference techniques can help in the problems of over-fitting and local optima \cite{ref:abcsysbio}, compared to optimisation based methods.
+Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression can be hard.
+
+Such methods introduce computation-intensive sampling-rejection pattern, where massive number of samples are drawn, followed by computations.
+As the sampling process is independent from each other in a certain period, the parallel execution becomes feasible and is expected to accelerated the parameter inference in a large scale.
+
+This dissertation includes the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results.
+Model comparison results are also included using the same method.
+Additionally, a performance test illustrates the strong-scaling performance of the algorithm.
 

### 2020-08-05 MSc-dissertation.tex
New para

@@
  [describe the process and how cells and cytokines are involved]
+
+Compared to mammals, zebrafish can regenerate spinal cord after injury, and possibly recover swimming function.
+They are widely used in regenerative ability researches.
+Experiments found that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}.
+Tsarouchas et al.
+\cite{ref:Tsarouchas} studied the the complex interactions happen at the lesion site.
+The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration.
+Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, affect the inflammation, which are dynamically controlled by immunes cells, i.e.
+cytokines works as media to control the regeneration.
+Here cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.
+
+An interaction map can be drawn, according to the evidences found in \cite{ref:Tsarouchas}.
+After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells: macrophage, neutrophil and keratinocyte.
+The dynamical system consists of complex interactions between cells and cytokines, e.g.
+the increasing tnf-$\alpha$
 
@@
         \hline
- Parameter & Definition & Units\\ [0.5ex] 
+        Parameter            & Definition                                       & Units                    \\ [0.5ex]
         \hline\hline
         $\lambda_N$          & Self-increase rate of neutrophil                 & $cell/h$                 \\
- $\kappa_{N\beta}$ & Promoting effect coefficient by il-1$\beta$ & $cell/(unit\cdotp h)$\\
+        $\kappa_{N\beta}$    & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
         $\mu_N$              & Coefficient of negative feedback of $N$          & $h^{-1}$                 \\
- $\nu_{N\Phi}$ & Coefficient of inhibition of both $N$ and $\Phi$& $cell^{-1}\cdotp h^{-1}$ \\
+        $\nu_{N\Phi}$        & Coefficient of inhibition of both $N$ and $\Phi$ & $cell^{-1}\cdotp h^{-1}$ \\
         \hline
@@
         \hline
-     Parameter & Definition & Units\\ [0.5ex] 
+        Parameter         & Definition                                                     & Units                    \\ [0.5ex]
         \hline\hline
@@
         Parameter            & Estimated mean  & Estimated mean      \\
-      & value (uniform) &value (log-uniform) \\[0.5ex] 
+                             & value (uniform) & value (log-uniform) \\[0.5ex]
         \hline\hline
@@
         $s_{\alpha\Phi}$     & 17.0            & 5.83                \\
-     $\mu_\alpha$ & 11.9 & 2.43\\
+        $\mu_\alpha$         & 11.9            & 2.43                \\
         \hline
@@
         \hline
-     $s_{\beta N}$ &6.5536  $unit/(cell\cdotp h)$ \\
+        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
         $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
-     $\mu_\beta$ &0.5212  $h^{-1}$ \\
+        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
         \hline
@@
         \hline
-     Parameter & Value & Unit\\ [0.5ex] 
+        Parameter            & Value & Unit                     \\ [0.5ex]
         \hline\hline
         $\lambda_N$          & 2.20  & $cell/h$                 \\
-     $\kappa_{N\beta}$ & 3.96 & $cell/(unit\cdotp h)$\\
-     $\mu_N$ & 1.72 & $h^{-1}$\\
+        $\kappa_{N\beta}$    & 3.96  & $cell/(unit\cdotp h)$    \\
+        $\mu_N$              & 1.72  & $h^{-1}$                 \\
         $\nu_{N\Phi}$        & 0.220 & $cell^{-1}\cdotp h^{-1}$ \\
@@
 \begin{thebibliography}{100}
+
+    \bibitem{axonal} Becker, T., Wullimann, M.F., Becker, C.G., Bernhardt, R.R.
+and Schachner, M., 1997.
+Axonal regrowth after spinal cord transection in adult zebrafish.
+Journal of Comparative Neurology, 377(4), pp.577-595.
 

### 2020-08-05 MSc-dissertation.tex
New para

@@
 The dynamical system consists of complex interactions between cells and cytokines, e.g.
-the increasing tnf-$\alpha$
+il-1$\beta$ promotes macrophages' increase.
+and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
+Consequently, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration.
+Generally, il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect, and tnf-$\alpha$ promotes the regeneration all the time.
+The presence if the cytokines also affects the number of cell, while as sources number of cells also affects cytokines' expression.
+Interactions between different kinds of cells are also considered in the system.
+
+This is a dynamic system suitable for mathematical modelling, as a set of equations can be written to represent the dynamics as possible.
+Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions iniside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and our own hypothesis which may include different interactions.
+
+The biological background of works in this dissertation is mostly related the the experimental evidences, quantitative data and proposed hypothesis in \cite{ref:Tsarouchas}.
+The published data on the quantitative measurement of cells and cytokines make the modelling and further analysis possible.
+Some other interactions proposed in our modelling stages are enlightened from Anderson et al.
+\cite{Anderson}.
 
@@
  [general topics of mathematical models: how to build model according to interactions; how to calibrate/ parameterise the model; how to solve the model; dynamics; how to evaluate the model]
+
+ Mathematical models can express real-life questions in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques.
+Mathematical models can be used for prediction, classification and other tasks and thus widely used to solve practical problems.
+However, for a model that describe the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models.
+
+ According to
 
@@
 
+    \bibitem{Anderson} Anderson, W.D., Makadia, H.K., Greenhalgh, A.D., Schwaber, J.S., David, S.
+and Vadigepalli, R., 2015.
+Computational modeling of cytokine signaling in microglia.
+Molecular BioSystems, 11(12), pp.3332-3346.
+
     \bibitem{ref:pyabc} Klinger, E., Rickert, D. and Hasenauer, J., 2018. pyABC: distributed, likelihood-free inference. Bioinformatics, 34(20), pp.3591-3593.

### 2020-08-08 MSc-dissertation.tex
New para

@@
 Mathematical models can express real-life questions in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques.
-Mathematical models can be used for prediction, classification and other tasks and thus widely used to solve practical problems.
+Mathematical models can be used for prediction, classification, testing our understanding or hypothesis and many other tasks, thus they are widely used to solve practical problems.
 However, for a model that describe the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models.
-
- According to
+In this case, we wish to use modelling techniques to test our inference method, our hypothesis and hopefully an well-fitted models can be found and compared.
+
+According to our understanding of the biological mechanism in the regeneration, hypothesis should firstly proposed, in a way which can represent the real  mechanism as much as possible.
+Usually a real-life problem, e.g.
+the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered, An elaborate description of the real cases using a mathematical model sometimes can be impossible; assumptions and hypothesis are proposed to make necessary simplification and abstract for a final reasonably detailed model.
+Some qualitative models are simple but illustrative for the real-life observed patterns, e.g.
+Lotka-Volterra (L-V) equation for predator and prey models; they are usually analytically solvable and some properties are used to make predictions (e.g.
+steady states in ecosystem).
+If a more precise model is desired, usually more effects or interactions are considered and added as terms to a re-constructed model, e.g.
+environmental factors, a changing environmental capacity and predation from other species.
+
+Models like L-V equation are mostly analytically solvable when analysing properties of the model, however, analytical solution for many models are not feasible, because of more complex dynamics, or the model itself is not a deterministic model.
+In the areas that are not analytically feasible, usually computational techniques are applied for approximate solutions.
+Solvers for different models are used to simulate the system or find roots for some spacial points.
+
+For the regeneration process that we wanted to model, differential equations are our preference.
+Non-linear ordinary differential equations are used in this dissertation, where the only independent variable is time $t$ denoting the post-lesion time (hours), but further models can also consider other models e.g.
+stochastic models and partial differential equations.
+For ordinary differential equations (ODE) system, there are several available computational solvers using different algorithms, implemented in MATLAB\footnote[1] {\url{https://uk.mathworks.com/help/matlab/math/choose-an-ode-solver.html}}, Python\footnote[2]{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html}} and many other languages.
+These solvers help to simulate the trajectories of the dependent variables, given initial values and a set of fixed parameters.
+The goodness of `fit' then can be measured by compared the simulated data and observed real data.
+
+Often further researches on the modelled problems require the model to be deterministic on its parameters.
+Parameters in the dynamic system models usually have their practical significance, e.g.
+describing reaction speed, environmental capacity or the birth rate.
+There are many approaches to derive these parameters from data or experience.
+For a specific problem and a widely used models, the parameters' values can be read from previous literatures; however in many others cases, we need to find parameter's values specifically for our model and problem, where numerical values for parameters are more suitable for practical analysis using ODE solvers or other tasks.
+
+To find the best parameters' values given observed data, either optimisation-based approaches (e.g.
+least-square (LS) fit) or inference approaches can be used.
+Optimisation-based approaches may face two major drawbacks: over-fitting and local optima \cite{ref:abcsysbio}.
+Bayesian inference approaches gain popularity in recent years, which can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
+Local optima can still be a major concern, and many efforts have be proposed for a more through explore of the parameter space.
 
@@
  [model comparison, ABC SMC for model comparison]
+
+ Bayesian inference approaches put Bayes Rule in the centre:
+
+ \begin{align}
+    \label{eq:bayes}
+    p(\theta|D) = \frac{p(D|\theta)p(\theta)}{P(D)}
+\end{align}
+
+where $p$ is the probability density function, $\theta$ denotes the parameter set, $D$ is the given observed data.
+
+Alternatively, as $p(D)$ is fixed once the observed data is given, the rule can be rewritten as
+
+\begin{align}
+    \label{eq:bayes2}
+    p(\theta|D) \propto l(D|\theta)p(\theta)
+\end{align}
+
+where $l$ represents the likelihood function.
+In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target.
+Prior $p(\theta)$ should express our initial believes or estimates on the parameter valeus before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers.
+Once the data is given, $\theta$ is the only parameter in the likelihood function.
+Some optimisation-based methods try to find an estimation of parameter set $\theta$ s.t.
+maximise the likelihood, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative.
+
+Based on Eqn \ref{eq:bayes2}, to find the posterior of parameter set, prior can be set according to the problem context and our understanding and cover a reasonable ranges; computational methods can be used to approximate likelihood function, e.g.
+Markov Chain Monte Carlo (MCMC) \cite{ref:MCMC}.
+As another alterative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood, but focused on simulating the data from the given model and observed data, then used the simulated data to approximate the desired posterior.
+ABC-rejection \cite{ABC_rejection} is a simple example of the ABC process, and further algorithms e.g.
+sequential Monte-Carlo (ABC SMC)\cite{Toni} are proposed to improve the efficiency.
+
+General steps for ABC algorithm includes the following (ABC-rejection):
+
+\begin{enumerate}
+    \item Draw a sample $\theta^*$ from parameters' prior distribution. $\theta^*$ is a vector that consists values for each parameter
+    \item Plug $\theta^*$ into the model, and use the model to generate simulated data $D^*$, where $D^*$ is of the same form as the observed data $D$, e.g. includes same summary statistics
+    \item The discrepancy between the simulated data and observed data is measured by a distance function $d(D^*, D)$.
+For a pre-fixed threshold value $\epsilon$, if $d(D^*, D)<\epsilon$ then the sample $\theta^*$ is accepted
+    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution can be approximated by these samples
+\end{enumerate}
+
+In our study of the regeneration model, full trajectories of the dependent variables are used as the data to be compared, as our goal is to find a model and a parameter set that can intuitively recover the real process as much as possible.
+The mean values at each time point for each dependent variable is calculated to form $D$, and for each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$.
+Using other statistics to summary the raw data are also possible bur care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
+
+Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of deceasing threshold values $\epsilon_t$, and derive a new population (called generation) of accepted sampled from the previous accepted samples (called particles).
+An example is ABC SMC \cite{Toni}, which can significantly reduce the required samples compared to ABC-rejection \cite{ref:disease}:
+
+\begin{enumerate}
+    \item 
+\end{enumerate}
 
@@
 
+    \bibitem{ref:MCMC} Gilks, W.R., Richardson, S.
+and Spiegelhalter, D.J., 1996.
+Markov Chain Monte Carlo in Practice Chapman \& Hall: London.
+Markov Chain Monte Carlo in practice.
+Chapman and Hall, London.
+
+    \bibitem{ABC_rejection} Tavaré, S., Balding, D.J., Griffiths, R.C. and Donnelly, P., 1997. Inferring coalescence times from DNA sequence data. Genetics, 145(2), pp.505-518.
+
+    \bibitem{summaryD} Joyce, P. and Marjoram, P., 2008. Approximately sufficient statistics and Bayesian computation. Statistical applications in genetics and molecular biology, 7(1).
+
+    \bibitem{summaryD2} Nunes, M.A.
+and Balding, D.J., 2010.
+On optimal selection of summary statistics for approximate Bayesian computation.
+Statistical applications in genetics and molecular biology, 9(1).
+
     \bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C.

### 2020-08-08 MSc-dissertation.tex
New para

@@
 \begin{enumerate}
-    \item 
+    \item Set a threshold schedule $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ where $T$ is the total number of populations and $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ is a descending sequence; set the number of particles in each population as $N$, set the population index $t=1$
+    \item if $t=1$, draw samples from prior distribution until $N$ particles are accepted under threshold $\epsilon_1$, i.e. each of $N$ particles s.t. $d(D,D^*)<\epsilon_1$
+
+    Else ($t>1$), draw samples from previous population $\{\theta^*_{t-1}\}$ with weights $\{w_{t-1}\}$, perturb the particles using a perturbation kernel $K(\theta^*|\theta)$ and obtain $\theta^{**}\sim K(\theta^*|\theta)$
+    \item Repeat step 2 until $N$ particles are accept under threshold $\epsilon_t$
+    \item Calculate the weights of the accepted particles. For particle $i$ 
+    \begin{align}
+        \label{eq:weight}
+        w^i_t =\begin{cases}
+            1 & \text{if t=1} \\
+            \frac{p(\theta^i_t)}{\sum_{j=1}^{N} w^j_{t-1}K(\theta^j_t|\theta^j_{t-1})} & \text{if t $\neq$ 1} 
+        \end{cases}
+    \end{align}
+
+    \item Normalise $\{w^i_t\}$, increase population index $t$ by 1
+    \item Repeat step 2, 3 and 4 until $t>N$
+    \item Output the accepted particles $\{\theta^*_{t}\}$ and their weights $\{w^i_t\}$ in each generation $t$
+
 \end{enumerate}
+
+An illustrative graph of the ABC SMC procedures can be found in Figure \ref{fig:smc}.
+
+\begin{figure}[t]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
+    \end{center}
+
+    \caption[ABC SMC sampling process]%
+    {ABC SMC sampling process.
+$\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$.
+$d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data.
+for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
+    \label{fig:smc}
+
+\end{figure}
 
@@
 
-\begin{figure}[t!]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
-    \end{center}
-
-    \caption[ABC SMC sampling process]%
-    {ABC SMC sampling process.
+% \begin{figure}[t!]
+%     \begin{center}
+%         \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
+%     \end{center}
+
+%     \caption[ABC SMC sampling process]%
+%     {ABC SMC sampling process.
 $\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$.
@@
 for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
-    \label{fig:smc}
-
-\end{figure}
+%     \label{fig:smc}
+
+% \end{figure}
 

### 2020-08-09 MSc-dissertation.tex
New section

@@
 An illustrative graph of the ABC SMC procedures can be found in Figure \ref{fig:smc}.
+In implementations, many factors can affect the efficiency and goodness of fit, some of them are discussed in the Chapter 4.
+Rather than a pre-fixed threshold schedule, median or quantile based epsilon schedule can be save total sampling size and thus improve acceptance rates of particles \cite{threshold}.
+Some other improvements include using adaptive population size for each generation \cite{population}, using adaptive distance function \cite{ref:adpt_dis} and switching kernels \cite{ref:kernel}, etc.
+In our implementations these options were tried when applicable and suitable to compare the behaviours of the algorithm.
 
@@
 
+In addition, ABC SMC is also capable of model selection tasks \cite{model_compare}.
+Instead of draw samples from prior of a single model, we can draw samples from multiple models $m_i$ and approximate $p(m_i, \theta|D)$, which is the joint probability distribution of model $m_i$ and its accepted particles $\theta$.
+In each generation the marginal distribution $p(m_i|D)$ can be calculated as the accepted ${\theta}$ have been obtained, thus models can be ranked according to their probabilities.
+
+The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modeling problems, especially in the area of biology \cite{ref:abcsysbio, ref:compare, ref:disease}.
+But ABC is not limited to these areas, as it can be generally used in inference tasks to give parameter estimates or model rankings.
+E.g., it has been successfully applied in cosmology studies \cite{cosmology}.
+
 \section{Software tools}
@@
  [workflow and developing process]
+
+ As we decided to apply ABC SMC on our regeneration models for its high-dimensional parameter estimation and possibly comparison of different models, a code development and experiments workflow was expected to be built.
+The ABC SMC algorithm needs a whole set of mathematical environments (e.g.
+class of distributions, models and particles) and closely linked sampling and fitting algorithms (e.g.
+KDE fitting and Gaussian sampling), thus suitable platforms with related packages are preferred.
+Regarding this, several packages in \verb|R| and \verb|Python| were studied.
+
+ A software packages that integrally implement ABC SMC is preferred, and more options in the settings of the algorithm, more customisable features can make it better.
+Moreover, as the ABC SMC itself is a computationally intensive task that consists of massive parallel potentials, an ideal software is expected to support as least one kind of parallel options e.g.
+multi-core, GPU accelerating and clusters.
+Otherwise, a modification or rewrite of the software package are expected to produce a reliable input-output framework for our parameter estimation and model selection experiments.
+
+ Preliminarily \verb|pyABC| \cite{ref:pyabc} in \verb|Python| was chosen for our first implementations.
 
@@
 
+    \bibitem{threshold} Silk, D., Filippi, S.
+and Stumpf, M.P., 2012.
+Optimizing threshold-schedules for approximate Bayesian computation sequential Monte Carlo samplers: applications to molecular systems.
+arXiv preprint arXiv:1210.3296.
+
+    \bibitem{population} Klinger, E.
+and Hasenauer, J., 2017, September.
+A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo.
+In International Conference on Computational Methods in Systems Biology (pp.
+128-144).
+Springer, Cham.
+
+    \bibitem{model_compare} Toni, T. and Stumpf, M.P., 2010. Simulation-based model selection for dynamical systems in systems and population biology. Bioinformatics, 26(1), pp.104-110.
+
+    \bibitem{cosmology} Weyant, A., Schafer, C.
+and Wood-Vasey, W.M., 2013.
+Likelihood-free cosmological inference with type Ia supernovae: approximate Bayesian computation for a complete treatment of uncertainty.
+The Astrophysical Journal, 764(2), p.116.
+
     \bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C.

### 2020-08-09 MSc-dissertation.tex
New structure

--- 2020-08-08 MSc-dissertation.tex
@@
 
-\begin{titlepage}
-    \vspace*{2in}
-    % an acknowledgements section is completely optional but if you decide
-    % not to include it you should still include an empty {titlepage}
-    % environment as this initialises things like section and page numbering.
-    \section*{Acknowledgements}
+% \begin{titlepage}
+%     \vspace*{2in}
+%     % an acknowledgements section is completely optional but if you decide
+%     % not to include it you should still include an empty {titlepage}
+%     % environment as this initialises things like section and page numbering.
+%     \section*{Acknowledgements}
 
-    This template is a slightly modified version of the one developed by
-    Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
-    acknowledgement follows:
+%     This template is a slightly modified version of the one developed by
+%     Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
+%     acknowledgement follows:
 
-    {\em This template has been produced with help from many former students who
-    have shown different ways of doing things. Please make suggestions for
-    further improvements.}
+%     {\em This template has been produced with help from many former students who
+%     have shown different ways of doing things. Please make suggestions for
+%     further improvements.}
 
-\end{titlepage}
+% \end{titlepage}
 
@@
 
-\chapter{Introduction}
+\input{introduction/introduction.tex}
 
- % This should contain a description of your project and the problem you
- % are trying to solve. Where appropriate you should also include
- % references to work which has already been done on your topic and
- % anything else which lets you set your work in context.
+\input{background/background.tex}
 
- % One of the things you will need to do is to ensure that you have a
- % suitable list of references.  To do this you should see \cite{ref:lam}
- % or some other suitable reference.  Note the format of the citation used
- % here is the style favoured in this department.  Here is another
- % reference \cite{ref:bloggs} for good measure.
+\input{modelling/modelling.tex}
 
- [background and motivation]
+\input{implementation/implementation.tex}
 
- [what is done]
+\input{performance/performance.tex}
 
-While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions of cells and chemicals in a biological system.
-The objective process in this dissertation is an example of complex biology system found in tissue regeneration.
-Zebrafish has the ability to regenerate its spinal cord after injury; at the injury-site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.
+\input{conclusion/future_works.tex}
 
-Our interest lies in the modeling of the regeneration process that happens in the injury-site, especially in the parameter estimation and comparison of different models.
-According to available researches \cite{ref:Tsarouchas}, a mathematical model is proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments.
-However, we would like to explore the the `best fit' if the model which comes with a set of parameter estimates given the observed data.
-Also, we wish to compare different it with different models and possibly find the best model to describe the data without losing biological and mathematical properties in the regeneration process.
-
-Fo the parameter estimation tasks, Bayesian inference techniques can help in the problems of over-fitting and local optima \cite{ref:abcsysbio}, compared to optimisation based methods.
-Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression can be hard.
-
-Such methods introduce computation-intensive sampling-rejection pattern, where massive number of samples are drawn, followed by computations.
-As the sampling process is independent from each other in a certain period, the parallel execution becomes feasible and is expected to accelerated the parameter inference in a large scale.
-
-This dissertation includes the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results.
-Model comparison results are also included using the same method.
-Additionally, a performance test illustrates the strong-scaling performance of the algorithm.
-
-\chapter{Background}
-
-\section{Zebrafish spinal cord regeneration}
-
- [describe the process and how cells and cytokines are involved]
-
-Compared to mammals, zebrafish can regenerate spinal cord after injury, and possibly recover swimming function.
-They are widely used in regenerative ability researches.
-Experiments found that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}.
-Tsarouchas et al.
-\cite{ref:Tsarouchas} studied the the complex interactions happen at the lesion site.
-The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration.
-Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, affect the inflammation, which are dynamically controlled by immunes cells, i.e.
-cytokines works as media to control the regeneration.
-Here cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.
-
-An interaction map can be drawn, according to the evidences found in \cite{ref:Tsarouchas}.
-After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells: macrophage, neutrophil and keratinocyte.
-The dynamical system consists of complex interactions between cells and cytokines, e.g.
-il-1$\beta$ promotes macrophages' increase.
-and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
-Consequently, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration.
-Generally, il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect, and tnf-$\alpha$ promotes the regeneration all the time.
-The presence if the cytokines also affects the number of cell, while as sources number of cells also affects cytokines' expression.
-Interactions between different kinds of cells are also considered in the system.
-
-This is a dynamic system suitable for mathematical modelling, as a set of equations can be written to represent the dynamics as possible.
-Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions iniside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and our own hypothesis which may include different interactions.
-
-The biological background of works in this dissertation is mostly related the the experimental evidences, quantitative data and proposed hypothesis in \cite{ref:Tsarouchas}.
-The published data on the quantitative measurement of cells and cytokines make the modelling and further analysis possible.
-Some other interactions proposed in our modelling stages are enlightened from Anderson et al.
-\cite{Anderson}.
-
-\section{Mathematical modelling}
-
- [how this can be modelled as dynamic systems]
-
- [general topics of mathematical models: how to build model according to interactions; how to calibrate/ parameterise the model; how to solve the model; dynamics; how to evaluate the model]
-
-Mathematical models can express real-life questions in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques.
-Mathematical models can be used for prediction, classification, testing our understanding or hypothesis and many other tasks, thus they are widely used to solve practical problems.
-However, for a model that describe the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models.
-In this case, we wish to use modelling techniques to test our inference method, our hypothesis and hopefully an well-fitted models can be found and compared.
-
-According to our understanding of the biological mechanism in the regeneration, hypothesis should firstly proposed, in a way which can represent the real  mechanism as much as possible.
-Usually a real-life problem, e.g.
-the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered, An elaborate description of the real cases using a mathematical model sometimes can be impossible; assumptions and hypothesis are proposed to make necessary simplification and abstract for a final reasonably detailed model.
-Some qualitative models are simple but illustrative for the real-life observed patterns, e.g.
-Lotka-Volterra (L-V) equation for predator and prey models; they are usually analytically solvable and some properties are used to make predictions (e.g.
-steady states in ecosystem).
-If a more precise model is desired, usually more effects or interactions are considered and added as terms to a re-constructed model, e.g.
-environmental factors, a changing environmental capacity and predation from other species.
-
-Models like L-V equation are mostly analytically solvable when analysing properties of the model, however, analytical solution for many models are not feasible, because of more complex dynamics, or the model itself is not a deterministic model.
-In the areas that are not analytically feasible, usually computational techniques are applied for approximate solutions.
-Solvers for different models are used to simulate the system or find roots for some spacial points.
-
-For the regeneration process that we wanted to model, differential equations are our preference.
-Non-linear ordinary differential equations are used in this dissertation, where the only independent variable is time $t$ denoting the post-lesion time (hours), but further models can also consider other models e.g.
-stochastic models and partial differential equations.
-For ordinary differential equations (ODE) system, there are several available computational solvers using different algorithms, implemented in MATLAB\footnote[1] {\url{https://uk.mathworks.com/help/matlab/math/choose-an-ode-solver.html}}, Python\footnote[2]{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html}} and many other languages.
-These solvers help to simulate the trajectories of the dependent variables, given initial values and a set of fixed parameters.
-The goodness of `fit' then can be measured by compared the simulated data and observed real data.
-
-Often further researches on the modelled problems require the model to be deterministic on its parameters.
-Parameters in the dynamic system models usually have their practical significance, e.g.
-describing reaction speed, environmental capacity or the birth rate.
-There are many approaches to derive these parameters from data or experience.
-For a specific problem and a widely used models, the parameters' values can be read from previous literatures; however in many others cases, we need to find parameter's values specifically for our model and problem, where numerical values for parameters are more suitable for practical analysis using ODE solvers or other tasks.
-
-To find the best parameters' values given observed data, either optimisation-based approaches (e.g.
-least-square (LS) fit) or inference approaches can be used.
-Optimisation-based approaches may face two major drawbacks: over-fitting and local optima \cite{ref:abcsysbio}.
-Bayesian inference approaches gain popularity in recent years, which can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
-Local optima can still be a major concern, and many efforts have be proposed for a more through explore of the parameter space.
-
-\section{Bayesian inference}
-
- [how to parameterise the model]
-
- [how to infer the parameter given data]
-
- [likelihood-free inference and information about ABC SMC]
-
- [model comparison, ABC SMC for model comparison]
-
- Bayesian inference approaches put Bayes Rule in the centre:
-
- \begin{align}
-    \label{eq:bayes}
-    p(\theta|D) = \frac{p(D|\theta)p(\theta)}{P(D)}
-\end{align}
-
-where $p$ is the probability density function, $\theta$ denotes the parameter set, $D$ is the given observed data.
-
-Alternatively, as $p(D)$ is fixed once the observed data is given, the rule can be rewritten as
-
-\begin{align}
-    \label{eq:bayes2}
-    p(\theta|D) \propto l(D|\theta)p(\theta)
-\end{align}
-
-where $l$ represents the likelihood function.
-In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target.
-Prior $p(\theta)$ should express our initial believes or estimates on the parameter valeus before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers.
-Once the data is given, $\theta$ is the only parameter in the likelihood function.
-Some optimisation-based methods try to find an estimation of parameter set $\theta$ s.t.
-maximise the likelihood, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative.
-
-Based on Eqn \ref{eq:bayes2}, to find the posterior of parameter set, prior can be set according to the problem context and our understanding and cover a reasonable ranges; computational methods can be used to approximate likelihood function, e.g.
-Markov Chain Monte Carlo (MCMC) \cite{ref:MCMC}.
-As another alterative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood, but focused on simulating the data from the given model and observed data, then used the simulated data to approximate the desired posterior.
-ABC-rejection \cite{ABC_rejection} is a simple example of the ABC process, and further algorithms e.g.
-sequential Monte-Carlo (ABC SMC)\cite{Toni} are proposed to improve the efficiency.
-
-General steps for ABC algorithm includes the following (ABC-rejection):
-
-\begin{enumerate}
-    \item Draw a sample $\theta^*$ from parameters' prior distribution. $\theta^*$ is a vector that consists values for each parameter
-    \item Plug $\theta^*$ into the model, and use the model to generate simulated data $D^*$, where $D^*$ is of the same form as the observed data $D$, e.g. includes same summary statistics
-    \item The discrepancy between the simulated data and observed data is measured by a distance function $d(D^*, D)$.
-For a pre-fixed threshold value $\epsilon$, if $d(D^*, D)<\epsilon$ then the sample $\theta^*$ is accepted
-    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution can be approximated by these samples
-\end{enumerate}
-
-In our study of the regeneration model, full trajectories of the dependent variables are used as the data to be compared, as our goal is to find a model and a parameter set that can intuitively recover the real process as much as possible.
-The mean values at each time point for each dependent variable is calculated to form $D$, and for each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$.
-Using other statistics to summary the raw data are also possible bur care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
-
-Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of deceasing threshold values $\epsilon_t$, and derive a new population (called generation) of accepted sampled from the previous accepted samples (called particles).
-An example is ABC SMC \cite{Toni}, which can significantly reduce the required samples compared to ABC-rejection \cite{ref:disease}:
-
-\begin{enumerate}
-    \item 
-\end{enumerate}
-
-\section{Software tools}
-
- [existing tools and software for this task]
-
- [workflow and developing process]
-
-\chapter{Mathematical modelling}
-
-Mathematical models can describe different kinds of dynamic systems, and thus can be used as a guide for prediction and analysis.
-An ideal model in our case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  data features in different time point.
-
-Proposed models for the regeneration process are in the form of ordinary differential equations (ODE), specifically the time differential form.
-Terms in the ODE are mostly explainable and corresponds interaction paths.
-
-\section{Observed data}
-
-Our models is built on the basis of the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement data include the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
-As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
-According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration' (compared to the value at first time point).
-
-It is noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations is mean of measurement data at each time point, assuming that measured data is Gaussian-like distributed.
-To validate this, the distribution of the measured data points is plotted and examined to see if the mean value can represent the distribution.
-The result is that at most time points the measurement values are Gaussian-distributed, although some distributions are skewed.
-One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
-Mean value can summarise most data and thus is still used as the target observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}.
-
-\begin{figure}
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/obs_curve.png}}
-    \end{center}
-
-    \caption[Mean of the observed data]%
-    {Mean of the observed data for neutrophil ($N$), macrophage ($\Phi$), il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean}
-    \label{fig:obs_data}
-
-\end{figure}
-
-\section{Hypothesis and Models}
-
-5 models in total are proposed according to different hypothesis.
-At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
-After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can represent the observed regeneration process better or can be used to test our hypothesis.
-
-All these models assume the involved interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for the inference task.
-Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data available.
-
-\subsection{Basic model}
-
-A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the inference framework code.
-An interaction map of the model is shown in Figure \ref{fig:m1}.
-This model is a simplification a the process described in \cite{ref:Tsarouchas} with some minor interactions ignored.
-To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
-number of cells as `cell', and denote the unit of $\beta$ and $\alpha$ i.e.
-relative mRNA expression as `unit' for simplicity.
-
-\begin{figure}
-    \begin{center}
-        \resizebox{0.4\hsize}{!}{\includegraphics{fig/model1.png}}
-    \end{center}
-
-    \caption[Interactions modelled in the basic model]%
-    {Interactions modelled in the basic model (model 1) based on Tsarouchas et al.\cite{ref:Tsarouchas}.
-Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
-    \label{fig:m1}
-
-\end{figure}
-
-It is assumed that there are negative feedbacks for all the variables. MORE DESCRIBE.
-
-\begin{align}
-    \label{eq:model1}
-    \begin{split}
-        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
-        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-    \end{split}
-\end{align}
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|c c c|}
-        \hline
-        Parameter            & Definition                                       & Units                    \\ [0.5ex]
-        \hline\hline
-        $\lambda_N$          & Self-increase rate of neutrophil                 & $cell/h$                 \\
-        $\kappa_{N\beta}$    & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
-        $\mu_N$              & Coefficient of negative feedback of $N$          & $h^{-1}$                 \\
-        $\nu_{N\Phi}$        & Coefficient of inhibition of both $N$ and $\Phi$ & $cell^{-1}\cdotp h^{-1}$ \\
-        \hline
-        $\lambda_\Phi$       & Self-increase rate of macrophage                 & $cell/h$                 \\
-        $\kappa_{\Phi\beta}$ & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
-        $\mu_\Phi$           & Coefficient of negative feedback of $\Phi$       & $h^{-1}$                 \\
-        \hline
-        $s_{\beta N}$        & Production rate from $N$                         & $unit/(cell\cdotp h)$    \\
-        $i_{\beta\Phi}$      & Coefficient of inhibition to the production      & $cell^{-1}$              \\
-        $\mu_\beta$          & Coefficient of negative feedback of $\beta$      & $h^{-1}$                 \\
-        \hline
-        $s_{\alpha\Phi}$     & Production rate from $\Phi$                      & $unit/(cell\cdotp h)$    \\
-        $\mu_\alpha$         & Coefficient of negative feedback of $\alpha$     & $h^{-1}$                 \\
-        \hline
-    \end{tabular}
-    \caption{Parameters introduced in the basic model (model 1) [REMOVE unit]}
-    \label{table:m1}
-\end{table}
-
-\subsection{Alternative models}
-
-\paragraph{Model 2 and model 3}
-
-As the observed data indicate, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
-Regarding this, the self-increase parameter $\lambda$ cannot be constant, thus a exponentially decaying $\lambda$ term is introduced and model 2 is proposed.
-(Eqn.
-\ref{eq:model2}).
-Also, the inhibition of il-1$\beta$ production, i.e.
-term $i_{\beta\Phi}\Phi$ is considered to be ignored for a more simple model, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
-This case corresponds to model 3, written as Eqn.
-\ref{eq:model3}.
-
-Model 2 and model 3 introduce one extra parameter $a$, which is a coefficient in the exponentially decaying $\lambda$, determining the decay speed, with the unit $h^{-1}$.
-
-\begin{align}
-    \label{eq:model2}
-    \begin{split}
-        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
-        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-    \end{split}
-\end{align}
-
-\begin{align}
-    \label{eq:model3}
-    \begin{split}
-        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
-        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-    \end{split}
-\end{align}
-
-Model 3 can be regarded as a simplification of model 2, as it can be treated as model 2 with parameter $i_{\beta\Phi}=0$.
-Among the proposed three models, model 1 is a naive one that was proposed at very first time and used as an ODE `template' to build and test parameter inference framework.
-As the implementation was successful, model 2 and 3 was proposed as we were trying to calibrate some terms and find a better model.
-After fitting, model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points (which relate to the steady state).
-Model 3 makes a small simplification on model 2 and is theoretically less `general' than model 2.
-
-\paragraph{Model 4 and model 5}
-
-After the first model selection experiment, it was found that some significant features presented in the observed data were not recovered by any of the existing model.
-To resolve this, attempts were tried to introduce more interactions within the dynamic system considering the biological and  mathematical context.
-Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or adding a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
-\ref{eq:model4}) and model 5 (Eqn.
-\ref{eq:model5}).
-
-\begin{align}
-    \label{eq:model4}
-    \begin{split}
-        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
-        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha+d_{\beta\alpha}\beta
-    \end{split}
-\end{align}
-
-\begin{align}
-    \label{eq:model5}
-    \begin{split}
-        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
-        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=(s_{\alpha\Phi}+f_{\beta\alpha}\beta)\Phi-\mu_\alpha\alpha
-    \end{split}
-\end{align}
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|c c c|}
-        \hline
-        Parameter         & Definition                                                     & Units                    \\ [0.5ex]
-        \hline\hline
-        $d_{\beta\alpha}$ & Coefficient of promoting effect from $\beta$                   & $h^{-1}$                 \\
-        \hline
-        $f_{\beta\alpha}$ & Coefficient of promoting the production of $\alpha$ by $\beta$ & $cell^{-1}\cdotp h^{-1}$ \\
-        \hline
-    \end{tabular}
-    \caption{Parameters introduced in model 4 and 5}
-    \label{table:m45}
-\end{table}
-
-\subsection{Model evaluation and comparison}
-
-[general topics on evaluating a model]
-
-[bayes factor for model selection]
-
-\subsection{Limitations}
-
-[hypothesis]
-
-[available data]
-
-[model misspecification]
-
-% \begin{figure}
-
-% \begin{center}
-% \resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
-% \end{center}
-
-% \caption{The University Crest}
-% \label{fig:eucrest}
-
-% \end{figure}
-
-% see the man page for dvips for details of the special command which is
-% much more powerful than is shown here. It allows offsets in the
-% horizontal and vertical and scaling in x and y.
-
-% choosing suitable values for offset and scale can be a tiring matter
-% of trial and error.
-
-% note that labels do not need to include a description of the object
-% they are labelling but it can be helpful, eg \label{fig:figurename}.
-
-% You can use a label on a figure to refer to it later. The university
-% crest is in \ref{fig:eucrest}. Note that you should not use phrases like
-% ``the figure above'' or ``the following figure'' since Latex may move
-% the figure relative to the text if it cannot be fitted onto the current page.
-
-\chapter{Parameter estimation and model comparison}
-
-Estimation of  parameters in the proposed models is the main task of this section, which involves the implementation of ABC SMC and several experiments.
-ABC SMC-based model selection is another task that can be easily conducted with the developed parameter inference framework.
-
-The build and test of code were under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
-Some other code e.g.
-shell script and \verb|R| code are also used to perform the experiments and analyse the results.
-Software and system environment used in our implementation can be found at Appendix A.
-The build and test were performed on local computers and HPC facilities available within EPCC.
-
-\section{Implementations and code}
-
- [ABC implementations details, e.g. distance, population, ODE related functions]
-
- [how the code is developed and built]
-
-According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
-\verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}.
-\verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks.
-Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.
-
-The inference framework code for the project was developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task.
-The scaling-up experiment was performed on Cirrus, and the profiling and its analysis were performed on local machine.
-Some development and debug were on the remote machines to ensure experiments run correctly on compute node.
-
-The ODE model-related code and utilities were firstly developed to enable reusable functions, variables and data calls.
-It includes the code for the ODE models, ODE solver, data structures, data format transform functions etc.
-By doing this in a separate code file, we can change options and activate ABC SMC easily in the implementation code file, without creating definition and duplicated code for the models and parameter priors.
-
-Some modifications were made to the source code of \verb|pyABC|.
-For example, the laboratorial measured data were not considered `complete' for modeling and analysis purpose: the available time points for cells and cytokines are different, e.g.
-cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl.
-These missing values were treated as `ignored' when calculating the distance between observed data and simulated data.
-To cope with this built-in distance function code is modified.
-
-Analysis code was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results.
-For different tasks there are also separated analysis code file.
-Other analysis tools used in the project include built-in database visualisation tool, R code and Microsoft Excel.
-
-The project code was managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also made the cross-hardware synchronisation of code versions convenient.
-
-After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (udsing synthetic data), and code files for parameter inference (using experimental data).
-Model selection can be easily done by adding more model to the implementation code file.
-
-% \section{Parameter estimation}
-
-% [implementation details and options/ settings that are affecting the ABC]
-
-\section{ABC SMC and model hyperparameters experiments}
-
-As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental observed data.
-
-In the first part, synthetic data is used with known true parameter values.
-The algorithm implementation is evaluated by the efficiency and goodness of resultant model under different implementation options and ABC SMC hyperparameters.
-
-Then according to findings in the first part, several options with good performance are tried in the inference using experimental observed data (Figure \ref{fig:obs_data}).
-To obtain a more accurate and general results, these experiments are repeated for 3 or more times.
-
-    [including the options that are tested/ tried/ not adopted]
-
-    [including experiments results and discussions]
-
-    [how do these options correspond to biological model and how to set them accordingly for proposed model]
-
-The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options.
-To firstly test the ability of inference and observe the results of different options, synthetic data with known true parameter values are used as target data.
-The true value of parameters that are used to generate synthetic data is listed in Appendix B.1.
-These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data (Figure \ref{fig:infer_back_data}), the target data is then generated using these parameter values via ODE solver in \verb|Python|.
-
-\begin{figure}[h!]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
-    \end{center}
-
-    \caption[Synthetic data generated with known parameter values]%
-    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot).
-Known parameter values are obtained from a least square fitting of the observed data}
-    \label{fig:infer_back_data}
-
-\end{figure}
-
-[TABLE for known parameter values]
-
-[why use synthetic data]
-
-Among them the following topics are studied using the synthetic data.
-
-\subsection{Perturbation kernels}
-
-Perturbation kernels work in the sampling process.
-In each generation $t$, samples are firstly taken from the previous population $\{\theta_{t-1}\}$ with weights $\{w_{t-1}\}$, then perturbed using the perturbation kernel $K(\theta|\theta^*)$.
-We keep sampling until $N$ particles are accepted, where $N$ is the pre-set population size.
-After that the new weights are calculated and normalised.
-
-\begin{figure}[t!]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
-    \end{center}
-
-    \caption[ABC SMC sampling process]%
-    {ABC SMC sampling process.
-$\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$.
-$d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data.
-for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
-    \label{fig:smc}
-
-\end{figure}
-
-The perturbation kernel is called in the sampling of every particles and determines how the new perturbed particle is chosen, thus it is influential to both computation complexity and the resultant posterior distribution.
-Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations, or could generate a population with a higher acceptance rate if the successive epsilon values are close; A kernel with wide variance, or spreading out in a large space could help in resolving the local optimal problem by a more thoroughly exploration of the parameter space, however it can be more computation-intensive and result in a lower acceptance rate.
-A desired optimal kernel should balance the trade-offs; their property and criteria is discussed in \cite{ref:kernel}.
-
-There are several common choice of perturbation kernels.
-Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.
-A covariance matrix $\Sigma_t$ of accepted particles is calculated form previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$.
-It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in relecting the true posterior structure \cite{ref:kernel}.
-It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks.
-The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `narrower distributed' perturbation result.
-
-Local M-NN kernel provided by \verb|pyABC| is also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.
-
-The kernel experiments is designed to explore the efficiency of SMC on our dynamical systems.
-Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rate would be our preference.
-The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters.
-The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference.
-As the threshold schedule is fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit i.e.
-recover of target data trends/features and errors of the inferred parameters compared to true values is not discussed.
-
-\subsubsection{Kernel experiment results}
-
-To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e.
-$\epsilon_t$ set to 10.0.
-From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared.
-Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
-has higher acceptance rates among almost all generations. Local M-NN with M=750 is the slowest kernel.
-
-For local M-NN kernels, generally a greater M will lead to lower acceptance rate and
-more required particles in each generation.
-As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve.
-Consequently, if greater M is test, then it will have even lower acceptance rates; the maximum M=2000 (whole population is considered) will have the lowest acceptance rates.
-
-\begin{figure}
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel1.png}}
-    \end{center}
-
-    \caption[Total required samples of different kernels]%
-    {Total required samples of different kernels after 20 populations (2000 particles in each population).
-Different color represents different generations (bottom to top: population 1 to population 20)}
-    \label{fig:kernel1}
-
-    \vspace*{\floatsep}
-
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
-    \end{center}
-
-    \caption[Acceptance rates of different kernels]%
-    {Acceptance rates of different kernels in 20 populations. Each population has 2000 particles}
-    \label{fig:acceptance1}
-
-\end{figure}
-
-Multivariate normal kernel has a performance between local M-NN M=250 and M=100.
-It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
-`Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
-Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1.
-Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
-
-It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
-This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
-local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold.
-In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
-Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}.
-One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
-
-\subsubsection{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
-local M-NN with M$\leq 250$ (for population size of 2000, i.e.
-12.5\% of the population) and multivariate normal with scaling.
-Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.
-
-% \begin{figure}
-%     \begin{center}
-%     \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
-%     \end{center}
-
-%     \caption{Acceptance rates of different kernels}
-%     \label{fig:acceptance1}
-
-% \end{figure}
-
-\subsection{Adaptive functions and factors}
-
-[how adaptive distance work]
-
-SMC can be more adaptive by introducing adaptive distance function and adaptive population; besides, factors can be applied to manually `normalise' the data.
-
-The trivial distance function is set to Euclidean distance (2-norm distance)
-
-\begin{align}
-    \label{eq:dis}
-    D=\sqrt{\sum_i \Delta x_i^2}
-\end{align}
-
-where $i$ is the index of data points and $\Delta x_i$ is the discrepancy between  observed data and simulated data at data point $i$, i.e.
-$\Delta x_i = x_{i, simulated}-x_{i, observed}$.
-In this case data points are 12 time points of four variables i.e.
-48 data points in total.
-
-A weighted 2-norm distance can be written as
-
-\begin{align}
-    \label{dis_w}
-    D=\sqrt{\sum_i w_i \Delta x_i^2}
-\end{align}
-
-where a weight $w_i$ is assigned to every data point.
-If data point $i$ have a higher weight, then the data value is regeared to be more informative, i.e.
-giving more help in inferring the true posteriors; weights can be either pre-set according to prior knowledge of the problem, or using adaptive method to be dynamically calculated according to , known as adaptive distance.
-
-Adaptive distance is changing the weights of all data points after each generation iteration, trying to assign informative data points higher weights.
-Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}
-
-\begin{align}
-    \label{dis_f}
-    D=\sqrt{\sum_i f_iw_i \Delta x_i^2}
-\end{align}
-
-Factor is helpful as an option for efficiency.
-When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data point, or behave like a normalisation that can balance the scales of different summarised statistics (here is the 48 mean values).
-There are two appliance case studied: (1) factors used as an normalisation option and (2) factors used to give more focus on some features of the observed data.
-For (2), as we noticed that the main features (e.g.
-rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points i.e.
-0 - 30 hpl, factors are tried in the later parameter estimation of real data where a more accurate fit of the curve is desired.
-
-Adaptive population \cite{ref:adpt_pop} can adapt the population size of each generation according to the mean correlation of variation error of the previous population.
-In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.
-
-    [what is factor]
-
-    [what is adaptive distance]
-
-\subsubsection{Experiments results}
-
-[FIGURE here]
-
-Two types of factors and adaptive distance function are tested by running the same number of generations (20 generation, 2000 particles per generation).
-From FIGURE, by applying factors the total required samples are less, and the adaptive distance requires much less samples; their acceptance rates are also higher than the standard one.
-It seems that adaptive function and factors could help to converge more quickly.
-
-However considering the result after 20 generation (FIGURE), the resultant model gives the contrary preference.
-Applying factors and using adaptive distance can lead the resultant model less accurate (after the same generation iterations): they have wider inter-quantile range and may requires more generations to converge.
-
-\subsubsection{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases.
-However, as the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, a direct compare of the efficiency under the same threshold schedule is unfeasible; after the same number of generations they can reach a small $\epsilon_t$ but still the approximated posterior is not as accurate as the standard implementation.
-AS a result, our implementation of the parameter inference will only consider standard distance function and factors are only tested for improvements (FIGURE).
-
-\subsection{Data size, prior distribution and population}
-
-[experiments plan]
-
-Some other hyperparameters, e.g.
-feed-in data size, prior distribution, number of populations and number of particles in each population were also of our interest.
-Experiments were designed to explore how these options affect the goodness of fit and efficiency of the implementation.
-
-Regarding the dynamic systems model, SMC was applied with two data size options, three prior distribution (uniform distribution, wider uniform distribution and log-uniform distribution).
-The population options i.e.
-population size and number of populations were also tried with different values.
-These experiments intended to give suggestions on the later SMC implementation on experimental data.
-
-\subsubsection{Experiments results}
-
-The results from data size and prior distribution range is shown in FIGURE.
-The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement.
-Although much less (60\%) data is used, the total required sampling numbers is not much less (XX\%).
-Although the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data (FIGURE), cares should be taken when using less data point, or using other more summative statistics, e.g.
-mean, standard deviation, peak values, starting and ending values etc.
-As shown in \cite{ref:disease}, less data can lead to a model with more variance, under-fitting or missing some local features e.g.
-local peak.
-
-The prior range and distribution can largely affect the execution of SMC (FIGURE and FIGURE).
-Samples are taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space.
-It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.
-
-    [discuss uniform and log uniform when result ready]
-
-\begin{figure}
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/size1.png}}
-    \end{center}
-
-    \caption[Data size and prior distribution range experiment]%
-    {Data size and prior distribution range experiment. (a) Total required samples, (b) Epsilon values under median epsilon schedule}
-    \label{fig:size}
-
-\end{figure}
-
-[FIGURE HERE]
-
-\subsubsection{Summary} This subsection aims to preliminarily explore ABC SMC settings using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options.
-Conclusion can be drawn to choose efficient and proper options of ABC SMC to be applied on the target models with experimental data.
-It also provides reference for a general inference task using SMC with high dimensional parameter space.
-
-\section{Parameter estimation and model comparison}
-
-\subsection{Model 1, 2 and 3}
-
-[separated ABC run on model 1, 2 and 3]
-
-Using the suggested options in the previous section, the target data (Figure \ref{fig:obs_data}) is prepared as input for the SMC inference framework.
-Regarding the result, the population size and number of generations are adjusted several times to obtain an general informative posterior which should be stable across repeated runs and avoid local optimal as much as possible.
-
-Parameter estimation of model 1, 2 and 3 was tried with different prior distribution setting: distribution range [0, 25] and [0, 75], amd distribution type uniform and log-uniform.
-All these ABC SMC runs uses 2,000 particles per population and 30 generations.
-
-Log-uniform-shaped prior seems to give a better fit of the model, and wider prior range [0, 75] does not give a more accurate result.
-Then distribution range [0, 25] was tried, in case some parameter have a true posterior laying between [25, 50].
-
-The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}.
-For model 3, the acceptance rates and epsilon path are presented in Figure \ref{fig:result123_2}.
-For model 3, the estimated parameters' posterior is shown in Figure \ref{fig:para1}.
-Figure \ref{fig:result123} plots the estimated curve of models, using the mean value of each parameter's approximated posterior; also 1000 particles (each particle is a parameter set) were sampled and used to generate simulated trajectories of the four variable, and the mean (blue) and 25th to 75th percentile range (grey) are calculated from the 1000 simulated data.
-
-Before applying model comparison methods, some features of the three model can still be compared.
-From simulated curves, It can be seen that all the inter-quartile ranges are tight around the mean value curve, which indicates the approximated posterior are concentrated to some degree; the peak-shaped posterior distributions (Figure \ref{fig:para1}) agrees with this.
-All epsilon values converges to a steady level, and model 3 has a lower final epsilon value and consequently the simulated trajectory is more close to the observed data (model 3, Figure \ref{fig:result123}).
-
-The acceptance rates is fluctuating and have a gradually increase trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated.
-Compared to model 1, model 2 gives a better fit of the decreasing trend at the second half of the observed data, by using a exponentially decaying self-increase rate $\lambda_N$ instead of a constant.
-The resultant simulated data from model 2 and 3 are in the similar trends and the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'.
-We expect model 3 to be the best model as it reaches a threshold value and intuitively gives a better fit, although the fit could not be considered generally a good representation of the biological process that we want to model, as the simulated data are significantly biased from the observed data for tnf-$\alpha$.
-A sharp peak at 4 hpl is observed from the measurement (Figure \ref{fig:obs_data}) but no similar features are represented by any of the three models.
-
-\begin{figure}
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve123.png}}
-    \end{center}
-
-    \caption[Simulated trajectory from the 20th population of model 1, 2 and 3]%
-    {Simulated trajectory from the last population of model 1, 2 and 3.
-Error bar indicates SEM.
-500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from}
-    \label{fig:result123}
-
-    \vspace*{\floatsep}
-
-    \begin{center}
-        \resizebox{0.6\hsize}{!}{\includegraphics{fig/eps_acc123.png}}
-    \end{center}
-
-    \caption{Epsilon trends and acceptance rates of model 1, 2 and 3}
-    \label{fig:result123_2}
-
-\end{figure}
-
-\begin{figure}
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/para1234.png}}
-    \end{center}
-
-    \caption[Estimated posterior distribution of parameters in model 3]%
-    {Estimated posterior distribution of parameters in model 3. Shaded range indicates 25\%--75\% quantile of the population}
-    \label{fig:para1}
-
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/cmp1.png}}
-    \end{center}
-
-    \caption[Model comparison of model 1, 2 and 3]%
-    {Model comparison of model 1, 2 and 3. From population 8 to population 30, model3 always wins}
-    \label{fig:cmp1}
-
-\end{figure}
-
-[model comparison among model 1, 2, 3]
-
-Next a model selection experiment with different prior was conducted, using the same ABC-SMC framework.
-As in the parameter inference, two distributions (uniform and log-uniform) are tested with three different interval ([0, 25], [0, 50] and [0, 75]).
-The results show that model 3 wins with nearly 100\% model probability at most runs after 30 generations, except log-uniform [0, 75].
-All the model probability results are of great discrepancy (e.g.
-100\% model 3 and 0\% for model 1 and 2), so Bayesian factor is not calculated.
-
-Figure \ref{fig:cmp1} shows the model selection process, under log-uniform [0, 50] prior distribution.
-Model 3 gains total advantage over model 1 and 2 after 7th population.
-Together with conclusions above, model 3 is considered to be the the best model to represent the target data so far.
-
-\begin{table}[t!]
-    \centering
-    \begin{tabular}{|c c c|}
-        \hline
-        Parameter            & Estimated mean  & Estimated mean      \\
-                             & value (uniform) & value (log-uniform) \\[0.5ex]
-        \hline\hline
-        $\lambda_N$          & 14.1            & 13.3                \\
-        $a$                  & 9.67            & 0.0170              \\
-        $\kappa_{N\beta}$    & 18.0            & 0.259               \\
-        $\mu_N$              & 13.2            & 0.0180              \\
-        $\nu_{N\Phi}$        & 0.742           & 0.131               \\
-        \hline
-        $\kappa_{\Phi\beta}$ & 0.239           & 0.156               \\
-        $\mu_\Phi$           & 0.154           & 0.0350              \\
-        \hline
-        $s_{\beta N}$        & 6.75            & 1.21                \\
-        $\mu_\beta$          & 6.38            & 1.32                \\
-        \hline
-        $s_{\alpha\Phi}$     & 17.0            & 5.83                \\
-        $\mu_\alpha$         & 11.9            & 2.43                \\
-        \hline
-    \end{tabular}
-    \caption[Estimated parameter values of model 3]
-    {Estimated parameter values of model 3, using uniform prior distribution and log-uniform prior distribution respectively}
-    \label{table:estimated1}
-\end{table}
-
-Compared to uniform distributed prior, we found that log-uniform distributed prior is more suitable in the inference.
-Log-uniform results in an obviously better fit and narrower inter-quantile interval (Figure \ref{fig:result123} and \ref{fig:resultCurve_uni}), and the final epsilon value is also smaller.
-Log-uniform tends to assign values that are close to the left boundary (i.e.
-zero in our case) higher density, i.e.
-they are more likely to be sampled in the first population.
-As a result, the estimated parameter posteriors are expected to have smaller mean and/or skewed to left of the x-axis.
-The estimated values proved this (see \ref{table:estimated1}; also observed in other models and prior interval experiments).
-It indicates that some of the true parameter tend to have small values that are close to zero, and using log-uniform distribution could results in a faster and more satisfying inference.
-Regarding this, further experiments all use log-uniform distribution only.
-
-\subsection{Model 4 and 5}
-
-[features that derive model 4 and 5]
-
-We observed that the existing models cannot well represent the trajectory of relative expression of tnf-$\alpha$: from experimental data, a rapid increase of tnf-$\alpha$ expression is observed in 0--4 hpl, followed by a rapid drop; all the models cannot well recover this feature and most test results see a less-rapid increase followed by a gradually decree and the observed peak value is not reached (e.g.
-Figure \ref{fig:result123}).
-Also in some other case e.g.
-Figure \ref{fig:resultCurve_uni}, tnf-$\alpha$ trajectories is well fitted at the cost of under-fitting $\Phi$.
-Hence efforts had been spent to propose more alternative models that can possibly solve this problem.
-
-The two proposed alternative models is based on model 3 and try to add extra term in the tnf-$\alpha$ equation $\mathrm{d} \alpha/\mathrm{d} t$.
-As macrophage is the main source of tnf-$\alpha$ production, and a similar sharp increase-and-drop trend is observed in the il-1$\beta$ trajectory, two hypothesis and corresponding terms are propose as follows.
-
-\paragraph{Model 4} It is assumed that the expression of il-1$\beta$ would have a promoting effect, representing by a phenomenological term $d_{\beta\alpha}\beta$ and $d_{\beta\alpha}$ is a new model parameter (positive constant).
-The promoting effect here is regarded to have the equivalent effect as directly promoting by il-1$\beta$, but the underlying mechanism is unclear so far.
-In mathematical view, this additional term can accelerate the expression of tnf-$\alpha$ in the first few time points and help to recover the peak-shaped trajectory.
-The proposed model is written as Equation \ref{eq:model4}.
-
-\paragraph{Model 5} Alternatively, we considered promotions to the tnf-$\alpha$ production, i.e.
-$s_{\alpha\Phi}\Phi$.
-It is assumed that il-$\beta$ could accelerate the production process.
-An additional term $f_{\beta\alpha}\beta$ is introduced to production rate, with $f_{\beta\alpha}$ being a new model parameter (positive constant).
-This model meets the biological context considering the source of tnf-$\alpha$.
-The proposed model is written as Equation \ref{eq:model5}.
-
-[further comparison of model 3, 4 and 5]
-
-Model 4 and 5 are base on model 3, so in this phase experiments of these three models are conducted for separately parameter inference and overall model comparison.
-Based on our experience in conducted experiments, we used the following setting for parameter inference:
-
-\begin{itemize}
-    \item population size: 2000 and 5000
-    \item number of populations: 30 generations
-    \item prior: [0, 50] log-uniform
-    \item perturbation kernel: multivariate normal kernel
-\end{itemize}
-
-Factors (Equation \ref{dis_f}) are also tried to find if we can force the inference framework to give more importance on the first few data points.
-In this experiment, for each trajectory first 8 data points are assigned higher factors (0.75), and the rest 4 data points are assigned with lower factors (0.25).
-All these runs are conducted on remote machines and the output database files are retrieved form analysis.
-
-Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 2000.
-It can be seen that the addressed problem in the trajectory of tnf-$\alpha$ is partially relieved in model 4 and 5: compared to model 3, a peak appears around 4 hpl; it can represent more features in the observed data and consequently the final reached epsilon (after 30 populations) is smaller than that of model 3, which proves that our proposed modifications in model 4 and 5 are helpful in recovering more features in observed data.
-
-\begin{figure}
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve345.png}}
-    \end{center}
-
-    \caption{Simulated data from the last population of model 3, 4 and 5}
-    \label{fig:resultCurve345}
-
-\end{figure}
-
-By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
-Model 1, 2 and 3 con also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- is not fitted.
-Some features, e.g.
-fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.
-
-After applying factors, the results were not largely different.
-As expected, applying factors made the data points at first half (where exists most of the fluctuations) more `focused', however the inferred models are close to one without factors, which could indicate that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.
-
-When we are trying more populations, an `over-fitting' behaviour is observed at the trajectory of macrophage (FIGURE).
-In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).
-
-Next, we performed a model selection on model 3, 4 and 5.
-FIGURE shows the model probabilities at different generations.
-From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred.
-Again, as DISCUSSED in the model comparison of model 1, 2 and 3, we considered the the violently changing model probability is related to the inference ability at certain epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
-Given the present result, model 5 is regarded as the best model.
-Some further experiments tried larger population size (10,000 particles per population) and more populations (36 populations), but the resultant model did not have a significantly better fit of the observed data, although much more computation resources are used.
-
-A inferred parameter posterior distribution of model 5 is shown in FIGURE, the estimated mean values is shown in TABLE. SOME DESCRIPTION.
-
-\subsection{Sensitivity of parameters}
-
-[why last PCs]
-
-[which parameters are most sensitive]
-
-[compared to credible intervals plot]
-
-[what does that mean]
-
-A quick sensitivity analysis of the dynamical systems models can be quantified using principle component analysis (PCA) \cite{Toni}, where the first PCs corresponds to sloppy parameters while the last PCs corresponds to stiff parameters \cite{sensitivity}.
-PCA can only provide rough approximated sensitivity behaviour.
-As models 5 won in the model selection, a further interest is to explore the parameter-to-model sensitivity
-
-FIGURE shows the results of PCA on the last population of inferred model5.
-The last PC (PC12) is mostly consists of the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive.
-If we conclude the last 5 PCs (FIGURE), then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$.
-As a conclusion, model 5 is sensitive to changes in the above parameters, which all come from the equation of neutrophil and macrophage i.e.
-equations of the cells; it may suggests that given the observed data, the inferred model is more sensitive to cells' dynamic rather than cytokines' dynamics.
-
-Also, some visualisations of the last population agreed with conclusions from PCA.
-Approximated posteriors of the last population plotted in FIGURE and credible ranges of parameters across populations in FIGURE shows that the 5 parameters identified by PCA is indeed `stiff' parameters.
-
-\chapter{Performance experiments}
-
-The performance experiments were designed to explore the parallel performance of ABC SMC inference framework which were used in the parameter estimation and model selection tasks.
-Usually ABC SMC is a time-consuming and computation intensive task and ideally executed on large clusters.
-The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
-
-    [MORE meanings of the performance study]
-
-\section{Scaling-up}
-
-First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
-
-\begin{itemize}
-    \item Prior distribution: default to log-uniform distribution [$1\times 10^{-6}$, 50] for all the 12 parameters
-    \item threshold schedule: median epsilon
-    \item No factors, no adaptive distance or adaptive population applied
-    \item Population size is 2000, with 20 generations
-\end{itemize}
-
-[HOW PYABC parallelise the sampling]
-
-For HPC systems like Cirrus, \verb|pyabc| uses \verb|multiprocessing| for multi-core parallel sampling.
-By default if the number of cores is not specified, it will automatically read the number of available cores and use them all.
-Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
-
-The program is executed on Cirrus, using 8, 16, 24, 36, 54 and 72 cores respectively.
-Each run is repeated 10 times.
-The average execution time, required sampling numbers are recorded.
-Hyperthreading is enabled when using 54 and 72 cores.
-The access to the node that contains computation cores is exclusive, such that the execution would not be affected by other programs of operations.
-
-The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the must time-consuming part.
-The rest part of the program is mostly not parallelised, e.g.
-database I/O and reductions operations.
-The sampling process involves sampling, perturbation and test of the acceptance criteria, all of which are computation-intensive.
-
-In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources is limited[REF].
-The performance experiment result could provide a reference that illustrate that how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
-
-\subsubsection{Results}
-
-[scaling-up performance: speed-up and efficiency]
-
-[large variation in required sampling numbers]
-
-[possible reasons]
-
-The recorded execution time using different number of cores is shown in FIGURE.
-According to that, speed-up and efficiency can be calculated.
-Here 8 cores is regarded as the baseline for speed-up and efficiency, as the serial version (using only one core) can take quite a long time to finish 10 repeats.
-When increasing number of cores, the execution time drops fast at first but the decreasing rate (dropping speed, i.e.
-the derivative of the execution time against time) is gradually reduced to zero; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (FIGURE).
-
-Denote the number of cores used in the ABC SMC runs as variable $p$.
-The speedup curve shows a linear trend when $p\leq 36$; when $p\geq 36$ the speedup stays nearly unchanged.
-$p=36$ is an inflection point, where 36 is the maximum numbers available physical cores in a compute node of Cirrus.
-A constant drop of efficiency is observed when increasing $p$; smaller $p$ ($p<36$) gives an efficiency higher than 65\% but greater $p$ e.g.
-$p=72$ results in a low efficiency (34\%).
-
-A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations.
-Moreover, we found that the execution time is not ideally proportional to the total required samples (FIGURE).
-A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs.
-Some runs that are stuck in a local optimum and require smaller target epsilon will take huge amount of samples to move away from the local optimum.
-This have been observed when the distribution of posterior is plotted before and after jumping away from local modes (FIGURE and more explain).
-Also, time spent in each sample using sanme number of cores is not constant in our case, which might be a result of the parallel schedule; ideally a serial run using only one core will see a nearly constant time spent in every particle.
-
-Due to the unstable sampling numbers, our interests switched to the per-particle performance, where the average sampling numbers per second (FIGURE) and average time spent in one particle (FIGURE) under different of cores ($p$) are plotted.
-The log-like average time per particle curve shows that using more cores can improve the sampling efficiency, but when $p$ is at a high level (e.g.
-36 in our case), the improvement will be less significant.
-
-\section{Profiling}
-
-The performance could also be analysed given a profiling report.
-The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck, according to which we could find the hot-spot of program and given possible suggestions on improving the performance.
-
-In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
-
-\subsubsection{Results}
-
-[profiling results: hot-spot and possible improvements]
+\input{conclusion/conclusion.tex}
 
@@
 
-\chapter{Future works}
-
- [something not done as scheduled]
-
- [interesting topics to look inside]
-
- [moving to generalisations]
-
-\chapter{Conclusions}
-
- [what is studied]
-
- [to what extend]
-
- [how good is the result]
-
- [summary of the findings and advise]
-
- [thanks]
-
-\appendix
-% the appendix command just changes heading styles for appendices.
-
-\chapter{System and software environment}
-
-\section{ABC SMC implementation}
-
-\subsection{Local machine}
-
-Local development machine is a Mac laptop, running on macOS 10.15.6. The environment of the development is listed in Table \ref{table:local_macine}.
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|c c|}
-        \hline
-        Environment                 & Version                       \\ [0.5ex]
-        \hline\hline
-        $\lambda_N$                 & 2.1989                        \\
-        $\kappa_{N\beta}$           & 3.9627                        \\
-        $\mu_N$                     & 1.7219                        \\
-        $\nu_{N\Phi}$  0.2195       & $cell^{-1}\cdotp h^{-1}$      \\
-        \hline
-        $\lambda_\Phi$              & 1.3146  $cell/h$              \\
-        $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$         \\
-        $\mu_\Phi$  0.1454          & $h^{-1}$                      \\
-        \hline
-        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
-        $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
-        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
-        \hline
-        $s_{\alpha\Phi}$  10.2416   & $unit/(cell\cdotp h)$         \\
-        $\mu_\alpha$ 19.6642        & $h^{-1}$                      \\
-        [1ex]
-        \hline
-    \end{tabular}
-    \caption{Environment on local machine}
-    \label{table:local_macine}
-\end{table}
-
-\subsection{Remote machine}
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|c c|}
-        \hline
-        Environment                 & Version                       \\ [0.5ex]
-        \hline\hline
-        $\lambda_N$                 & 2.1989                        \\
-        $\kappa_{N\beta}$           & 3.9627                        \\
-        $\mu_N$                     & 1.7219                        \\
-        $\nu_{N\Phi}$  0.2195       & $cell^{-1}\cdotp h^{-1}$      \\
-        \hline
-        $\lambda_\Phi$              & 1.3146  $cell/h$              \\
-        $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$         \\
-        $\mu_\Phi$  0.1454          & $h^{-1}$                      \\
-        \hline
-        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
-        $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
-        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
-        \hline
-        $s_{\alpha\Phi}$  10.2416   & $unit/(cell\cdotp h)$         \\
-        $\mu_\alpha$ 19.6642        & $h^{-1}$                      \\
-        [1ex]
-        \hline
-    \end{tabular}
-    \caption{Environment on remote machine}
-    \label{table:remote_macine}
-\end{table}
-
-\section{Data analysis}
-
-\chapter{Data and settings}
-
-\section{Infer-back experiments}
-
-\subsection{Parameter values used to generate synthetic data}
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|c c c|}
-        \hline
-        Parameter            & Value & Unit                     \\ [0.5ex]
-        \hline\hline
-        $\lambda_N$          & 2.20  & $cell/h$                 \\
-        $\kappa_{N\beta}$    & 3.96  & $cell/(unit\cdotp h)$    \\
-        $\mu_N$              & 1.72  & $h^{-1}$                 \\
-        $\nu_{N\Phi}$        & 0.220 & $cell^{-1}\cdotp h^{-1}$ \\
-        \hline
-        $\lambda_\Phi$       & 1.31  & $cell/h$                 \\
-        $\kappa_{\Phi\beta}$ & 0.124 & $cell/(unit\cdotp h)$    \\
-        $\mu_\Phi$           & 0.145 & $h^{-1}$                 \\
-        \hline
-        $s_{\beta N}$        & 6.554 & $unit/(cell\cdotp h)$    \\
-        $i_{\beta\Phi}$      & 1.71  & $cell^{-1}$              \\
-        $\mu_\beta$          & 0.521 & $h^{-1}$                 \\
-        \hline
-        $s_{\alpha\Phi}$     & 10.2  & $unit/(cell\cdotp h)$    \\
-        $\mu_\alpha$         & 19.1  & $h^{-1}$                 \\
-        \hline
-    \end{tabular}
-    \caption{Parameter values used for model 1}
-    \label{table:m1}
-\end{table}
-
-\subsection{Kernel experiment: median epsilon schedule}
-
-\begin{figure}[h!]
-
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel2.png}}
-    \end{center}
-
-    \caption[Total sampling size of different kernels, using median epsilon strategy]
-    {Total sampling size of different kernels, using median epsilon strategy. Different color represents different generations (bottom to top: population 1 to population 20)}
-    \label{fig:kernel2}
-
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance2.png}}
-    \end{center}
-
-    \caption[Acceptance rates of different kernels, using median epsilon strategy]
-    {Acceptance rates of different kernels, using median epsilon strategy. Each population has 2000 particles}
-    \label{fig:acceptance2}
-
-\end{figure}
-
-\section{Parameter inference and model comparison results}
-
-\begin{figure}[H]
-
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve_uni.png}}
-    \end{center}
-
-    \caption{Total sampling size}
-    \label{fig:resultCurve_uni}
-
-\end{figure}
+\input{appendix/appendix.tex}
 

### 2020-08-09 appendix/appendix.tex
New structure

\appendix
% the appendix command just changes heading styles for appendices.

\chapter{System and software environment}

\section{ABC SMC implementation}

\subsection{Local machine}

Local development machine is a Mac laptop, running on macOS 10.15.6. The environment of the development is listed in Table \ref{table:local_macine}.

\begin{table}[h!]
    \centering
    \begin{tabular}{|c c|}
        \hline
        Environment                 & Version                       \\ [0.5ex]
        \hline\hline
        $\lambda_N$                 & 2.1989                        \\
        $\kappa_{N\beta}$           & 3.9627                        \\
        $\mu_N$                     & 1.7219                        \\
        $\nu_{N\Phi}$  0.2195       & $cell^{-1}\cdotp h^{-1}$      \\
        \hline
        $\lambda_\Phi$              & 1.3146  $cell/h$              \\
        $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$         \\
        $\mu_\Phi$  0.1454          & $h^{-1}$                      \\
        \hline
        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
        $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
        \hline
        $s_{\alpha\Phi}$  10.2416   & $unit/(cell\cdotp h)$         \\
        $\mu_\alpha$ 19.6642        & $h^{-1}$                      \\
        [1ex]
        \hline
    \end{tabular}
    \caption{Environment on local machine}
    \label{table:local_macine}
\end{table}

\subsection{Remote machine}

\begin{table}[h!]
    \centering
    \begin{tabular}{|c c|}
        \hline
        Environment                 & Version                       \\ [0.5ex]
        \hline\hline
        $\lambda_N$                 & 2.1989                        \\
        $\kappa_{N\beta}$           & 3.9627                        \\
        $\mu_N$                     & 1.7219                        \\
        $\nu_{N\Phi}$  0.2195       & $cell^{-1}\cdotp h^{-1}$      \\
        \hline
        $\lambda_\Phi$              & 1.3146  $cell/h$              \\
        $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$         \\
        $\mu_\Phi$  0.1454          & $h^{-1}$                      \\
        \hline
        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
        $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
        \hline
        $s_{\alpha\Phi}$  10.2416   & $unit/(cell\cdotp h)$         \\
        $\mu_\alpha$ 19.6642        & $h^{-1}$                      \\
        [1ex]
        \hline
    \end{tabular}
    \caption{Environment on remote machine}
    \label{table:remote_macine}
\end{table}

\section{Data analysis}

\chapter{Data and settings}

\section{Infer-back experiments}

\subsection{Parameter values used to generate synthetic data}

\begin{table}[h!]
    \centering
    \begin{tabular}{|c c c|}
        \hline
        Parameter            & Value & Unit                     \\ [0.5ex]
        \hline\hline
        $\lambda_N$          & 2.20  & $cell/h$                 \\
        $\kappa_{N\beta}$    & 3.96  & $cell/(unit\cdotp h)$    \\
        $\mu_N$              & 1.72  & $h^{-1}$                 \\
        $\nu_{N\Phi}$        & 0.220 & $cell^{-1}\cdotp h^{-1}$ \\
        \hline
        $\lambda_\Phi$       & 1.31  & $cell/h$                 \\
        $\kappa_{\Phi\beta}$ & 0.124 & $cell/(unit\cdotp h)$    \\
        $\mu_\Phi$           & 0.145 & $h^{-1}$                 \\
        \hline
        $s_{\beta N}$        & 6.554 & $unit/(cell\cdotp h)$    \\
        $i_{\beta\Phi}$      & 1.71  & $cell^{-1}$              \\
        $\mu_\beta$          & 0.521 & $h^{-1}$                 \\
        \hline
        $s_{\alpha\Phi}$     & 10.2  & $unit/(cell\cdotp h)$    \\
        $\mu_\alpha$         & 19.1  & $h^{-1}$                 \\
        \hline
    \end{tabular}
    \caption{Parameter values used for model 1}
    \label{table:m1}
\end{table}

\subsection{Kernel experiment: median epsilon schedule}

\begin{figure}[h!]

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel2.png}}
    \end{center}

    \caption[Total sampling size of different kernels, using median epsilon strategy]
    {Total sampling size of different kernels, using median epsilon strategy. Different color represents different generations (bottom to top: population 1 to population 20)}
    \label{fig:kernel2}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance2.png}}
    \end{center}

    \caption[Acceptance rates of different kernels, using median epsilon strategy]
    {Acceptance rates of different kernels, using median epsilon strategy. Each population has 2000 particles}
    \label{fig:acceptance2}

\end{figure}

\section{Parameter inference and model comparison results}

\begin{figure}[H]

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve_uni.png}}
    \end{center}

    \caption{Total sampling size}
    \label{fig:resultCurve_uni}

\end{figure}

### 2020-08-09 background/background.tex
New structure

\chapter{Background}

\section{Zebrafish spinal cord regeneration}

 [describe the process and how cells and cytokines are involved]

Compared to mammals, zebrafish can regenerate spinal cord after injury, and possibly recover swimming function. They are widely used in regenerative ability researches. Experiments found that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}. Tsarouchas et al. \cite{ref:Tsarouchas} studied the the complex interactions happen at the lesion site. The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration. Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, affect the inflammation, which are dynamically controlled by immunes cells, i.e. cytokines works as media to control the regeneration. Here cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.

An interaction map can be drawn, according to the evidences found in \cite{ref:Tsarouchas}. After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells: macrophage, neutrophil and keratinocyte. The dynamical system consists of complex interactions between cells and cytokines, e.g. il-1$\beta$ promotes macrophages' increase. and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA. Consequently, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration. Generally, il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect, and tnf-$\alpha$ promotes the regeneration all the time. The presence if the cytokines also affects the number of cell, while as sources number of cells also affects cytokines' expression. Interactions between different kinds of cells are also considered in the system.

This is a dynamic system suitable for mathematical modelling, as a set of equations can be written to represent the dynamics as possible. Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions iniside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and our own hypothesis which may include different interactions.

The biological background of works in this dissertation is mostly related the the experimental evidences, quantitative data and proposed hypothesis in \cite{ref:Tsarouchas}. The published data on the quantitative measurement of cells and cytokines make the modelling and further analysis possible. Some other interactions proposed in our modelling stages are enlightened from Anderson et al. \cite{Anderson}.

\section{Mathematical modelling}

 [how this can be modelled as dynamic systems]

 [general topics of mathematical models: how to build model according to interactions; how to calibrate/ parameterise the model; how to solve the model; dynamics; how to evaluate the model]

Mathematical models can express real-life questions in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques. Mathematical models can be used for prediction, classification, testing our understanding or hypothesis and many other tasks, thus they are widely used to solve practical problems. However, for a model that describe the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models. In this case, we wish to use modelling techniques to test our inference method, our hypothesis and hopefully an well-fitted models can be found and compared.

According to our understanding of the biological mechanism in the regeneration, hypothesis should firstly proposed, in a way which can represent the real  mechanism as much as possible. Usually a real-life problem, e.g. the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered, An elaborate description of the real cases using a mathematical model sometimes can be impossible; assumptions and hypothesis are proposed to make necessary simplification and abstract for a final reasonably detailed model. Some qualitative models are simple but illustrative for the real-life observed patterns, e.g. Lotka-Volterra (L-V) equation for predator and prey models; they are usually analytically solvable and some properties are used to make predictions (e.g. steady states in ecosystem). If a more precise model is desired, usually more effects or interactions are considered and added as terms to a re-constructed model, e.g. environmental factors, a changing environmental capacity and predation from other species.

Models like L-V equation are mostly analytically solvable when analysing properties of the model, however, analytical solution for many models are not feasible, because of more complex dynamics, or the model itself is not a deterministic model. In the areas that are not analytically feasible, usually computational techniques are applied for approximate solutions. Solvers for different models are used to simulate the system or find roots for some spacial points.

For the regeneration process that we wanted to model, differential equations are our preference. Non-linear ordinary differential equations are used in this dissertation, where the only independent variable is time $t$ denoting the post-lesion time (hours), but further models can also consider other models e.g. stochastic models and partial differential equations. For ordinary differential equations (ODE) system, there are several available computational solvers using different algorithms, implemented in MATLAB\footnote[1] {\url{https://uk.mathworks.com/help/matlab/math/choose-an-ode-solver.html}}, Python\footnote[2]{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html}} and many other languages. These solvers help to simulate the trajectories of the dependent variables, given initial values and a set of fixed parameters. The goodness of `fit' then can be measured by compared the simulated data and observed real data.

Often further researches on the modelled problems require the model to be deterministic on its parameters. Parameters in the dynamic system models usually have their practical significance, e.g. describing reaction speed, environmental capacity or the birth rate. There are many approaches to derive these parameters from data or experience. For a specific problem and a widely used models, the parameters' values can be read from previous literatures; however in many others cases, we need to find parameter's values specifically for our model and problem, where numerical values for parameters are more suitable for practical analysis using ODE solvers or other tasks.

To find the best parameters' values given observed data, either optimisation-based approaches (e.g. least-square (LS) fit) or inference approaches can be used. Optimisation-based approaches may face two major drawbacks: over-fitting and local optima \cite{ref:abcsysbio}. Bayesian inference approaches gain popularity in recent years, which can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values. Local optima can still be a major concern, and many efforts have be proposed for a more through explore of the parameter space.

\section{Bayesian inference}

 [how to parameterise the model]

 [how to infer the parameter given data]

 [likelihood-free inference and information about ABC SMC]

 [model comparison, ABC SMC for model comparison]

 Bayesian inference approaches put Bayes Rule in the centre:

 \begin{align}
    \label{eq:bayes}
    p(\theta|D) = \frac{p(D|\theta)p(\theta)}{P(D)}
\end{align}

where $p$ is the probability density function, $\theta$ denotes the parameter set, $D$ is the given observed data.

Alternatively, as $p(D)$ is fixed once the observed data is given, the rule can be rewritten as

\begin{align}
    \label{eq:bayes2}
    p(\theta|D) \propto l(D|\theta)p(\theta)
\end{align}

where $l$ represents the likelihood function. In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target. Prior $p(\theta)$ should express our initial believes or estimates on the parameter valeus before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers. Once the data is given, $\theta$ is the only parameter in the likelihood function. Some optimisation-based methods try to find an estimation of parameter set $\theta$ s.t. maximise the likelihood, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative.

Based on Eqn \ref{eq:bayes2}, to find the posterior of parameter set, prior can be set according to the problem context and our understanding and cover a reasonable ranges; computational methods can be used to approximate likelihood function, e.g. Markov Chain Monte Carlo (MCMC) \cite{ref:MCMC}. As another alterative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood, but focused on simulating the data from the given model and observed data, then used the simulated data to approximate the desired posterior. ABC-rejection \cite{ABC_rejection} is a simple example of the ABC process, and further algorithms e.g. sequential Monte-Carlo (ABC SMC)\cite{Toni} are proposed to improve the efficiency.

General steps for ABC algorithm includes the following (ABC-rejection):

\begin{enumerate}
    \item Draw a sample $\theta^*$ from parameters' prior distribution. $\theta^*$ is a vector that consists values for each parameter
    \item Plug $\theta^*$ into the model, and use the model to generate simulated data $D^*$, where $D^*$ is of the same form as the observed data $D$, e.g. includes same summary statistics
    \item The discrepancy between the simulated data and observed data is measured by a distance function $d(D^*, D)$. For a pre-fixed threshold value $\epsilon$, if $d(D^*, D)<\epsilon$ then the sample $\theta^*$ is accepted
    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution can be approximated by these samples
\end{enumerate}

In our study of the regeneration model, full trajectories of the dependent variables are used as the data to be compared, as our goal is to find a model and a parameter set that can intuitively recover the real process as much as possible. The mean values at each time point for each dependent variable is calculated to form $D$, and for each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$. Using other statistics to summary the raw data are also possible bur care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.

Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of deceasing threshold values $\epsilon_t$, and derive a new population (called generation) of accepted sampled from the previous accepted samples (called particles). An example is ABC SMC \cite{Toni}, which can significantly reduce the required samples compared to ABC-rejection \cite{ref:disease}:

\begin{enumerate}
    \item Set a threshold schedule $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ where $T$ is the total number of populations and $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ is a descending sequence; set the number of particles in each population as $N$, set the population index $t=1$
    \item if $t=1$, draw samples from prior distribution until $N$ particles are accepted under threshold $\epsilon_1$, i.e. each of $N$ particles s.t. $d(D,D^*)<\epsilon_1$

    Else ($t>1$), draw samples from previous population $\{\theta^*_{t-1}\}$ with weights $\{w_{t-1}\}$, perturb the particles using a perturbation kernel $K(\theta^*|\theta)$ and obtain $\theta^{**}\sim K(\theta^*|\theta)$
    \item Repeat step 2 until $N$ particles are accept under threshold $\epsilon_t$
    \item Calculate the weights of the accepted particles. For particle $i$ 
    \begin{align}
        \label{eq:weight}
        w^i_t =\begin{cases}
            1 & \text{if t=1} \\
            \frac{p(\theta^i_t)}{\sum_{j=1}^{N} w^j_{t-1}K(\theta^j_t|\theta^j_{t-1})} & \text{if t $\neq$ 1} 
        \end{cases}
    \end{align}

    \item Normalise $\{w^i_t\}$, increase population index $t$ by 1
    \item Repeat step 2, 3 and 4 until $t>N$
    \item Output the accepted particles $\{\theta^*_{t}\}$ and their weights $\{w^i_t\}$ in each generation $t$

\end{enumerate}

An illustrative graph of the ABC SMC procedures can be found in Figure \ref{fig:smc}. In implementations, many factors can affect the efficiency and goodness of fit, some of them are discussed in the Chapter 4. Rather than a pre-fixed threshold schedule, median or quantile based epsilon schedule can be save total sampling size and thus improve acceptance rates of particles \cite{threshold}. Some other improvements include using adaptive population size for each generation \cite{population}, using adaptive distance function \cite{ref:adpt_dis} and switching kernels \cite{ref:kernel}, etc. In our implementations these options were tried when applicable and suitable to compare the behaviours of the algorithm.

\begin{figure}[t]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
    \end{center}

    \caption[ABC SMC sampling process]%
    {ABC SMC sampling process. $\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$. $d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data. for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
    \label{fig:smc}

\end{figure}

In addition, ABC SMC is also capable of model selection tasks \cite{model_compare}. Instead of draw samples from prior of a single model, we can draw samples from multiple models $m_i$ and approximate $p(m_i, \theta|D)$, which is the joint probability distribution of model $m_i$ and its accepted particles $\theta$. In each generation the marginal distribution $p(m_i|D)$ can be calculated as the accepted ${\theta}$ have been obtained, thus models can be ranked according to their probabilities.

The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modeling problems, especially in the area of biology \cite{ref:abcsysbio, ref:compare, ref:disease}. But ABC is not limited to these areas, as it can be generally used in inference tasks to give parameter estimates or model rankings. E.g., it has been successfully applied in cosmology studies \cite{cosmology}.

\section{Software tools}

 [existing tools and software for this task]

 [workflow and developing process]

 As we decided to apply ABC SMC on our regeneration models for its high-dimensional parameter estimation and possibly comparison of different models, a code development and experiments workflow was expected to be built. The ABC SMC algorithm needs a whole set of mathematical environments (e.g. class of distributions, models and particles) and closely linked sampling and fitting algorithms (e.g. KDE fitting and Gaussian sampling), thus suitable platforms with related packages are preferred. Regarding this, several packages in \verb|R| and \verb|Python| were studied.

 A software packages that integrally implement ABC SMC is preferred, and more options in the settings of the algorithm, more customisable features can make it better. Moreover, as the ABC SMC itself is a computationally intensive task that consists of massive parallel potentials, an ideal software is expected to support as least one kind of parallel options e.g. multi-core, GPU accelerating and clusters. Otherwise, a modification or rewrite of the software package are expected to produce a reliable input-output framework for our parameter estimation and model selection experiments.

 Preliminarily \verb|pyABC| \cite{ref:pyabc} in \verb|Python| was chosen for our first implementations.

### 2020-08-09 conclusion/conclusion.tex
New structure

--- 2020-07-11 __latexindent_temp.tex
@@
-\documentclass[12pt,a4paper]{report}
-
-\usepackage{epcc}
-\usepackage{graphics}
-\usepackage{mathtools}
-
-% This example file shows how a thesis can be laid out using Latex. It
-% does not use any special local features so should be portable to other
-% places.
-% 
-% When producing draft copies of a thesis you may want to print only
-% selected pages of the thesis. To do this use the command
-% 
-% dvips -f -p 4 -n 3 myfile.dvi | lpr
-% 
-% where -p 4 means start printing at page 4 (ie the page that will be
-% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
-% This example will print pages 4, 5 and 6.
-% 
-% If you want to print the thesis and also save paper you can print more
-% than one page on each sheet of paper. Use the command
-% 
-% dvips -f myfile.dvi | psnup -2 | lpr
-% 
-% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
-%
-% To produce a PDF version you can create a PostScript copy first
-%
-% dvips -f myfile.dvi > myfile.pdf
-%
-% and then convert it
-%
-% distill myfile.ps
-%
-% or you can go straight to PDF
-%
-% pdflatex myfile
-%
-% Note that pdflatex expects all included figures to be in PDF too. See
-% the includegraphics command below.
-
-% This document contains many cross-references and forward references,
-% eg in constructing a table of contents, so Latex may need to be run
-% twice to get all the references correct. If you need to run Latex twice
-% you may get the warning:
-% 
-% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.
-
-% the following 4 lines are the content of the smallmargins.sty file
-% but including them explicitly makes this more portable.
-%AC%\oddsidemargin=0.1in
-%AC%\topmargin=-0.5in
-%AC%\textheight=9in
-%AC%\textwidth=6.25in
-
-%AC%\parskip 10pt
-%AC%\parindent 0in
-
-\begin{document}
-
-%AC%\pagestyle{myheadings}
-%AC%\markright{D.~S.~Henty}
-
-%\title{A Latex thesis example}
-%\author{D.~S.~Henty}
-%\date{\today}
-
-%\maketitle
-
-\pagenumbering{roman}
-
-\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
-\author{Chaolin Han}
-\date{\today}
-
-\makeEPCCtitle
-
-\thispagestyle{empty}
-
-\vspace{11cm}
-
-\begin{center}
-
-\large{MSc in High Performance Computing}
-
-\large{The University of Edinburgh}
-
-\large{Year of Presentation: 2020}
-
-\end{center}
-
-\newpage
-
-\begin{abstract}
-This is the bit where you summarise what is in your thesis.
-\end{abstract}
-
-\pagenumbering{roman}
-
-\tableofcontents
-\listoftables
-\listoffigures
-
-\begin{titlepage}
-\vspace*{2in}
-% an acknowledgements section is completely optional but if you decide
-% not to include it you should still include an empty {titlepage}
-% environment as this initialises things like section and page numbering.
-\section*{Acknowledgements}
-
-This template is a slightly modified version of the one developed by
-Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
-acknowledgement follows:
-
-{\em This template has been produced with help from many former students who
-have shown different ways of doing things. Please make suggestions for
-further improvements.}
-
-\end{titlepage}
-
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-This should contain a description of your project and the problem you
-are trying to solve. Where appropriate you should also include
-references to work which has already been done on your topic and
-anything else which lets you set your work in context.
-
-One of the things you will need to do is to ensure that you have a
-suitable list of references.  To do this you should see \cite{ref:lam}
-or some other suitable reference.  Note the format of the citation used
-here is the style favoured in this department.  Here is another
-reference \cite{ref:bloggs} for good measure.
-
-You will also want to make sure you have no spelling or grammatical
-mistakes. To help idwentify spelling mistukes you caan use the commands
-{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
-that spelling mistakes are not the only errors which can occur. Spelling
-checkers will not find errors which are, in fact, valid words such as
-{\em there} for {\em their}, nor will they find repeated words which
-sometimes occur if your concentration is broken when typing. {\bf There
-is no substitute for thorough proof reading!}
-
-\chapter{Background}
-
-\section{Zebrafish spinal cord regeneration}
-
-\section{Mathematical modelling}
-
-\subsection{Hard bits}
-You might want to include an equation here:
-
-\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
-\label{equation:delsplit}
-% note that the label is optional but allows you to refer to this
-% equation later.
-\end{equation}
-
-\subsection{Even harder bits}
-This might be one of the places where you might want to refer to
-equation \ref{equation:delsplit}. You will usually need to use the
-Latex command twice to make cross-references like this work properly.
-The cross-reference information is stored in the {\em .aux} file so don't
-delete it.
-
-\subsubsection{Numbering}
-You can keep subdividing but eventually you get to a level where
-numbering stops. This text is in a subsubsection which is not numbered
-by default.
-
-\paragraph{More on numbering:}
-
-This text is in a paragraph which is also not numbered by default and
-the ``title'' of the paragraph is not on a separate line.
-If you want to increase the depth to which sections are numbered you
-should see the section on setting the secnumdepth counter in the manual.
-
-\section{Bayesian inference}
-
-\section{Software tools}
-
-\chapter{Mathematical modelling}
-
-Mathematical models can describe different kinds of dynamic system, and can be used as a guide to prediction and analysis.
-An ideal model in this case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  trend against time.
-
-Models for the regeneration process is in the form of ordinary differential equations, specifically the time differential form.
-Terms in the ODE are mostly explainable and corresponds interaction paths.
-
-\section{Observed data}
-
-Our models is built regarding the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement includes the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
-As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
-According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration'.
-
-It is noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations is mean of measurement, assuming that measured data is Gaussian-like distributed.
-The distribution of the measured data points is plotted and examined to see if the measurement mean could represent the distribution.
-The result is that at most time points the measurement values are Gaussian-distributed, some distributions are skewed.
-One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
-Mean value can summarise most data measures and thus is still used as the target summarised observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}
-
-\begin{figure}
-
-\begin{center}
-\resizebox{1.0\hsize}{!}{\includegraphics{fig/obs_curve.png}}
-\end{center}
-
-\caption{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
-\label{fig:obs_data}
-
-\end{figure}
-
-\section{Hypothesis and Models}
-
-5 models in total are proposed according to different hypothesis.
-At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
-After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can be more close to the true process, recover more features of the observed data or test our hypothesis.
-
-All these models assumes the interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for parameter estimation.
-Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data.
-
-\subsection{Basic model}
-
-A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
-
-\begin{equation*}\label{eq:pareto mle2}
-    \begin{multlined}
-        \frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
-\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-   \end{multlined}
-  \end{equation*}
-
-\subsection{Alternative models}
-
-\subsection{Model evaluation and comparison}
-
-% \begin{figure}
-
-% \begin{center}
-% \resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
-% \end{center}
-
-% \caption{The University Crest}
-% \label{fig:eucrest}
-
-% \end{figure}
-
-% see the man page for dvips for details of the special command which is
-% much more powerful than is shown here. It allows offsets in the
-% horizontal and vertical and scaling in x and y.
-
-% choosing suitable values for offset and scale can be a tiring matter
-% of trial and error.
-
-% note that labels do not need to include a description of the object
-% they are labelling but it can be helpful, eg \label{fig:figurename}.
-
-% You can use a label on a figure to refer to it later. The university
-% crest is in \ref{fig:eucrest}. Note that you should not use phrases like
-% ``the figure above'' or ``the following figure'' since Latex may move
-% the figure relative to the text if it cannot be fitted onto the current page.
-
-\chapter{Implementations and Experiments}
-
-\section{Parameter estimation}
-Here are some results.
-
-\section{Model comparison}
-Here are some results.
-
-\section{Performance experiments}
-
-The performance experiments are designed to explore the parallel performance of ABC SMC implementations.
-Usually ABC SMC is a time-consuming and computation intensive task and usually is executed on large clusters.
-The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
-
-\subsection{Scaling-up}
-
-First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
-
-\begin{itemize}
-    \item Prior distribution: default to log-uniform distribution [$1\times 10^{-6}$, 50] for all the 12 parameters
-    \item threshold schedule: median epsilon
-    \item No factors, no adaptive distance or adaptive population applied
-    \item Population size is 2000, with 20 generations
-\end{itemize}
-
-HOW PYABC parallelise the sampling
-
-For HPC systems like Cirrus, \verb|pyabc| uses \verb|multiprocessing| for multi-core parallel sampling.
-By default if the number of cores is not specified, it will automatically read the number of available cores and use them all.
-Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
-
-The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54, 72 cores respectively.
-Each run is repeated 5 times.
-The average execution time, required sampling numbers are recorded.
-Hyperthreading is enabled when using 54 and 72 cores.
-The access to the node that contains computation cores is exclusive, such that the execution would not be affected by other programs of operations.
-
-The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the must time-consuming part.
-The rest part of the program is mostly not parallelised, e.g.
-database I/O and reductions operations.
-The sampling process involves sampling, perturbation and test of the acceptance criteria, all of which are computation-intensive.
-
-In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources is limited[REF].
-The performance experiment result could provide a reference that illustrate that how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
-
-\subsection{Profiling}
-
-The performance could also be analysed given a profiling report.
-The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck, according to which we could find the hot-spot of program and given possible suggestions on improving the performance.
-
-In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
-
-\chapter{Results Analysis}
-
-\section{ABC SMC results}
-
-\section{Performance experiments}
-
-\section{Discussions}
-This is the place to put your conclusions about your work. You can
-split it into different sections if appropriate. You may want to include
-a section of future work which could be carried out to continue your
-research.
-
-\chapter{Future works}
-
 \chapter{Conclusions}
 
-\appendix
-% the appendix command just changes heading styles for appendices.
+ [what is studied]
 
-\chapter{Stuff which is too detailed}
+ [to what extend]
 
-Appendices should contain all the material which is considered too
-detailed to be included in the main bod but which is, nevertheless,
-important enough to be included in the thesis.
+ [how good is the result]
 
-\chapter{Stuff which no-one will read}
+ [summary of the findings and advise]
 
-Some people include in their thesis a lot of detail, particularly
-computer code, which no-one will ever read. You should be careful that
-anything like this you include should contain some element of uniqueness
-which justifies its inclusion.
-
-\begin{thebibliography}{100}
-
-\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
-and Reference Manual.} Addison Wesley. pp242.
-
-\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
-in Environments} Int. Journal of Silly Findings. pp 23-29.
-
-\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N.
-and Feng, Y., 2018.
-Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
-Nature communications, 9(1), pp.1-17.
-
-\end{thebibliography}
-
-\end{document}
+ [thanks]

### 2020-08-09 implementation/implementation.tex
New structure

\chapter{Parameter estimation and model comparison}

Estimation of  parameters in the proposed models is the main task of this section, which involves the implementation of ABC SMC and several experiments. ABC SMC-based model selection is another task that can be easily conducted with the developed parameter inference framework.

The build and test of code were under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}. Some other code e.g. shell script and \verb|R| code are also used to perform the experiments and analyse the results. Software and system environment used in our implementation can be found at Appendix A. The build and test were performed on local computers and HPC facilities available within EPCC.

\section{Implementations and code}

 [ABC implementations details, e.g. distance, population, ODE related functions]

 [how the code is developed and built]

According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project. \verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}. \verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks. Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.

The inference framework code for the project was developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task. The scaling-up experiment was performed on Cirrus, and the profiling and its analysis were performed on local machine. Some development and debug were on the remote machines to ensure experiments run correctly on compute node.

The ODE model-related code and utilities were firstly developed to enable reusable functions, variables and data calls. It includes the code for the ODE models, ODE solver, data structures, data format transform functions etc. By doing this in a separate code file, we can change options and activate ABC SMC easily in the implementation code file, without creating definition and duplicated code for the models and parameter priors.

Some modifications were made to the source code of \verb|pyABC|. For example, the laboratorial measured data were not considered `complete' for modeling and analysis purpose: the available time points for cells and cytokines are different, e.g. cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl. These missing values were treated as `ignored' when calculating the distance between observed data and simulated data. To cope with this built-in distance function code is modified.

Analysis code was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results. For different tasks there are also separated analysis code file. Other analysis tools used in the project include built-in database visualisation tool, R code and Microsoft Excel.

The project code was managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also made the cross-hardware synchronisation of code versions convenient.

After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (udsing synthetic data), and code files for parameter inference (using experimental data). Model selection can be easily done by adding more model to the implementation code file.

% \section{Parameter estimation}

% [implementation details and options/ settings that are affecting the ABC]

\section{ABC SMC and model hyperparameters experiments}

As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental observed data.

In the first part, synthetic data is used with known true parameter values. The algorithm implementation is evaluated by the efficiency and goodness of resultant model under different implementation options and ABC SMC hyperparameters.

Then according to findings in the first part, several options with good performance are tried in the inference using experimental observed data (Figure \ref{fig:obs_data}). To obtain a more accurate and general results, these experiments are repeated for 3 or more times.

    [including the options that are tested/ tried/ not adopted]

    [including experiments results and discussions]

    [how do these options correspond to biological model and how to set them accordingly for proposed model]

The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options. To firstly test the ability of inference and observe the results of different options, synthetic data with known true parameter values are used as target data. The true value of parameters that are used to generate synthetic data is listed in Appendix B.1. These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data (Figure \ref{fig:infer_back_data}), the target data is then generated using these parameter values via ODE solver in \verb|Python|.

\begin{figure}[h!]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
    \end{center}

    \caption[Synthetic data generated with known parameter values]%
    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot). Known parameter values are obtained from a least square fitting of the observed data}
    \label{fig:infer_back_data}

\end{figure}

[TABLE for known parameter values]

[why use synthetic data]

Among them the following topics are studied using the synthetic data.

\subsection{Perturbation kernels}

Perturbation kernels work in the sampling process. In each generation $t$, samples are firstly taken from the previous population $\{\theta_{t-1}\}$ with weights $\{w_{t-1}\}$, then perturbed using the perturbation kernel $K(\theta|\theta^*)$. We keep sampling until $N$ particles are accepted, where $N$ is the pre-set population size. After that the new weights are calculated and normalised.

% \begin{figure}[t!]
%     \begin{center}
%         \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
%     \end{center}

%     \caption[ABC SMC sampling process]%
%     {ABC SMC sampling process. $\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$. $d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data. for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
%     \label{fig:smc}

% \end{figure}

The perturbation kernel is called in the sampling of every particles and determines how the new perturbed particle is chosen, thus it is influential to both computation complexity and the resultant posterior distribution. Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations, or could generate a population with a higher acceptance rate if the successive epsilon values are close; A kernel with wide variance, or spreading out in a large space could help in resolving the local optimal problem by a more thoroughly exploration of the parameter space, however it can be more computation-intensive and result in a lower acceptance rate. A desired optimal kernel should balance the trade-offs; their property and criteria is discussed in \cite{ref:kernel}.

There are several common choice of perturbation kernels. Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models. A covariance matrix $\Sigma_t$ of accepted particles is calculated form previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$. It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in relecting the true posterior structure \cite{ref:kernel}. It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks. The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `narrower distributed' perturbation result.

Local M-NN kernel provided by \verb|pyABC| is also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.

The kernel experiments is designed to explore the efficiency of SMC on our dynamical systems. Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rate would be our preference. The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters. The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference. As the threshold schedule is fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit i.e. recover of target data trends/features and errors of the inferred parameters compared to true values is not discussed.

\subsubsection{Kernel experiment results}

To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e. $\epsilon_t$ set to 10.0. From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared. Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
has higher acceptance rates among almost all generations. Local M-NN with M=750 is the slowest kernel.

For local M-NN kernels, generally a greater M will lead to lower acceptance rate and
more required particles in each generation. As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve. Consequently, if greater M is test, then it will have even lower acceptance rates; the maximum M=2000 (whole population is considered) will have the lowest acceptance rates.

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel1.png}}
    \end{center}

    \caption[Total required samples of different kernels]%
    {Total required samples of different kernels after 20 populations (2000 particles in each population). Different color represents different generations (bottom to top: population 1 to population 20)}
    \label{fig:kernel1}

    \vspace*{\floatsep}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
    \end{center}

    \caption[Acceptance rates of different kernels]%
    {Acceptance rates of different kernels in 20 populations. Each population has 2000 particles}
    \label{fig:acceptance1}

\end{figure}

Multivariate normal kernel has a performance between local M-NN M=250 and M=100. It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}. `Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance. Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1. Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).

It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution. This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g. local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold. In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes. Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}. One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.

\subsubsection{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g. local M-NN with M$\leq 250$ (for population size of 2000, i.e. 12.5\% of the population) and multivariate normal with scaling. Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.

% \begin{figure}
%     \begin{center}
%     \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
%     \end{center}

%     \caption{Acceptance rates of different kernels}
%     \label{fig:acceptance1}

% \end{figure}

\subsection{Adaptive functions and factors}

[how adaptive distance work]

SMC can be more adaptive by introducing adaptive distance function and adaptive population; besides, factors can be applied to manually `normalise' the data.

The trivial distance function is set to Euclidean distance (2-norm distance)

\begin{align}
    \label{eq:dis}
    D=\sqrt{\sum_i \Delta x_i^2}
\end{align}

where $i$ is the index of data points and $\Delta x_i$ is the discrepancy between  observed data and simulated data at data point $i$, i.e. $\Delta x_i = x_{i, simulated}-x_{i, observed}$. In this case data points are 12 time points of four variables i.e. 48 data points in total.

A weighted 2-norm distance can be written as

\begin{align}
    \label{dis_w}
    D=\sqrt{\sum_i w_i \Delta x_i^2}
\end{align}

where a weight $w_i$ is assigned to every data point. If data point $i$ have a higher weight, then the data value is regeared to be more informative, i.e. giving more help in inferring the true posteriors; weights can be either pre-set according to prior knowledge of the problem, or using adaptive method to be dynamically calculated according to , known as adaptive distance.

Adaptive distance is changing the weights of all data points after each generation iteration, trying to assign informative data points higher weights. Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}

\begin{align}
    \label{dis_f}
    D=\sqrt{\sum_i f_iw_i \Delta x_i^2}
\end{align}

Factor is helpful as an option for efficiency. When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data point, or behave like a normalisation that can balance the scales of different summarised statistics (here is the 48 mean values). There are two appliance case studied: (1) factors used as an normalisation option and (2) factors used to give more focus on some features of the observed data. For (2), as we noticed that the main features (e.g. rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points i.e. 0 - 30 hpl, factors are tried in the later parameter estimation of real data where a more accurate fit of the curve is desired.

Adaptive population \cite{ref:adpt_pop} can adapt the population size of each generation according to the mean correlation of variation error of the previous population. In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.

    [what is factor]

    [what is adaptive distance]

\subsubsection{Experiments results}

[FIGURE here]

Two types of factors and adaptive distance function are tested by running the same number of generations (20 generation, 2000 particles per generation). From FIGURE, by applying factors the total required samples are less, and the adaptive distance requires much less samples; their acceptance rates are also higher than the standard one. It seems that adaptive function and factors could help to converge more quickly.

However considering the result after 20 generation (FIGURE), the resultant model gives the contrary preference. Applying factors and using adaptive distance can lead the resultant model less accurate (after the same generation iterations): they have wider inter-quantile range and may requires more generations to converge.

\subsubsection{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases. However, as the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, a direct compare of the efficiency under the same threshold schedule is unfeasible; after the same number of generations they can reach a small $\epsilon_t$ but still the approximated posterior is not as accurate as the standard implementation. AS a result, our implementation of the parameter inference will only consider standard distance function and factors are only tested for improvements (FIGURE).

\subsection{Data size, prior distribution and population}

[experiments plan]

Some other hyperparameters, e.g. feed-in data size, prior distribution, number of populations and number of particles in each population were also of our interest. Experiments were designed to explore how these options affect the goodness of fit and efficiency of the implementation.

Regarding the dynamic systems model, SMC was applied with two data size options, three prior distribution (uniform distribution, wider uniform distribution and log-uniform distribution). The population options i.e. population size and number of populations were also tried with different values. These experiments intended to give suggestions on the later SMC implementation on experimental data.

\subsubsection{Experiments results}

The results from data size and prior distribution range is shown in FIGURE. The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement. Although much less (60\%) data is used, the total required sampling numbers is not much less (XX\%). Although the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data (FIGURE), cares should be taken when using less data point, or using other more summative statistics, e.g. mean, standard deviation, peak values, starting and ending values etc. As shown in \cite{ref:disease}, less data can lead to a model with more variance, under-fitting or missing some local features e.g. local peak.

The prior range and distribution can largely affect the execution of SMC (FIGURE and FIGURE). Samples are taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space. It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.

    [discuss uniform and log uniform when result ready]

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/size1.png}}
    \end{center}

    \caption[Data size and prior distribution range experiment]%
    {Data size and prior distribution range experiment. (a) Total required samples, (b) Epsilon values under median epsilon schedule}
    \label{fig:size}

\end{figure}

[FIGURE HERE]

\subsubsection{Summary} This subsection aims to preliminarily explore ABC SMC settings using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options. Conclusion can be drawn to choose efficient and proper options of ABC SMC to be applied on the target models with experimental data. It also provides reference for a general inference task using SMC with high dimensional parameter space.

\section{Parameter estimation and model comparison}

\subsection{Model 1, 2 and 3}

[separated ABC run on model 1, 2 and 3]

Using the suggested options in the previous section, the target data (Figure \ref{fig:obs_data}) is prepared as input for the SMC inference framework. Regarding the result, the population size and number of generations are adjusted several times to obtain an general informative posterior which should be stable across repeated runs and avoid local optimal as much as possible.

Parameter estimation of model 1, 2 and 3 was tried with different prior distribution setting: distribution range [0, 25] and [0, 75], amd distribution type uniform and log-uniform. All these ABC SMC runs uses 2,000 particles per population and 30 generations.

Log-uniform-shaped prior seems to give a better fit of the model, and wider prior range [0, 75] does not give a more accurate result. Then distribution range [0, 25] was tried, in case some parameter have a true posterior laying between [25, 50].

The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}. For model 3, the acceptance rates and epsilon path are presented in Figure \ref{fig:result123_2}. For model 3, the estimated parameters' posterior is shown in Figure \ref{fig:para1}. Figure \ref{fig:result123} plots the estimated curve of models, using the mean value of each parameter's approximated posterior; also 1000 particles (each particle is a parameter set) were sampled and used to generate simulated trajectories of the four variable, and the mean (blue) and 25th to 75th percentile range (grey) are calculated from the 1000 simulated data.

Before applying model comparison methods, some features of the three model can still be compared. From simulated curves, It can be seen that all the inter-quartile ranges are tight around the mean value curve, which indicates the approximated posterior are concentrated to some degree; the peak-shaped posterior distributions (Figure \ref{fig:para1}) agrees with this. All epsilon values converges to a steady level, and model 3 has a lower final epsilon value and consequently the simulated trajectory is more close to the observed data (model 3, Figure \ref{fig:result123}).

The acceptance rates is fluctuating and have a gradually increase trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated. Compared to model 1, model 2 gives a better fit of the decreasing trend at the second half of the observed data, by using a exponentially decaying self-increase rate $\lambda_N$ instead of a constant. The resultant simulated data from model 2 and 3 are in the similar trends and the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'. We expect model 3 to be the best model as it reaches a threshold value and intuitively gives a better fit, although the fit could not be considered generally a good representation of the biological process that we want to model, as the simulated data are significantly biased from the observed data for tnf-$\alpha$. A sharp peak at 4 hpl is observed from the measurement (Figure \ref{fig:obs_data}) but no similar features are represented by any of the three models.

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve123.png}}
    \end{center}

    \caption[Simulated trajectory from the 20th population of model 1, 2 and 3]%
    {Simulated trajectory from the last population of model 1, 2 and 3. Error bar indicates SEM. 500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from}
    \label{fig:result123}

    \vspace*{\floatsep}

    \begin{center}
        \resizebox{0.6\hsize}{!}{\includegraphics{fig/eps_acc123.png}}
    \end{center}

    \caption{Epsilon trends and acceptance rates of model 1, 2 and 3}
    \label{fig:result123_2}

\end{figure}

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/para1234.png}}
    \end{center}

    \caption[Estimated posterior distribution of parameters in model 3]%
    {Estimated posterior distribution of parameters in model 3. Shaded range indicates 25\%--75\% quantile of the population}
    \label{fig:para1}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/cmp1.png}}
    \end{center}

    \caption[Model comparison of model 1, 2 and 3]%
    {Model comparison of model 1, 2 and 3. From population 8 to population 30, model3 always wins}
    \label{fig:cmp1}

\end{figure}

[model comparison among model 1, 2, 3]

Next a model selection experiment with different prior was conducted, using the same ABC-SMC framework. As in the parameter inference, two distributions (uniform and log-uniform) are tested with three different interval ([0, 25], [0, 50] and [0, 75]). The results show that model 3 wins with nearly 100\% model probability at most runs after 30 generations, except log-uniform [0, 75]. All the model probability results are of great discrepancy (e.g. 100\% model 3 and 0\% for model 1 and 2), so Bayesian factor is not calculated.

Figure \ref{fig:cmp1} shows the model selection process, under log-uniform [0, 50] prior distribution. Model 3 gains total advantage over model 1 and 2 after 7th population. Together with conclusions above, model 3 is considered to be the the best model to represent the target data so far.

\begin{table}[t!]
    \centering
    \begin{tabular}{|c c c|}
        \hline
        Parameter            & Estimated mean  & Estimated mean      \\
                             & value (uniform) & value (log-uniform) \\[0.5ex]
        \hline\hline
        $\lambda_N$          & 14.1            & 13.3                \\
        $a$                  & 9.67            & 0.0170              \\
        $\kappa_{N\beta}$    & 18.0            & 0.259               \\
        $\mu_N$              & 13.2            & 0.0180              \\
        $\nu_{N\Phi}$        & 0.742           & 0.131               \\
        \hline
        $\kappa_{\Phi\beta}$ & 0.239           & 0.156               \\
        $\mu_\Phi$           & 0.154           & 0.0350              \\
        \hline
        $s_{\beta N}$        & 6.75            & 1.21                \\
        $\mu_\beta$          & 6.38            & 1.32                \\
        \hline
        $s_{\alpha\Phi}$     & 17.0            & 5.83                \\
        $\mu_\alpha$         & 11.9            & 2.43                \\
        \hline
    \end{tabular}
    \caption[Estimated parameter values of model 3]
    {Estimated parameter values of model 3, using uniform prior distribution and log-uniform prior distribution respectively}
    \label{table:estimated1}
\end{table}

Compared to uniform distributed prior, we found that log-uniform distributed prior is more suitable in the inference. Log-uniform results in an obviously better fit and narrower inter-quantile interval (Figure \ref{fig:result123} and \ref{fig:resultCurve_uni}), and the final epsilon value is also smaller. Log-uniform tends to assign values that are close to the left boundary (i.e. zero in our case) higher density, i.e. they are more likely to be sampled in the first population. As a result, the estimated parameter posteriors are expected to have smaller mean and/or skewed to left of the x-axis. The estimated values proved this (see \ref{table:estimated1}; also observed in other models and prior interval experiments). It indicates that some of the true parameter tend to have small values that are close to zero, and using log-uniform distribution could results in a faster and more satisfying inference. Regarding this, further experiments all use log-uniform distribution only.

\subsection{Model 4 and 5}

[features that derive model 4 and 5]

We observed that the existing models cannot well represent the trajectory of relative expression of tnf-$\alpha$: from experimental data, a rapid increase of tnf-$\alpha$ expression is observed in 0--4 hpl, followed by a rapid drop; all the models cannot well recover this feature and most test results see a less-rapid increase followed by a gradually decree and the observed peak value is not reached (e.g. Figure \ref{fig:result123}). Also in some other case e.g. Figure \ref{fig:resultCurve_uni}, tnf-$\alpha$ trajectories is well fitted at the cost of under-fitting $\Phi$. Hence efforts had been spent to propose more alternative models that can possibly solve this problem.

The two proposed alternative models is based on model 3 and try to add extra term in the tnf-$\alpha$ equation $\mathrm{d} \alpha/\mathrm{d} t$. As macrophage is the main source of tnf-$\alpha$ production, and a similar sharp increase-and-drop trend is observed in the il-1$\beta$ trajectory, two hypothesis and corresponding terms are propose as follows.

\paragraph{Model 4} It is assumed that the expression of il-1$\beta$ would have a promoting effect, representing by a phenomenological term $d_{\beta\alpha}\beta$ and $d_{\beta\alpha}$ is a new model parameter (positive constant). The promoting effect here is regarded to have the equivalent effect as directly promoting by il-1$\beta$, but the underlying mechanism is unclear so far. In mathematical view, this additional term can accelerate the expression of tnf-$\alpha$ in the first few time points and help to recover the peak-shaped trajectory. The proposed model is written as Equation \ref{eq:model4}.

\paragraph{Model 5} Alternatively, we considered promotions to the tnf-$\alpha$ production, i.e. $s_{\alpha\Phi}\Phi$. It is assumed that il-$\beta$ could accelerate the production process. An additional term $f_{\beta\alpha}\beta$ is introduced to production rate, with $f_{\beta\alpha}$ being a new model parameter (positive constant). This model meets the biological context considering the source of tnf-$\alpha$. The proposed model is written as Equation \ref{eq:model5}.

[further comparison of model 3, 4 and 5]

Model 4 and 5 are base on model 3, so in this phase experiments of these three models are conducted for separately parameter inference and overall model comparison. Based on our experience in conducted experiments, we used the following setting for parameter inference:

\begin{itemize}
    \item population size: 2000 and 5000
    \item number of populations: 30 generations
    \item prior: [0, 50] log-uniform
    \item perturbation kernel: multivariate normal kernel
\end{itemize}

Factors (Equation \ref{dis_f}) are also tried to find if we can force the inference framework to give more importance on the first few data points. In this experiment, for each trajectory first 8 data points are assigned higher factors (0.75), and the rest 4 data points are assigned with lower factors (0.25). All these runs are conducted on remote machines and the output database files are retrieved form analysis.

Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 2000. It can be seen that the addressed problem in the trajectory of tnf-$\alpha$ is partially relieved in model 4 and 5: compared to model 3, a peak appears around 4 hpl; it can represent more features in the observed data and consequently the final reached epsilon (after 30 populations) is smaller than that of model 3, which proves that our proposed modifications in model 4 and 5 are helpful in recovering more features in observed data.

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve345.png}}
    \end{center}

    \caption{Simulated data from the last population of model 3, 4 and 5}
    \label{fig:resultCurve345}

\end{figure}

By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features. Model 1, 2 and 3 con also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- is not fitted. Some features, e.g. fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.

After applying factors, the results were not largely different. As expected, applying factors made the data points at first half (where exists most of the fluctuations) more `focused', however the inferred models are close to one without factors, which could indicate that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.

When we are trying more populations, an `over-fitting' behaviour is observed at the trajectory of macrophage (FIGURE). In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).

Next, we performed a model selection on model 3, 4 and 5. FIGURE shows the model probabilities at different generations. From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred. Again, as DISCUSSED in the model comparison of model 1, 2 and 3, we considered the the violently changing model probability is related to the inference ability at certain epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space. Given the present result, model 5 is regarded as the best model. Some further experiments tried larger population size (10,000 particles per population) and more populations (36 populations), but the resultant model did not have a significantly better fit of the observed data, although much more computation resources are used.

A inferred parameter posterior distribution of model 5 is shown in FIGURE, the estimated mean values is shown in TABLE. SOME DESCRIPTION.

\subsection{Sensitivity of parameters}

[why last PCs]

[which parameters are most sensitive]

[compared to credible intervals plot]

[what does that mean]

A quick sensitivity analysis of the dynamical systems models can be quantified using principle component analysis (PCA) \cite{Toni}, where the first PCs corresponds to sloppy parameters while the last PCs corresponds to stiff parameters \cite{sensitivity}. PCA can only provide rough approximated sensitivity behaviour. As models 5 won in the model selection, a further interest is to explore the parameter-to-model sensitivity

FIGURE shows the results of PCA on the last population of inferred model5. The last PC (PC12) is mostly consists of the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive. If we conclude the last 5 PCs (FIGURE), then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$. As a conclusion, model 5 is sensitive to changes in the above parameters, which all come from the equation of neutrophil and macrophage i.e. equations of the cells; it may suggests that given the observed data, the inferred model is more sensitive to cells' dynamic rather than cytokines' dynamics.

Also, some visualisations of the last population agreed with conclusions from PCA. Approximated posteriors of the last population plotted in FIGURE and credible ranges of parameters across populations in FIGURE shows that the 5 parameters identified by PCA is indeed `stiff' parameters.

### 2020-08-09 introduction/introduction.tex
New structure

\chapter{Introduction}

 % This should contain a description of your project and the problem you
 % are trying to solve. Where appropriate you should also include
 % references to work which has already been done on your topic and
 % anything else which lets you set your work in context.

 % One of the things you will need to do is to ensure that you have a
 % suitable list of references.  To do this you should see \cite{ref:lam}
 % or some other suitable reference.  Note the format of the citation used
 % here is the style favoured in this department.  Here is another
 % reference \cite{ref:bloggs} for good measure.

 [background and motivation]

 [what is done]

While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions of cells and chemicals in a biological system. The objective process in this dissertation is an example of complex biology system found in tissue regeneration. Zebrafish has the ability to regenerate its spinal cord after injury; at the injury-site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.

Our interest lies in the modeling of the regeneration process that happens in the injury-site, especially in the parameter estimation and comparison of different models. According to available researches \cite{ref:Tsarouchas}, a mathematical model is proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments. However, we would like to explore the the `best fit' if the model which comes with a set of parameter estimates given the observed data. Also, we wish to compare different it with different models and possibly find the best model to describe the data without losing biological and mathematical properties in the regeneration process.

Fo the parameter estimation tasks, Bayesian inference techniques can help in the problems of over-fitting and local optima \cite{ref:abcsysbio}, compared to optimisation based methods. Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression can be hard.

Such methods introduce computation-intensive sampling-rejection pattern, where massive number of samples are drawn, followed by computations. As the sampling process is independent from each other in a certain period, the parallel execution becomes feasible and is expected to accelerated the parameter inference in a large scale.

This dissertation includes the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results. Model comparison results are also included using the same method. Additionally, a performance test illustrates the strong-scaling performance of the algorithm.

### 2020-08-09 modelling/modelling.tex
New structure

--- 2020-07-11 __latexindent_temp.tex
@@
-\documentclass[12pt,a4paper]{report}
-
-\usepackage{epcc}
-\usepackage{graphics}
-\usepackage{mathtools}
-
-% This example file shows how a thesis can be laid out using Latex. It
-% does not use any special local features so should be portable to other
-% places.
-% 
-% When producing draft copies of a thesis you may want to print only
-% selected pages of the thesis. To do this use the command
-% 
-% dvips -f -p 4 -n 3 myfile.dvi | lpr
-% 
-% where -p 4 means start printing at page 4 (ie the page that will be
-% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
-% This example will print pages 4, 5 and 6.
-% 
-% If you want to print the thesis and also save paper you can print more
-% than one page on each sheet of paper. Use the command
-% 
-% dvips -f myfile.dvi | psnup -2 | lpr
-% 
-% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
-%
-% To produce a PDF version you can create a PostScript copy first
-%
-% dvips -f myfile.dvi > myfile.pdf
-%
-% and then convert it
-%
-% distill myfile.ps
-%
-% or you can go straight to PDF
-%
-% pdflatex myfile
-%
-% Note that pdflatex expects all included figures to be in PDF too. See
-% the includegraphics command below.
-
-% This document contains many cross-references and forward references,
-% eg in constructing a table of contents, so Latex may need to be run
-% twice to get all the references correct. If you need to run Latex twice
-% you may get the warning:
-% 
-% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.
-
-% the following 4 lines are the content of the smallmargins.sty file
-% but including them explicitly makes this more portable.
-%AC%\oddsidemargin=0.1in
-%AC%\topmargin=-0.5in
-%AC%\textheight=9in
-%AC%\textwidth=6.25in
-
-%AC%\parskip 10pt
-%AC%\parindent 0in
-
-\begin{document}
-
-%AC%\pagestyle{myheadings}
-%AC%\markright{D.~S.~Henty}
-
-%\title{A Latex thesis example}
-%\author{D.~S.~Henty}
-%\date{\today}
-
-%\maketitle
-
-\pagenumbering{roman}
-
-\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
-\author{Chaolin Han}
-\date{\today}
-
-\makeEPCCtitle
-
-\thispagestyle{empty}
-
-\vspace{11cm}
-
-\begin{center}
-
-\large{MSc in High Performance Computing}
-
-\large{The University of Edinburgh}
-
-\large{Year of Presentation: 2020}
-
-\end{center}
-
-\newpage
-
-\begin{abstract}
-This is the bit where you summarise what is in your thesis.
-\end{abstract}
-
-\pagenumbering{roman}
-
-\tableofcontents
-\listoftables
-\listoffigures
-
-\begin{titlepage}
-\vspace*{2in}
-% an acknowledgements section is completely optional but if you decide
-% not to include it you should still include an empty {titlepage}
-% environment as this initialises things like section and page numbering.
-\section*{Acknowledgements}
-
-This template is a slightly modified version of the one developed by
-Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
-acknowledgement follows:
-
-{\em This template has been produced with help from many former students who
-have shown different ways of doing things. Please make suggestions for
-further improvements.}
-
-\end{titlepage}
-
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-This should contain a description of your project and the problem you
-are trying to solve. Where appropriate you should also include
-references to work which has already been done on your topic and
-anything else which lets you set your work in context.
-
-One of the things you will need to do is to ensure that you have a
-suitable list of references.  To do this you should see \cite{ref:lam}
-or some other suitable reference.  Note the format of the citation used
-here is the style favoured in this department.  Here is another
-reference \cite{ref:bloggs} for good measure.
-
-You will also want to make sure you have no spelling or grammatical
-mistakes. To help idwentify spelling mistukes you caan use the commands
-{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
-that spelling mistakes are not the only errors which can occur. Spelling
-checkers will not find errors which are, in fact, valid words such as
-{\em there} for {\em their}, nor will they find repeated words which
-sometimes occur if your concentration is broken when typing. {\bf There
-is no substitute for thorough proof reading!}
-
-\chapter{Background}
-
-\section{Zebrafish spinal cord regeneration}
-
-\section{Mathematical modelling}
-
-\subsection{Hard bits}
-You might want to include an equation here:
-
-\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
-\label{equation:delsplit}
-% note that the label is optional but allows you to refer to this
-% equation later.
-\end{equation}
-
-\subsection{Even harder bits}
-This might be one of the places where you might want to refer to
-equation \ref{equation:delsplit}. You will usually need to use the
-Latex command twice to make cross-references like this work properly.
-The cross-reference information is stored in the {\em .aux} file so don't
-delete it.
-
-\subsubsection{Numbering}
-You can keep subdividing but eventually you get to a level where
-numbering stops. This text is in a subsubsection which is not numbered
-by default.
-
-\paragraph{More on numbering:}
-
-This text is in a paragraph which is also not numbered by default and
-the ``title'' of the paragraph is not on a separate line.
-If you want to increase the depth to which sections are numbered you
-should see the section on setting the secnumdepth counter in the manual.
-
-\section{Bayesian inference}
-
-\section{Software tools}
-
 \chapter{Mathematical modelling}
 
-Mathematical models can describe different kinds of dynamic system, and can be used as a guide to prediction and analysis.
-An ideal model in this case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  trend against time.
+Mathematical models can describe different kinds of dynamic systems, and thus can be used as a guide for prediction and analysis.
+An ideal model in our case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  data features in different time point.
 
-Models for the regeneration process is in the form of ordinary differential equations, specifically the time differential form.
+Proposed models for the regeneration process are in the form of ordinary differential equations (ODE), specifically the time differential form.
 Terms in the ODE are mostly explainable and corresponds interaction paths.
@@
 
-Our models is built regarding the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement includes the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
+Our models is built on the basis of the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
+The measurement data include the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
 As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
-According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration'.
+According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$).
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration' (compared to the value at first time point).
 
 It is noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations is mean of measurement, assuming that measured data is Gaussian-like distributed.
-The distribution of the measured data points is plotted and examined to see if the measurement mean could represent the distribution.
-The result is that at most time points the measurement values are Gaussian-distributed, some distributions are skewed.
+The summary statistic used for the parameter estimations is mean of measurement data at each time point, assuming that measured data is Gaussian-like distributed.
+To validate this, the distribution of the measured data points is plotted and examined to see if the mean value can represent the distribution.
+The result is that at most time points the measurement values are Gaussian-distributed, although some distributions are skewed.
 One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
-Mean value can summarise most data measures and thus is still used as the target summarised observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}
+Mean value can summarise most data and thus is still used as the target observed data.
+A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}.
 
 \begin{figure}
-
     \begin{center}
@@
 
-\caption{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
+    \caption[Mean of the observed data]%
+    {Mean of the observed data for neutrophil ($N$), macrophage ($\Phi$), il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean}
     \label{fig:obs_data}
@@
 At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
-After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can be more close to the true process, recover more features of the observed data or test our hypothesis.
+After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can represent the observed regeneration process better or can be used to test our hypothesis.
 
-All these models assumes the interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for parameter estimation.
-Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data.
+All these models assume the involved interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for the inference task.
+Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data available.
 
@@
 
-A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
+A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the inference framework code.
+An interaction map of the model is shown in Figure \ref{fig:m1}.
+This model is a simplification a the process described in \cite{ref:Tsarouchas} with some minor interactions ignored.
+To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
+number of cells as `cell', and denote the unit of $\beta$ and $\alpha$ i.e.
+relative mRNA expression as `unit' for simplicity.
 
-\begin{equation*}\label{eq:pareto mle2}
-    \begin{multlined}
-        \frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
-\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-   \end{multlined}
-  \end{equation*}
+\begin{figure}
+    \begin{center}
+        \resizebox{0.4\hsize}{!}{\includegraphics{fig/model1.png}}
+    \end{center}
+
+    \caption[Interactions modelled in the basic model]%
+    {Interactions modelled in the basic model (model 1) based on Tsarouchas et al.\cite{ref:Tsarouchas}.
+Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
+    \label{fig:m1}
+
+\end{figure}
+
+It is assumed that there are negative feedbacks for all the variables. MORE DESCRIBE.
+
+\begin{align}
+    \label{eq:model1}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c c|}
+        \hline
+        Parameter            & Definition                                       & Units                    \\ [0.5ex]
+        \hline\hline
+        $\lambda_N$          & Self-increase rate of neutrophil                 & $cell/h$                 \\
+        $\kappa_{N\beta}$    & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
+        $\mu_N$              & Coefficient of negative feedback of $N$          & $h^{-1}$                 \\
+        $\nu_{N\Phi}$        & Coefficient of inhibition of both $N$ and $\Phi$ & $cell^{-1}\cdotp h^{-1}$ \\
+        \hline
+        $\lambda_\Phi$       & Self-increase rate of macrophage                 & $cell/h$                 \\
+        $\kappa_{\Phi\beta}$ & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
+        $\mu_\Phi$           & Coefficient of negative feedback of $\Phi$       & $h^{-1}$                 \\
+        \hline
+        $s_{\beta N}$        & Production rate from $N$                         & $unit/(cell\cdotp h)$    \\
+        $i_{\beta\Phi}$      & Coefficient of inhibition to the production      & $cell^{-1}$              \\
+        $\mu_\beta$          & Coefficient of negative feedback of $\beta$      & $h^{-1}$                 \\
+        \hline
+        $s_{\alpha\Phi}$     & Production rate from $\Phi$                      & $unit/(cell\cdotp h)$    \\
+        $\mu_\alpha$         & Coefficient of negative feedback of $\alpha$     & $h^{-1}$                 \\
+        \hline
+    \end{tabular}
+    \caption{Parameters introduced in the basic model (model 1) [REMOVE unit]}
+    \label{table:m1}
+\end{table}
 
@@
 
+\paragraph{Model 2 and model 3}
+
+As the observed data indicate, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
+Regarding this, the self-increase parameter $\lambda$ cannot be constant, thus a exponentially decaying $\lambda$ term is introduced and model 2 is proposed.
+(Eqn.
+\ref{eq:model2}).
+Also, the inhibition of il-1$\beta$ production, i.e.
+term $i_{\beta\Phi}\Phi$ is considered to be ignored for a more simple model, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
+This case corresponds to model 3, written as Eqn.
+\ref{eq:model3}.
+
+Model 2 and model 3 introduce one extra parameter $a$, which is a coefficient in the exponentially decaying $\lambda$, determining the decay speed, with the unit $h^{-1}$.
+
+\begin{align}
+    \label{eq:model2}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{align}
+    \label{eq:model3}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+Model 3 can be regarded as a simplification of model 2, as it can be treated as model 2 with parameter $i_{\beta\Phi}=0$.
+Among the proposed three models, model 1 is a naive one that was proposed at very first time and used as an ODE `template' to build and test parameter inference framework.
+As the implementation was successful, model 2 and 3 was proposed as we were trying to calibrate some terms and find a better model.
+After fitting, model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points (which relate to the steady state).
+Model 3 makes a small simplification on model 2 and is theoretically less `general' than model 2.
+
+\paragraph{Model 4 and model 5}
+
+After the first model selection experiment, it was found that some significant features presented in the observed data were not recovered by any of the existing model.
+To resolve this, attempts were tried to introduce more interactions within the dynamic system considering the biological and  mathematical context.
+Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or adding a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
+\ref{eq:model4}) and model 5 (Eqn.
+\ref{eq:model5}).
+
+\begin{align}
+    \label{eq:model4}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha+d_{\beta\alpha}\beta
+    \end{split}
+\end{align}
+
+\begin{align}
+    \label{eq:model5}
+    \begin{split}
+        &\frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_Ne^{-at}+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
+        &\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
+        &\frac{\mathrm{d} \beta}{\mathrm{d} t}=s_{\beta N}N-\mu_\beta\beta\\
+        &\frac{\mathrm{d} \alpha}{\mathrm{d} t}=(s_{\alpha\Phi}+f_{\beta\alpha}\beta)\Phi-\mu_\alpha\alpha
+    \end{split}
+\end{align}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c c c|}
+        \hline
+        Parameter         & Definition                                                     & Units                    \\ [0.5ex]
+        \hline\hline
+        $d_{\beta\alpha}$ & Coefficient of promoting effect from $\beta$                   & $h^{-1}$                 \\
+        \hline
+        $f_{\beta\alpha}$ & Coefficient of promoting the production of $\alpha$ by $\beta$ & $cell^{-1}\cdotp h^{-1}$ \\
+        \hline
+    \end{tabular}
+    \caption{Parameters introduced in model 4 and 5}
+    \label{table:m45}
+\end{table}
+
 \subsection{Model evaluation and comparison}
+
+[general topics on evaluating a model]
+
+[bayes factor for model selection]
+
+\subsection{Limitations}
+
+[hypothesis]
+
+[available data]
+
+[model misspecification]
 
@@
 % the figure relative to the text if it cannot be fitted onto the current page.
-
-\chapter{Implementations and Experiments}
-
-\section{Parameter estimation}
-Here are some results.
-
-\section{Model comparison}
-Here are some results.
-
-\section{Performance experiments}
-
-The performance experiments are designed to explore the parallel performance of ABC SMC implementations.
-Usually ABC SMC is a time-consuming and computation intensive task and usually is executed on large clusters.
-The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
-
-\subsection{Scaling-up}
-
-First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
-
-\begin{itemize}
-    \item Prior distribution: default to log-uniform distribution [$1\times 10^{-6}$, 50] for all the 12 parameters
-    \item threshold schedule: median epsilon
-    \item No factors, no adaptive distance or adaptive population applied
-    \item Population size is 2000, with 20 generations
-\end{itemize}
-
-HOW PYABC parallelise the sampling
-
-For HPC systems like Cirrus, \verb|pyabc| uses \verb|multiprocessing| for multi-core parallel sampling.
-By default if the number of cores is not specified, it will automatically read the number of available cores and use them all.
-Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
-
-The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54, 72 cores respectively.
-Each run is repeated 5 times.
-The average execution time, required sampling numbers are recorded.
-Hyperthreading is enabled when using 54 and 72 cores.
-The access to the node that contains computation cores is exclusive, such that the execution would not be affected by other programs of operations.
-
-The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the must time-consuming part.
-The rest part of the program is mostly not parallelised, e.g.
-database I/O and reductions operations.
-The sampling process involves sampling, perturbation and test of the acceptance criteria, all of which are computation-intensive.
-
-In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources is limited[REF].
-The performance experiment result could provide a reference that illustrate that how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
-
-\subsection{Profiling}
-
-The performance could also be analysed given a profiling report.
-The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck, according to which we could find the hot-spot of program and given possible suggestions on improving the performance.
-
-In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
-
-\chapter{Results Analysis}
-
-\section{ABC SMC results}
-
-\section{Performance experiments}
-
-\section{Discussions}
-This is the place to put your conclusions about your work. You can
-split it into different sections if appropriate. You may want to include
-a section of future work which could be carried out to continue your
-research.
-
-\chapter{Future works}
-
-\chapter{Conclusions}
-
-\appendix
-% the appendix command just changes heading styles for appendices.
-
-\chapter{Stuff which is too detailed}
-
-Appendices should contain all the material which is considered too
-detailed to be included in the main bod but which is, nevertheless,
-important enough to be included in the thesis.
-
-\chapter{Stuff which no-one will read}
-
-Some people include in their thesis a lot of detail, particularly
-computer code, which no-one will ever read. You should be careful that
-anything like this you include should contain some element of uniqueness
-which justifies its inclusion.
-
-\begin{thebibliography}{100}
-
-\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
-and Reference Manual.} Addison Wesley. pp242.
-
-\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
-in Environments} Int. Journal of Silly Findings. pp 23-29.
-
-\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N.
-and Feng, Y., 2018.
-Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
-Nature communications, 9(1), pp.1-17.
-
-\end{thebibliography}
-
-\end{document}

### 2020-08-09 performance/performance.tex
New structure

--- 2020-07-11 __latexindent_temp.tex
@@
-\documentclass[12pt,a4paper]{report}
+\chapter{Performance experiments}
 
-\usepackage{epcc}
-\usepackage{graphics}
-\usepackage{mathtools}
-
-% This example file shows how a thesis can be laid out using Latex. It
-% does not use any special local features so should be portable to other
-% places.
-% 
-% When producing draft copies of a thesis you may want to print only
-% selected pages of the thesis. To do this use the command
-% 
-% dvips -f -p 4 -n 3 myfile.dvi | lpr
-% 
-% where -p 4 means start printing at page 4 (ie the page that will be
-% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
-% This example will print pages 4, 5 and 6.
-% 
-% If you want to print the thesis and also save paper you can print more
-% than one page on each sheet of paper. Use the command
-% 
-% dvips -f myfile.dvi | psnup -2 | lpr
-% 
-% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
-%
-% To produce a PDF version you can create a PostScript copy first
-%
-% dvips -f myfile.dvi > myfile.pdf
-%
-% and then convert it
-%
-% distill myfile.ps
-%
-% or you can go straight to PDF
-%
-% pdflatex myfile
-%
-% Note that pdflatex expects all included figures to be in PDF too. See
-% the includegraphics command below.
-
-% This document contains many cross-references and forward references,
-% eg in constructing a table of contents, so Latex may need to be run
-% twice to get all the references correct. If you need to run Latex twice
-% you may get the warning:
-% 
-% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.
-
-% the following 4 lines are the content of the smallmargins.sty file
-% but including them explicitly makes this more portable.
-%AC%\oddsidemargin=0.1in
-%AC%\topmargin=-0.5in
-%AC%\textheight=9in
-%AC%\textwidth=6.25in
-
-%AC%\parskip 10pt
-%AC%\parindent 0in
-
-\begin{document}
-
-%AC%\pagestyle{myheadings}
-%AC%\markright{D.~S.~Henty}
-
-%\title{A Latex thesis example}
-%\author{D.~S.~Henty}
-%\date{\today}
-
-%\maketitle
-
-\pagenumbering{roman}
-
-\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}
-\author{Chaolin Han}
-\date{\today}
-
-\makeEPCCtitle
-
-\thispagestyle{empty}
-
-\vspace{11cm}
-
-\begin{center}
-
-\large{MSc in High Performance Computing}
-
-\large{The University of Edinburgh}
-
-\large{Year of Presentation: 2020}
-
-\end{center}
-
-\newpage
-
-\begin{abstract}
-This is the bit where you summarise what is in your thesis.
-\end{abstract}
-
-\pagenumbering{roman}
-
-\tableofcontents
-\listoftables
-\listoffigures
-
-\begin{titlepage}
-\vspace*{2in}
-% an acknowledgements section is completely optional but if you decide
-% not to include it you should still include an empty {titlepage}
-% environment as this initialises things like section and page numbering.
-\section*{Acknowledgements}
-
-This template is a slightly modified version of the one developed by
-Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
-acknowledgement follows:
-
-{\em This template has been produced with help from many former students who
-have shown different ways of doing things. Please make suggestions for
-further improvements.}
-
-\end{titlepage}
-
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-This should contain a description of your project and the problem you
-are trying to solve. Where appropriate you should also include
-references to work which has already been done on your topic and
-anything else which lets you set your work in context.
-
-One of the things you will need to do is to ensure that you have a
-suitable list of references.  To do this you should see \cite{ref:lam}
-or some other suitable reference.  Note the format of the citation used
-here is the style favoured in this department.  Here is another
-reference \cite{ref:bloggs} for good measure.
-
-You will also want to make sure you have no spelling or grammatical
-mistakes. To help idwentify spelling mistukes you caan use the commands
-{\em ispell} or {\em spell}. See the appropriate manual pages. Remember
-that spelling mistakes are not the only errors which can occur. Spelling
-checkers will not find errors which are, in fact, valid words such as
-{\em there} for {\em their}, nor will they find repeated words which
-sometimes occur if your concentration is broken when typing. {\bf There
-is no substitute for thorough proof reading!}
-
-\chapter{Background}
-
-\section{Zebrafish spinal cord regeneration}
-
-\section{Mathematical modelling}
-
-\subsection{Hard bits}
-You might want to include an equation here:
-
-\begin{equation} \delta N_{\nu} = (\delta N_{\nu})_{ex} + (\delta N_{\nu})_{au} 
-\label{equation:delsplit}
-% note that the label is optional but allows you to refer to this
-% equation later.
-\end{equation}
-
-\subsection{Even harder bits}
-This might be one of the places where you might want to refer to
-equation \ref{equation:delsplit}. You will usually need to use the
-Latex command twice to make cross-references like this work properly.
-The cross-reference information is stored in the {\em .aux} file so don't
-delete it.
-
-\subsubsection{Numbering}
-You can keep subdividing but eventually you get to a level where
-numbering stops. This text is in a subsubsection which is not numbered
-by default.
-
-\paragraph{More on numbering:}
-
-This text is in a paragraph which is also not numbered by default and
-the ``title'' of the paragraph is not on a separate line.
-If you want to increase the depth to which sections are numbered you
-should see the section on setting the secnumdepth counter in the manual.
-
-\section{Bayesian inference}
-
-\section{Software tools}
-
-\chapter{Mathematical modelling}
-
-Mathematical models can describe different kinds of dynamic system, and can be used as a guide to prediction and analysis.
-An ideal model in this case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  trend against time.
-
-Models for the regeneration process is in the form of ordinary differential equations, specifically the time differential form.
-Terms in the ODE are mostly explainable and corresponds interaction paths.
-
-\section{Observed data}
-
-Our models is built regarding the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement includes the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
-As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
-According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (il-1$\beta$) and $\alpha$ (tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration'.
-
-It is noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations is mean of measurement, assuming that measured data is Gaussian-like distributed.
-The distribution of the measured data points is plotted and examined to see if the measurement mean could represent the distribution.
-The result is that at most time points the measurement values are Gaussian-distributed, some distributions are skewed.
-One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
-Mean value can summarise most data measures and thus is still used as the target summarised observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}
-
-\begin{figure}
-
-\begin{center}
-\resizebox{1.0\hsize}{!}{\includegraphics{fig/obs_curve.png}}
-\end{center}
-
-\caption{Mean of the observed data for neutrophil $N$, macrophage $\Phi$, il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean} 
-\label{fig:obs_data}
-
-\end{figure}
-
-\section{Hypothesis and Models}
-
-5 models in total are proposed according to different hypothesis.
-At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
-After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can be more close to the true process, recover more features of the observed data or test our hypothesis.
-
-All these models assumes the interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for parameter estimation.
-Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data.
-
-\subsection{Basic model}
-
-A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the code.
-
-\begin{equation*}\label{eq:pareto mle2}
-    \begin{multlined}
-        \frac{\mathrm{d} N}{\mathrm{d} t}=\lambda_N+\kappa_{N\beta}\beta-\mu_NN-\nu_{N\Phi}N\Phi\\
-\frac{\mathrm{d} \Phi}{\mathrm{d} t}=\lambda_\Phi+\kappa_{\Phi\beta}\beta-\mu_\Phi\Phi\\
-\frac{\mathrm{d} \beta}{\mathrm{d} t}=\frac{s_{\beta N}N}{1+i_{\beta\Phi}\Phi}-\mu_\beta\beta\\
-\frac{\mathrm{d} \alpha}{\mathrm{d} t}=s_{\alpha\Phi}\Phi-\mu_\alpha\alpha
-   \end{multlined}
-  \end{equation*}
-
-\subsection{Alternative models}
-
-\subsection{Model evaluation and comparison}
-
-% \begin{figure}
-
-% \begin{center}
-% \resizebox{0.30\hsize}{!}{\includegraphics{logos/crest_bw}}
-% \end{center}
-
-% \caption{The University Crest}
-% \label{fig:eucrest}
-
-% \end{figure}
-
-% see the man page for dvips for details of the special command which is
-% much more powerful than is shown here. It allows offsets in the
-% horizontal and vertical and scaling in x and y.
-
-% choosing suitable values for offset and scale can be a tiring matter
-% of trial and error.
-
-% note that labels do not need to include a description of the object
-% they are labelling but it can be helpful, eg \label{fig:figurename}.
-
-% You can use a label on a figure to refer to it later. The university
-% crest is in \ref{fig:eucrest}. Note that you should not use phrases like
-% ``the figure above'' or ``the following figure'' since Latex may move
-% the figure relative to the text if it cannot be fitted onto the current page.
-
-\chapter{Implementations and Experiments}
-
-\section{Parameter estimation}
-Here are some results.
-
-\section{Model comparison}
-Here are some results.
-
-\section{Performance experiments}
-
-The performance experiments are designed to explore the parallel performance of ABC SMC implementations.
-Usually ABC SMC is a time-consuming and computation intensive task and usually is executed on large clusters.
+The performance experiments were designed to explore the parallel performance of ABC SMC inference framework which were used in the parameter estimation and model selection tasks.
+Usually ABC SMC is a time-consuming and computation intensive task and ideally executed on large clusters.
 The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
 
-\subsection{Scaling-up}
+    [MORE meanings of the performance study]
+
+\section{Scaling-up}
 
@@
 
-The program is executed on Cirrus, using 8, 12, 16, 24, 36, 54, 72 cores respectively.
-Each run is repeated 5 times.
+The program is executed on Cirrus, using 8, 16, 24, 36, 54 and 72 cores respectively.
+Each run is repeated 10 times.
 The average execution time, required sampling numbers are recorded.
@@
 
-\subsection{Profiling}
+\subsubsection{Results}
+
+[scaling-up performance: speed-up and efficiency]
+
+[large variation in required sampling numbers]
+
+[possible reasons]
+
+The recorded execution time using different number of cores is shown in FIGURE.
+According to that, speed-up and efficiency can be calculated.
+Here 8 cores is regarded as the baseline for speed-up and efficiency, as the serial version (using only one core) can take quite a long time to finish 10 repeats.
+When increasing number of cores, the execution time drops fast at first but the decreasing rate (dropping speed, i.e.
+the derivative of the execution time against time) is gradually reduced to zero; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (FIGURE).
+
+Denote the number of cores used in the ABC SMC runs as variable $p$.
+The speedup curve shows a linear trend when $p\leq 36$; when $p\geq 36$ the speedup stays nearly unchanged.
+$p=36$ is an inflection point, where 36 is the maximum numbers available physical cores in a compute node of Cirrus.
+A constant drop of efficiency is observed when increasing $p$; smaller $p$ ($p<36$) gives an efficiency higher than 65\% but greater $p$ e.g.
+$p=72$ results in a low efficiency (34\%).
+
+A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations.
+Moreover, we found that the execution time is not ideally proportional to the total required samples (FIGURE).
+A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs.
+Some runs that are stuck in a local optimum and require smaller target epsilon will take huge amount of samples to move away from the local optimum.
+This have been observed when the distribution of posterior is plotted before and after jumping away from local modes (FIGURE and more explain).
+Also, time spent in each sample using sanme number of cores is not constant in our case, which might be a result of the parallel schedule; ideally a serial run using only one core will see a nearly constant time spent in every particle.
+
+Due to the unstable sampling numbers, our interests switched to the per-particle performance, where the average sampling numbers per second (FIGURE) and average time spent in one particle (FIGURE) under different of cores ($p$) are plotted.
+The log-like average time per particle curve shows that using more cores can improve the sampling efficiency, but when $p$ is at a high level (e.g.
+36 in our case), the improvement will be less significant.
+
+\section{Profiling}
 
@@
 
-\chapter{Results Analysis}
+\subsubsection{Results}
 
-\section{ABC SMC results}
-
-\section{Performance experiments}
-
-\section{Discussions}
-This is the place to put your conclusions about your work. You can
-split it into different sections if appropriate. You may want to include
-a section of future work which could be carried out to continue your
-research.
-
-\chapter{Future works}
-
-\chapter{Conclusions}
-
-\appendix
-% the appendix command just changes heading styles for appendices.
-
-\chapter{Stuff which is too detailed}
-
-Appendices should contain all the material which is considered too
-detailed to be included in the main bod but which is, nevertheless,
-important enough to be included in the thesis.
-
-\chapter{Stuff which no-one will read}
-
-Some people include in their thesis a lot of detail, particularly
-computer code, which no-one will ever read. You should be careful that
-anything like this you include should contain some element of uniqueness
-which justifies its inclusion.
-
-\begin{thebibliography}{100}
-
-\bibitem{ref:lam} L.Lamport. {\em 1986 Latex User's Guide
-and Reference Manual.} Addison Wesley. pp242.
-
-\bibitem{ref:bloggs} F.Bloggs. {\em 1993 Latex Users do it
-in Environments} Int. Journal of Silly Findings. pp 23-29.
-
-\bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N.
-and Feng, Y., 2018.
-Dynamic control of proinflam- matory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
-Nature communications, 9(1), pp.1-17.
-
-\end{thebibliography}
-
-\end{document}
+[profiling results: hot-spot and possible improvements]

### 2020-08-09 MSc-dissertation.tex
Bib restructure

@@
 
-\begin{thebibliography}{100}
+\bibliographystyle{ieeetr}
 
-    \bibitem{axonal} Becker, T., Wullimann, M.F., Becker, C.G., Bernhardt, R.R.
-and Schachner, M., 1997.
-Axonal regrowth after spinal cord transection in adult zebrafish.
-Journal of Comparative Neurology, 377(4), pp.577-595.
-
-    \bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N.
-and Feng, Y., 2018.
-Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration.
-Nature communications, 9(1), pp.1-17.
-
-    \bibitem{Anderson} Anderson, W.D., Makadia, H.K., Greenhalgh, A.D., Schwaber, J.S., David, S.
-and Vadigepalli, R., 2015.
-Computational modeling of cytokine signaling in microglia.
-Molecular BioSystems, 11(12), pp.3332-3346.
-
-    \bibitem{ref:pyabc} Klinger, E., Rickert, D. and Hasenauer, J., 2018. pyABC: distributed, likelihood-free inference. Bioinformatics, 34(20), pp.3591-3593.
-
-    \bibitem{ref:kernel} Filippi, S., Barnes, C.
-P., Cornebise, J., and Stumpf, M..H.
-(2013).
-On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo.
-Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107.
-
-    \bibitem{ref:abcsysbio} Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P., and Stumpf, M.P.H., 2014.
-A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation.
-Nature Protocols.
-9 (2) pp.
-439–456.
-
-    \bibitem{ref:MCMC} Gilks, W.R., Richardson, S.
-and Spiegelhalter, D.J., 1996.
-Markov Chain Monte Carlo in Practice Chapman \& Hall: London.
-Markov Chain Monte Carlo in practice.
-Chapman and Hall, London.
-
-    \bibitem{ABC_rejection} Tavaré, S., Balding, D.J., Griffiths, R.C. and Donnelly, P., 1997. Inferring coalescence times from DNA sequence data. Genetics, 145(2), pp.505-518.
-
-    \bibitem{summaryD} Joyce, P. and Marjoram, P., 2008. Approximately sufficient statistics and Bayesian computation. Statistical applications in genetics and molecular biology, 7(1).
-
-    \bibitem{summaryD2} Nunes, M.A.
-and Balding, D.J., 2010.
-On optimal selection of summary statistics for approximate Bayesian computation.
-Statistical applications in genetics and molecular biology, 9(1).
-
-    \bibitem{threshold} Silk, D., Filippi, S.
-and Stumpf, M.P., 2012.
-Optimizing threshold-schedules for approximate Bayesian computation sequential Monte Carlo samplers: applications to molecular systems.
-arXiv preprint arXiv:1210.3296.
-
-    \bibitem{population} Klinger, E.
-and Hasenauer, J., 2017, September.
-A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo.
-In International Conference on Computational Methods in Systems Biology (pp.
-128-144).
-Springer, Cham.
-
-    \bibitem{model_compare} Toni, T. and Stumpf, M.P., 2010. Simulation-based model selection for dynamical systems in systems and population biology. Bioinformatics, 26(1), pp.104-110.
-
-    \bibitem{cosmology} Weyant, A., Schafer, C.
-and Wood-Vasey, W.M., 2013.
-Likelihood-free cosmological inference with type Ia supernovae: approximate Bayesian computation for a complete treatment of uncertainty.
-The Astrophysical Journal, 764(2), p.116.
-
-    \bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C.
-2017.
-Comparing two sequentialMonte Carlo samplers for exact andapproximate Bayesian inference on biologicalmodels.J.
-R.
-Soc.
-Interface14: 20170340.
-
-    \bibitem{ref:disease} Minter, A. and Retkute, R., 2019. Approximate Bayesian Computation for infectious disease modelling. Epidemics, 29, p.100368.
-
-    \bibitem{ref:adpt_dis} Prangle, D., 2017. Adapting the ABC distance function. Bayesian Analysis, 12(1), pp.289-309.
-
-    \bibitem{ref:adpt_pop} Klinger, E.
-and Hasenauer, J., 2017, September.
-A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo.
-In International Conference on Computational Methods in Systems Biology (pp.
-128-144).
-Springer, Cham.
-
-    \bibitem{Toni} Toni, T., Welch, D., Strelkowa, N., Ipsen, A.
-and Stumpf, M.P., 2009.
-Approximate Bayesian computation scheme for parameter inference and model selection in dynamical systems.
-Journal of the Royal Society Interface, 6(31), pp.187-202.
-
-    \bibitem{sensitivity} Gutenkunst, R.N., Waterfall, J.J., Casey, F.P., Brown, K.S., Myers, C.R.
-and Sethna, J.P., 2007.
-Universally sloppy parameter sensitivities in systems biology models.
-PLoS Comput Biol, 3(10), p.e189.
-
-\end{thebibliography}
+\bibliography{bibtex/ref}
 

### 2020-08-09 background/background.tex
Bib restructure

@@
 In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target.
-Prior $p(\theta)$ should express our initial believes or estimates on the parameter valeus before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers.
+Prior $p(\theta)$ should express our initial believes or estimates on the parameter values before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers.
 Once the data is given, $\theta$ is the only parameter in the likelihood function.

### 2020-08-09 implementation/implementation.tex
Bib restructure

@@
 According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
-\verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}.
+\verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{population}.
 \verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks.
@@
 
-Adaptive population \cite{ref:adpt_pop} can adapt the population size of each generation according to the mean correlation of variation error of the previous population.
+Adaptive population \cite{population} can adapt the population size of each generation according to the mean correlation of variation error of the previous population.
 In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.

### 2020-08-09 background/background.tex
Fix order

@@
 
-The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modeling problems, especially in the area of biology \cite{ref:abcsysbio, ref:compare, ref:disease}.
+The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modeling problems, especially in the area of biology \cite{ref:abcsysbio, ref:disease, ref:compare}.
 But ABC is not limited to these areas, as it can be generally used in inference tasks to give parameter estimates or model rankings.

### 2020-08-09 appendix/appendix.tex
Bib restructure
Appendix fix

@@
 
-\begin{table}[h!]
+\begin{table}[H]
     \centering
@@
         \hline\hline
-        $\lambda_N$                 & 2.1989                        \\
-        $\kappa_{N\beta}$           & 3.9627                        \\
-        $\mu_N$                     & 1.7219                        \\
-        $\nu_{N\Phi}$  0.2195       & $cell^{-1}\cdotp h^{-1}$      \\
-        \hline
-        $\lambda_\Phi$              & 1.3146  $cell/h$              \\
-        $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$         \\
-        $\mu_\Phi$  0.1454          & $h^{-1}$                      \\
-        \hline
-        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
-        $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
-        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
-        \hline
-        $s_{\alpha\Phi}$  10.2416   & $unit/(cell\cdotp h)$         \\
-        $\mu_\alpha$ 19.6642        & $h^{-1}$                      \\
-        [1ex]
+        Operating system   & macOS Catalina 10.15.6 \\
+        PyCharm (IDE)      & Professional 202.2     \\
+        Python interpretor & 3.7.0                  \\
+        IPython            & 7.12.0                 \\
+        Clang              & 6.0 (clang-600.0.57)   \\
         \hline
@@
 
-\subsection{Remote machine}
+Version requirements for some critical site packages (\verb|Python|) used in the code are listed in Table \ref{table:local_package}.
 
-\begin{table}[h!]
+\begin{table}
     \centering
@@
         \hline\hline
-        $\lambda_N$                 & 2.1989                        \\
-        $\kappa_{N\beta}$           & 3.9627                        \\
-        $\mu_N$                     & 1.7219                        \\
-        $\nu_{N\Phi}$  0.2195       & $cell^{-1}\cdotp h^{-1}$      \\
+        pyABC       & $\simeq$ 0.10.3 \\
+        NumPy       & $\simeq$ 1.18.4 \\
+        SciPy       & $\simeq$ 1.4.1  \\
+        pandas      & $\simeq$ 1.1.0  \\
+        matplotlib  & $\simeq$ 3.0.1  \\
+        bokeh       & $=$ 1.4.0       \\
         \hline
-        $\lambda_\Phi$              & 1.3146  $cell/h$              \\
-        $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$         \\
-        $\mu_\Phi$  0.1454          & $h^{-1}$                      \\
+    \end{tabular}
+    \caption{Site packages on local machine}
+    \label{table:local_package}
+\end{table}
+
+Local development and some runs of small size were done under the hardware listed in Table \ref{table:local_hardware}.
+When running ABC SMC, the program can make used of 8 python process (Hyper-Threading enabled) and run under 3.8GHz (maximal) for a long time under proper cooling, proved that the program could efficiently use the computation resources for a local personal computer.
+It is noted that the program took no advantage of graphics card, as \verb|multiprocessing| along can out exploit the resources inside CPU.
+The execution time and other performance data presented in Chapter 5 were obtained using remote machines but not local machine.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c c|}
         \hline
-        $s_{\beta N}$               & 6.5536  $unit/(cell\cdotp h)$ \\
-        $i_{\beta\Phi}$  1.7062     & $cell^{-1}$                   \\
-        $\mu_\beta$                 & 0.5212  $h^{-1}$              \\
+        Hardware        & Detail                                \\ [0.5ex]
+        \hline\hline
+        CPU             & Quad-Core Intel Core i5 8259U 2.3 GHz \\
+        Architecture    & 64 Bit                                \\
+        Hyper-Threading & Supported                             \\
+        Turbo Boost     & Max 3.8 GHz                           \\
+        Memory          & 16 GB                                  \\
+        Graphics card   & Iris Plus Graphics 655                \\
         \hline
-        $s_{\alpha\Phi}$  10.2416   & $unit/(cell\cdotp h)$         \\
-        $\mu_\alpha$ 19.6642        & $h^{-1}$                      \\
-        [1ex]
+    \end{tabular}
+    \caption{Hardwares on local machine}
+    \label{table:local_hardware}
+\end{table}
+
+\subsection{Remote machine}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c c|}
+        \hline
+        Environment                 & Version                       \\ [0.5ex]
+        \hline\hline
+        Placeholder                 & 2\\
         \hline
@@
     \caption{Parameter values used for model 1}
-    \label{table:m1}
+    \label{table:synthetic_data}
 \end{table}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
 

### 2020-08-09 background/background.tex
Appendix fix

@@
 
- Preliminarily \verb|pyABC| \cite{ref:pyabc} in \verb|Python| was chosen for our first implementations.
+ Preliminarily \verb|pyABC| \cite{ref:pyabc} in \verb|Python| was chosen for implementations.
+\verb|pyABC| supports multiple types of parallelisation, e.g., multicore processing and distributed cluster, and supports resume stored runs, which made our experiments easier.
+It is well documented and flexible in customisations of the algorithm.
+
+ Some other ABC SMC related packages were also examined and considered as backups, include ELFI \footnote{\url{https://elfi.readthedocs.io/en/latest/}}, ABC-SysBio \cite{ref:abcsysbio}, EasyABC\footnote{\url{https://cran.r-project.org/web/packages/EasyABC/index.html}}, GpABC \footnote{\url{https://github.com/tanhevg/GpABC.jl}}, etc.
+These packages may have their advantages (e.g.
+ABC-SysBio supports CUDA acceleration) and considered as alternative implementation methods, or can be used to compare performance difference.
+
+ Additionally, some other softwares were also need for solving ODE, results analysis and visualisations. Some software and hardware environments are listed in Appendix A.

### 2020-08-10 background/background.tex
C1 and C2 read

@@
 
- [describe the process and how cells and cytokines are involved]
+%  [describe the process and how cells and cytokines are involved]
 
 Compared to mammals, zebrafish can regenerate spinal cord after injury, and possibly recover swimming function.
-They are widely used in regenerative ability researches.
+They are widely used in regenerative ability researches due to this ability.
 Experiments found that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}.
@@
 The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration.
-Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, affect the inflammation, which are dynamically controlled by immunes cells, i.e.
-cytokines works as media to control the regeneration.
-Here cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.
+Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, are related the inflammation, and they are dynamically controlled by immunes cells, i.e.
+cytokines works as media to control the axonal regeneration.
+Cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.
 
-An interaction map can be drawn, according to the evidences found in \cite{ref:Tsarouchas}.
-After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells: macrophage, neutrophil and keratinocyte.
-The dynamical system consists of complex interactions between cells and cytokines, e.g.
-il-1$\beta$ promotes macrophages' increase.
-and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
-Consequently, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration.
-Generally, il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect, and tnf-$\alpha$ promotes the regeneration all the time.
-The presence if the cytokines also affects the number of cell, while as sources number of cells also affects cytokines' expression.
+An interaction map can be drawn, according to the evidences found by Tsarouchas et al.
+\cite{ref:Tsarouchas}.
+After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells: macrophage, neutrophil.
+This dynamical system consists of complex interactions between cells and cytokines, For example, il-1$\beta$ promotes macrophages' increase, and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
+Consequently as media, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration, where generally il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect; tnf-$\alpha$ promotes the regeneration all the time.
+The presence of the cytokines also affects the number of immune cells, while as sources, more cells lead to more expression of cytokines.
 Interactions between different kinds of cells are also considered in the system.
 
-This is a dynamic system suitable for mathematical modelling, as a set of equations can be written to represent the dynamics as possible.
-Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions iniside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and our own hypothesis which may include different interactions.
+This dynamic system is suitable for mathematical modelling, where a set of equations can be written to represent the dynamics of different objects e.g.
+cells and cytokines.
+Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions iniside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and furthermore based our own hypothesis which may include different interactions.
 
@@
 
- [how this can be modelled as dynamic systems]
+%  [how this can be modelled as dynamic systems]
 
- [general topics of mathematical models: how to build model according to interactions; how to calibrate/ parameterise the model; how to solve the model; dynamics; how to evaluate the model]
+%  [general topics of mathematical models: how to build model according to interactions; how to calibrate/ parameterise the model; how to solve the model; dynamics; how to evaluate the model]
 
-Mathematical models can express real-life questions in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques.
+Mathematical models can express real-life problems in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques.
 Mathematical models can be used for prediction, classification, testing our understanding or hypothesis and many other tasks, thus they are widely used to solve practical problems.
-However, for a model that describe the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models.
-In this case, we wish to use modelling techniques to test our inference method, our hypothesis and hopefully an well-fitted models can be found and compared.
+However, for a model that describes the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models.
+In this case, we wish to use modelling techniques to implement inference methods, test our hypothesis and hopefully find and compare well-fitted models.
 
-According to our understanding of the biological mechanism in the regeneration, hypothesis should firstly proposed, in a way which can represent the real  mechanism as much as possible.
+According to our understanding of the biological mechanism in the regeneration, hypothesis should firstly proposed in a way that can represent the real  mechanism as much as possible.
 Usually a real-life problem, e.g.
-the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered, An elaborate description of the real cases using a mathematical model sometimes can be impossible; assumptions and hypothesis are proposed to make necessary simplification and abstract for a final reasonably detailed model.
+the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered; an elaborate description of the real cases using a overly complex mathematical model sometimes can be impossible; assumptions and hypothesis are often proposed to make necessary simplification and abstract for a final reasonably detailed model.
 Some qualitative models are simple but illustrative for the real-life observed patterns, e.g.
@@
 steady states in ecosystem).
-If a more precise model is desired, usually more effects or interactions are considered and added as terms to a re-constructed model, e.g.
-environmental factors, a changing environmental capacity and predation from other species.
+If a more precise model is desired, usually more effects or interactions are considered and added as terms of equations to a re-constructed model, e.g.
+considering environmental factors, a changing environmental capacity and predation from other species in L-V model.
 
-Models like L-V equation are mostly analytically solvable when analysing properties of the model, however, analytical solution for many models are not feasible, because of more complex dynamics, or the model itself is not a deterministic model.
-In the areas that are not analytically feasible, usually computational techniques are applied for approximate solutions.
-Solvers for different models are used to simulate the system or find roots for some spacial points.
+Models like L-V equation are mostly analytically solvable when analysing properties of the model, however, analytical solution for many other general models are not feasible, because of more complex dynamics, or the model itself is not a deterministic model.
+In the areas where solutions are not analytically feasible, computational techniques are usually applied for approximate solutions.
+Solvers for different models had been proposed to simulate the system or find roots for some spacial points, such as Explicit Runge-Kutta \cite{dormand1980family} and LSODA.
 
 For the regeneration process that we wanted to model, differential equations are our preference.
-Non-linear ordinary differential equations are used in this dissertation, where the only independent variable is time $t$ denoting the post-lesion time (hours), but further models can also consider other models e.g.
+Non-linear ordinary differential equations are used in this dissertation, where the only independent variable is time $t$, denoting the post-lesion time (hours).
+Further models can also consider other models in other forms, e.g.
 stochastic models and partial differential equations.
@@
 Often further researches on the modelled problems require the model to be deterministic on its parameters.
-Parameters in the dynamic system models usually have their practical significance, e.g.
+Parameters in a dynamic system models usually have their practical significance, e.g.
 describing reaction speed, environmental capacity or the birth rate.
 There are many approaches to derive these parameters from data or experience.
-For a specific problem and a widely used models, the parameters' values can be read from previous literatures; however in many others cases, we need to find parameter's values specifically for our model and problem, where numerical values for parameters are more suitable for practical analysis using ODE solvers or other tasks.
+For a popular topic and a widely used models, the parameters' values can be read from previous literatures; however in many others cases, we need to find parameter's values specifically for our model, and numerical values for parameters are mostly preferred, as they suitable for practical analysis and comparison.
 
@@
 Bayesian inference approaches gain popularity in recent years, which can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
-Local optima can still be a major concern, and many efforts have be proposed for a more through explore of the parameter space.
+Local optima can still be a concern, and many efforts have be proposed for a more through explore of the parameter space to avoid this.
 
@@
 
- [how to parameterise the model]
+%  [how to parameterise the model]
 
- [how to infer the parameter given data]
+%  [how to infer the parameter given data]
 
- [likelihood-free inference and information about ABC SMC]
+%  [likelihood-free inference and information about ABC SMC]
 
- [model comparison, ABC SMC for model comparison]
+%  [model comparison, ABC SMC for model comparison]
 
- Bayesian inference approaches put Bayes Rule in the centre:
+ Bayesian inference approaches for parameter estimation put Bayes Rule in the centre:
 
@@
     \label{eq:bayes}
-    p(\theta|D) = \frac{p(D|\theta)p(\theta)}{P(D)}
+    p(\theta|D) = \frac{p(D|\theta)p(\theta)}{p(D)}
 \end{align}
@@
 where $l$ represents the likelihood function.
-In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target.
-Prior $p(\theta)$ should express our initial believes or estimates on the parameter values before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers.
+In other words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target in parameter estimation task.
+Prior $p(\theta)$ should express our initial believes or estimates on the parameter values before observing the data and conducting experiments; likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability values.
 Once the data is given, $\theta$ is the only parameter in the likelihood function.
-Some optimisation-based methods try to find an estimation of parameter set $\theta$ s.t.
-maximise the likelihood, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative.
+Some optimisation-based methods try to find estimates of parameter set $\theta$ s.t.
+maximise the likelihood, however, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative of it.
 
@@
 Markov Chain Monte Carlo (MCMC) \cite{ref:MCMC}.
-As another alterative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood, but focused on simulating the data from the given model and observed data, then used the simulated data to approximate the desired posterior.
+As an alterative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood function, but focused on simulating the data from the given model and observed data, then used the simulated data to approximate the desired posterior.
 ABC-rejection \cite{ABC_rejection} is a simple example of the ABC process, and further algorithms e.g.
-sequential Monte-Carlo (ABC SMC)\cite{Toni} are proposed to improve the efficiency.
+sequential Monte-Carlo (ABC SMC)\cite{Toni} were proposed to improve the efficiency.
 
@@
 
-In our study of the regeneration model, full trajectories of the dependent variables are used as the data to be compared, as our goal is to find a model and a parameter set that can intuitively recover the real process as much as possible.
+In our study of the regeneration model, full trajectories of the dependent variables are used as the data to be compared, as our goal is to find a model with a parameter set that can intuitively recover the real process as much as possible.
 The mean values at each time point for each dependent variable is calculated to form $D$, and for each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$.
-Using other statistics to summary the raw data are also possible bur care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
+Using other statistics to summary the raw data are also possible but care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
 
-Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of deceasing threshold values $\epsilon_t$, and derive a new population (called generation) of accepted sampled from the previous accepted samples (called particles).
+Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of deceasing threshold values $\epsilon_t$, and derive new populations (called generation) of samples from the accepted samples (called particles) in previous population.
 An example is ABC SMC \cite{Toni}, which can significantly reduce the required samples compared to ABC-rejection \cite{ref:disease}:
@@
 In implementations, many factors can affect the efficiency and goodness of fit, some of them are discussed in the Chapter 4.
-Rather than a pre-fixed threshold schedule, median or quantile based epsilon schedule can be save total sampling size and thus improve acceptance rates of particles \cite{threshold}.
+Rather than a pre-fixed threshold schedule, median or quantile based epsilon schedule can save the total sampling size and thus improve acceptance rates of particles \cite{threshold}.
 Some other improvements include using adaptive population size for each generation \cite{population}, using adaptive distance function \cite{ref:adpt_dis} and switching kernels \cite{ref:kernel}, etc.
-In our implementations these options were tried when applicable and suitable to compare the behaviours of the algorithm.
+In our implementations these options were tried and compared when applicable and suitable.
 
@@
 Instead of draw samples from prior of a single model, we can draw samples from multiple models $m_i$ and approximate $p(m_i, \theta|D)$, which is the joint probability distribution of model $m_i$ and its accepted particles $\theta$.
-In each generation the marginal distribution $p(m_i|D)$ can be calculated as the accepted ${\theta}$ have been obtained, thus models can be ranked according to their probabilities.
+In each generation the marginal distribution $p(m_i|D)$ can be calculated as the accepted ${\theta}$ is known, thus models can be ranked according to their probabilities.
 
@@
 
- [existing tools and software for this task]
+%  [existing tools and software for this task]
 
- [workflow and developing process]
+%  [workflow and developing process]
 
- As we decided to apply ABC SMC on our regeneration models for its high-dimensional parameter estimation and possibly comparison of different models, a code development and experiments workflow was expected to be built.
+ As we decided to apply ABC SMC on our regeneration models for its high-dimensional parameter estimation and comparison of different models, a code development and experiments workflow was expected to be built.
 The ABC SMC algorithm needs a whole set of mathematical environments (e.g.
 class of distributions, models and particles) and closely linked sampling and fitting algorithms (e.g.
-KDE fitting and Gaussian sampling), thus suitable platforms with related packages are preferred.
+KDE fitting and Gaussian sampling), thus suitable platforms with related software packages are preferred.
 Regarding this, several packages in \verb|R| and \verb|Python| were studied.
 
- A software packages that integrally implement ABC SMC is preferred, and more options in the settings of the algorithm, more customisable features can make it better.
-Moreover, as the ABC SMC itself is a computationally intensive task that consists of massive parallel potentials, an ideal software is expected to support as least one kind of parallel options e.g.
+ A software packages that integrally implement ABC SMC is preferred, and more options in the settings of the algorithm, more customisable features make it better.
+Moreover, as the ABC SMC itself is a computationally intensive task that contains massive parallel potentials, an ideal software package is expected to support as least one kind of parallel options e.g.
 multi-core, GPU accelerating and clusters.
-Otherwise, a modification or rewrite of the software package are expected to produce a reliable input-output framework for our parameter estimation and model selection experiments.
+Otherwise, a modification or rewrite of the software package are expected to produce a reliable input-output framework for our parameter estimation and model selection experiments, where much more work would be introduced.
 
- Preliminarily \verb|pyABC| \cite{ref:pyabc} in \verb|Python| was chosen for implementations.
+ Preliminarily \verb|pyABC| packages \cite{ref:pyabc} in \verb|Python| was chosen for implementations.
 \verb|pyABC| supports multiple types of parallelisation, e.g., multicore processing and distributed cluster, and supports resume stored runs, which made our experiments easier.
@@
 
- Additionally, some other softwares were also need for solving ODE, results analysis and visualisations. Some software and hardware environments are listed in Appendix A.
+ Additionally, some other softwares were also need for solving ODE, results analysis and visualisations. These software and hardware environments are listed in Appendix A.

### 2020-08-10 introduction/introduction.tex
C1 and C2 read

@@
 
- [background and motivation]
+%  [background and motivation]
 
- [what is done]
+%  [what is done]
 
-While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions of cells and chemicals in a biological system.
-The objective process in this dissertation is an example of complex biology system found in tissue regeneration.
-Zebrafish has the ability to regenerate its spinal cord after injury; at the injury-site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.
+While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions and effects of cells and chemicals in a biological system.
+The objective in this dissertation is an example of complex biology system found in tissue regeneration: zebrafish spinal cord regeneration.
+Zebrafish has the ability to regenerate its spinal cord after injury; at the lesion site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.
 
-Our interest lies in the modeling of the regeneration process that happens in the injury-site, especially in the parameter estimation and comparison of different models.
-According to available researches \cite{ref:Tsarouchas}, a mathematical model is proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments.
-However, we would like to explore the the `best fit' if the model which comes with a set of parameter estimates given the observed data.
-Also, we wish to compare different it with different models and possibly find the best model to describe the data without losing biological and mathematical properties in the regeneration process.
+Our interest lies in the modeling of the regeneration process that happens around the lesion site, especially in the model parameter and comparison of different models.
+According to available researches \cite{ref:Tsarouchas}, a mathematical model was firstly proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments.
+However, we would like to explore the the `best fit' of the model which comes with a set of parameter estimates, given the observed data.
+Also, we wish to compare different it with other alternative models and possibly find the best model to describe the observed data.
+In these steps, models are expected to retain mathematical and biological significance, where differential equations are suitable to describe this system.
 
-Fo the parameter estimation tasks, Bayesian inference techniques can help in the problems of over-fitting and local optima \cite{ref:abcsysbio}, compared to optimisation based methods.
-Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression can be hard.
+Fo the parameter estimation tasks, Bayesian inference techniques can help in the problems of over-fitting and local optima \cite{ref:abcsysbio}, compared to optimisation-based methods.
+Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression for the dynamic system can be hard.
 
-Such methods introduce computation-intensive sampling-rejection pattern, where massive number of samples are drawn, followed by computations.
-As the sampling process is independent from each other in a certain period, the parallel execution becomes feasible and is expected to accelerated the parameter inference in a large scale.
+Such methods introduce computation-intensive sampling-and-test patterns, where massive number of samples are drawn and followed by complex computations.
+As the sampling is independent from each other in a certain period, the parallel implementation of the algorithm becomes feasible and is expected to accelerate the parameter inference in a large scale when using HPC resources.
 
 This dissertation includes the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results.
-Model comparison results are also included using the same method.
-Additionally, a performance test illustrates the strong-scaling performance of the algorithm.
+ABC SMC was also applied in model comparison.
+Additionally, performance experiments illustrated the strong-scaling performance of the algorithm.

### 2020-08-10 background/background.tex
Some fix

@@
 Often further researches on the modelled problems require the model to be deterministic on its parameters.
-Parameters in a dynamic system models usually have their practical significance, e.g.
+Parameters in dynamic system models usually have their practical significance, e.g.
 describing reaction speed, environmental capacity or the birth rate.
@@
 To find the best parameters' values given observed data, either optimisation-based approaches (e.g.
-least-square (LS) fit) or inference approaches can be used.
+least-square fit (LS), maximal likelihood estimation (MLE)) or inference-based approaches can be used.
 Optimisation-based approaches may face two major drawbacks: over-fitting and local optima \cite{ref:abcsysbio}.
-Bayesian inference approaches gain popularity in recent years, which can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
-Local optima can still be a concern, and many efforts have be proposed for a more through explore of the parameter space to avoid this.
+Bayesian inference approaches gain popularity in recent years, as it can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
+Local optima can still be a concern, and many efforts have be proposed for a more through explore of the parameter space to avoid local modes.
 
@@
  Bayesian inference approaches for parameter estimation put Bayes Rule in the centre:
-
  \begin{align}
@@
 \end{align}
-
 where $p$ is the probability density function, $\theta$ denotes the parameter set, $D$ is the given observed data.
@@
 Alternatively, as $p(D)$ is fixed once the observed data is given, the rule can be rewritten as 
-
 \begin{align}
@@
 \end{align}
-
 where $l$ represents the likelihood function.
@@
 Some optimisation-based methods try to find estimates of parameter set $\theta$ s.t.
-maximise the likelihood, however, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative of it.
+maximise the likelihood, however, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone calculating and comparing values of it.
 
@@
 For a pre-fixed threshold value $\epsilon$, if $d(D^*, D)<\epsilon$ then the sample $\theta^*$ is accepted
-    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution can be approximated by these samples
+    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution then can be approximated by these samples
 \end{enumerate}
@@
     \item Set a threshold schedule $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ where $T$ is the total number of populations and $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ is a descending sequence; set the number of particles in each population as $N$, set the population index $t=1$
-    \item if $t=1$, draw samples from prior distribution until $N$ particles are accepted under threshold $\epsilon_1$, i.e. each of $N$ particles s.t. $d(D,D^*)<\epsilon_1$
+    \item if $t=1$, draw samples from prior distribution until $N$ particles are accepted under threshold $\epsilon_1$, i.e. each of $N$ particles generates data $D^*$ s.t. $d(D,D^*)<\epsilon_1$
 
-    Else ($t>1$), draw samples from previous population $\{\theta^*_{t-1}\}$ with weights $\{w_{t-1}\}$, perturb the particles using a perturbation kernel $K(\theta^*|\theta)$ and obtain $\theta^{**}\sim K(\theta^*|\theta)$
+    If $t>1$, draw samples from previous population $\{\theta^*_{t-1}\}$ with weights $\{w_{t-1}\}$, perturb the particles using a perturbation kernel $K(\theta^*|\theta)$ and obtain $\theta^{**}\sim K(\theta^*|\theta)$
     \item Repeat step 2 until $N$ particles are accept under threshold $\epsilon_t$
-    \item Calculate the weights of the accepted particles. For particle $i$ 
+    \item Calculate the weights of the accepted particles. For particle $i$, the weight is 
     \begin{align}

### 2020-08-10 modelling/modelling.tex
Some fix

@@
 
-Mathematical models can describe different kinds of dynamic systems, and thus can be used as a guide for prediction and analysis.
-An ideal model in our case, can represent reasonable interactions/effects between cells and cytokines and recover the observed  data features in different time point.
+Mathematical models can describe different kinds of dynamic systems, and thus can be used in prediction, analysis abd other tasks.
+An ideal model in our case, can represent reasonable interactions or effects between immune cells and cytokines and reproduce features and characteristics that were observed in experimental data.
 
-Proposed models for the regeneration process are in the form of ordinary differential equations (ODE), specifically the time differential form.
-Terms in the ODE are mostly explainable and corresponds interaction paths.
+Proposed models for the regeneration process are in the form of non-linear ordinary differential equations (ODE).
+The only independent variable is post-lesion time $t$, and the equations consists of derivatives of dependent variables (number of cells, mRNA expression of cytokines) with respect to $t$.
+The equations were mostly an interpretation of the proposed interaction maps, with terms correspond to interaction paths.
 
@@
 
-Our models is built on the basis of the existing experiments data form Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement data include the number of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative concentration of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
-As proposed in \cite{ref:Tsarouchas}, neutrophil and macrophage play important roles in the promotion of spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the mediation.
-According to this, our current models focus on the changes of four variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$).
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative concentration' (compared to the value at first time point).
+Our models were built on the basis of the existing experimental findings from Tsarouchas et al.\cite{ref:Tsarouchas}.
+The measurement data included the numbers of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative mRNA expression of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
+As proposed by Tsarouchas et al., neutrophil and macrophage play important roles in the spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the media.
+According to this, our initial models focused on the changes of four dependent variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$) that are most important in the regeneration.
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative mRNA expression' (compared to the value at the initla time point i.e.
+0 hpl).
 

### 2020-08-11 conclusion/conclusion.tex
New para

@@
 
+ To conclude, ABC SMC was successfully applied on the models of zebrafish spinal cord regeneration.
+These models were all ODE equations that described the interactions and effects between cells and proteins in the lesion site.
+These equations contains four dependent variable and more than 10 model parameters.
+Parameter estimation for high-dimensional parameter space is a challenging work for many dynamic system models, as insufficient explore of parameter space can easily lead to local optimal.
+Using a large population size in ABC SMC in our models could partially relief the concerns in local modes, and give considerable parameter estimates (FIGURE).
+Also ABC SMC successfully helped in the model comparison and thus can be used for suggestions of possibly best model, although there were uncertainties remained in the process and a large number of population size was need.
+
+ ABC SMC implementation is a complex process with respect to both algorithm settings and code development.
+The ability of parameter inference and robustness of algorithm were firstly studied using synthetic data where the true parameter value was known, and experiments were performed to find proper hyperparameter and implementation settings of ABC SMC.
+Later inference on the observed data followed the suggestions from the synthetic data experiments and acceptable results were produced.
+
+ The inference program was mostly run on multi-core machines, and additional performance
+
  [how good is the result]

### 2020-08-11 conclusion/conclusion.tex
New sevtion

@@
-\chapter{Conclusions}
+\chapter{Conclusion}
 
@@
  [to what extend]
+
+ [how good is the result]
 
@@
 
- The inference program was mostly run on multi-core machines, and additional performance
-
- [how good is the result]
+ The inference program was mostly run on multi-core machines, and additional scaling-up performance test illustrated the scaling behaviour.
+When the number of running \verb|Python| process exceeds the number of physical cores, hyperthreading would be enabled but gains not significant improvements with respect to execution time, while increase of average sampling size per second was still observed.
 
@@
 
+ From this application, some suggestions on ABC SMC implementation could be drawn.
+
  [thanks]

### 2020-08-11 conclusion/conclusion.tex
New para

@@
-\chapter{Conclusion}
+\chapter{Conclusion and future works}
 
- [what is studied]
+%  [what is studied]
 
- [to what extend]
+%  [to what extend]
 
- [how good is the result]
+%  [how good is the result]
 
@@
 
- [summary of the findings and advise]
+%  [summary of the findings and advise]
 
  From this application, some suggestions on ABC SMC implementation could be drawn.
+How to set the prior could be the most tricky decision for a high dimensional parameter inference.
+The distribution and the interval of each parameter can affect the inference results significantly and improper prior cold lead to poor fit.
+The epsilon trajectory and joint distribution of parameters can be used to identity possible local modes; if stuck, it takes much more samples than usual to jump out out local optimal.
+To avoid this a large population size and a proper epsilon schedule may help.for the uncertainties observed in model selection and some duplicated runs with high variance, a longer run with more generation could help to identify whether the convergency is consistent such that a reproducible inferred results can be obtained.
 
- [thanks]
+%  [thanks]
+
+In terms of the future works, some more comparison with other algorisms could be of our interest.
+Least square fitting using MCMC \cite{ref:MCMC} and some other exact inference approaches could be performed on the same model and observed data, as the standard deviation at each measurement point is known.
+By comparing to these methods, differences in the inference results, algorithm robustness and required computational resources could be revealed and possibly help us in choosing an inference algorism that can balance the trade-off between efficiency and a reasonable good `fit'.
+Also there other implementation options of ABC SMC worth trying, e.g.
+CUDA acceleration and \verb|Julia| implementations\footnote[1]{\url{https://github.com/tanhevg/GpABC.jl}} \footnote[2]{\url{https://github.com/marcjwilliams1/ApproxBayes.jl}}.
+
+In the aspect of models, some further models can be studied.
+If there were some more experimental evidence on the proposed interaction map, we could proposed more precise models, or calibrate some terms in our existing models.
+Also, models of other forms, e.g.
+PDE and stochastic models, could also be studied if a more complex model is needed for the dynamic system.
+Simplification is also possible of existing models

### 2020-08-12 modelling/modelling.tex
New fix

@@
 Our models were built on the basis of the existing experimental findings from Tsarouchas et al.\cite{ref:Tsarouchas}.
-The measurement data included the numbers of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative mRNA expression of four cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
+The available measurement data included the numbers of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative mRNA expression of four kinds of cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
 As proposed by Tsarouchas et al., neutrophil and macrophage play important roles in the spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the media.
-According to this, our initial models focused on the changes of four dependent variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$) that are most important in the regeneration.
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is of the unit `relative mRNA expression' (compared to the value at the initla time point i.e.
-0 hpl).
+According to this, our initial models focused on the changes of four dependent variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$) that are most important and determinative in the regeneration.
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ was recorded as `relative mRNA expression' (compared to the value at the initla time point i.e.
+0 hpl) and considered unitless.
 
-It is noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations is mean of measurement data at each time point, assuming that measured data is Gaussian-like distributed.
-To validate this, the distribution of the measured data points is plotted and examined to see if the mean value can represent the distribution.
-The result is that at most time points the measurement values are Gaussian-distributed, although some distributions are skewed.
-One abnormal distribution is observed at time point 120 h post-lesion (hpl) for macrophage where there are two concentrations.
-Mean value can summarise most data and thus is still used as the target observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}.
+It was noted that the variance of the measured data is relatively high.
+The summary statistic used for the parameter estimations was mean of measurement data at each time point, assuming that measured data was Gaussian-like distributed.
+To validate this, the distribution of the measured data points was plotted and examined to see if the mean value can represent the distribution.
+The result was that at most time points the measurement values are Gaussian-distributed, although some distributions were skewed.
+One abnormal distribution was observed at time point 120 hpl for macrophage where there are two concentrations.
+Despite this, mean value could summarise most data and thus was still used as the target observed data.
+A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}, which is our target trajectories for the models to fit.
 
@@
 
-\section{Hypothesis and Models}
+\section{Hypothesis and models}
 
@@
 
-A preliminary model is proposed according to \cite{ref:Tsarouchas} and used to build and test the inference framework code.
+A preliminary model (Eqn.
+\ref{eq:model1}) was proposed according to \cite{ref:Tsarouchas} and then mainly used to build and test the code of inference framework.
 An interaction map of the model is shown in Figure \ref{fig:m1}.
@@
 To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
-number of cells as `cell', and denote the unit of $\beta$ and $\alpha$ i.e.
-relative mRNA expression as `unit' for simplicity.
+number of cells as `cell', and denote $\beta$ and $\alpha$'s relative mRNA expression as unitless for simplicity.
+This model included 12 parameters, all of which should be positive real number.
+A description of these parameters can be found in Table \ref{table:m1}.
 
@@
 
-It is assumed that there are negative feedbacks for all the variables. MORE DESCRIBE.
+It was assumed that there were negative feedbacks the cells and cytokines, i.e. overcrowding inhibits the increasing rates.
+
+MORE DESCRIBE.
 

### 2020-08-12 introduction/introduction.tex
Fix

@@
 
-This dissertation includes the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results.
-ABC SMC was also applied in model comparison.
+This dissertation mainly focuses on the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results.
 Additionally, performance experiments illustrated the strong-scaling performance of the algorithm.
+Chapter 2 talks about the background information and theories, Chapter 3 shows how the mathematical models are derived from hypothesis and Chapter 4 describes the ABC SMC implementations and experiments on the models in detail.
+A brief scaling-up performance experiment is included in Chapter 5, and Chapter 6 summarises the conclusions and states the future works on this topic.

### 2020-08-12 modelling/modelling.tex
Fix

@@
 
-It was assumed that there were negative feedbacks the cells and cytokines, i.e. overcrowding inhibits the increasing rates.
-
-MORE DESCRIBE.
-
 \begin{align}
@@
 
-\begin{table}[h!]
+It was assumed that there are negative feedbacks the cells and cytokines, i.e.
+overcrowding inhibits the increasing rates.
+As sources, macrophage produces tnf-$\alpha$ and neutrophil produces il-1$\beta$, however, macrophage was assumed to have negative effect on the il-1$\beta$, term $1+i_{\beta\Phi}\Phi$ in Eqn.
+\ref{eq:model1} was used to represent this.
+Despite of self-driven increase of immune cells, il-1$\beta$ was assumed to promote the increase of both cells.
+Neutrophil was also assumed to be inhibited by the total immune cells presented in the lesion site, term $\nu_{N\Phi}N\Phi$ was used to denote this.
+
+\begin{table}[t]
     \centering
@@
         $\lambda_N$          & Self-increase rate of neutrophil                 & $cell/h$                 \\
-        $\kappa_{N\beta}$    & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
+        $\kappa_{N\beta}$    & Promoting effect coefficient by il-1$\beta$      & $cell/h$                 \\
         $\mu_N$              & Coefficient of negative feedback of $N$          & $h^{-1}$                 \\
@@
         $\lambda_\Phi$       & Self-increase rate of macrophage                 & $cell/h$                 \\
-        $\kappa_{\Phi\beta}$ & Promoting effect coefficient by il-1$\beta$      & $cell/(unit\cdotp h)$    \\
+        $\kappa_{\Phi\beta}$ & Promoting effect coefficient by il-1$\beta$      & $cell/h$                 \\
         $\mu_\Phi$           & Coefficient of negative feedback of $\Phi$       & $h^{-1}$                 \\
         \hline
-        $s_{\beta N}$        & Production rate from $N$                         & $unit/(cell\cdotp h)$    \\
+        $s_{\beta N}$        & Production rate from $N$                         & $cell^{-1}\cdotp h^{-1}$ \\
         $i_{\beta\Phi}$      & Coefficient of inhibition to the production      & $cell^{-1}$              \\
@@
         \hline
-        $s_{\alpha\Phi}$     & Production rate from $\Phi$                      & $unit/(cell\cdotp h)$    \\
+        $s_{\alpha\Phi}$     & Production rate from $\Phi$                      & $cell^{-1}\cdotp h^{-1}$ \\
         $\mu_\alpha$         & Coefficient of negative feedback of $\alpha$     & $h^{-1}$                 \\
@@
 
-As the observed data indicate, this dynamic system has a steady state where the  inflammation is resolved and immune cells should not be present at the injury site.
-Regarding this, the self-increase parameter $\lambda$ cannot be constant, thus a exponentially decaying $\lambda$ term is introduced and model 2 is proposed.
-(Eqn.
+As the observed data indicated, this dynamic system has a steady state where after a long time ($\geq 120$ hpl), the  inflammation is resolved and immune cells should not be present at the injury site any more.
+Regarding this, the parameter for self-increase rate for the macrophage and neutrophil i.e.
+$\lambda_\Phi$ and $\lambda_N$  cannot be constant.
+An exponentially decaying $\lambda$ term is introduced in the new model (model 2, Eqn.
 \ref{eq:model2}).
@@
 
-Model 2 and model 3 introduce one extra parameter $a$, which is a coefficient in the exponentially decaying $\lambda$, determining the decay speed, with the unit $h^{-1}$.
+Model 2 and model 3 introduced one extra parameter $a$ and removed $\lambda_\Phi$, which is a coefficient in the exponentially decaying $\lambda_N$, determining the decay speed, with the unit $h^{-1}$.
 
@@
 Model 3 can be regarded as a simplification of model 2, as it can be treated as model 2 with parameter $i_{\beta\Phi}=0$.
-Among the proposed three models, model 1 is a naive one that was proposed at very first time and used as an ODE `template' to build and test parameter inference framework.
-As the implementation was successful, model 2 and 3 was proposed as we were trying to calibrate some terms and find a better model.
-After fitting, model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points (which relate to the steady state).
-Model 3 makes a small simplification on model 2 and is theoretically less `general' than model 2.
+Among the proposed three models, model 1 is a naive one that was proposed at very first time and used as an ODE `template' to build and test our parameter inference framework.
+As the implementation was successful, model 2 and 3 was proposed, as we were trying to calibrate some terms and find better models.
+After fitting, model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points (which were related to the steady states of immune cells).
+Model 3 made a small simplification based on model 2 and thus it was theoretically less `general' than model 2.
 
@@
 
-After the first model selection experiment, it was found that some significant features presented in the observed data were not recovered by any of the existing model.
-To resolve this, attempts were tried to introduce more interactions within the dynamic system considering the biological and  mathematical context.
-Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term (which means the same effect as directly promoting but the underlying mechanism is unclean) or adding a term that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
+After the first model selection experiment among the exiting 3 models, it was found that some significant features presented in the observed data were not recovered by any of the model.
+To resolve this, attempts were tried to introduce additional reasonable interactions within the dynamic system, considering the biological and  mathematical context.
+Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term $d_{\beta\alpha}\beta$ (which means the same effect as directly promoting but the underlying mechanism is unclean), or adding a term $f_{\beta\alpha}$ that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
 \ref{eq:model4}) and model 5 (Eqn.

### 2020-08-13 background/background.tex
Fix
New figure

@@
 Experiments found that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}.
-Tsarouchas et al.
-\cite{ref:Tsarouchas} studied the the complex interactions happen at the lesion site.
-The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration.
+Tsarouchas et al.\cite{ref:Tsarouchas} studied the the complex interactions happen at the lesion site.
+The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration\cite{ref:Tsarouchas}.
 Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, are related the inflammation, and they are dynamically controlled by immunes cells, i.e.
-cytokines works as media to control the axonal regeneration.
+immune cells produce and control the cytokines-mediated inflammation \cite{ref:Tsarouchas}.
 Cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.
 
-An interaction map can be drawn, according to the evidences found by Tsarouchas et al.
-\cite{ref:Tsarouchas}.
-After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells: macrophage, neutrophil.
-This dynamical system consists of complex interactions between cells and cytokines, For example, il-1$\beta$ promotes macrophages' increase, and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
-Consequently as media, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration, where generally il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect; tnf-$\alpha$ promotes the regeneration all the time.
-The presence of the cytokines also affects the number of immune cells, while as sources, more cells lead to more expression of cytokines.
-Interactions between different kinds of cells are also considered in the system.
+An interaction map (Figure \ref{fig:map}) can be drawn according to the evidences found in experiments.\cite{ref:Tsarouchas}.
+After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells.
+This dynamical system consists of complex interactions and dynamical feedback between cells and cytokines, For example, il-1$\beta$ promotes macrophages' increase, and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
+
+Consequently, tnf-$\alpha$ and il-1$\beta$ plays determinative roles in the axonal regeneration.
+il-1$\beta$ promotes the regeneration at early stages, but later has an inhibition effect; tnf-$\alpha$ promotes the regeneration all the time.
+The presence of the cytokines also affects the number of immune cells, while as sources, more immune cells produce more cytokines.
+Interactions between different types of cells are also considered in the system (e.g., the number of neutrophil is affected by macrophages).
+
+\begin{figure}[t]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/network.png}}
+    \end{center}
+
+    \caption[Working model of the influence of the innate immune system on axonal regrowth]%
+    {Working model of the influence of the innate immune system on axonal regrowth, from Tsarouchas et al.\cite{ref:Tsarouchas}.
+Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
+    \label{fig:map}
+
+\end{figure}
 
 This dynamic system is suitable for mathematical modelling, where a set of equations can be written to represent the dynamics of different objects e.g.
-cells and cytokines.
-Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions iniside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and furthermore based our own hypothesis which may include different interactions.
+each types of cells and cytokines.
+Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions inside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and furthermore based our own hypothesis which may include different interactions.
 
@@
 Some other interactions proposed in our modelling stages are enlightened from Anderson et al.
-\cite{Anderson}.
+\cite{Anderson}, where intracellular cytokine signaling networks were established and computationally modelled.
 

### 2020-08-13 introduction/introduction.tex
New figure

@@
 
-While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions and effects of cells and chemicals in a biological system.
-The objective in this dissertation is an example of complex biology system found in tissue regeneration: zebrafish spinal cord regeneration.
-Zebrafish has the ability to regenerate its spinal cord after injury; at the lesion site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.
+% While we are interested in the tissue regeneration among different species, mathematical models and techniques can help us to resolve complex interactions and effects of cells and chemicals in a biological system.
 
-Our interest lies in the modeling of the regeneration process that happens around the lesion site, especially in the model parameter and comparison of different models.
+% The objective in this dissertation is an example of complex biology system found in tissue regeneration: zebrafish spinal cord regeneration.
+
+Zebrafish has the ability to regenerate its spinal cord after injury; tissue regeneration observed at the injury site is a complex biological process with interacting parts and dynamical feedback between different types of cells and molecules.
+These cells and molecules are involved in the inflammation and repair in a dynamical way and a complex interacting network can be drawn according to it.
+
+% at the lesion site various immune cells are presented and complex interactions are found while they are regulating inflammation and repair.
+
+% [think about what the appropriate context is to start the introduction with.
+You could for example talk about regeneration of tissues being a complex biological process with many interacting parts and dynamical feedback between different cell types and signalling molecules.
+Mathematical models can help to resolve this complexity.
+An important problem is how to parameterise such models from available data.
+
+% or you could talk about mathematical models being useful in biology in general, then transition to the example of tissue regeneration.]
+
+Mathematical models and techniques can help us to resolve the complexity in this dynamic system, however, how to derive the parameter estimates of the models remains a challenging work.
+In this dissertation, our interest lies in the modeling of the regeneration process that happens around the lesion site, especially in the parameter estimates and comparison of possible models.
 According to available researches \cite{ref:Tsarouchas}, a mathematical model was firstly proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments.
-However, we would like to explore the the `best fit' of the model which comes with a set of parameter estimates, given the observed data.
-Also, we wish to compare different it with other alternative models and possibly find the best model to describe the observed data.
-In these steps, models are expected to retain mathematical and biological significance, where differential equations are suitable to describe this system.
+We would like to find the best parameter sets of the model using the published data, and ranking models under certain metrics and possibly find the best-fit model.
+In these steps, models are expected to retain mathematical and biological significance regarding the complex interactions and differential equations are suitable for this.
 
-Fo the parameter estimation tasks, Bayesian inference techniques can help in the problems of over-fitting and local optima \cite{ref:abcsysbio}, compared to optimisation-based methods.
+% Also, we wish to compare different it with other alternative models and possibly find the best model to describe the observed data.
+
+Fo the parameter estimation tasks, optimisation-based methods usually have trouble in determining global optima from local optimal, and noise in the observed data may lead to over-fitting.
+Compared to optimisation-based methods, Bayesian inference techniques can help to relief these two concerns\cite{ref:abcsysbio}.
 Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression for the dynamic system can be hard.

### 2020-08-13 modelling/modelling.tex
New figure

@@
 
-\subsection{Model evaluation and comparison}
+% \subsection{Model evaluation and comparison}
 
-[general topics on evaluating a model]
+% [general topics on evaluating a model]
 
-[bayes factor for model selection]
+% There exists several metrics to evaluate the model based on the observed data.
+A widely used method is to simulate the data from the model, then calculate the discrepancy between simulated data and observed data.
+P-norm distance functions are usually used in this method
+
+% [bayes factor for model selection]
 
@@
 
-[hypothesis]
+% [hypothesis]
 
-[available data]
+% [available data]
 
-[model misspecification]
+% [model misspecification]
+
+Some limitations were included in Section 3.1, where the observed data have a relative high standard deviation at certain measurement points; more measurement repeats or more measurement points could help to represent more accurate trajectories of the modelled variables and thus more possible models can be proposed.
+On the other hand, more accurate observed are alway beneficial to mathematical modelling, especially to complex dynamic systems.
+
+As stated, our models are largely based on the experimental evidence from Tsarouchas et al.\cite{ref:Tsarouchas}.
+Some hypothesised interactions may differ from real case, and there may exist undiscovered mechanisms.
+These could lead to model misspecification, where the model could fit the data but with wrong interaction paths, or the model can hardly fit the data and recover some local features observed in the data.
 

### 2020-08-14 implementation/implementation.tex
New figure
New fig and table

@@
 
-    [what is factor]
+% [what is factor]
 
-    [what is adaptive distance]
+% [what is adaptive distance]
 
@@
 
-[FIGURE here]
+% [FIGURE here]
+
+\begin{figure}[ht]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/ib.png}}
+    \end{center}
+
+    \caption[Factors and adaptive distance experiments]%
+    {Factors and adaptive distance experiments. Left: Total required samples. Right: acceptance rates in each generation}
+    \label{fig:factor}
+
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/ib_factor.png}}
+    \end{center}
+
+    \caption[CAPTION]%
+    {DESCRIBE}
+    \label{fig:factor_sim}
+
+\end{figure}
 
@@
 Applying factors and using adaptive distance can lead the resultant model less accurate (after the same generation iterations): they have wider inter-quantile range and may requires more generations to converge.
+TODO
 
@@
 
+\begin{figure}[ht]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model5_para.png}}
+    \end{center}
+
+    \caption[Estimated posterior distribution of parameters in model 5]%
+    {Estimated posterior distribution of parameters in model 5. Shaded range indicates 25\%--75\% quantile of the population}
+    \label{fig:model5_para}
+
+\end{figure}
+
+\begin{figure}[ht]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/inferno5.png}}
+    \end{center}
+    \caption[Joint density distribution of each parameter pair]%
+    {Joint distribution of each parameter pair. MORE DESCRIPTION}
+    \label{fig:model5_para_joint}
+\end{figure}
+
 By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
@@
 fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.
+
+\begin{table}[t!]
+    \centering
+    \begin{tabular}{|c c c|}
+        \hline
+        Parameter            & Mean value & Inter-quartile interval \\[0.5ex]
+        \hline\hline
+        $\lambda_N$          & 13.5       & [12.9, 14.2]            \\
+        $a$                  & 0.0175     & [0.0145, 0.0207]        \\
+        $\kappa_{N\beta}$    & 0.176      & [0.109, 0.234]          \\
+        $\mu_N$              & 0.0239     & [0.0113, 0.0332]        \\
+        $\nu_{N\Phi}$        & 0.125      & [0.115, 0.134]          \\
+        \hline
+        $\kappa_{\Phi\beta}$ & 0.123      & [0.115, 0.131]          \\
+        $\mu_\Phi$           & 0.0234     & [0.0199, 0.0267]        \\
+        \hline
+        $s_{\beta N}$        & 2.38       & [2.02, 2.73]            \\
+        $\mu_\beta$          & 2.67       & [2.24, 3.10]            \\
+        \hline
+        $s_{\alpha\Phi}$     & 0.574      & [0.295, 0.786]          \\
+        $\mu_\alpha$         & 8.91       & [6.07, 11.1]            \\
+        $f_{\beta\alpha}$    & 0.577      & [0.361, 0.738]          \\
+        \hline
+    \end{tabular}
+    \caption[Estimated parameter values of model 3]
+    {Estimated parameter values of model 3, using uniform prior distribution and log-uniform prior distribution respectively}
+    \label{table:estimated5}
+\end{table}
 

### 2020-08-14 implementation/implementation.tex
New figures

@@
     \end{tabular}
-    \caption[Estimated parameter values of model 3]
-    {Estimated parameter values of model 3, using uniform prior distribution and log-uniform prior distribution respectively}
+    \caption[Inferred parameter values of model 5]
+    {Inferred parameter values of model 5. MORE DETAIL(SUPER maxt-5)}
     \label{table:estimated5}
@@
 In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).
+
+\begin{figure}[h]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/overfit.png}}
+    \end{center}
+
+    \caption[Simulated trajectory from the last population of model 5, with factors applied]{Simulated trajectory from the last population of model 5, with factors applied. MORE DETAIL}
+    \label{fig:overfit}
+\end{figure}
 
@@
 PCA can only provide rough approximated sensitivity behaviour.
-As models 5 won in the model selection, a further interest is to explore the parameter-to-model sensitivity
+As models 5 won in the model selection, a further interest is to explore the parameter-to-model sensitivity.
 
-FIGURE shows the results of PCA on the last population of inferred model5.
+\begin{figure}[t]
+    \begin{center}
+        \resizebox{0.8\hsize}{!}{\includegraphics{fig/pc_pie_all.png}}
+    \end{center}
+
+    \caption[TODO]{PCA results of the last population of model 5. MORE DESCRIPTION}
+    \label{fig:pc}
+\end{figure}
+
+\begin{figure}[h]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/pc.png}}
+    \end{center}
+
+    \caption[PCA results of the last population of model 5]{PCA results of the last population of model 5. MORE DESCRIPTION}
+    \label{fig:pc}
+\end{figure}
+
+FIGURE shows the results of PCA on the last population of inferred model 5.
 The last PC (PC12) is mostly consists of the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive.

### 2020-08-15 appendix/appendix.tex
New para
New fix

@@
 When running ABC SMC, the program can make used of 8 python process (Hyper-Threading enabled) and run under 3.8GHz (maximal) for a long time under proper cooling, proved that the program could efficiently use the computation resources for a local personal computer.
-It is noted that the program took no advantage of graphics card, as \verb|multiprocessing| along can out exploit the resources inside CPU.
+It is noted that the program took no advantage of graphics card, as \verb|multiprocessing| along can only exploit the resources inside CPU.
 The execution time and other performance data presented in Chapter 5 were obtained using remote machines but not local machine.
@@
     \centering
-    \begin{tabular}{|c c c|}
+    \begin{tabular}{|c c|}
         \hline
-        Parameter            & Value & Unit                     \\ [0.5ex]
+        Parameter            & Value \\[0.5ex]
         \hline\hline
-        $\lambda_N$          & 2.20  & $cell/h$                 \\
-        $\kappa_{N\beta}$    & 3.96  & $cell/(unit\cdotp h)$    \\
-        $\mu_N$              & 1.72  & $h^{-1}$                 \\
-        $\nu_{N\Phi}$        & 0.220 & $cell^{-1}\cdotp h^{-1}$ \\
+        $\lambda_N$          & 2.20  \\
+        $\kappa_{N\beta}$    & 3.96  \\
+        $\mu_N$              & 1.72  \\
+        $\nu_{N\Phi}$        & 0.219 \\
         \hline
-        $\lambda_\Phi$       & 1.31  & $cell/h$                 \\
-        $\kappa_{\Phi\beta}$ & 0.124 & $cell/(unit\cdotp h)$    \\
-        $\mu_\Phi$           & 0.145 & $h^{-1}$                 \\
+        $\lambda_\Phi$       & 1.31 \\
+        $\kappa_{\Phi\beta}$ & 0.124 \\
+        $\mu_\Phi$           & 0.145 \\
         \hline
-        $s_{\beta N}$        & 6.554 & $unit/(cell\cdotp h)$    \\
-        $i_{\beta\Phi}$      & 1.71  & $cell^{-1}$              \\
-        $\mu_\beta$          & 0.521 & $h^{-1}$                 \\
+        $s_{\beta N}$        & 6.55  \\
+        $i_{\beta\Phi}$ & 1.71 \\
+        $\mu_\beta$          & 0.521  \\
         \hline
-        $s_{\alpha\Phi}$     & 10.2  & $unit/(cell\cdotp h)$    \\
-        $\mu_\alpha$         & 19.1  & $h^{-1}$                 \\
+        $s_{\alpha\Phi}$     & 10.2 \\
+        $\mu_\alpha$         & 19.7  \\
         \hline
     \end{tabular}
-    \caption{Parameter values used for model 1}
-    \label{table:synthetic_data}
+    \caption[TODO]
+    {Known values TODO}
+    \label{table:known_values}
 \end{table}
@@
 \end{figure}
+
+\begin{figure}[ht]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/inferno5.png}}
+    \end{center}
+    \caption[Joint density distribution of each parameter pair]%
+    {Joint distribution of each parameter pair. MORE DESCRIPTION}
+    \label{fig:model5_para_joint}
+\end{figure}

### 2020-08-15 implementation/implementation.tex
New fix

@@
 
-Estimation of  parameters in the proposed models is the main task of this section, which involves the implementation of ABC SMC and several experiments.
-ABC SMC-based model selection is another task that can be easily conducted with the developed parameter inference framework.
+For the proposed models, exploration of the uncertainties in the parameter estimations is our main task in this chapter, which involved the application of ABC SMC and several experiments.
+After that, the parameter inference framework was also used for model selection.
 
-The build and test of code were under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
-Some other code e.g.
-shell script and \verb|R| code are also used to perform the experiments and analyse the results.
-Software and system environment used in our implementation can be found at Appendix A.
+The build and test of the code were under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}.
+Some other code, e.g.
+shell scripts and \verb|R| code were also used to perform the experiments and result analysis.
+Software and system environment used in our implementation can be found in Appendix A.
 The build and test were performed on local computers and HPC facilities available within EPCC.
 
-\section{Implementations and code}
+\section{Code and implementations}
 
- [ABC implementations details, e.g. distance, population, ODE related functions]
+%  [ABC implementations details, e.g. distance, population, ODE related functions]
 
- [how the code is developed and built]
+%  [how the code is developed and built]
 
-According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
-\verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{population}.
-\verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks.
-Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.
+According to our preliminary studies, a ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
+\verb|pyABC| is a popular SMC software packages\cite{ref:pyabc} which has been used form many ABC SMC inference applications\cite{population}.
+\verb|pyABC| provides an open-source framework for likelihood-free inference, which is a \verb|NumPy| implementation of ABC SMC algorithm and a useful tool-box for our own inference and analysis tasks.
+Besides, it supports multi-core execution implemented using \verb|multiprocessing| built-in package, and it made our further studies in the scaling-up performance possible.
 
-The inference framework code for the project was developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task.
-The scaling-up experiment was performed on Cirrus, and the profiling and its analysis were performed on local machine.
-Some development and debug were on the remote machines to ensure experiments run correctly on compute node.
+%  which is of our interests and suitable for the followed performance experiments.
+
+The inference framework code for the project was developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first using a small input, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection tasks.
+The scaling-up experiment was performed on Cirrus.
+The profiling and its analysis were performed on local machine.
+Some necessary development and debug were performed on the remote machines to ensure that experiments run correctly on compute node.
 
 The ODE model-related code and utilities were firstly developed to enable reusable functions, variables and data calls.
-It includes the code for the ODE models, ODE solver, data structures, data format transform functions etc.
-By doing this in a separate code file, we can change options and activate ABC SMC easily in the implementation code file, without creating definition and duplicated code for the models and parameter priors.
+It included the code for the ODE equations, ODE solver, data structures and data format transform functions, etc.
+By doing this in separate code files, we could change implementation options and activate ABC SMC easily in another `trigger' file, without creating definition and duplicated code for the models, observed data and parameter priors.
 
-Some modifications were made to the source code of \verb|pyABC|.
-For example, the laboratorial measured data were not considered `complete' for modeling and analysis purpose: the available time points for cells and cytokines are different, e.g.
+Some missing features were found when using \verb|pyABC|, therefore some modifications were made to the source code of \verb|pyABC|.
+
+For example, the laboratorial measured data were not considered `complete' or `tidy' for modeling and analysis purpose: the available time points for cells and cytokines were different, e.g.
 cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl.
-These missing values were treated as `ignored' when calculating the distance between observed data and simulated data.
-To cope with this built-in distance function code is modified.
+These missing values were treated as `ignored' when calculating the distance between observed data and simulated data, and represented by \verb|NaN|s.
+However, in some distance calculations \verb|NaN| was not acceptable; to cope with this, built-in distance function was modified.
+Other modified code include some visualisation and result representation enhancements
 
-Analysis code was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results.
+Analysis of the program output was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results.
 For different tasks there are also separated analysis code file.
@@
 
-\section{ABC SMC and model hyperparameters experiments}
+\section{Hyperparameters and implementation options experiments}
 
-As the key focus of this project is on the parameter estimations of dynamical system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental observed data.
+As the key focus of this project was on the parameter estimations of dynamical system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental observed data.
 
@@
 
-    [including the options that are tested/ tried/ not adopted]
+    % [including the options that are tested/ tried/ not adopted]
 
-    [including experiments results and discussions]
+    % [including experiments results and discussions]
 
-    [how do these options correspond to biological model and how to set them accordingly for proposed model]
+    % [how do these options correspond to biological model and how to set them accordingly for proposed model]
+
+    % [why use synthetic data]
 
@@
 To firstly test the ability of inference and observe the results of different options, synthetic data with known true parameter values are used as target data.
-The true value of parameters that are used to generate synthetic data is listed in Appendix B.1.
-These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data (Figure \ref{fig:infer_back_data}), the target data is then generated using these parameter values via ODE solver in \verb|Python|.
+by doing this, we could directly compare the inferred parameters to the true value, and compare the simulated trajectories to the true trajectories and then observe the result to see whether the inference is successful and how well the model fit the data.
+Also, by setting the same final threshold value, the efficiency under different options can be compared and help us to choose proper options when applying ABC SMC on the observed data.
+The true value of parameters that were used to generate synthetic data is listed in Table \ref{table:known_values}.
 
-\begin{figure}[h!]
+The true values were obtained by fitting model 1 onto the experimental data (Figure \ref{fig:infer_back_data}) using least-square optimisation method.
+Available least-square fit tools in \verb|SciPy| were used.
+The fit algorithm was Trust Region Reflective\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.least_squares.html}}.
+Target data was then generated using these parameter values via ODE integrate solver\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.odeint.html}}.
+
+\begin{figure}[ht]
     \begin{center}
@@
 \end{figure}
-
-[TABLE for known parameter values]
-
-[why use synthetic data]
 
@@
 
-[how adaptive distance work]
+% [how adaptive distance work]
 
@@
 
-\subsubsection{Experiments results}
+\subsubsection{Experiment results}
 
@@
 
-\begin{figure}[ht]
+Two types of factor and adaptive distance function were tested by running the same number of generations (20 generation, 2000 particles per generation).
+The first type of factor is 25:75  factor, where first half of the trajectory data was assigned higher factor (75), and second half of the data was assigned lower factor (25); this could make the first half if the data more `important' in the inference.
+The second factor was range factor, which tried to normalise the trajectory of each dependent variable according to its data range.
+
+From Figure \ref{fig:factor}, different types of factor gave different performance.
+25:75 factor required much more samples in the first generation (orange bar in the left of Figure \ref{fig:factor}) and also did not improve much in later generations, thus it is the slowest one.
+The first half of observed data contained more features and information, assigning it with higher factor would make the algorithm think that it is more important to fit the first half and thus focus on the hard part, and consequently more particles were needed in the inference process.
+The acceptance rates of 25:75 factor among different generations were also generally slightly lower than the standard.
+By contrast, range factor required less total samples and gave generally higher acceptance rates than standard, which made it considerably faster.
+Adaptive distance gave the best efficiency, as it significantly improved the acceptance rates and resulted in even less required samples (right, Figure \ref{fig:factor}).
+
+\begin{figure}[t!]
     \begin{center}
@@
 
-Two types of factors and adaptive distance function are tested by running the same number of generations (20 generation, 2000 particles per generation).
-From FIGURE, by applying factors the total required samples are less, and the adaptive distance requires much less samples; their acceptance rates are also higher than the standard one.
+However, the result of factors adaptive distance could not be directly compared to standard run, because it changed the distance measurement by adapting weights or manually setting factors.
+Factors were normalise to make it more close the standard distance measurement.
+As a result, the same final threshold value did not guarantee that adaptive distance can reach the same level of fit as others.
+After 20 populations, adaptive distance had a poor fit (Figure \ref{fig:factor_sim}), while the standard run and two types of factors had similar tight fits of the data, which all converge to the true posterior.
+Additional in our implementations, variance factor (which tries to normalise the data according to the dependent variables' variance) was also tries and gave very similar performance as range factor.
+
+To conclude, because the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, the direct compare of the efficiency under the same threshold schedule was unfeasible; we could only compare the results after the same number of generations.
+Factors gave nearly the same inferred results but required different number of samples, hence cares must betaken when choosing factor schemes.
+Adaptive distance did not result in a accurate result as others.
+As a result, our further implementations used standard distance function, and factors were tried for possible improvements.
+
+% by applying factors the total required samples are less, and the adaptive distance requires much less samples; their acceptance rates are also higher than the standard one.
 It seems that adaptive function and factors could help to converge more quickly.
 
-However considering the result after 20 generation (FIGURE), the resultant model gives the contrary preference.
+% However considering the result after 20 generation (Figure \ref{fig:factor_sim}), the resultant model gives the contrary preference.
 Applying factors and using adaptive distance can lead the resultant model less accurate (after the same generation iterations): they have wider inter-quantile range and may requires more generations to converge.
@@
 
-\subsubsection{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases.
+% \subsubsection{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases.
 However, as the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, a direct compare of the efficiency under the same threshold schedule is unfeasible; after the same number of generations they can reach a small $\epsilon_t$ but still the approximated posterior is not as accurate as the standard implementation.
@@
 
-Regarding the dynamic systems model, SMC was applied with two data size options, three prior distribution (uniform distribution, wider uniform distribution and log-uniform distribution).
+Regarding the existing dynamic systems model, SMC was applied with two data size options, three prior distribution (uniform distribution, wider uniform distribution and log-uniform distribution).
 The population options i.e.
@@
 
-The prior range and distribution can largely affect the execution of SMC (FIGURE and FIGURE).
-Samples are taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space.
-It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.
-
-    [discuss uniform and log uniform when result ready]
-
-\begin{figure}
+\begin{figure}[t!]
     \begin{center}
@@
 
-[FIGURE HERE]
+The prior range and distribution can largely affect the execution of SMC (FIGURE and FIGURE).
+Samples are taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space.
+It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.
+
+%     [discuss uniform and log uniform when result ready]
+
+% [FIGURE HERE]
 
@@
 
-    \caption[Model comparison of model 1, 2 and 3]%
-    {Model comparison of model 1, 2 and 3. From population 8 to population 30, model3 always wins}
+    \caption[ABC SMC based model comparison of model 1, 2 and 3]%
+    {ABC SMC based model comparison of model 1, 2 and 3. From population 8 to population 30, model3 always wins}
     \label{fig:cmp1}
@@
 
-\begin{figure}[ht]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/inferno5.png}}
-    \end{center}
-    \caption[Joint density distribution of each parameter pair]%
-    {Joint distribution of each parameter pair. MORE DESCRIPTION}
-    \label{fig:model5_para_joint}
-\end{figure}
-
 By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
@@
 
+\begin{figure}[H]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model345cmp.png}}
+    \end{center}
+
+    \caption[ABC SMC based model comparison of model 3, 4 and 5]{ABC SMC based model comparison of model 3, 4 and 5. MORE DETAIL}
+    \label{fig:overfit}
+\end{figure}
+
 Next, we performed a model selection on model 3, 4 and 5.
@@
 
-\begin{figure}[t]
+\begin{figure}[H]
     \begin{center}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \begin{center}
@@
     \caption[PCA results of the last population of model 5]{PCA results of the last population of model 5. MORE DESCRIPTION}
-    \label{fig:pc}
+    \label{fig:pc_pie}
 \end{figure}

### 2020-08-15 implementation/implementation.tex
New fix

@@
 
-After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (udsing synthetic data), and code files for parameter inference (using experimental data).
+After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (using synthetic data), and code files for parameter inference (using experimental data).
 Model selection can be easily done by adding more model to the implementation code file.
@@
 
-Among them the following topics are studied using the synthetic data.
+The following topics were studied using the synthetic data and ABC SMC inference framework.
 
@@
 After that the new weights are calculated and normalised.
+Figure \ref{fig:smc} illustrates the process.
 
@@
 
-It seems that a more `local' kernel can give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
+A more `local' kernel was more likely to give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
 This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
@@
 12.5\% of the population) and multivariate normal with scaling.
-Also multivariate M-NN is worth trying, although it is not built in \verb|pyABC| and requires additional implementations.
+Also, multivariate M-NN is worth trying, although it was not built in \verb|pyABC| and requires additional implementations.
 
@@
 
-[experiments plan]
+% [experiments plan]
 
@@
 
-The results from data size and prior distribution range is shown in FIGURE.
-The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement.
-Although much less (60\%) data is used, the total required sampling numbers is not much less (XX\%).
-Although the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data (FIGURE), cares should be taken when using less data point, or using other more summative statistics, e.g.
-mean, standard deviation, peak values, starting and ending values etc.
-As shown in \cite{ref:disease}, less data can lead to a model with more variance, under-fitting or missing some local features e.g.
-local peak.
-
-\begin{figure}[t!]
+\begin{figure}[H]
     \begin{center}
@@
 
-The prior range and distribution can largely affect the execution of SMC (FIGURE and FIGURE).
-Samples are taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space.
+The results from data size and prior distribution range is shown in Figure \ref{fig:size}.
+The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement.
+Although much less (60\%) data was used, the total required sampling numbers was only 4.47\% less.
+This indicated that the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data, as they finally reached a similar epsilon threshold ((b), Figure \ref{fig:size}), but cares should still be taken when using less data point, or using other more summative statistics (e.g.
+mean, standard deviation, peak values, starting and ending values etc.) as the target data: as shown in \cite{ref:disease}, less data can lead to a model with more variance, under-fitting or missing some local features such as local peaks.
+
+The prior range and distribution can largely affect the inference progress (Figure \ref{fig:size}.
+(a)).
+Samples were drawn from a high dimensional parameter space in each sampling process, a wider parameter prior interval in only one dimension can cause the parameter space to be much bigger, and consequently the exploration and sampling of parameter vectors would take much more efforts.
+Figure \ref{fig:size}.
+(b) also indicates that the convergency is slower for wider prior ranges.
+As a result, a proper prior interval based on our experience and belief with seeing the data should be set as narrow as possible, as inference are largely affected by the prior interval and its length; improper prior could even make the algorithm miss the true posterior \cite{ref:abcsysbio}.
+
+How to specify prior can be tricky from problem to problem.
+For our task in the parameter inference of regeneration model, biologically motivated priors were used, which were general reflection of biological constrains.
+Also the units of the parameters help us in identify the possible scale of the parameters.
+
+% taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space.
 It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.
@@
 
-\subsubsection{Summary} This subsection aims to preliminarily explore ABC SMC settings using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options.
+\subsubsection{Choise of options}
+
+This subsection aims to preliminarily explore options in ABC SMC implementations.
+From the above results, we adopted multivariate normal kernel as the kernel for our real inference on observed data; Local M-NN kernels were also planned to be tested.
+In the real inference, 25:75 factors were also tried to see if it could help in recovering more features and resulting a better fit.
+because the improvement by using less data was inapparent, we did not try to change the target summary data size.
+Prior intervals were explored explicitly and gradually shrunk for our models.
+
+Some other hyperparameter, e.g.
+number of generations, population size (i.e.
+number of particles per generation) were not addressed in this subsection.
+In the real inference, population size was set to be greater than 2000 to avoid insufficient exploration of the previous population; in some case than high variance or uncertainties in the results were observed, population size as large as 20,000 were used.
+If the epsilon did not show a convergency trend, the run could be extended to more generations until a stable result was observed.
+
+% using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options.
 Conclusion can be drawn to choose efficient and proper options of ABC SMC to be applied on the target models with experimental data.
@@
     \caption[ABC SMC based model comparison of model 3, 4 and 5]{ABC SMC based model comparison of model 3, 4 and 5. MORE DETAIL}
-    \label{fig:overfit}
+    \label{fig:model345cmp}
 \end{figure}
 
-Next, we performed a model selection on model 3, 4 and 5.
-FIGURE shows the model probabilities at different generations.
+Next, we performed a model comparison on model 3, 4 and 5, using the same ABC SMC inference framework. Figure \ref{fig:model345cmp} shows the model probabilities at different generations.
+
+Model 4 did not overweight other models across all the generations; model 3 gained advantage before generation 15 except generation 5.
+After generation 18 ($\epsilon_{18} = 16.98$), model 5 gained absolute advantage and won in further generations.
+Also, repeated runs showed similar trends: although uncertainties were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
+$\epsilon_t<16$.
+
+Based on the model comparison results, we could conclude that model 5 was the best model among the tested models. Uncertainties were observed in early generations,
+
 From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred.

### 2020-08-16 implementation/implementation.tex
New content

@@
 From the above results, we adopted multivariate normal kernel as the kernel for our real inference on observed data; Local M-NN kernels were also planned to be tested.
-In the real inference, 25:75 factors were also tried to see if it could help in recovering more features and resulting a better fit.
+In the real inference, 25:75 factor were also tried to see if it could help in recovering more features and resulting a better fit.
 because the improvement by using less data was inapparent, we did not try to change the target summary data size.
@@
 
-[separated ABC run on model 1, 2 and 3]
+% [separated ABC run on model 1, 2 and 3]
 
@@
 
+By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
+Model 1, 2 and 3 con also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- is not fitted.
+Some features, e.g.
+fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.
+
+After applying factors, the results were not largely different.
+As expected, applying factors made the data points at first half (where exists most of the fluctuations) more `focused', however the inferred models are close to one without factors, which could indicate that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.
+
+When we are trying more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}).
+In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).
+MORE
+
+\begin{figure}[H]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/overfit.png}}
+    \end{center}
+
+    \caption[Simulated trajectory from the last population of model 5, with factors applied]{Simulated trajectory from the last population of model 5, with factors applied. MORE DETAIL}
+    \label{fig:overfit}
+
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model345cmp.png}}
+    \end{center}
+
+    \caption[ABC SMC based model comparison of model 3, 4 and 5]{ABC SMC based model comparison of model 3, 4 and 5. MORE DETAIL}
+    \label{fig:model345cmp}
+\end{figure}
+
+Next, we performed a model comparison on model 3, 4 and 5, using the same ABC SMC inference framework. Figure \ref{fig:model345cmp} shows the model probabilities at different generations.
+
+Model 4 did not overweight other models across all the generations; model 3 gained advantage before generation 15 except generation 5.
+After generation 18 ($\epsilon_{18} = 16.98$), model 5 gained absolute advantage and won in further generations.
+Also, repeated runs showed similar trends: although uncertainties were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
+$\epsilon_t<16$.
+
+Based on the model comparison results, we could conclude that model 5 was the best model among the tested models.
+Uncertainties were observed in early generations, where similar patterns were also observed in the model comparison of model 1, 2 and 3.
+We considered the the violently changing model probability in early generations is related to the inference ability at high epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
+However as we only focused on the best model with small final threshold $\epsilon_T$, the easiest way is to extended the number of generation until epsilon converges, or increase the population size for a more statistically confident probability.
+
+% From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred.
+Again, as DISCUSSED in the model comparison of model 1, 2 and 3, Given the present result, model 5 is regarded as the best model.
+
+% Some further experiments tried larger population size (10,000 particles per population) and more populations (36 populations), but the resultant model did not have a significantly better fit of the observed data, although much more computation resources are used.
+
+For the best model (model 5), the inferred parameter posterior distribution of model 5 is shown in Figure \ref{fig:model5_para}, the estimated mean values is shown in Table \ref{table:estimated5}.
+The (pair-wise) joint density distribution is shown in Figure \ref{fig:model5_para_joint}.
+SOME DETAIL.
+
 \begin{figure}[ht]
@@
 
-By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
-Model 1, 2 and 3 con also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- is not fitted.
-Some features, e.g.
-fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.
-
-\begin{table}[t!]
+\begin{table}[H]
     \centering
@@
 
-After applying factors, the results were not largely different.
-As expected, applying factors made the data points at first half (where exists most of the fluctuations) more `focused', however the inferred models are close to one without factors, which could indicate that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.
-
-When we are trying more populations, an `over-fitting' behaviour is observed at the trajectory of macrophage (FIGURE).
-In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).
-
-\begin{figure}[h]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/overfit.png}}
-    \end{center}
-
-    \caption[Simulated trajectory from the last population of model 5, with factors applied]{Simulated trajectory from the last population of model 5, with factors applied. MORE DETAIL}
-    \label{fig:overfit}
-\end{figure}
-
-\begin{figure}[H]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model345cmp.png}}
-    \end{center}
-
-    \caption[ABC SMC based model comparison of model 3, 4 and 5]{ABC SMC based model comparison of model 3, 4 and 5. MORE DETAIL}
-    \label{fig:model345cmp}
-\end{figure}
-
-Next, we performed a model comparison on model 3, 4 and 5, using the same ABC SMC inference framework. Figure \ref{fig:model345cmp} shows the model probabilities at different generations.
-
-Model 4 did not overweight other models across all the generations; model 3 gained advantage before generation 15 except generation 5.
-After generation 18 ($\epsilon_{18} = 16.98$), model 5 gained absolute advantage and won in further generations.
-Also, repeated runs showed similar trends: although uncertainties were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
-$\epsilon_t<16$.
-
-Based on the model comparison results, we could conclude that model 5 was the best model among the tested models. Uncertainties were observed in early generations,
-
-From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred.
-Again, as DISCUSSED in the model comparison of model 1, 2 and 3, we considered the the violently changing model probability is related to the inference ability at certain epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
-Given the present result, model 5 is regarded as the best model.
-Some further experiments tried larger population size (10,000 particles per population) and more populations (36 populations), but the resultant model did not have a significantly better fit of the observed data, although much more computation resources are used.
-
-A inferred parameter posterior distribution of model 5 is shown in FIGURE, the estimated mean values is shown in TABLE. SOME DESCRIPTION.
-
 \subsection{Sensitivity of parameters}
 
-[why last PCs]
+% [why last PCs]
 
-[which parameters are most sensitive]
+% [which parameters are most sensitive]
 
-[compared to credible intervals plot]
+% [compared to credible intervals plot]
 
-[what does that mean]
+% [what does that mean]
 
@@
 PCA can only provide rough approximated sensitivity behaviour.
-As models 5 won in the model selection, a further interest is to explore the parameter-to-model sensitivity.
-
-\begin{figure}[H]
-    \begin{center}
-        \resizebox{0.8\hsize}{!}{\includegraphics{fig/pc_pie_all.png}}
-    \end{center}
-
-    \caption[TODO]{PCA results of the last population of model 5. MORE DESCRIPTION}
-    \label{fig:pc}
-\end{figure}
+By identifying the component loadings, the PCs can be visualised in pie charts and used to indicate the sensitivity.
 
@@
     \caption[PCA results of the last population of model 5]{PCA results of the last population of model 5. MORE DESCRIPTION}
+    \label{fig:pc}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \resizebox{0.8\hsize}{!}{\includegraphics{fig/pc_pie_all.png}}
+    \end{center}
+
+    \caption[TODO]{PCA results of the last population of model 5. MORE DESCRIPTION}
     \label{fig:pc_pie}
@@
 
-FIGURE shows the results of PCA on the last population of inferred model 5.
-The last PC (PC12) is mostly consists of the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive.
-If we conclude the last 5 PCs (FIGURE), then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$.
+As models 5 won in the model selection, a further interest was to explore the parameter-to-model sensitivity of the resultant model.
+PCA was performed on the last generation of inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC.
+The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less less than 5\% of the variance each.
+PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs, thus they were consider to extend much narrower regions on the posterior distributions and could be used to identify parameter that the model is sensitive to.
+
+Figure \ref{fig:pc_pie} shows the fractions of the PCs explained by individual parameters, and a sum of the fractions in the last 8 PCs, i.e.
+PC 5 to PC 12.
+The last PC (PC12) is mostly the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive.
+If we conclude the last 8 PCs, it can be found that the biggest portions ($v_{N\Phi}$, $a$, $\kappa_{N\beta}$, $\kappa_{\Phi\beta}$, $\mu_N$, $\mu_\Phi$, $\nu_{N\Phi}$) mostly come from the equation of neutrophil and macrophage i.e.
+equations of the cells (the un-exploded portion in Figure \ref{fig:pc_pie} a); if we look at the last 8 PCs individually, the same pattern can be observed.
+This might suggest that the model is more sensitive to cells' dynamic rather than cytokines' dynamics.
+
+The credible intervals (Figure \ref{fig:credible}) of each parameter across generations also agreed with the above conclusion from PCA.
+The stiff parameters, e.g.
+$a$ and $\mu_N$, extend narrow credible intervals, while sloppy parameters, e.g.
+$\mu_\alpha$ and $s_{\alpha\Phi}$, have wider credible intervals across the last few generations.
+
+% then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$.
 As a conclusion, model 5 is sensitive to changes in the above parameters, which all come from the equation of neutrophil and macrophage i.e.
@@
 
-Also, some visualisations of the last population agreed with conclusions from PCA.
+% Also, some visualisations of the last population agreed with conclusions from PCA.
 Approximated posteriors of the last population plotted in FIGURE and credible ranges of parameters across populations in FIGURE shows that the 5 parameters identified by PCA is indeed `stiff' parameters.

### 2020-08-16 performance/performance.tex
New figures
New fix
New content

@@
 
-    [MORE meanings of the performance study]
+% [MORE meanings of the performance study]
 
@@
 
-[HOW PYABC parallelise the sampling]
+% [HOW PYABC parallelise the sampling]
 
@@
 
-[scaling-up performance: speed-up and efficiency]
+% [scaling-up performance: speed-up and efficiency]
 
-[large variation in required sampling numbers]
+% [large variation in required sampling numbers]
 
-[possible reasons]
+% [possible reasons]
 
-The recorded execution time using different number of cores is shown in FIGURE.
-According to that, speed-up and efficiency can be calculated.
+\begin{figure}[h]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/speedup.png}}
+    \end{center}
+
+    \caption[TODO]{Speedup. MORE DESCRIPTION}
+    \label{fig:speedup}
+\end{figure}
+
+The recorded execution time using different number of cores were used to calculate the speedup of the program (Figure \ref{fig:speedup}).
 Here 8 cores is regarded as the baseline for speed-up and efficiency, as the serial version (using only one core) can take quite a long time to finish 10 repeats.
 When increasing number of cores, the execution time drops fast at first but the decreasing rate (dropping speed, i.e.
-the derivative of the execution time against time) is gradually reduced to zero; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (FIGURE).
+the derivative of the execution time against time) is gradually reduced to zero; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (Table \ref{table:performance}).
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c c c c c|}
+        \hline
+        Number of cores & Time (second) & Avg. required samples & Speedup & Efficiency \\ [0.5ex]
+        \hline\hline
+        8               & 6617          & 445406.8              & 1       & 1          \\
+        16              & 3926          & 590221.0              & 1.61    & 0.807      \\
+        24              & 2921          & 573339.8              & 2.17    & 0.723      \\
+        36              & 2096          & 446948.7              & 3.02    & 0.672      \\
+        54              & 2076          & 640404.9              & 3.05    & 0.452      \\
+        72              & 2071          & 768096.0              & 3.06    & 0.340      \\
+        \hline
+    \end{tabular}
+    \caption{Performance data}
+    \label{table:performance}
+\end{table}
 
@@
 
-A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations.
-Moreover, we found that the execution time is not ideally proportional to the total required samples (FIGURE).
+\begin{figure}[h]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/sample_per_time.png}}
+    \end{center}
+
+    \caption[TODO]{Samples per second. MORE DESCRIPTION}
+    \label{fig:sample_per_sec}
+\end{figure}
+
+A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations, where they were supposed to have relatively similar average required samples such that we can compare the performance directly via speedup or efficiency.
+
+Due to the variant required samples, our interests switched to the per-second performance, where the average sampling numbers per second under different of number of cores ($p$) are plotted (Figure \ref{fig:sample_per_sec}).
+The plot shows that even theoretical speedup stops growing when $p>36$, the average sampling numbers per second performance still get improved when using more than 36 cores.
+It proves that the implementation used here scales reasonably well without encountering bottleneck when using multi-core parallelisation inside one physical node.
+
+The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path.
+Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
+For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated
+
+will take huge amount of samples to move away from the local optimum
+
+Moreover, we found that the execution time was not ideally proportional to the sampling tires in that period of time, for both different repeated runs and different generations in a single run.
+To illustrate this, a single run from $p=24$ was selected and visualised.
+
 A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs.
@@
 
-Due to the unstable sampling numbers, our interests switched to the per-particle performance, where the average sampling numbers per second (FIGURE) and average time spent in one particle (FIGURE) under different of cores ($p$) are plotted.
-The log-like average time per particle curve shows that using more cores can improve the sampling efficiency, but when $p$ is at a high level (e.g.
-36 in our case), the improvement will be less significant.
+% \section{Profiling}
 
-\section{Profiling}
-
-The performance could also be analysed given a profiling report.
+% The performance could also be analysed given a profiling report.
 The second experiment profiles the program to reveal the detailed time consumption for each operation and the possible bottleneck, according to which we could find the hot-spot of program and given possible suggestions on improving the performance.
 
-In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
+% In this case, profile tools \verb|cProfile| and \verb|yappi| is used in PyCharm IDE.
 
-\subsubsection{Results}
+% \subsubsection{Results}
 
-[profiling results: hot-spot and possible improvements]
+% [profiling results: hot-spot and possible improvements]

### 2020-08-16 performance/performance.tex
Almost finished

@@
 
+\section{Discussions}
+
 A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations, where they were supposed to have relatively similar average required samples such that we can compare the performance directly via speedup or efficiency.
@@
 
+\begin{figure}[ht]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/local_modes.png}}
+    \end{center}
+
+    \caption[TODO]{Local modes. MORE DESCRIPTION}
+    \label{fig:local_modes}
+\end{figure}
+
 The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path.
 Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
-For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated
+For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated in the local optimum.
+In this case, huge amount of sample will be drawn to find enough accepted particles, and new generation will be able to jump out of the local optimum and have different concentrations.
 
-will take huge amount of samples to move away from the local optimum
+To illustrate this, Two independent run, namely Run A and Run B, were chosen from the experiment of $p=24$ and visualised in Figure \ref{fig:local_modes} and Figure \ref{fig:local_para}.
+Thess two run used identical options of the algorithm.
+Figure \ref{fig:local_modes} (a) shows the comparison of the required samples, where Run A converged to a local mode, and in population 11 took much more sample tires to move away from the local optimum (Figure \ref{fig:local_modes} (b)).
+Figure \ref{fig:local_para} compares the posterior density distribution before (t=9, cyan) and after (t=13, red) the jumping away: significant movement of concentrations in posterior density distributions are observed, especially in parameters from ODE 1, 2 and 4.
+An interning movement is that there are two concentrations in $\kappa_{\Phi\beta}$ and $\mu_\Phi$ when t=9, and the majority of particles are concentrated in the second concentration with high parameter values; when t=13, the second concentration distribution disappears and all particles are concentrated near zero.
+
+\begin{figure}[t]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/local_para.png}}
+    \end{center}
+
+    \caption[TODO]{Parameter density distribution. MORE DESCRIPTION}
+    \label{fig:local_para}
+\end{figure}
 
 Moreover, we found that the execution time was not ideally proportional to the sampling tires in that period of time, for both different repeated runs and different generations in a single run.
-To illustrate this, a single run from $p=24$ was selected and visualised.
+From the same example, Figure \ref{fig:local_modes} (b) and (c) show the non-linear relationship between execution time and sample tries in different generations.
+Apart from 1st and 2nd generation, generation 3 to 20 show a general linear relationship i.e.
+$T_t\propto n_t$, where $t$ is the population index, $T_t$ is the execution time of generation $t$ and $n_t$ is the required samples in generation $t$.
+However, this relationship does not hold for the first two generation.
+Generation 1 is to draw sample from prior, but it is done with high efficiency as only small amount of time are taken; generation 2 takes unusual large amount of the, but it does not require much more samples.
+The abnormal execution time of generation 1 and two are observed in all other runs, where for $p=24$ generation 1 usually takes c.
+16s and generation 2 takes c.
+1600s.
+This is believed to be related to the initialisation of the algorithm.
 
-A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs.
+Additionally, by observing CPU frequency and load curve in-time while running the program, we found that the parallelised part is limited the the sampling, calculation and criterion test process; some reduction calculations and preparation work for the next generation (e.g.
+calculation of the new epsilon values, normalisation of data, fit of multivariate kernel) are serial.
+Database I/O is also serial.
+
+% A possible explain to these phenomenons can be that the median epsilon schedule results in different threshold strategies due to the randomness in the sampling process, thus results in different approximate convergency paths for different runs.
 Some runs that are stuck in a local optimum and require smaller target epsilon will take huge amount of samples to move away from the local optimum.

### 2020-08-16 implementation/implementation.tex
New fix

@@
     \caption[Factors and adaptive distance experiments]%
-    {Factors and adaptive distance experiments. Left: Total required samples. Right: acceptance rates in each generation}
+    {Factors and adaptive distance experiments.
+Left: total required samples.
+Different color represents different generations (bottom totop: population 1 to population 20).
+Right: acceptance rates in each generation.
+}
     \label{fig:factor}
@@
 
-    \caption[CAPTION]%
-    {DESCRIBE}
+    \caption[Simulated trajectories ]%
+    {Simulated trajectories from the last population in the factor and adaptive distance experiment.
+500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from.
+25:75 factor showed very similar curve as the range factor.
+    }
     \label{fig:factor_sim}
@@
     \caption[Data size and prior distribution range experiment]%
-    {Data size and prior distribution range experiment. (a) Total required samples, (b) Epsilon values under median epsilon schedule}
+    {Data size and prior distribution range experiment. Left: total required samples. Right: epsilon values under median epsilon schedule}
     \label{fig:size}
@@
 
-    \caption{Simulated data from the last population of model 3, 4 and 5}
+    \caption{Simulated data from the last population of model 3, 4 and 5. Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
     \label{fig:resultCurve345}
@@
 When we are trying more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}).
-In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl).
-MORE
+In this case, the discrepancy of simulated data and observed data mainly comes from the last 2 data point (72 hpl and 120 hpl) of macrophage.
+As additional importance was assigned the the first few data points of each trajectory, the inference algorithm tended to preferentially fit first data points and thus relative large discrepancy at the tails were acceptable.
 
@@
 
-    \caption[Simulated trajectory from the last population of model 5, with factors applied]{Simulated trajectory from the last population of model 5, with factors applied. MORE DETAIL}
+    \caption[Simulated trajectory from the 30th population of model 5, with 25:75 factor applied]{Simulated trajectory from the 30th population of model 5, with 25:75 factor applied.
+Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small).}
     \label{fig:overfit}
+\end{figure}
 
+Next, we performed a model comparison on model 3, 4 and 5, using the same ABC SMC inference framework. Figure \ref{fig:model345cmp} shows the model probabilities at different generations.
+
+\begin{figure}[H]
     \begin{center}
@@
 
-    \caption[ABC SMC based model comparison of model 3, 4 and 5]{ABC SMC based model comparison of model 3, 4 and 5. MORE DETAIL}
+    \caption[Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison]{Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison.
+Model 5 also wins in further populations (population 25 to 30)}
     \label{fig:model345cmp}
 \end{figure}
-
-Next, we performed a model comparison on model 3, 4 and 5, using the same ABC SMC inference framework. Figure \ref{fig:model345cmp} shows the model probabilities at different generations.
 
@@
 For the best model (model 5), the inferred parameter posterior distribution of model 5 is shown in Figure \ref{fig:model5_para}, the estimated mean values is shown in Table \ref{table:estimated5}.
-The (pair-wise) joint density distribution is shown in Figure \ref{fig:model5_para_joint}.
-SOME DETAIL.
+The pair-wise joint density distribution is shown in Figure \ref{fig:model5_para_joint}.
+Most of inferred posteriors show a bell-shaped density distribution, indicating that they are gaussian distributed.
+A thin and concentrated shape indicate a well inferred parameter, e.g.
+$\mu{\Phi}$ and $\kappa_{\Phi\beta}$.
+Some distributions, however, are skewed to zero, e.g.
+$\mu_N$ ans $s_{\alpha\Phi}$.
+
+By looking at the joint density plot (Figure \ref{fig:model5_para_joint}), some correlated patterns are observed.
+For example, an obvious linear relationship on the distributions of $\mu_\beta$ and $s_{\beta\Phi}$.
+This leads us to consider that reducing the number of parameters in the model is possible.
+Highly correlated parameters can be reduced, e.g.
+$s_{\beta\Phi}$ can be expressed as $A\mu_\beta+B$ where $A$ and $B$ can be calculated from the joint distribution by fitting a linear regression.
 
@@
 
-    \caption[Estimated posterior distribution of parameters in model 5]%
-    {Estimated posterior distribution of parameters in model 5. Shaded range indicates 25\%--75\% quantile of the population}
+    \caption[Inferred posterior distribution of parameters in model 5]%
+    {Inferred posterior distribution of parameters in model 5. Shaded range indicates 25\%--75\% quantile of the population}
     \label{fig:model5_para}
@@
     \caption[Inferred parameter values of model 5]
-    {Inferred parameter values of model 5. MORE DETAIL(SUPER maxt-5)}
+    {Inferred parameter values of model 5}
     \label{table:estimated5}
@@
 
-    \caption[PCA results of the last population of model 5]{PCA results of the last population of model 5. MORE DESCRIPTION}
+    \caption[Explained total variance in original parameters by each PC]{Explained total variance in original parameters by each PC.
+From PC 5 on.
+little variance is explained by each PC, which is enlarged in Right}
     \label{fig:pc}
@@
 
-    \caption[TODO]{PCA results of the last population of model 5. MORE DESCRIPTION}
+    \caption[TODO]{Fraction of the length of PCs explained by parameters}
     \label{fig:pc_pie}

### 2020-08-16 appendix/appendix.tex
New content
Grammar fix

@@
 
+Cirrus hardware can be found in \url{https://www.cirrus.ac.uk/about/hardware.html}.
+The software environment used in the development and experiments are listed in Table \ref{table:cirrus_soft}.
+Requirement of site packages is the sanme as that on local macine (Table \ref{table:local_package})
+
 \begin{table}[H]
@@
         \hline\hline
-        Placeholder                 & 2\\
+        Operating system      & Red Hat Enterprise Linux 8.1 \\
+        miniconda environment & 4.8.3                        \\
+        Python interpreter    & 3.7.7                        \\
+        gcc                   & 6.3.0                        \\
         \hline
     \end{tabular}
-    \caption{Environment on remote machine}
-    \label{table:remote_macine}
+    \caption{Environment on local machine}
+    \label{table:cirrus_soft}
 \end{table}
 
-\section{Data analysis}
+% \section{Data analysis}
 
-\chapter{Data and settings}
+\chapter{Data and supplementary figures}
 
@@
     \end{tabular}
-    \caption[TODO]
-    {Known values TODO}
+    \caption[Parameter values used to generate synthetic data for model 1]
+    {Parameter values used to generate synthetic data for model 1. They were obtained from a preliminarily least-square fitting}
     \label{table:known_values}
@@
     \caption[Acceptance rates of different kernels, using median epsilon strategy]
-    {Acceptance rates of different kernels, using median epsilon strategy. Each population has 2000 particles}
+    {Acceptance rates of different kernels, using median epsilon strategy.
+Each population has 2000 particles.
+The total length of populations is different because some runs reached the final $\epsilon_T$ earlier than others}
     \label{fig:acceptance2}
@@
 
-    \caption{Total sampling size}
+    \caption{Simulated data from the last population of model 1, 2 and 3, using uniform prior.  Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
     \label{fig:resultCurve_uni}
@@
     \caption[Joint density distribution of each parameter pair]%
-    {Joint distribution of each parameter pair. MORE DESCRIPTION}
+    {Joint distribution of each parameter pair.
+The diagonal is KDE fit of each individual parameter, the upper triangle contains scatter plot of the particles and the lower triangle is the joint density distribution for each parameter pair}
     \label{fig:model5_para_joint}
 \end{figure}
+
+\begin{figure}[ht]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/credible.png}}
+    \end{center}
+    \caption[Credible intervals of parameters in model 5]%
+    {Credible intervals of parameters in model 5 across all generations}
+    \label{fig:credible}
+\end{figure}

### 2020-08-16 background/background.tex
Grammar fix

@@
 
-Compared to mammals, zebrafish can regenerate spinal cord after injury, and possibly recover swimming function.
+Compared to mammals, zebrafish can regenerate the spinal cord after injury, and possibly recover swimming function.
 They are widely used in regenerative ability researches due to this ability.
-Experiments found that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}.
-Tsarouchas et al.\cite{ref:Tsarouchas} studied the the complex interactions happen at the lesion site.
-The inflammation affects the regeneration, as if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration\cite{ref:Tsarouchas}.
+Experiments discovered that although being cut off, the axonal of nerve cells can regenerate and bridge across the lesion site, which critically determined whether zebrafish can regain the swimming function \cite{axonal}.
+Tsarouchas et al.
+\cite{ref:Tsarouchas} studied the complex interactions that happen at the lesion site.
+The inflammation affects the regeneration: if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration\cite{ref:Tsarouchas}.
 Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, are related the inflammation, and they are dynamically controlled by immunes cells, i.e.
 immune cells produce and control the cytokines-mediated inflammation \cite{ref:Tsarouchas}.
-Cytokines are small proteins, work as messenger to pass signal across different cells, which are similar to hormones and neurotransmitter.
+Cytokines are small proteins, work as messengers to pass signal across different cells, which are similar to hormones and neurotransmitter.
 
 An interaction map (Figure \ref{fig:map}) can be drawn according to the evidences found in experiments.\cite{ref:Tsarouchas}.
-After injury, a massive immune response was observed, with the increasing numbers of several kinds of cells.
-This dynamical system consists of complex interactions and dynamical feedback between cells and cytokines, For example, il-1$\beta$ promotes macrophages' increase, and as a major source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
+After the injury, a massive immune response was observed, with the increasing numbers of several kinds of cells.
+This dynamical system consists of complex interactions and dynamical feedback between cells and cytokines.
+For example, il-1$\beta$ promotes macrophages' increase.
+As a primary source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
 
@@
 The presence of the cytokines also affects the number of immune cells, while as sources, more immune cells produce more cytokines.
-Interactions between different types of cells are also considered in the system (e.g., the number of neutrophil is affected by macrophages).
+Interactions between different types of cells are also considered in the system (e.g., the number of neutrophils is affected by macrophages).
 
@@
     \caption[Working model of the influence of the innate immune system on axonal regrowth]%
-    {Working model of the influence of the innate immune system on axonal regrowth, from Tsarouchas et al.\cite{ref:Tsarouchas}.
-Lines ended with arrow represent promoting effect, lines ended with T-connectors represent inhibition}
+    {Working model of the influence of the innate immune system on axonal regrowth, from Tsarouchas et al.
+\cite{ref:Tsarouchas}.
+Lines ended with arrows represent promoting effect, lines ended with T-connectors represent inhibition}
     \label{fig:map}
@@
 
-This dynamic system is suitable for mathematical modelling, where a set of equations can be written to represent the dynamics of different objects e.g.
-each types of cells and cytokines.
-Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions inside the system, based on the proposed hypothesis \cite{ref:Tsarouchas}, and furthermore based our own hypothesis which may include different interactions.
+This dynamic system is suitable for mathematical modelling, where a set of equations can be written to represent the dynamics of different objects, e.g.
+each type of cells and cytokines.
+Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions inside the system, based on the proposed hypothesis \cite{ref:Tsarouchas},  furthermore based our hypothesis which may include different interactions.
 
-The biological background of works in this dissertation is mostly related the the experimental evidences, quantitative data and proposed hypothesis in \cite{ref:Tsarouchas}.
+The biological background of works in this dissertation is mostly related to the experimental evidence, quantitative data and proposed hypothesis in \cite{ref:Tsarouchas}.
 The published data on the quantitative measurement of cells and cytokines make the modelling and further analysis possible.
 Some other interactions proposed in our modelling stages are enlightened from Anderson et al.
-\cite{Anderson}, where intracellular cytokine signaling networks were established and computationally modelled.
+\cite{Anderson}, where intracellular cytokine signalling networks were established and computationally modelled.
 
@@
 Mathematical models can express real-life problems in mathematical forms, and provide qualitative or quantitative inside-look of the question through various mathematical tools and techniques.
-Mathematical models can be used for prediction, classification, testing our understanding or hypothesis and many other tasks, thus they are widely used to solve practical problems.
-However, for a model that describes the zebrafish spinal cord regeneration process, we were less interested to the model's predictive functionality; the main focus is on the explore of uncertainty in parameter estimates and comparison with alternative models.
-In this case, we wish to use modelling techniques to implement inference methods, test our hypothesis and hopefully find and compare well-fitted models.
+They can be used for prediction, classification, testing our understanding or hypothesis and many other tasks; thus, they are widely used to solve practical problems.
+However, for a model that describes the zebrafish spinal cord regeneration process, we were less interested in the model's predictive functionality; the main focus is on the exploration of uncertainties in parameter estimates and comparison with alternative models.
+In this case, we wish to use modelling techniques and implement inference methods, test our hypothesis and hopefully find and compare well-fitted models.
 
-According to our understanding of the biological mechanism in the regeneration, hypothesis should firstly proposed in a way that can represent the real  mechanism as much as possible.
-Usually a real-life problem, e.g.
-the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered; an elaborate description of the real cases using a overly complex mathematical model sometimes can be impossible; assumptions and hypothesis are often proposed to make necessary simplification and abstract for a final reasonably detailed model.
-Some qualitative models are simple but illustrative for the real-life observed patterns, e.g.
-Lotka-Volterra (L-V) equation for predator and prey models; they are usually analytically solvable and some properties are used to make predictions (e.g.
-steady states in ecosystem).
-If a more precise model is desired, usually more effects or interactions are considered and added as terms of equations to a re-constructed model, e.g.
+According to our understanding of the biological mechanism in the regeneration, hypotheses should be proposed in a way that can represent real mechanism as much as possible.
+Usually, a real-life problem, e.g.
+the spinal cord regeneration, consists of complex dynamics and interactions with many other environmental factors to be considered.
+An elaborate description of the real cases using an overly complicated mathematical model sometimes can be impossible; assumptions and hypotheses are often proposed to make necessary abstract and simplification for a final reasonably detailed model.
+Some qualitative models are illustrative but straightforward for the observed patterns, e.g.
+Lotka-Volterra (L-V) equation for predator and prey models.
+They are usually analytically solvable, and their properties are used to make predictions (e.g.
+steady states in the ecosystem).
+If a more precise model is desired, usually more influences or interactions are considered and added as terms of equations to a re-constructed model, e.g.
 considering environmental factors, a changing environmental capacity and predation from other species in L-V model.
 
-Models like L-V equation are mostly analytically solvable when analysing properties of the model, however, analytical solution for many other general models are not feasible, because of more complex dynamics, or the model itself is not a deterministic model.
+Models like the L-V equation are mostly analytically solvable when analysing properties of the model.
+However, analytical solutions for many other general models are not feasible because of more complicated dynamics, or the model itself is not a deterministic model.
 In the areas where solutions are not analytically feasible, computational techniques are usually applied for approximate solutions.
-Solvers for different models had been proposed to simulate the system or find roots for some spacial points, such as Explicit Runge-Kutta \cite{dormand1980family} and LSODA.
+Solvers for different models had been proposed to simulate the system or find roots for some special points, such as Explicit Runge-Kutta \cite{dormand1980family} and LSODA.
 
@@
 These solvers help to simulate the trajectories of the dependent variables, given initial values and a set of fixed parameters.
-The goodness of `fit' then can be measured by compared the simulated data and observed real data.
+The decency of fit can be measured by compared the simulated data and observed real data.
 
@@
 There are many approaches to derive these parameters from data or experience.
-For a popular topic and a widely used models, the parameters' values can be read from previous literatures; however in many others cases, we need to find parameter's values specifically for our model, and numerical values for parameters are mostly preferred, as they suitable for practical analysis and comparison.
+For a popular topic and widely used models, parameters' values can be read from previous literature.
+However, in many other cases, we need to find values explicitly for our model, where numerical values for parameters are mostly preferred as they are suitable for practical analysis and comparison.
 
 To find the best parameters' values given observed data, either optimisation-based approaches (e.g.
-least-square fit (LS), maximal likelihood estimation (MLE)) or inference-based approaches can be used.
+least-squares fit (LS), maximal likelihood estimation (MLE)) or inference-based approaches can be applied.
 Optimisation-based approaches may face two major drawbacks: over-fitting and local optima \cite{ref:abcsysbio}.
-Bayesian inference approaches gain popularity in recent years, as it can relief the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
-Local optima can still be a concern, and many efforts have be proposed for a more through explore of the parameter space to avoid local modes.
+Bayesian inference approaches gain popularity in recent years, as it can relieve the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
+Local optima can still be a concern, and many efforts have been proposed for a more thorough exploration of the parameter space to avoid local modes.
 
@@
 
-Alternatively, as $p(D)$ is fixed once the observed data is given, the rule can be rewritten as 
+Alternatively, as $p(D)$ is fixed once the observed data is provided, the rule can be rewritten as 
 \begin{align}
@@
 where $l$ represents the likelihood function.
-In other words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target in parameter estimation task.
+In other words, Eqn.
+\ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target in the parameter estimation task.
 Prior $p(\theta)$ should express our initial believes or estimates on the parameter values before observing the data and conducting experiments; likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability values.
 Once the data is given, $\theta$ is the only parameter in the likelihood function.
-Some optimisation-based methods try to find estimates of parameter set $\theta$ s.t.
-maximise the likelihood, however, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone calculating and comparing values of it.
+Some optimisation-based methods attempt to find estimates of parameter values $\theta$ s.t.
+maximise the likelihood.
+However, such methods will struggle when facing a problem where it is hard to specify an analytical expression for likelihood function, let alone calculating and comparing values of it.
 
-Based on Eqn \ref{eq:bayes2}, to find the posterior of parameter set, prior can be set according to the problem context and our understanding and cover a reasonable ranges; computational methods can be used to approximate likelihood function, e.g.
+Based on Eqn.
+\ref{eq:bayes2}, to find the posterior of the parameter set, prior can be set according to the problem context and our understanding and cover a reasonable range.
+Computational methods can be used to approximate likelihood function, e.g.
 Markov Chain Monte Carlo (MCMC) \cite{ref:MCMC}.
-As an alterative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood function, but focused on simulating the data from the given model and observed data, then used the simulated data to approximate the desired posterior.
+As an alternative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood function but focus on simulating the data from the given model and observed data, then use the simulated data to approximate the true posterior.
 ABC-rejection \cite{ABC_rejection} is a simple example of the ABC process, and further algorithms e.g.
@@
 \begin{enumerate}
-    \item Draw a sample $\theta^*$ from parameters' prior distribution. $\theta^*$ is a vector that consists values for each parameter
+    \item Draw a sample $\theta^*$ from parameters' prior distribution. $\theta^*$ is a vector that consists of values for each parameter
     \item Plug $\theta^*$ into the model, and use the model to generate simulated data $D^*$, where $D^*$ is of the same form as the observed data $D$, e.g. includes same summary statistics
@@
 In our study of the regeneration model, full trajectories of the dependent variables are used as the data to be compared, as our goal is to find a model with a parameter set that can intuitively recover the real process as much as possible.
-The mean values at each time point for each dependent variable is calculated to form $D$, and for each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$.
-Using other statistics to summary the raw data are also possible but care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
+The mean values at each time point for each dependent variable are calculated to form $D$.
+For each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$.
+Using other statistics to summarise the raw data are also possible, but care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
 
-Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of deceasing threshold values $\epsilon_t$, and derive new populations (called generation) of samples from the accepted samples (called particles) in previous population.
+Instead of fixing the threshold value, the algorithm can be more adaptive by using a sequence of decreasing threshold values $\epsilon_t$, and derive new populations (called generation) of samples from the accepted samples (called particles) in the previous population.
 An example is ABC SMC \cite{Toni}, which can significantly reduce the required samples compared to ABC-rejection \cite{ref:disease}:
@@
 An illustrative graph of the ABC SMC procedures can be found in Figure \ref{fig:smc}.
-In implementations, many factors can affect the efficiency and goodness of fit, some of them are discussed in the Chapter 4.
+In implementations, many factors can affect the efficiency and goodness of fit; some of them are discussed in Chapter 4.
 Rather than a pre-fixed threshold schedule, median or quantile based epsilon schedule can save the total sampling size and thus improve acceptance rates of particles \cite{threshold}.
 Some other improvements include using adaptive population size for each generation \cite{population}, using adaptive distance function \cite{ref:adpt_dis} and switching kernels \cite{ref:kernel}, etc.
-In our implementations these options were tried and compared when applicable and suitable.
+In our implementations, these options were tried and compared when applicable.
 
@@
 Instead of draw samples from prior of a single model, we can draw samples from multiple models $m_i$ and approximate $p(m_i, \theta|D)$, which is the joint probability distribution of model $m_i$ and its accepted particles $\theta$.
-In each generation the marginal distribution $p(m_i|D)$ can be calculated as the accepted ${\theta}$ is known, thus models can be ranked according to their probabilities.
+In each generation, the marginal distribution $p(m_i|D)$ is calculated as the accepted ${\theta}$ is known; thus models can be ranked according to their probabilities.
 
-The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modeling problems, especially in the area of biology \cite{ref:abcsysbio, ref:disease, ref:compare}.
-But ABC is not limited to these areas, as it can be generally used in inference tasks to give parameter estimates or model rankings.
+The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modelling problems, especially in the area of biology \cite{ref:abcsysbio, ref:disease, ref:compare}.
+However, ABC is not restricted to these areas, as it can be generally used in inference tasks to give parameter estimates or model rankings.
 E.g., it has been successfully applied in cosmology studies \cite{cosmology}.
@@
 
- A software packages that integrally implement ABC SMC is preferred, and more options in the settings of the algorithm, more customisable features make it better.
-Moreover, as the ABC SMC itself is a computationally intensive task that contains massive parallel potentials, an ideal software package is expected to support as least one kind of parallel options e.g.
+Software packages that integrally implement ABC SMC are preferred, and more options in the settings of the algorithm, more customisable features make it better.
+Moreover, as the ABC SMC itself is a computationally intensive task that contains massive parallel potentials, an ideal software package is expected to support as least one kind of parallel options, e.g.
 multi-core, GPU accelerating and clusters.
-Otherwise, a modification or rewrite of the software package are expected to produce a reliable input-output framework for our parameter estimation and model selection experiments, where much more work would be introduced.
+Otherwise, a modification or rewrite of the software package is expected to produce a reliable input-output framework for our parameter estimation and model selection experiments, where much more work would be introduced.
 
 Preliminarily \verb|pyABC| packages \cite{ref:pyabc} in \verb|Python| was chosen for implementations.
-\verb|pyABC| supports multiple types of parallelisation, e.g., multicore processing and distributed cluster, and supports resume stored runs, which made our experiments easier.
+\verb|pyABC| supports multiple types of parallelisation, e.g., multi-core processing and distributed cluster, and supports resume stored runs, which made our experiments easier.
 It is well documented and flexible in customisations of the algorithm.
@@
 These packages may have their advantages (e.g.
-ABC-SysBio supports CUDA acceleration) and considered as alternative implementation methods, or can be used to compare performance difference.
+ABC-SysBio supports CUDA acceleration) and considered as alternative implementation methods, or can be used to compare performance difference in future works.
 
- Additionally, some other softwares were also need for solving ODE, results analysis and visualisations. These software and hardware environments are listed in Appendix A.
+Additionally, some other software tools were also needed in ODE solving, result analysis and visualisations. The software and hardware environments used in this dissertation are listed in Appendix A.

### 2020-08-16 introduction/introduction.tex
Grammar fix

@@
 
-Zebrafish has the ability to regenerate its spinal cord after injury; tissue regeneration observed at the injury site is a complex biological process with interacting parts and dynamical feedback between different types of cells and molecules.
-These cells and molecules are involved in the inflammation and repair in a dynamical way and a complex interacting network can be drawn according to it.
+Zebrafish can regenerate its spinal cord after injury; tissue regeneration observed at the injury site is a complex biological process with interacting parts and dynamical feedback between different types of cells and molecules.
+These cells and molecules are involved in the inflammation and repair dynamically, and a complex interacting network can be drawn according to it.
 
@@
 
-Mathematical models and techniques can help us to resolve the complexity in this dynamic system, however, how to derive the parameter estimates of the models remains a challenging work.
-In this dissertation, our interest lies in the modeling of the regeneration process that happens around the lesion site, especially in the parameter estimates and comparison of possible models.
-According to available researches \cite{ref:Tsarouchas}, a mathematical model was firstly proposed but the addressed interactions remains to be precisely defined, and quantitatively examined in experiments.
-We would like to find the best parameter sets of the model using the published data, and ranking models under certain metrics and possibly find the best-fit model.
-In these steps, models are expected to retain mathematical and biological significance regarding the complex interactions and differential equations are suitable for this.
+Mathematical models and techniques can help us to resolve the complexity in this dynamic system; however, how to derive the parameter estimates of the models remains a challenging work.
+In this dissertation, our interest lies in the modelling of the regeneration process that happens around the lesion site, especially in the parameter estimates and comparison of possible models.
+According to available researches \cite{ref:Tsarouchas}, a mathematical model was first proposed, but the addressed interactions remain to be precisely defined, and quantitatively examined in experiments.
+We want to find the best parameter sets of the model using the published data, and ranking models under specific metrics and possibly find the best-fit model.
+In these steps, models are expected to retain mathematical and biological significance regarding the complex interactions, and differential equations are suitable for this.
 
@@
 
-Fo the parameter estimation tasks, optimisation-based methods usually have trouble in determining global optima from local optimal, and noise in the observed data may lead to over-fitting.
-Compared to optimisation-based methods, Bayesian inference techniques can help to relief these two concerns\cite{ref:abcsysbio}.
+For the parameter estimation tasks, optimisation-based methods usually have trouble in determining global optima from local optimal, and noise in the observed data may lead to over-fitting.
+Compared to optimisation-based methods, Bayesian inference techniques can help to relieve these two concerns\cite{ref:abcsysbio}.
 Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression for the dynamic system can be hard.
 
-Such methods introduce computation-intensive sampling-and-test patterns, where massive number of samples are drawn and followed by complex computations.
-As the sampling is independent from each other in a certain period, the parallel implementation of the algorithm becomes feasible and is expected to accelerate the parameter inference in a large scale when using HPC resources.
+Such methods introduce computation-intensive sampling-and-test patterns, where a massive number of samples are drawn and followed by complex computations.
+As the sampling is independent of each other in a certain period, the parallel implementation of the algorithm becomes feasible and is expected to accelerate the parameter inference on a large scale when using HPC resources.
 
@@
 Additionally, performance experiments illustrated the strong-scaling performance of the algorithm.
-Chapter 2 talks about the background information and theories, Chapter 3 shows how the mathematical models are derived from hypothesis and Chapter 4 describes the ABC SMC implementations and experiments on the models in detail.
+Chapter 2 talks about the background information and theories, Chapter 3 shows how the mathematical models are derived from hypothesis, and Chapter 4 describes the ABC SMC implementations and experiments on the models in detail.
 A brief scaling-up performance experiment is included in Chapter 5, and Chapter 6 summarises the conclusions and states the future works on this topic.

### 2020-08-16 performance/performance.tex
Grammar fix

@@
 
-    \caption[TODO]{Speedup. MORE DESCRIPTION}
+    \caption[Speedup of the program]{Speedup of the program}
     \label{fig:speedup}
@@
 
-    \caption[TODO]{Samples per second. MORE DESCRIPTION}
+    \caption[Average samples in a second, under different number of cores]{Average samples in one second, under different number of cores}
     \label{fig:sample_per_sec}
@@
 
+The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path.
+Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
+For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated in the local optimum.
+In this case, huge amount of sample will be drawn to find enough accepted particles, and new generation will be able to jump out of the local optimum and have different concentrations.
+
 \begin{figure}[ht]
@@
 
-    \caption[TODO]{Local modes. MORE DESCRIPTION}
+    \caption[Local optimum and abnormal pattern]{(a) Required sampling size for two example runs.
+Different colour represents different generation (from bottom to top: generation 1 to generation 20).
+Run A took huge amount of time in population 11.
+(b) Required sample in each generation of Run A.
+(c) Time-consumption in each generation of Run A}
     \label{fig:local_modes}
 \end{figure}
-
-The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path.
-Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
-For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated in the local optimum.
-In this case, huge amount of sample will be drawn to find enough accepted particles, and new generation will be able to jump out of the local optimum and have different concentrations.
 
@@
 
-    \caption[TODO]{Parameter density distribution. MORE DESCRIPTION}
+    \caption[Parameter density distribution before and after the concentration movement in generation 11]{Parameter density distribution before (generation 9, cyan) and after (generation 13, red) the concentration movement in generation 11}
     \label{fig:local_para}

### 2020-08-16 modelling/modelling.tex
Grammar fix

@@
 
-Mathematical models can describe different kinds of dynamic systems, and thus can be used in prediction, analysis abd other tasks.
+Mathematical models can describe different kinds of dynamic systems, and thus can be used in prediction, analysis and other tasks.
 An ideal model in our case, can represent reasonable interactions or effects between immune cells and cytokines and reproduce features and characteristics that were observed in experimental data.
 
-Proposed models for the regeneration process are in the form of non-linear ordinary differential equations (ODE).
-The only independent variable is post-lesion time $t$, and the equations consists of derivatives of dependent variables (number of cells, mRNA expression of cytokines) with respect to $t$.
+Proposed models for the regeneration process were in the form of non-linear ordinary differential equations (ODE).
+The only independent variable is post-lesion time $t$.
+The equations consist of four derivatives of dependent variables (number of cells, mRNA expression of cytokines) with respect to $t$.
 The equations were mostly an interpretation of the proposed interaction maps, with terms correspond to interaction paths.
@@
 
-Our models were built on the basis of the existing experimental findings from Tsarouchas et al.\cite{ref:Tsarouchas}.
+Our models were built based on the existing experimental findings from Tsarouchas et al.
+\cite{ref:Tsarouchas}.
 The available measurement data included the numbers of three kinds of cells (neutrophil $N$, macrophage $\Phi$ and microglia) and the relative mRNA expression of four kinds of cytokines (il-1$\beta$, tnf-$\alpha$, tgf-$\beta$1a and tgf-$\beta$3).
-As proposed by Tsarouchas et al., neutrophil and macrophage play important roles in the spinal cord regeneration with il-1$\beta$ and tnf-$\alpha$ being the media.
-According to this, our initial models focused on the changes of four dependent variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$) that are most important and determinative in the regeneration.
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ was recorded as `relative mRNA expression' (compared to the value at the initla time point i.e.
+As proposed by Tsarouchas et al., neutrophil and macrophage play essential roles in controlling the il-1$\beta$ and tnf-$\alpha$ mediated inflammation.
+Our initial models focused on the changes of four dependent variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$) that are most important and determinative in the regeneration.
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is recorded as `relative mRNA expression' (compared to the value at the initial time point, i.e.
 0 hpl) and considered unitless.
@@
 It was noted that the variance of the measured data is relatively high.
-The summary statistic used for the parameter estimations was mean of measurement data at each time point, assuming that measured data was Gaussian-like distributed.
-To validate this, the distribution of the measured data points was plotted and examined to see if the mean value can represent the distribution.
+The summary statistic was the mean of measurement data at each time point, assuming that measured data was Gaussian-like distributed.
+To validate this, the distribution of the measured data points was plotted.
 The result was that at most time points the measurement values are Gaussian-distributed, although some distributions were skewed.
 One abnormal distribution was observed at time point 120 hpl for macrophage where there are two concentrations.
-Despite this, mean value could summarise most data and thus was still used as the target observed data.
-A plot of the mean of the four variables is shown in Figure \ref{fig:obs_data}, which is our target trajectories for the models to fit.
+Despite this, the mean value could summarise most data and thus was still used as the target summary data.
+The mean value at each time point is shown in Figure \ref{fig:obs_data}, which is our target trajectories for the models to fit.
 
@@
     \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/obs_curve.png}}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/observed.png}}
     \end{center}
@@
     \caption[Mean of the observed data]%
-    {Mean of the observed data for neutrophil ($N$), macrophage ($\Phi$), il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}. Error bars indicate standard error of mean}
+    {Mean of the observed data for neutrophil ($N$), macrophage ($\Phi$), il-1$\beta$ and tnf-$\alpha$, from experiment results of \cite{ref:Tsarouchas}.
+Error bars indicate standard error of mean (SEM).
+For $N$ and $\Phi$ the SEM is relatively small}
     \label{fig:obs_data}
@@
 
-\section{Hypothesis and models}
+\section{Hypotheses and models}
 
-5 models in total are proposed according to different hypothesis.
-At first our tests and implementations of ABC SMC for parameter estimations use only the basic model for developing propose.
-After the parameter estimation framework is built and tested, more models are proposed, in order to calibrate and adjust the basic model such that it can represent the observed regeneration process better or can be used to test our hypothesis.
+Five models in total are proposed according to different hypotheses.
+At first, our tests and implementations of ABC SMC for parameter estimations used only the basic model for developing propose.
+After the parameter inference framework was built and tested, more models were proposed in order to calibrate and improve the basic model such that the observed regeneration process can be better represented.
 
-All these models assume the involved interactions is within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$) and use the data presented in Figure \ref{fig:obs_data} for the inference task.
-Interactions or effects from other cells or cytokines is not considered as there might not be corresponding data available.
+All these models assumed that the involved interactions are within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$), and used the data presented in Figure \ref{fig:obs_data} for the inference task.
+Interactions or influences from other cells or cytokines are not considered as there might not be corresponding data available.
+If such data or experimental findings are available in the future, we can consider them by then as long as these interactions are indeed necessary for our model.
 
@@
 An interaction map of the model is shown in Figure \ref{fig:m1}.
-This model is a simplification a the process described in \cite{ref:Tsarouchas} with some minor interactions ignored.
-To describe the parameters' units, we denote the unit of $N$ and $\Phi$ i.e.
-number of cells as `cell', and denote $\beta$ and $\alpha$'s relative mRNA expression as unitless for simplicity.
-This model included 12 parameters, all of which should be positive real number.
+This model is a simplification of the process described in \cite{ref:Tsarouchas} (Figure \ref{fig:map}) with some minor interactions ignored.
+This model included 12 parameters, all of which should be positive real numbers.
 A description of these parameters can be found in Table \ref{table:m1}.
@@
 overcrowding inhibits the increasing rates.
-As sources, macrophage produces tnf-$\alpha$ and neutrophil produces il-1$\beta$, however, macrophage was assumed to have negative effect on the il-1$\beta$, term $1+i_{\beta\Phi}\Phi$ in Eqn.
+As sources, macrophage produces tnf-$\alpha$ and neutrophil produces il-1$\beta$; however, macrophage was assumed to have a negative effect on the il-1$\beta$, term $1+i_{\beta\Phi}\Phi$ in Eqn.
 \ref{eq:model1} was used to represent this.
-Despite of self-driven increase of immune cells, il-1$\beta$ was assumed to promote the increase of both cells.
-Neutrophil was also assumed to be inhibited by the total immune cells presented in the lesion site, term $\nu_{N\Phi}N\Phi$ was used to denote this.
+Despite the self-driven increase of immune cells, il-1$\beta$ was assumed to promote the increase of both cells.
+Neutrophil was also assumed to be inhibited by the total immune cells presented in the lesion site, the term $\nu_{N\Phi}N\Phi$ was used to denote this.
 
@@
 
-As the observed data indicated, this dynamic system has a steady state where after a long time ($\geq 120$ hpl), the  inflammation is resolved and immune cells should not be present at the injury site any more.
-Regarding this, the parameter for self-increase rate for the macrophage and neutrophil i.e.
+As the observed data indicated, this dynamic system has a steady-state where after a long time ($\geq 120$ hpl), the inflammation is resolved, and immune cells should not be present at the injury site any more.
+Regarding this, the parameter for self-increase rate for the macrophage and neutrophil, i.e.
 $\lambda_\Phi$ and $\lambda_N$  cannot be constant.
@@
 Also, the inhibition of il-1$\beta$ production, i.e.
-term $i_{\beta\Phi}\Phi$ is considered to be ignored for a more simple model, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophil and the negative feedback from itself.
+the term $i_{\beta\Phi}\Phi$ is considered to be ignored for a more simple model, in which case the relative expression of il-1$\beta$ is only affected by the number of neutrophils and the negative feedback from itself.
 This case corresponds to model 3, written as Eqn.
@@
 Model 3 can be regarded as a simplification of model 2, as it can be treated as model 2 with parameter $i_{\beta\Phi}=0$.
-Among the proposed three models, model 1 is a naive one that was proposed at very first time and used as an ODE `template' to build and test our parameter inference framework.
+Among the proposed three models, model 1 is a naive one that was proposed at the very first time and used as an ODE `template' to build and test our parameter inference framework.
 As the implementation was successful, model 2 and 3 was proposed, as we were trying to calibrate some terms and find better models.
 After fitting, model 2 is supposed to be better than model 1 as it corrects the problem that appears at the final time points (which were related to the steady states of immune cells).
-Model 3 made a small simplification based on model 2 and thus it was theoretically less `general' than model 2.
+Model 3 made a small simplification based on model 2, and thus it was theoretically less `general' than model 2.
 
@@
 
-After the first model selection experiment among the exiting 3 models, it was found that some significant features presented in the observed data were not recovered by any of the model.
+After the first model selection experiment among the exiting three models, it was found that any of the models did not recover some significant features presented in the observed data.
 To resolve this, attempts were tried to introduce additional reasonable interactions within the dynamic system, considering the biological and mathematical context.
-Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term $d_{\beta\alpha}\beta$ (which means the same effect as directly promoting but the underlying mechanism is unclean), or adding a term $f_{\beta\alpha}$ that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
+Extra promoting effect to the expression of tnf-$\alpha$ was considered, by either adding a phenomenological term $d_{\beta\alpha}\beta$ (which means the same effect as directly promoting but the underlying mechanism is unclear), or adding a term $f_{\beta\alpha}$ that represents a promoting effect to the production process of tnf-$\alpha$, namely model 4 (Eqn.
 \ref{eq:model4}) and model 5 (Eqn.

### 2020-08-16 implementation/implementation.tex
Grammar fix

@@
 
-According to our preliminary studies, a ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
+%  which is of our interests and suitable for the followed performance experiments.
+
+According to our preliminary studies, an ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project.
 \verb|pyABC| is a popular SMC software packages\cite{ref:pyabc} which has been used form many ABC SMC inference applications\cite{population}.
-\verb|pyABC| provides an open-source framework for likelihood-free inference, which is a \verb|NumPy| implementation of ABC SMC algorithm and a useful tool-box for our own inference and analysis tasks.
+\verb|pyABC| provides an open-source framework for likelihood-free inference, which is a \verb|NumPy| implementation of ABC SMC algorithm and a useful tool-box for our inference and analysis tasks.
 Besides, it supports multi-core execution implemented using \verb|multiprocessing| built-in package, and it made our further studies in the scaling-up performance possible.
 
-%  which is of our interests and suitable for the followed performance experiments.
+The inference framework code for the project was developed and tested in a local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first using a small input, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection tasks.
+The scaling-up experiment was performed on Cirrus.
+The profiling and its analysis were performed on a local machine.
+Some necessary development and debug were performed on the remote machines to ensure that experiments run correctly on the compute node.
 
-The inference framework code for the project was developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first using a small input, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection tasks.
-The scaling-up experiment was performed on Cirrus.
-The profiling and its analysis were performed on local machine.
-Some necessary development and debug were performed on the remote machines to ensure that experiments run correctly on compute node.
-
-The ODE model-related code and utilities were firstly developed to enable reusable functions, variables and data calls.
+The ODE model-related code and utilities were first developed to enable reusable functions, variables and data calls.
 It included the code for the ODE equations, ODE solver, data structures and data format transform functions, etc.
@@
 
-Some missing features were found when using \verb|pyABC|, therefore some modifications were made to the source code of \verb|pyABC|.
+Some missing features were found when using \verb|pyABC|; therefore some modifications were made to the source code of \verb|pyABC|.
 
-For example, the laboratorial measured data were not considered `complete' or `tidy' for modeling and analysis purpose: the available time points for cells and cytokines were different, e.g.
+For example, the laboratory-measured data were not considered `complete' or `tidy' for modelling and analysis purpose: the available time points for cells and cytokines were different, e.g.
 cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl.
 These missing values were treated as `ignored' when calculating the distance between observed data and simulated data, and represented by \verb|NaN|s.
-However, in some distance calculations \verb|NaN| was not acceptable; to cope with this, built-in distance function was modified.
-Other modified code include some visualisation and result representation enhancements
+However, in some distance calculations, \verb|NaN| was not acceptable.
+To cope with this, the built-in distance function was modified.
+Other modified code included some visualisation and result representation enhancements.
 
 Analysis of the program output was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results.
-For different tasks there are also separated analysis code file.
+For different tasks, there are also separated analysis code file.
 Other analysis tools used in the project include built-in database visualisation tool, R code and Microsoft Excel.
@@
 After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (using synthetic data), and code files for parameter inference (using experimental data).
-Model selection can be easily done by adding more model to the implementation code file.
+Model selection can be easily performed by adding more model to the implementation code file.
 
@@
 
-As the key focus of this project was on the parameter estimations of dynamical system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental observed data.
+As the key focus of this project was on the parameter estimations of a dynamic system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental data.
 
-In the first part, synthetic data is used with known true parameter values.
-The algorithm implementation is evaluated by the efficiency and goodness of resultant model under different implementation options and ABC SMC hyperparameters.
+In the first part, synthetic data (Table \ref{table:known_values}) was used with known true parameter values.
+The algorithm implementation was evaluated by the efficiency and goodness of the resultant model under different implementation options and ABC SMC hyperparameters.
 
-Then according to findings in the first part, several options with good performance are tried in the inference using experimental observed data (Figure \ref{fig:obs_data}).
-To obtain a more accurate and general results, these experiments are repeated for 3 or more times.
+Then according to findings in the first part, several options with considerable performance were tried in the actual inference with observed data (Figure \ref{fig:obs_data}).
+To obtain more accurate and general results, these experiments were repeated for three or more times.
 
@@
 
-The ABC SMC inference can have largely different performance when using different hyperparameters and implementation options.
-To firstly test the ability of inference and observe the results of different options, synthetic data with known true parameter values are used as target data.
-by doing this, we could directly compare the inferred parameters to the true value, and compare the simulated trajectories to the true trajectories and then observe the result to see whether the inference is successful and how well the model fit the data.
-Also, by setting the same final threshold value, the efficiency under different options can be compared and help us to choose proper options when applying ABC SMC on the observed data.
-The true value of parameters that were used to generate synthetic data is listed in Table \ref{table:known_values}.
+    The ABC SMC inference can have widely different performance when using different hyperparameters and implementation options.
+To firstly test the ability of inference and observe the results of different options, synthetic trajectories with known true parameter values were used as target data.
+By doing this, we could directly compare the inferred parameters to the true values and compare the simulated trajectories to the true trajectories, then observe the result to see whether the inference is successful and how well the model fit the data.
+Also, by setting the same final threshold value, the efficiency under different options can be compared and help us to choose proper options when applying ABC SMC.
 
-The true values were obtained by fitting model 1 onto the experimental data (Figure \ref{fig:infer_back_data}) using least-square optimisation method.
+    The true values (Table \ref{table:known_values}) were obtained by fitting model 1 onto the experimental data (Figure \ref{fig:infer_back_data}) using least-square optimisation method.
 Available least-square fit tools in \verb|SciPy| were used.
-The fit algorithm was Trust Region Reflective\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.least_squares.html}}.
+The fitting algorithm was Trust Region Reflective\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.least_squares.html}}.
 Target data was then generated using these parameter values via ODE integrate solver\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.odeint.html}}.
@@
 
-The following topics were studied using the synthetic data and ABC SMC inference framework.
+The following topics were studied using the synthetic data and the developed ABC SMC inference framework.
 
@@
 We keep sampling until $N$ particles are accepted, where $N$ is the pre-set population size.
-After that the new weights are calculated and normalised.
+After that, the new weights are calculated and normalised.
 Figure \ref{fig:smc} illustrates the process.
 
-% \begin{figure}[t!]
-%     \begin{center}
-%         \resizebox{1.0\hsize}{!}{\includegraphics{fig/smc.png}}
-%     \end{center}
+The perturbation kernel is called in the sampling of every particle and determines how the new perturbed particle is chosen; thus it is influential to both computation complexity and the resulting posterior distribution.
+Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations or could generate a population with a higher acceptance rate if the successive epsilon values are close.
+A kernel with a wide variance, or spreading out in a large space could help in resolving the local optima problem by a more thorough exploration of the parameter space.
+However, it can be more computation-intensive and result in a lower acceptance rate.
+A desired optimal kernel should balance the trade-offs; their properties and criteria were discussed in \cite{ref:kernel}.
 
-%     \caption[ABC SMC sampling process]%
-%     {ABC SMC sampling process.
-$\theta^*$ denotes a sample drawn from previous population, which is used to generate simulated data $D^*$.
-$d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data.
-for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
-%     \label{fig:smc}
+There are several common choices of perturbation kernels.
+Among them, the multivariate normal kernel and local M-nearest neighbour model is preferred to be applied to our models.
+A covariance matrix $\Sigma_t$ of accepted particles is calculated from the previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$.
+It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in reflecting the true posterior structure \cite{ref:kernel}.
+It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks.
+The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `thinner-distributed' perturbation result.
 
-% \end{figure}
+Local M-NN kernel provided by \verb|pyABC| was also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.
 
-The perturbation kernel is called in the sampling of every particles and determines how the new perturbed particle is chosen, thus it is influential to both computation complexity and the resultant posterior distribution.
-Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations, or could generate a population with a higher acceptance rate if the successive epsilon values are close; A kernel with wide variance, or spreading out in a large space could help in resolving the local optimal problem by a more thoroughly exploration of the parameter space, however it can be more computation-intensive and result in a lower acceptance rate.
-A desired optimal kernel should balance the trade-offs; their property and criteria is discussed in \cite{ref:kernel}.
-
-There are several common choice of perturbation kernels.
-Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.
-A covariance matrix $\Sigma_t$ of accepted particles is calculated form previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$.
-It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in relecting the true posterior structure \cite{ref:kernel}.
-It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks.
-The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `narrower distributed' perturbation result.
-
-Local M-NN kernel provided by \verb|pyABC| is also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.
-
-The kernel experiments is designed to explore the efficiency of SMC on our dynamical systems.
-Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rate would be our preference.
+The kernel experiment is designed to explore the efficiency of SMC on our dynamical systems.
+Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rates would be our preference.
 The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters.
 The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference.
-As the threshold schedule is fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit i.e.
-recover of target data trends/features and errors of the inferred parameters compared to true values is not discussed.
+As the threshold schedule was fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit, i.e.
+recover of target data trends/features and errors of the inferred parameters compared to true values were not discussed.
 
@@
 
-To compare kernels, a fixed schedule schedule is used with minimal epsilon i.e.
-$\epsilon_t$ set to 10.0.
-From the total required samples graph Figure \ref{fig:kernel1}, the efficiency can be compared.
+To compare kernels, a fixed schedule was used with the final epsilon $\epsilon_t=10.0$.
+From the total required samples plot (Figure \ref{fig:kernel1}) the efficiency can be compared.
 Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
@@
 
-For local M-NN kernels, generally a greater M will lead to lower acceptance rate and
-more required particles in each generation.
+For local M-NN kernels, generally, a higher M leads to lower acceptance rates and more required particles in each generation.
 As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve.
-Consequently, if greater M is test, then it will have even lower acceptance rates; the maximum M=2000 (whole population is considered) will have the lowest acceptance rates.
+Consequently, if greater M was tested, then it would have even lower acceptance rates; the maximum M=2000 (the whole population is considered) would have the lowest acceptance rates.
 
@@
 Multivariate normal kernel has a performance between local M-NN M=250 and M=100.
-It proves that rather than a trivial normal kernel, multivariate normal kernels is more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
-`Scaling' option can narrower the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
-Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of the model 1.
-Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ is also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
+It proves that rather than a trivial normal kernel, multivariate normal kernels are more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}.
+`Scaling' option can narrow the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance.
+Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of model 1.
+Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ was also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
 
-A more `local' kernel was more likely to give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
-This holds in most cases under a given target final threshold value $\epsilon_t$, however may not produce the proper result want: a more local kernel, e.g.
-local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted under given threshold.
-In other words, using local kernels the epsilon may quickly converge to to local optimal; if a even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
-Also, the shapes of posterior distribution will affect the performance of local kernels \cite{ref:kernel}.
+A more `local' kernel is more likely to give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
+This holds in most cases under a given target final threshold value $\epsilon_t$.
+However, it may not produce the proper result want: a more local kernel, e.g.
+local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted before even smaller threshold values is demanded (see Section 5.2 for more discussions).
 One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
 
-\subsubsection{Conclusion} Local kernels can be fast bur unstable in some case; for a general parameter inference of our models, multivariate kernels are preferred, as a good fit is our prior target; some local kernels are also worth trying after the multivariate normal kernel, e.g.
-local M-NN with M$\leq 250$ (for population size of 2000, i.e.
-12.5\% of the population) and multivariate normal with scaling.
-Also, multivariate M-NN is worth trying, although it was not built in \verb|pyABC| and requires additional implementations.
+% In other words, by using local kernels the epsilon may quickly converge to a local optimum; if an even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.
+
+% Also, the shapes of posterior distribution affect the performance of local kernels \cite{ref:kernel}.
+
+To conclude, Local kernels can be fast but unstable in some case.
+For a general parameter inference of our models, multivariate kernels were preferred, as a good fit is our prior target.
+Some local kernels are also worth trying after the multivariate normal kernel, e.g.
+local M-NN with M$\leq 250$ (for a population size of 2000, i.e.
+12.5\% of the population) and multivariate normal with scaling, but extra cares must be taken: although they are fast, moving in and moving out of local optimum make them more unstable in efficiency.
+
+% Also, multivariate M-NN is worth trying, although it was not built in \verb|pyABC| and requires additional implementations.
 
@@
 $\Delta x_i = x_{i, simulated}-x_{i, observed}$.
-In this case data points are 12 time points of four variables i.e.
+In this case, data points are 12 time points of four variables, i.e.
 48 data points in total.
@@
 where a weight $w_i$ is assigned to every data point.
-If data point $i$ have a higher weight, then the data value is regeared to be more informative, i.e.
-giving more help in inferring the true posteriors; weights can be either pre-set according to prior knowledge of the problem, or using adaptive method to be dynamically calculated according to , known as adaptive distance.
+If data point $i$ has a higher weight, then the data value is regeared to be more informative, i.e.
+giving more help in inferring the true posteriors.
+Weights can be either pre-set according to prior knowledge of the problem or updated using an adaptive method.
+This method is known as adaptive distance.
 
@@
 Factor is helpful as an option for efficiency.
-When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data point, or behave like a normalisation that can balance the scales of different summarised statistics (here is the 48 mean values).
-There are two appliance case studied: (1) factors used as an normalisation option and (2) factors used to give more focus on some features of the observed data.
+When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data points, or behave like a normalisation that can balance the scales of different summarised statistics (here are the 48 mean values).
+There are two appliance case studied: (1) factors used as a normalisation option and (2) factors used to give more focus on some features of the observed data.
 For (2), as we noticed that the main features (e.g.
-rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points i.e.
-0 - 30 hpl, factors are tried in the later parameter estimation of real data where a more accurate fit of the curve is desired.
+rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points, i.e.
+0 - 30 hpl, factors were tried in later experiments.
+Such factor scheme assigns data in the first half of the time point with higher factors (75) and data in the second half with lower factors (25).
 
 Adaptive population \cite{population} can adapt the population size of each generation according to the mean correlation of variation error of the previous population.
-In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.
+However, in our early tests,  the maximal allowed population size is set to 10,000, and the adaptive strategy always selected the upper bound (10,000) for new populations, as a result of wide parameter space or the wide posterior approximation shape.
+Additionally, some internal bugs were observed when applying it.
+Hence, the adaptive population was not tried in this project; however, it is worth trying in future works as an efficiency improvement.
 
@@
 
-Two types of factor and adaptive distance function were tested by running the same number of generations (20 generation, 2000 particles per generation).
-The first type of factor is 25:75  factor, where first half of the trajectory data was assigned higher factor (75), and second half of the data was assigned lower factor (25); this could make the first half if the data more `important' in the inference.
-The second factor was range factor, which tried to normalise the trajectory of each dependent variable according to its data range.
+Two types of factor and adaptive distance function were tested by running the same number of generations (20 generations, 2000 particles per generation).
+The first type of factor is 25:75  factor, where the first half of the trajectory data were assigned higher factor (75), and the second half of the data was assigned lower factor (25); this could make the first half of the data more `important' in the inference.
+The second factor was `range' factor, which tried to normalise the trajectory of each dependent variable according to its data range.
 
 From Figure \ref{fig:factor}, different types of factor gave different performance.
-25:75 factor required much more samples in the first generation (orange bar in the left of Figure \ref{fig:factor}) and also did not improve much in later generations, thus it is the slowest one.
+25:75 factor required much more samples in the first generation (orange bar in the left of Figure \ref{fig:factor}) and also did not improve much in later generations; thus, it is the slowest one.
 The first half of observed data contained more features and information, assigning it with higher factor would make the algorithm think that it is more important to fit the first half and thus focus on the hard part, and consequently more particles were needed in the inference process.
@@
 By contrast, range factor required less total samples and gave generally higher acceptance rates than standard, which made it considerably faster.
-Adaptive distance gave the best efficiency, as it significantly improved the acceptance rates and resulted in even less required samples (right, Figure \ref{fig:factor}).
+The adaptive distance gave the best efficiency, as it significantly improved the acceptance rates and resulted in even less required samples (right, Figure \ref{fig:factor}).
 
@@
 However, the result of factors adaptive distance could not be directly compared to standard run, because it changed the distance measurement by adapting weights or manually setting factors.
-Factors were normalise to make it more close the standard distance measurement.
+Factors were normalised to make it closer to the standard distance measurement.
 As a result, the same final threshold value did not guarantee that adaptive distance can reach the same level of fit as others.
 After 20 populations, adaptive distance had a poor fit (Figure \ref{fig:factor_sim}), while the standard run and two types of factors had similar tight fits of the data, which all converge to the true posterior.
-Additional in our implementations, variance factor (which tries to normalise the data according to the dependent variables' variance) was also tries and gave very similar performance as range factor.
+Additional in our implementations, variance factor (which tries to normalise the data according to the dependent variables' variance) was also tried and gave a very similar performance as the range factor.
 
-To conclude, because the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, the direct compare of the efficiency under the same threshold schedule was unfeasible; we could only compare the results after the same number of generations.
-Factors gave nearly the same inferred results but required different number of samples, hence cares must betaken when choosing factor schemes.
-Adaptive distance did not result in a accurate result as others.
-As a result, our further implementations used standard distance function, and factors were tried for possible improvements.
+To conclude, because the factors and weights are directly multiplied to $\Delta x$, and thus the metrics for distance are changed, the direct comparison of the efficiency under the same threshold schedule was unfeasible; we could only compare the results after the same number of generations.
+Factors gave nearly the same inferred results but required a different number of samples, hence cares must be taken when choosing factor schemes.
+Adaptive distance did not provide an accurate result as others.
+As a result, our further implementations used standard distance function, and factors were also tried for possible improvements.
 
@@
 
-Regarding the existing dynamic systems model, SMC was applied with two data size options, three prior distribution (uniform distribution, wider uniform distribution and log-uniform distribution).
-The population options i.e.
-population size and number of populations were also tried with different values.
-These experiments intended to give suggestions on the later SMC implementation on experimental data.
+Regarding the existing dynamic systems model, SMC was applied with two data size options and two ranges of the prior interval.
+The population options, i.e.
+population size and the number of populations were also tried with different values.
+These experiments intended to give suggestions on the later implementation on experimental data.
 
@@
 
-The results from data size and prior distribution range is shown in Figure \ref{fig:size}.
+The results from data size and prior distribution range are shown in Figure \ref{fig:size}.
 The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement.
 Although much less (60\%) data was used, the total required sampling numbers was only 4.47\% less.
-This indicated that the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data, as they finally reached a similar epsilon threshold ((b), Figure \ref{fig:size}), but cares should still be taken when using less data point, or using other more summative statistics (e.g.
-mean, standard deviation, peak values, starting and ending values etc.) as the target data: as shown in \cite{ref:disease}, less data can lead to a model with more variance, under-fitting or missing some local features such as local peaks.
+This indicated that the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data, as they finally reached a similar epsilon threshold (Figure \ref{fig:size}), but cares should still be taken when using less data points, or using other more summative statistics (e.g.
+mean, standard deviation, peak values, starting and ending values etc.): as shown in \cite{ref:disease}, fewer data can lead to a model with more variance, under-fitting or missing some local features such as local peaks.
 
@@
 (a)).
-Samples were drawn from a high dimensional parameter space in each sampling process, a wider parameter prior interval in only one dimension can cause the parameter space to be much bigger, and consequently the exploration and sampling of parameter vectors would take much more efforts.
-Figure \ref{fig:size}.
-(b) also indicates that the convergency is slower for wider prior ranges.
-As a result, a proper prior interval based on our experience and belief with seeing the data should be set as narrow as possible, as inference are largely affected by the prior interval and its length; improper prior could even make the algorithm miss the true posterior \cite{ref:abcsysbio}.
+Samples were drawn from a high dimensional parameter space in each sampling process.
+A wider parameter prior interval in only one dimension can cause the parameter space to be much bigger.
+Consequently, the exploration and sampling of parameter vectors would take much more efforts.
+Right plot in Figure \ref{fig:size} also indicates that the convergency is slower for wider prior ranges.
+As a result, a proper prior interval based on our experience and belief with seeing the data should be set as narrow as possible, as the inference is largely affected by the prior interval and its length.
+Improper prior could even make the algorithm miss the true posterior \cite{ref:abcsysbio}.
 
 How to specify prior can be tricky from problem to problem.
-For our task in the parameter inference of regeneration model, biologically motivated priors were used, which were general reflection of biological constrains.
-Also the units of the parameters help us in identify the possible scale of the parameters.
+For our task in the parameter inference of regeneration model, biologically motivated priors were used, which were a general reflection of biological constraints.
+Also, the units of the parameters helped us in identifying the possible scale of the parameters.
 
@@
 
-This subsection aims to preliminarily explore options in ABC SMC implementations.
+This subsection aims to explore options in ABC SMC implementations preliminarily.
 From the above results, we adopted multivariate normal kernel as the kernel for our real inference on observed data; Local M-NN kernels were also planned to be tested.
-In the real inference, 25:75 factor were also tried to see if it could help in recovering more features and resulting a better fit.
-because the improvement by using less data was inapparent, we did not try to change the target summary data size.
+In the real inference, 25:75 factor were also tried to see if it could help in recovering more features and resulting in a better fit.
+Because the improvement by using less data was inapparent, we did not try to change the target summary data size.
 Prior intervals were explored explicitly and gradually shrunk for our models.
 
-Some other hyperparameter, e.g.
-number of generations, population size (i.e.
-number of particles per generation) were not addressed in this subsection.
+Some other hyperparameters, e.g.
+the number of generations, population size (i.e.
+the number of particles per generation) were not addressed in this subsection.
 In the real inference, population size was set to be greater than 2000 to avoid insufficient exploration of the previous population; in some case than high variance or uncertainties in the results were observed, population size as large as 20,000 were used.
@@
 
-Using the suggested options in the previous section, the target data (Figure \ref{fig:obs_data}) is prepared as input for the SMC inference framework.
-Regarding the result, the population size and number of generations are adjusted several times to obtain an general informative posterior which should be stable across repeated runs and avoid local optimal as much as possible.
+Using the suggested options in the previous section, the target data (Figure \ref{fig:obs_data}) was prepared as input for the SMC inference framework.
+Regarding the result, the population size and number of generations were adjusted several times to obtain a general informative posterior which should be stable across repeated runs and avoid local optima as much as possible.
 
-Parameter estimation of model 1, 2 and 3 was tried with different prior distribution setting: distribution range [0, 25] and [0, 75], amd distribution type uniform and log-uniform.
-All these ABC SMC runs uses 2,000 particles per population and 30 generations.
+Two prior intervals ([0, 25] and [0, 75]) and two distribution shape (uniform and log-uniform) were tried in the parameter estimation of the first three models.
+All these ABC SMC runs used 2,000 particles per population and 30 generations.
 
-Log-uniform-shaped prior seems to give a better fit of the model, and wider prior range [0, 75] does not give a more accurate result.
-Then distribution range [0, 25] was tried, in case some parameter have a true posterior laying between [25, 50].
+In practice, we found that log-uniform-shaped prior gave a better fit of the model than uniform-shaped prior, and wider prior range [0, 75] does not give a more accurate result than [0, 25].
+Then prior interval [0, 50] was tried, in case some parameters have a true posterior laying between [25, 50].
 
-The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}.
-For model 3, the acceptance rates and epsilon path are presented in Figure \ref{fig:result123_2}.
-For model 3, the estimated parameters' posterior is shown in Figure \ref{fig:para1}.
-Figure \ref{fig:result123} plots the estimated curve of models, using the mean value of each parameter's approximated posterior; also 1000 particles (each particle is a parameter set) were sampled and used to generate simulated trajectories of the four variable, and the mean (blue) and 25th to 75th percentile range (grey) are calculated from the 1000 simulated data.
+The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}, which is considered to give a better fit than other prior options.
+For these models, the acceptance rates and epsilon path are presented in Figure \ref{fig:result123_2}, and the estimated parameters' posterior of model 3 is shown in Figure \ref{fig:para1}.
+Figure \ref{fig:result123} plots the simulated trajectories of the three models.
+One thousand particles (each particle is a parameter set) were sampled, and the mean (blue) and 25th to 75th percentile interval (grey) were calculated from the data simulated by these 1000 particles.
 
 Before applying model comparison methods, some features of the three model can still be compared.
-From simulated curves, It can be seen that all the inter-quartile ranges are tight around the mean value curve, which indicates the approximated posterior are concentrated to some degree; the peak-shaped posterior distributions (Figure \ref{fig:para1}) agrees with this.
-All epsilon values converges to a steady level, and model 3 has a lower final epsilon value and consequently the simulated trajectory is more close to the observed data (model 3, Figure \ref{fig:result123}).
+From simulated curves, It can be seen that all the inter-quartile ranges are tight around the mean value curve, which indicates the approximated posterior are concentrated; the bell-shaped posterior distributions with a single peak (Figure \ref{fig:para1}) agrees with this.
+All epsilon values converge to a steady level, and model 3 has a lower final epsilon value, and consequently, its simulated trajectory is more close to the observed data (model 3, Figure \ref{fig:result123}).
 
-The acceptance rates is fluctuating and have a gradually increase trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated.
-Compared to model 1, model 2 gives a better fit of the decreasing trend at the second half of the observed data, by using a exponentially decaying self-increase rate $\lambda_N$ instead of a constant.
-The resultant simulated data from model 2 and 3 are in the similar trends and the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'.
-We expect model 3 to be the best model as it reaches a threshold value and intuitively gives a better fit, although the fit could not be considered generally a good representation of the biological process that we want to model, as the simulated data are significantly biased from the observed data for tnf-$\alpha$.
-A sharp peak at 4 hpl is observed from the measurement (Figure \ref{fig:obs_data}) but no similar features are represented by any of the three models.
+The acceptance rates are fluctuating and have a gradually increasing trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated.
+Compared to model 1, model 2 gives a better fit for the decreasing trend in the second half of the observed data.
+This proves that using an exponentially decaying self-increase rate $\lambda_N$ instead of a constant term indeed help to improve the model.
+
+The resultant simulated data from model 2 and 3 have similar trends, while the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'.
+We expected model 3 to be the best as it reached a lower threshold value, and intuitively gives a better fit.
+However, the fit of all these models could not be considered as a generally good representation of the biological process that we want to approximate: for tnf-$\alpha$, the simulated data were significantly biased from the observed data.
+A sharp peak at 4 hpl was observed in the experimental data, but none of the three models represents similar features.
 
@@
 
-[model comparison among model 1, 2, 3]
+% [model comparison among model 1, 2, 3]
 

### 2020-08-16 modelling/modelling.tex
Grammar fix

@@
 
-Some limitations were included in Section 3.1, where the observed data have a relative high standard deviation at certain measurement points; more measurement repeats or more measurement points could help to represent more accurate trajectories of the modelled variables and thus more possible models can be proposed.
-On the other hand, more accurate observed are alway beneficial to mathematical modelling, especially to complex dynamic systems.
+Some limitations were included in Section 3.1, where the observed data have a relatively high standard deviation at certain measurement points.
+More measurement repeats or more measurement points could help to represent more accurate trajectories of the modelled variables, and thus more possible models can be proposed.
+On the other hand, more accurate observed data are always beneficial to mathematical modelling, especially to complex dynamic systems.
 
 As stated, our models are largely based on the experimental evidence from Tsarouchas et al.\cite{ref:Tsarouchas}.
-Some hypothesised interactions may differ from real case, and there may exist undiscovered mechanisms.
+Some hypothesised interactions may differ from a real case, and there may exist undiscovered mechanisms.
 These could lead to model misspecification, where the model could fit the data but with wrong interaction paths, or the model can hardly fit the data and recover some local features observed in the data.

### 2020-08-16 implementation/implementation.tex
Grammar fix

@@
 
-Next a model selection experiment with different prior was conducted, using the same ABC-SMC framework.
+Next, a model selection experiment with different prior was conducted, using the same ABC-SMC framework.
 As in the parameter inference, two distributions (uniform and log-uniform) are tested with three different interval ([0, 25], [0, 50] and [0, 75]).
-The results show that model 3 wins with nearly 100\% model probability at most runs after 30 generations, except log-uniform [0, 75].
+The results show that model 3 wins with 100\% model probability after 30 generations for most runs, except log-uniform [0, 75].
 All the model probability results are of great discrepancy (e.g.
@@
 
-Figure \ref{fig:cmp1} shows the model selection process, under log-uniform [0, 50] prior distribution.
-Model 3 gains total advantage over model 1 and 2 after 7th population.
-Together with conclusions above, model 3 is considered to be the the best model to represent the target data so far.
+Figure \ref{fig:cmp1} shows the model selection process, under log-uniform [0, 50] prior.
+Model 3 gains total advantage over model 1 and 2 after the seventh population.
+Together with the conclusions above, model 3 is considered to be the best model to represent the target data so far.
 
@@
 
-Compared to uniform distributed prior, we found that log-uniform distributed prior is more suitable in the inference.
-Log-uniform results in an obviously better fit and narrower inter-quantile interval (Figure \ref{fig:result123} and \ref{fig:resultCurve_uni}), and the final epsilon value is also smaller.
+Compared to uniform-distributed prior, we found that log-uniform distributed prior is more suitable in our parameter inference.
+Log-uniform resulted in an obviously better fit and narrower inter-quantile interval (Figure \ref{fig:result123} and \ref{fig:resultCurve_uni}), and the final epsilon value was also smaller.
 Log-uniform tends to assign values that are close to the left boundary (i.e.
-zero in our case) higher density, i.e.
+zero in our case) with higher density, i.e.
 they are more likely to be sampled in the first population.
-As a result, the estimated parameter posteriors are expected to have smaller mean and/or skewed to left of the x-axis.
-The estimated values proved this (see \ref{table:estimated1}; also observed in other models and prior interval experiments).
-It indicates that some of the true parameter tend to have small values that are close to zero, and using log-uniform distribution could results in a faster and more satisfying inference.
-Regarding this, further experiments all use log-uniform distribution only.
+As a result, the parameter estimates are expected to have smaller mean values or skewed to the left of the x-axis.
+The estimated values proved this ( \ref{table:estimated1}).
+This was also observed in other models and prior interval experiments.
+It indicated that some of the parameters tend to have small true values that are close to zero, and using log-uniform prior could result in a faster and more satisfying inference.
+Regarding this, further experiments all used log-uniform prior only.
 
@@
 
-[features that derive model 4 and 5]
+% [features that derive model 4 and 5]
 
-We observed that the existing models cannot well represent the trajectory of relative expression of tnf-$\alpha$: from experimental data, a rapid increase of tnf-$\alpha$ expression is observed in 0--4 hpl, followed by a rapid drop; all the models cannot well recover this feature and most test results see a less-rapid increase followed by a gradually decree and the observed peak value is not reached (e.g.
-Figure \ref{fig:result123}).
-Also in some other case e.g.
-Figure \ref{fig:resultCurve_uni}, tnf-$\alpha$ trajectories is well fitted at the cost of under-fitting $\Phi$.
-Hence efforts had been spent to propose more alternative models that can possibly solve this problem.
+We observed that the existing models could not well represent the trajectory of relative expression of tnf-$\alpha$.
+From experimental data, a rapid increase of tnf-$\alpha$ expression was observed in 0--4 hpl, followed by a rapid drop.
+All the models cannot well recover this feature, and most results saw a less-rapid increase followed by a gradual decrease, while the observed peak value was not reached (Figure \ref{fig:result123}).
+Also, in some other case, e.g.
+uniform runs (Figure \ref{fig:resultCurve_uni}), tnf-$\alpha$ trajectories was well fitted; however, at the cost of under-fitting $\Phi$ curve.
+Hence efforts had been spent to propose more alternative models that can potentially solve this problem.
 
-The two proposed alternative models is based on model 3 and try to add extra term in the tnf-$\alpha$ equation $\mathrm{d} \alpha/\mathrm{d} t$.
-As macrophage is the main source of tnf-$\alpha$ production, and a similar sharp increase-and-drop trend is observed in the il-1$\beta$ trajectory, two hypothesis and corresponding terms are propose as follows.
+The two proposed alternative models are based on model 3 and try to add extra term in the tnf-$\alpha$ equation $\mathrm{d} \alpha/\mathrm{d} t$.
+As macrophage is the main source of tnf-$\alpha$ production, and a similar sharp increase-and-drop trend is observed in the il-1$\beta$ trajectory, two hypotheses and corresponding terms are propose as follows.
 
@@
 An additional term $f_{\beta\alpha}\beta$ is introduced to production rate, with $f_{\beta\alpha}$ being a new model parameter (positive constant).
-This model meets the biological context considering the source of tnf-$\alpha$.
+This model meets the biological context, considering the source of tnf-$\alpha$.
 The proposed model is written as Equation \ref{eq:model5}.
 
-[further comparison of model 3, 4 and 5]
+% [further comparison of model 3, 4 and 5]
 
-Model 4 and 5 are base on model 3, so in this phase experiments of these three models are conducted for separately parameter inference and overall model comparison.
+Model 4 and 5 are based on model 3, so in this phase experiments of these three models were conducted for separately parameter inferences and overall model comparison.
 Based on our experience in conducted experiments, we used the following setting for parameter inference:
@@
 
-Factors (Equation \ref{dis_f}) are also tried to find if we can force the inference framework to give more importance on the first few data points.
-In this experiment, for each trajectory first 8 data points are assigned higher factors (0.75), and the rest 4 data points are assigned with lower factors (0.25).
-All these runs are conducted on remote machines and the output database files are retrieved form analysis.
+Factors (Equation \ref{dis_f}) were also tried to find if we can enforce the inference framework to give more importance to the first few data points.
+An optional factor scheme was proposed: for each trajectory, the first 8 data points were assigned higher factors (75), and the rest 4 data points are assigned with lower factors (25).
+
+All these experiments were conducted on remote machines, and the output database files were retrieved for analysis.
 
@@
 
-By intuition model 4 and 5 would be the best models sofar that can recover most of observed data features.
-Model 1, 2 and 3 con also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- is not fitted.
+By intuition model 4 and 5 would be the best models so far that can recover most of the observed data features.
+Model 1, 2 and 3 could also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- was not fitted.
 Some features, e.g.
-fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl are still not ideally fitted; all existing models give a steady or gradually decrease trends for all the four variables after 20 hpl, and fluctuations in the observed data are fitted by smooth and steady curves.
+fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl were still not ideally fitted.
+All existing models gave steady or gradually decreasing trends for all the four variables after 20 hpl, where fluctuations in the observed data are fitted by smooth and steady curves.
 
 After applying factors, the results were not largely different.
-As expected, applying factors made the data points at first half (where exists most of the fluctuations) more `focused', however the inferred models are close to one without factors, which could indicate that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.
+As expected, applying factors made the data points at first half (where exists most of the fluctuations) to be more `focused'.
+However, the inferred models were not largely different from ones without factors, which indicated that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.
 
-When we are trying more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}).
+After we tried more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}).
 In this case, the discrepancy of simulated data and observed data mainly comes from the last 2 data point (72 hpl and 120 hpl) of macrophage.
-As additional importance was assigned the the first few data points of each trajectory, the inference algorithm tended to preferentially fit first data points and thus relative large discrepancy at the tails were acceptable.
+As additional importance was assigned to the first few data points, the inference algorithm tended to fit first data points preferentially, and thus relatively large discrepancy at the tails were acceptable for it.
 
@@
 
-Model 4 did not overweight other models across all the generations; model 3 gained advantage before generation 15 except generation 5.
+Model 4 did not overweight other models across all the generations; model 3 gained an advantage before generation 15 except generation 5.
 After generation 18 ($\epsilon_{18} = 16.98$), model 5 gained absolute advantage and won in further generations.
@@
 Uncertainties were observed in early generations, where similar patterns were also observed in the model comparison of model 1, 2 and 3.
-We considered the the violently changing model probability in early generations is related to the inference ability at high epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
-However as we only focused on the best model with small final threshold $\epsilon_T$, the easiest way is to extended the number of generation until epsilon converges, or increase the population size for a more statistically confident probability.
+We considered the violently changing model probability in early generations is related to the inference ability at high epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
+However as we only focused on the best model with small final threshold $\epsilon_T$, the easiest way is to extend the number of generation until epsilon converges, or increase the population size for a more statistically confident probability.
 
@@
 The pair-wise joint density distribution is shown in Figure \ref{fig:model5_para_joint}.
-Most of inferred posteriors show a bell-shaped density distribution, indicating that they are gaussian distributed.
-A thin and concentrated shape indicate a well inferred parameter, e.g.
+Most of the inferred posteriors show a bell-shaped density distribution, indicating that they are Gaussian distributed.
+A thin and concentrated shape indicates a well-inferred parameter, e.g.
 $\mu{\Phi}$ and $\kappa_{\Phi\beta}$.
 Some distributions, however, are skewed to zero, e.g.
-$\mu_N$ ans $s_{\alpha\Phi}$.
+$\mu_N$ and $s_{\alpha\Phi}$.
 
@@
 
-A quick sensitivity analysis of the dynamical systems models can be quantified using principle component analysis (PCA) \cite{Toni}, where the first PCs corresponds to sloppy parameters while the last PCs corresponds to stiff parameters \cite{sensitivity}.
+A quick sensitivity analysis of the dynamical systems models can be quantified using principal component analysis (PCA) \cite{Toni}, where the first PCs corresponds to sloppy parameters while the last PCs corresponds to stiff parameters \cite{sensitivity}.
 PCA can only provide rough approximated sensitivity behaviour.
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[h]
     \begin{center}
@@
 
-As models 5 won in the model selection, a further interest was to explore the parameter-to-model sensitivity of the resultant model.
+As models 5 won in the model selection, further interest was to explore the parameter-to-model sensitivity of the resultant model.
 PCA was performed on the last generation of inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC.
-The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less less than 5\% of the variance each.
-PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs, thus they were consider to extend much narrower regions on the posterior distributions and could be used to identify parameter that the model is sensitive to.
+The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less than 5\% of the variance each.
+PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs.
+Thus, they were considered to extend much narrower regions on the posterior distributions and could be used to identify the parameters that the model is sensitive to.
 

### 2020-08-17 conclusion/conclusion.tex
New fix
Grammar fix

@@
 
- To conclude, ABC SMC was successfully applied on the models of zebrafish spinal cord regeneration.
+ To conclude, ABC SMC was successfully applied to the models of zebrafish spinal cord regeneration.
 These models were all ODE equations that described the interactions and effects between cells and proteins in the lesion site.
-These equations contains four dependent variable and more than 10 model parameters.
+These equations contain four dependent variable and more than 10 model parameters.
 Parameter estimation for high-dimensional parameter space is a challenging work for many dynamic system models, as insufficient explore of parameter space can easily lead to local optimal.
-Using a large population size in ABC SMC in our models could partially relief the concerns in local modes, and give considerable parameter estimates (FIGURE).
-Also ABC SMC successfully helped in the model comparison and thus can be used for suggestions of possibly best model, although there were uncertainties remained in the process and a large number of population size was need.
+Using a large population size in ABC SMC in our models could partially relief the concerns in local modes, and give considerable parameter estimates.
+We found that the first three models could not recover some local features observed in the early times, so further models -- model 4 and 5 -- were proposed.
+Additional interactions were introduced in the new model, and the results proved that they were helpful in resolving the under-fit.
 
- ABC SMC implementation is a complex process with respect to both algorithm settings and code development.
-The ability of parameter inference and robustness of algorithm were firstly studied using synthetic data where the true parameter value was known, and experiments were performed to find proper hyperparameter and implementation settings of ABC SMC.
-Later inference on the observed data followed the suggestions from the synthetic data experiments and acceptable results were produced.
+ Also, ABC SMC successfully helped in the model comparison and thus can be used for suggestions of the best model, although there were uncertainties remained in the process and a large number of population size needed.
+In our practice, model 5 won with absolute advantages when the threshold is converged to a small value.
+The approximated posteriors showed bell-shaped density distribution for most of the parameters (Figure \ref{fig:model5_para} and Figure \ref{fig:para1}), indicating that most parameters were well-inferred in our implementations.
+
+%  ABC SMC implementation is a complex process with respect to both algorithm settings and code development.
+The ability of parameter inference and robustness of algorithm was firstly studied using synthetic data where the true parameter value was known, and experiments were performed to find proper hyperparameter and implementation settings of ABC SMC.
+
+% Later inference on the observed data followed these suggestions from the synthetic data experiments and acceptable results were produced.
+These examples proved that ABC SMC was capable of the inference tasks for the proposed models with high-dimensional parameter space.
 
  The inference program was mostly run on multi-core machines, and additional scaling-up performance test illustrated the scaling behaviour.
-When the number of running \verb|Python| process exceeds the number of physical cores, hyperthreading would be enabled but gains not significant improvements with respect to execution time, while increase of average sampling size per second was still observed.
+When the number of running \verb|Python| process exceeds the number of physical cores, hyperthreading would be enabled, and performance improvement was still considerable (Figure \ref{fig:sample_per_sec}).
+Studies on the variant required samples revealed that some threshold paths could lead to local optimum and significantly affect the efficiency.
+A non-linear relationship between the required samples and execution time was also observed in the first two generations.
 
@@
  From this application, some suggestions on ABC SMC implementation could be drawn.
-How to set the prior could be the most tricky decision for a high dimensional parameter inference.
-The distribution and the interval of each parameter can affect the inference results significantly and improper prior cold lead to poor fit.
-The epsilon trajectory and joint distribution of parameters can be used to identity possible local modes; if stuck, it takes much more samples than usual to jump out out local optimal.
-To avoid this a large population size and a proper epsilon schedule may help.for the uncertainties observed in model selection and some duplicated runs with high variance, a longer run with more generation could help to identify whether the convergency is consistent such that a reproducible inferred results can be obtained.
+How to set the prior could be the most tricky decision for high dimensional parameter inference.
+The distribution and the interval of each parameter can affect the inference results significantly, and improper prior cold lead to poor fit.
+In our cases, log-uniform prior gave better fits than uniform prior.
+The threshold path and plot of required samples in each generation can be used to identify possible local modes; if stuck, it takes much more samples than usual to move out.
+To avoid this, large population size and a proper epsilon schedule may help.
+For the uncertainties observed in model selection and some duplicated runs with high variance, a longer run with more generation could help to identify whether the convergency is consistent such that a reproducible inferred results can be obtained.
 
@@
 In the aspect of models, some further models can be studied.
-If there were some more experimental evidence on the proposed interaction map, we could proposed more precise models, or calibrate some terms in our existing models.
+If there were some more experimental evidence on the proposed interaction map, we could propose more precise models, or calibrate some terms in our existing models.
 Also, models of other forms, e.g.
 PDE and stochastic models, could also be studied if a more complex model is needed for the dynamic system.
-Simplification is also possible of existing models
+Simplification is also possible of existing models.
+Some highly-correlated parameters could be reduced, as there exist some linear relationships between parameters.
+
+% Beyond this scope, we are interested in the

### 2020-08-17 implementation/implementation.tex
Grammar fix

@@
 After generation 18 ($\epsilon_{18} = 16.98$), model 5 gained absolute advantage and won in further generations.
-Also, repeated runs showed similar trends: although uncertainties were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
+Also, repeated runs showed similar trends: although uncertainties (e.g.
+violently changes mode probabilities) were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
 $\epsilon_t<16$.

### 2020-08-17 performance/performance.tex
Grammar fix

@@
 
-The performance experiments were designed to explore the parallel performance of ABC SMC inference framework which were used in the parameter estimation and model selection tasks.
-Usually ABC SMC is a time-consuming and computation intensive task and ideally executed on large clusters.
-The scheduling strategy, implementation details and many other factors can affect the parallel efficiency.
+The performance experiments were designed to explore the parallel performance of ABC SMC inference framework that was used in our parameter estimation and model selection tasks.
+Generally, ABC SMC is a time-consuming and computation-intensive task and ideally executed on large clusters.
+The scheduling strategy, implementation details, randomness in the algorithm and many other factors can affect the parallel efficiency.
+Thus, we designed a scaling-up experiment and studied a case where abnormal performance was observed.
 
@@
 
-First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABc SMC on model 5. The details of the ABC SMC settings is listed below
+First experiments are designed to illustrate the scaling-up performance. The program used here is an implementation of ABC SMC on model 5. The details of the ABC SMC settings are listed below
 
@@
 For HPC systems like Cirrus, \verb|pyabc| uses \verb|multiprocessing| for multi-core parallel sampling.
-By default if the number of cores is not specified, it will automatically read the number of available cores and use them all.
+By default, if the number of cores is not specified, it will automatically read the number of available cores and use them all.
 Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
 
-The program is executed on Cirrus, using 8, 16, 24, 36, 54 and 72 cores respectively.
-Each run is repeated 10 times.
-The average execution time, required sampling numbers are recorded.
-Hyperthreading is enabled when using 54 and 72 cores.
+The experiment was executed on Cirrus, using 8, 16, 24, 36, 54 and 72 cores respectively.
+Each run was repeated ten times and the average execution time, required sampling numbers are recorded.
+Hyperthreading was enabled when using 54 and 72 cores.
 The access to the node that contains computation cores is exclusive, such that the execution would not be affected by other programs of operations.
 
-The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the must time-consuming part.
+The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the most time-consuming part.
 The rest part of the program is mostly not parallelised, e.g.
 database I/O and reductions operations.
-The sampling process involves sampling, perturbation and test of the acceptance criteria, all of which are computation-intensive.
+The sampling process involves sampling, perturbation and test of the acceptance criteria, which are computation-intensive.
 
-In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources is limited[REF].
-The performance experiment result could provide a reference that illustrate that how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
+In practice, using ABC SMC to estimate the parameters of a given model could cost up to several hundred of hours if the computational resources are limited \cite{ref:compare}.
+The performance experiment result could provide a reference that illustrates how the efficiency changes when scaling-up or the trade-offs in computational resources' cost and their benefit.
 
@@
 
-The recorded execution time using different number of cores were used to calculate the speedup of the program (Figure \ref{fig:speedup}).
-Here 8 cores is regarded as the baseline for speed-up and efficiency, as the serial version (using only one core) can take quite a long time to finish 10 repeats.
-When increasing number of cores, the execution time drops fast at first but the decreasing rate (dropping speed, i.e.
-the derivative of the execution time against time) is gradually reduced to zero; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (Table \ref{table:performance}).
+The recorded execution time under different numbers of cores was used to calculate the speedup of the program (Figure \ref{fig:speedup}).
+Here 8-core run is regarded as the baseline for speedup and efficiency, as the serial version (using only one core) can take quite a long time to finish 10 repeats.
+When increasing number of cores, the execution time drops fast at first, but the decreasing trend became slower for large numbers of cores; 36, 54 and 72 cores gives nearly the same average execution time and consequently gains very close speedup values (Table \ref{table:performance}).
 
@@
 The speedup curve shows a linear trend when $p\leq 36$; when $p\geq 36$ the speedup stays nearly unchanged.
-$p=36$ is an inflection point, where 36 is the maximum numbers available physical cores in a compute node of Cirrus.
+$p=36$ is an inflexion point, where 36 is the maximum numbers available physical cores in a compute node of Cirrus.
 A constant drop of efficiency is observed when increasing $p$; smaller $p$ ($p<36$) gives an efficiency higher than 65\% but greater $p$ e.g.
@@
 
-Due to the variant required samples, our interests switched to the per-second performance, where the average sampling numbers per second under different of number of cores ($p$) are plotted (Figure \ref{fig:sample_per_sec}).
+Due to the variant required samples, our interests switched to the per-second performance, where the average sampling numbers per second under different of numbers of cores ($p$) are plotted (Figure \ref{fig:sample_per_sec}).
 The plot shows that even theoretical speedup stops growing when $p>36$, the average sampling numbers per second performance still get improved when using more than 36 cores.
@@
 The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path.
-Median epsilon schedule and randomness in the sampling process resulted in different sequences of threshold for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
-For example, Some runs temporarily converges to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated in the local optimum.
-In this case, huge amount of sample will be drawn to find enough accepted particles, and new generation will be able to jump out of the local optimum and have different concentrations.
+Median epsilon schedule and randomness in the sampling process resulted in different sequences of thresholds for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
+For example, Some runs temporarily converge to a local optimum, while the target epsilon in the next generation cannot possibly be reached if the new particles are still concentrated in the local optimum.
+In this case, a huge amount of sample will be drawn to find enough accepted particles, and the new generation will be able to jump out of the local optimum and have different concentrations.
 
@@
 To illustrate this, Two independent run, namely Run A and Run B, were chosen from the experiment of $p=24$ and visualised in Figure \ref{fig:local_modes} and Figure \ref{fig:local_para}.
-Thess two run used identical options of the algorithm.
+These two runs used identical options of the algorithm.
 Figure \ref{fig:local_modes} (a) shows the comparison of the required samples, where Run A converged to a local mode, and in population 11 took much more sample tires to move away from the local optimum (Figure \ref{fig:local_modes} (b)).
@@
 $T_t\propto n_t$, where $t$ is the population index, $T_t$ is the execution time of generation $t$ and $n_t$ is the required samples in generation $t$.
-However, this relationship does not hold for the first two generation.
-Generation 1 is to draw sample from prior, but it is done with high efficiency as only small amount of time are taken; generation 2 takes unusual large amount of the, but it does not require much more samples.
+However, this relationship does not hold for the first two generations.
+Generation 1 is to draw samples from prior, but it is done with high efficiency as only a small amount of time are taken; generation 2 takes an unusually large amount of the, but it does not require much more samples.
 The abnormal execution time of generation 1 and two are observed in all other runs, where for $p=24$ generation 1 usually takes c.
@@
 
-Additionally, by observing CPU frequency and load curve in-time while running the program, we found that the parallelised part is limited the the sampling, calculation and criterion test process; some reduction calculations and preparation work for the next generation (e.g.
+Additionally, by observing CPU frequency and load curve in-time while running the program, we found that the parallelised part is limited to the sampling, calculation and criterion test process.
+Some reduction calculations and preparation work for the next generation (e.g.
 calculation of the new epsilon values, normalisation of data, fit of multivariate kernel) are serial.
-Database I/O is also serial.
+The database I/O is also serial.
 

### 2020-08-17 MSc-dissertation.tex
New para
New fix
Fig fix

@@
 \begin{abstract}
-    This is the bit where you summarise what is in your thesis.
+    Zebrafish can regenerate the spinal cord after injury, and possibly recover swimming function.
+The regeneration is a complex biological process with many interacting parts.
+We modelled the dynamics of two cell and two molecules that participate in the regeneration, according to the existing experimental findings\cite{ref:Tsarouchas}, and wished to explore uncertainties in their parameter estimates.
+
+    As an Approximate Bayesian Computation (ABC) method, ABC SMC (Sequential Monte-Carlo) method was implemented on the proposed models to approximate the posteriors the parameters.
+We conducted some preliminary experiments to study the ability of inference and robustness of the algorithm before we performed the inference of parameters and model comparison using the same inference framework.
+The result shows well-inferred parameters and a preferred model with high model probability.
+
+    The algorithm implementation shows a decent scaling-up performance; however, problems with local optimum and some other uncertainties were also found.
+Suggestions on the implementation options and ABC SMC settings were made to avoid these problems.
 \end{abstract}
@@
 
+\newpage
+
 \pagenumbering{arabic}
@@
 \input{performance/performance.tex}
-
-\input{conclusion/future_works.tex}
 

### 2020-08-17 implementation/implementation.tex
Fig fix

@@
 
-    \caption{Simulated data from the last population of model 3, 4 and 5. Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
+    \caption[Simulated data from the last population of model 3, 4 and 5]{Simulated data from the last population of model 3, 4 and 5.
+Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
     \label{fig:resultCurve345}
@@
     \begin{center}
-        \resizebox{0.8\hsize}{!}{\includegraphics{fig/pc_pie_all.png}}
+        \resizebox{0.8\hsize}{!}{\includegraphics{fig/pc_pie.png}}
     \end{center}
 
-    \caption[TODO]{Fraction of the length of PCs explained by parameters}
+    \caption[Fraction of the length of PCs explained by parameters]{Fraction of the length of PCs explained by parameters. Portions less than 2\% are not labelled}
     \label{fig:pc_pie}

### 2020-08-18 appendix/appendix.tex
Fig fix
Big fix

@@
         Operating system   & macOS Catalina 10.15.6 \\
-        PyCharm (IDE)      & Professional 202.2     \\
+        PyCharm (IDE)      & Professional 2020.2    \\
         Python interpretor & 3.7.0                  \\
@@
 Version requirements for some critical site packages (\verb|Python|) used in the code are listed in Table \ref{table:local_package}.
+When replicating the experiments, higher version of these packages are acceptable, except \verb|bokeh|.
 
@@
         \hline\hline
-        pyABC       & $\simeq$ 0.10.3 \\
-        NumPy       & $\simeq$ 1.18.4 \\
-        SciPy       & $\simeq$ 1.4.1  \\
-        pandas      & $\simeq$ 1.1.0  \\
-        matplotlib  & $\simeq$ 3.0.1  \\
-        bokeh       & $=$ 1.4.0       \\
+        pyABC       & 0.10.3  \\
+        NumPy       & 1.18.4  \\
+        SciPy       & 1.4.1   \\
+        pandas      & 1.1.0   \\
+        matplotlib  & 3.0.1   \\
+        bokeh       & 1.4.0   \\
         \hline
@@
 When running ABC SMC, the program can make used of 8 python process (Hyper-Threading enabled) and run under 3.8GHz (maximal) for a long time under proper cooling, proved that the program could efficiently use the computation resources for a local personal computer.
-It is noted that the program took no advantage of graphics card, as \verb|multiprocessing| along can only exploit the resources inside CPU.
 The execution time and other performance data presented in Chapter 5 were obtained using remote machines but not local machine.
@@
 
-Cirrus hardware can be found in \url{https://www.cirrus.ac.uk/about/hardware.html}.
+Most experiments were performaed on Cirrus, its hardware can be found in Table \ref{table:remote_hardware}.
+The program can run in sigle compute-node which contains two processors\footnote{\url{https://www.cirrus.ac.uk/about/hardware.html}}.
 The software environment used in the development and experiments are listed in Table \ref{table:cirrus_soft}.
-Requirement of site packages is the sanme as that on local macine (Table \ref{table:local_package})
+Requirement of site packages is the same as that on local macine (Table \ref{table:local_package})
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c c|}
+        \hline
+        Hardware        & Detail                                          \\ [0.5ex]
+        \hline\hline
+        Processors      & 2.1 GHz, 18-core Intel Xeon E5-2695 (Broadwell) \\
+        Architecture    & x86-64                                          \\
+        Hyper-Threading & Supported                                       \\
+        Memory          & 256 GB NUMA, shared between two processors         \\
+        \hline
+    \end{tabular}
+    \caption{Cirrus hardware (compute node)}
+    \label{table:remote_hardware}
+\end{table}
 
@@
 
-\chapter{Data and supplementary figures}
+\chapter{Supplementary figures and data}
 
-\section{Infer-back experiments}
+% \section{Infer-back experiments}
 
-\subsection{Parameter values used to generate synthetic data}
+% \subsection{Parameter values used to generate synthetic data}
 
@@
 
-\subsection{Kernel experiment: median epsilon schedule}
+% \subsection{Kernel experiment: median epsilon schedule}
 
@@
 
-\section{Parameter inference and model comparison results}
+% \section{Parameter inference and model comparison results}
 
 \begin{figure}[H]
-
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve_uni.png}}
-    \end{center}
-
-    \caption{Simulated data from the last population of model 1, 2 and 3, using uniform prior.  Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
-    \label{fig:resultCurve_uni}
-
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/inferno5.png}}
-    \end{center}
-    \caption[Joint density distribution of each parameter pair]%
-    {Joint distribution of each parameter pair.
-The diagonal is KDE fit of each individual parameter, the upper triangle contains scatter plot of the particles and the lower triangle is the joint density distribution for each parameter pair}
-    \label{fig:model5_para_joint}
-\end{figure}
-
-\begin{figure}[ht]
     \begin{center}

### 2020-08-18 conclusion/conclusion.tex
Big fix

@@
  The inference program was mostly run on multi-core machines, and additional scaling-up performance test illustrated the scaling behaviour.
-When the number of running \verb|Python| process exceeds the number of physical cores, hyperthreading would be enabled, and performance improvement was still considerable (Figure \ref{fig:sample_per_sec}).
-Studies on the variant required samples revealed that some threshold paths could lead to local optimum and significantly affect the efficiency.
-A non-linear relationship between the required samples and execution time was also observed in the first two generations.
+
+ When the number of running \verb|Python| processes is less than the number of physical cores, a considerable scaling performance was observed.
+Studies on the variant required sample numbers revealed that some threshold paths could lead to local optimum and significantly affect the efficiency.
+A non-linear relationship between the required sample numbers and required execution time was also observed and made the detailed analysis complicated.
 
@@
 
- From this application, some suggestions on ABC SMC implementation could be drawn.
-How to set the prior could be the most tricky decision for high dimensional parameter inference.
-The distribution and the interval of each parameter can affect the inference results significantly, and improper prior cold lead to poor fit.
+ From this application, some suggestions on ABC SMC implementation could be drawn.The prior distribution and interval of each parameter can affect the inference results significantly, and improper prior cold lead to poor fit.
 In our cases, log-uniform prior gave better fits than uniform prior.
@@
 %  [thanks]
+
+In terms of the project progress regarding original project proposal, all the proposed subjects were explicitly studied as planned.
+We found that code development cost less time than expected, however, implementations and debugging cost more efforts than expected because of both uncertainties in both the inference ability of the algorithm for our problem, and the implementation options and algorithm hyperparameters.
+Further on our results, we studied some in-depth topics and derived more conclusions.
+Some interesting further topics were found in the studies, e.g.
+comparison with other algorithms and implementation optimisations, however, they were not conducted due the insufficient time.
+
+Also, we found that risk analysis and management throughout the project were helpful.
+Thanks to the planned mitigations and backup strategies, we were able to proceed with our project as planned when interruptions happened.
+We experienced local laptop broken and unavailability of remote machines, which consequently made our development and implementations harder.
+We were able to proceed with the project by using backup plans for local development and using Cirrus or ARCHER interchangeably during their downtime.
 

### 2020-08-18 implementation/implementation.tex
Big fix

@@
 
-    \caption[Total required samples of different kernels]%
+    \caption[Total required number of sample for different kernels]%
     {Total required samples of different kernels after 20 populations (2000 particles in each population).
@@
 Two types of factor and adaptive distance function were tested by running the same number of generations (20 generations, 2000 particles per generation).
-The first type of factor is 25:75  factor, where the first half of the trajectory data were assigned higher factor (75), and the second half of the data was assigned lower factor (25); this could make the first half of the data more `important' in the inference.
+The first type of factor is 75:25 factor, where the first half of the trajectory data were assigned higher factor (75), and the second half of the data was assigned lower factor (25); this could make the first half of the data more `important' in the inference.
 The second factor was `range' factor, which tried to normalise the trajectory of each dependent variable according to its data range.
@@
 From Figure \ref{fig:factor}, different types of factor gave different performance.
-25:75 factor required much more samples in the first generation (orange bar in the left of Figure \ref{fig:factor}) and also did not improve much in later generations; thus, it is the slowest one.
+75:25 factor required much more samples in the first generation (orange bar in the left of Figure \ref{fig:factor}) and also did not improve much in later generations; thus, it is the slowest one.
 The first half of observed data contained more features and information, assigning it with higher factor would make the algorithm think that it is more important to fit the first half and thus focus on the hard part, and consequently more particles were needed in the inference process.
-The acceptance rates of 25:75 factor among different generations were also generally slightly lower than the standard.
+The acceptance rates of 75:25 factor among different generations were also generally slightly lower than the standard.
 By contrast, range factor required less total samples and gave generally higher acceptance rates than standard, which made it considerably faster.
@@
     {Factors and adaptive distance experiments.
-Left: total required samples.
-Different color represents different generations (bottom totop: population 1 to population 20).
+Left: total required number of samples.
+Different color represents different generations (bottom to top: population 1 to population 20).
 Right: acceptance rates in each generation.
@@
 500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from.
-25:75 factor showed very similar curve as the range factor.
+75:25 factor showed very similar curve as the range factor.
     }
@@
     \caption[Data size and prior distribution range experiment]%
-    {Data size and prior distribution range experiment. Left: total required samples. Right: epsilon values under median epsilon schedule}
+    {Data size and prior distribution range experiment. Left: total required number of samples. Right: epsilon values under median epsilon schedule}
     \label{fig:size}
@@
 
-This subsection aims to explore options in ABC SMC implementations preliminarily.
+This section aims to explore options in ABC SMC implementations preliminarily.
 From the above results, we adopted multivariate normal kernel as the kernel for our real inference on observed data; Local M-NN kernels were also planned to be tested.
-In the real inference, 25:75 factor were also tried to see if it could help in recovering more features and resulting in a better fit.
-Because the improvement by using less data was inapparent, we did not try to change the target summary data size.
+In the real inference, 75:25 factor were also tried to see if it could help in recovering more features and resulting in a better fit.
 Prior intervals were explored explicitly and gradually shrunk for our models.
+
+% Because the improvement by using less data was inapparent, we did not try to change the target summary data size.
 
@@
 All these ABC SMC runs used 2,000 particles per population and 30 generations.
+The population size should be set at a high level (e.g.
+$\geq 1000$ in our case) when having models with many parameters to ensure a statistically adequate exploration.
+The convergency were usually observed around generation 20 in our previous experiments; the length of generation is set to 30 to ensure that most runs converge in the end.
 
@@
 We expected model 3 to be the best as it reached a lower threshold value, and intuitively gives a better fit.
-However, the fit of all these models could not be considered as a generally good representation of the biological process that we want to approximate: for tnf-$\alpha$, the simulated data were significantly biased from the observed data.
-A sharp peak at 4 hpl was observed in the experimental data, but none of the three models represents similar features.
+However, the fit of all these models could not be considered as a generally good representation of the biological process that we want to approximate: for tnf-$\alpha$, the simulated data and the observed data show largely different trends.
+A sharp peak at 4 hpl was observed in the experimental data, but none of the three models represents similar features (Figure \ref{fig:result123}).
 
@@
     \caption[Estimated posterior distribution of parameters in model 3]%
-    {Estimated posterior distribution of parameters in model 3. Shaded range indicates 25\%--75\% quantile of the population}
+    {Estimated posterior distribution of parameters in model 3. Shaded range indicates inter-quartile interval}
     \label{fig:para1}
@@
 Log-uniform tends to assign values that are close to the left boundary (i.e.
-zero in our case) with higher density, i.e.
-they are more likely to be sampled in the first population.
-As a result, the parameter estimates are expected to have smaller mean values or skewed to the left of the x-axis.
-The estimated values proved this ( \ref{table:estimated1}).
+zero in our case) with higher density.
+As a result, the parameter estimates from log-uniform prior are highly likely to have smaller mean values than that from uniform prior, and the inferred parameter values proved this (Table \ref{table:estimated1}).
 This was also observed in other models and prior interval experiments.
@@
 All the models cannot well recover this feature, and most results saw a less-rapid increase followed by a gradual decrease, while the observed peak value was not reached (Figure \ref{fig:result123}).
-Also, in some other case, e.g.
+
+Also, in some other experiment results, e.g.
 uniform runs (Figure \ref{fig:resultCurve_uni}), tnf-$\alpha$ trajectories was well fitted; however, at the cost of under-fitting $\Phi$ curve.
 Hence efforts had been spent to propose more alternative models that can potentially solve this problem.
+
+\begin{figure}[h]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve_uni.png}}
+    \end{center}
+
+    \caption[Simulated data from the last population of model 1, 2 and 3, using uniform prior]{Simulated data from the last population of model 1, 2 and 3, using uniform prior.
+Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
+    \label{fig:resultCurve_uni}
+\end{figure}
 
@@
 
-    \caption[Simulated trajectory from the 30th population of model 5, with 25:75 factor applied]{Simulated trajectory from the 30th population of model 5, with 25:75 factor applied.
+    \caption[Simulated trajectory from the 30th population of model 5, with 75:25 factor applied]{Simulated trajectory from the 30th population of model 5, with 75:25 factor applied.
 Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small).}
@@
 
-Model 4 did not overweight other models across all the generations; model 3 gained an advantage before generation 15 except generation 5.
-After generation 18 ($\epsilon_{18} = 16.98$), model 5 gained absolute advantage and won in further generations.
-Also, repeated runs showed similar trends: although uncertainties (e.g.
-violently changes mode probabilities) were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
+From the model probability plot (Figure \ref{fig:model345cmp}), we found that model 4 did not overweight other models across all the generations.
+Model 3 was comitative and gained advantages at most early generation.
+From generation 14 and onwards, model 5 saw a gradually increasing probability and gained 100\% model probability after generation 18 ($\epsilon_{18} = 16.98$) and won in further generations.
+Also, repeated runs showed similar trends, although uncertainties (e.g.
+drastically model probability changes) were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e.
 $\epsilon_t<16$.
@@
 Uncertainties were observed in early generations, where similar patterns were also observed in the model comparison of model 1, 2 and 3.
-We considered the violently changing model probability in early generations is related to the inference ability at high epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
+We considered the drastically changing model probability in early generations is related to the inference ability at high epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space.
 However as we only focused on the best model with small final threshold $\epsilon_T$, the easiest way is to extend the number of generation until epsilon converges, or increase the population size for a more statistically confident probability.
@@
 The pair-wise joint density distribution is shown in Figure \ref{fig:model5_para_joint}.
-Most of the inferred posteriors show a bell-shaped density distribution, indicating that they are Gaussian distributed.
+Most of the inferred posteriors show a bell-shaped density distribution with a single peak..
 A thin and concentrated shape indicates a well-inferred parameter, e.g.
@@
 $\mu_N$ and $s_{\alpha\Phi}$.
+Concentrations in the posteriors were presented, and using inter-quartiles (shaded interval in Figure \ref{fig:model5_para}) or mean values to summarise the posteriors is feasible in practice uses, e.g.
+simulating data (Figure \ref{fig:overfit}).
 
 By looking at the joint density plot (Figure \ref{fig:model5_para_joint}), some correlated patterns are observed.
-For example, an obvious linear relationship on the distributions of $\mu_\beta$ and $s_{\beta\Phi}$.
-This leads us to consider that reducing the number of parameters in the model is possible.
-Highly correlated parameters can be reduced, e.g.
-$s_{\beta\Phi}$ can be expressed as $A\mu_\beta+B$ where $A$ and $B$ can be calculated from the joint distribution by fitting a linear regression.
+For example, an obvious linear relationship is presented in the joint distribution of $\mu_\beta$ and $s_{\beta\Phi}$.
+This leads us to consider that we could possibly reduce the number of parameters, e.g.
+replace $s_{\beta\Phi}$ by $A\mu_\beta+B$ where $A$ and $B$ can be calculated from the joint distribution by fitting a linear regression.
+The reason for the highly correlated pattern, or `fat' distribution shapes could be that the provided data are lees informative about the parameters.
 
@@
     \caption[Inferred posterior distribution of parameters in model 5]%
-    {Inferred posterior distribution of parameters in model 5. Shaded range indicates 25\%--75\% quantile of the population}
+    {Inferred posterior distribution of parameters in model 5. Shaded range indicates inter-quartile  interval of the population}
     \label{fig:model5_para}
 
+\end{figure}
+
+\begin{figure}[h]
+    \begin{center}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/inferno5.png}}
+    \end{center}
+    \caption[Joint density distribution of each parameter pair]%
+    {Joint distribution of each parameter pair.
+The diagonal is KDE fit of each individual parameter, the upper triangle contains scatter plot of the particles and the lower triangle is the joint density distribution for each parameter pair}
+    \label{fig:model5_para_joint}
 \end{figure}
@@
 
-\begin{figure}[h]
+As models 5 won in the model selection, further interest was to explore the parameter-to-model sensitivity of the resultant model.
+PCA was performed on the last generation of inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC.
+The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less than 5\% of the variance each.
+PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs.
+Thus, they were considered to extend much narrower regions on the posterior distributions and could be used to identify the parameters that the model is sensitive to.
+
+\begin{figure}[ht!]
     \begin{center}
-        \resizebox{0.8\hsize}{!}{\includegraphics{fig/pc_pie.png}}
+        \resizebox{0.7\hsize}{!}{\includegraphics{fig/pc_pie.png}}
     \end{center}
@@
 \end{figure}
-
-As models 5 won in the model selection, further interest was to explore the parameter-to-model sensitivity of the resultant model.
-PCA was performed on the last generation of inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC.
-The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less than 5\% of the variance each.
-PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs.
-Thus, they were considered to extend much narrower regions on the posterior distributions and could be used to identify the parameters that the model is sensitive to.
 
@@
 
+To conclude, model 5 is most sensitive to the parameters in the equations for cells' dynamics (i.e. first two equations in Eqn.\ref{eq:model5}), especially to $a$ and $\mu_\Phi$.
+
 % then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$.

### 2020-08-18 modelling/modelling.tex
Big fix

@@
 \end{figure}
-
 \begin{align}
@@
 \end{align}
-
 It was assumed that there are negative feedbacks the cells and cytokines, i.e.
@@
 Model 2 and model 3 introduced one extra parameter $a$ and removed $\lambda_\Phi$, which is a coefficient in the exponentially decaying $\lambda_N$, determining the decay speed, with the unit $h^{-1}$.
-
 \begin{align}
@@
 \end{align}
-
 Model 3 can be regarded as a simplification of model 2, as it can be treated as model 2 with parameter $i_{\beta\Phi}=0$.
@@
 \ref{eq:model5}).
-
 \begin{align}
@@
 \end{align}
-
 \begin{table}[h!]

### 2020-08-18 performance/performance.tex
Fig fix
Big fix

@@
 
-    \caption[Speedup of the program]{Speedup of the program}
+    \caption[Speedup of the program]{Speedup of the program. Line in orange indicates the ideal linear speedup}
     \label{fig:speedup}
@@
         \hline
-        Number of cores & Time (second) & Avg. required samples & Speedup & Efficiency \\ [0.5ex]
+        Number of cores & Avg. time (second) & Avg. sample size & Speedup & Efficiency \\ [0.5ex]
         \hline\hline
@@
 
-\begin{figure}[h]
-    \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/sample_per_time.png}}
-    \end{center}
+% \begin{figure}[h]
+%     \begin{center}
+%         \resizebox{1.0\hsize}{!}{\includegraphics{fig/sample_per_time.png}}
+%     \end{center}
 
-    \caption[Average samples in a second, under different number of cores]{Average samples in one second, under different number of cores}
-    \label{fig:sample_per_sec}
-\end{figure}
+%     \caption[Average samples in a second, under different number of cores]{Average samples in one second, under different number of cores}
+%     \label{fig:sample_per_sec}
+% \end{figure}
 
@@
 
-Due to the variant required samples, our interests switched to the per-second performance, where the average sampling numbers per second under different of numbers of cores ($p$) are plotted (Figure \ref{fig:sample_per_sec}).
+Due to the variant total required samples drawn our attention to the per-second performance.
+Although the execution time was not improved by using 54 or 72 cores, the required number of samples were higher than that of $p=36$, which indicates slight improvements.
+However, the high variance made us hard to gave statistically significant conclusions on the per-second performance.
+
+% our interests switched to the per-second performance, where the average sampling numbers per second under different of numbers of cores ($p$) are plotted (Figure \ref{fig:sample_per_sec}).
 The plot shows that even theoretical speedup stops growing when $p>36$, the average sampling numbers per second performance still get improved when using more than 36 cores.
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
     \begin{center}
-        \resizebox{1.0\hsize}{!}{\includegraphics{fig/local_modes.png}}
+        \resizebox{1.0\hsize}{!}{\includegraphics{fig/local_modes_new.png}}
     \end{center}
 
-    \caption[Local optimum and abnormal pattern]{(a) Required sampling size for two example runs.
+    \caption[Local optimum and abnormal pattern]{(a) Required number of samples for two example runs (A and B).
 Different colour represents different generation (from bottom to top: generation 1 to generation 20).
 Run A took huge amount of time in population 11.
-(b) Required sample in each generation of Run A.
-(c) Time-consumption in each generation of Run A}
+(b) Execution time for each generation of Run A.
+(c) Scatter plot of required number of samples against execution time for each generation.
+Generation 1 and two outliers, i.e.
+generation 11 and generation 2, are marked in different colours from other generations}
     \label{fig:local_modes}
@@
 
-\begin{figure}[t]
+\begin{figure}[ht!]
     \begin{center}
@@
 Moreover, we found that the execution time was not ideally proportional to the sampling tires in that period of time, for both different repeated runs and different generations in a single run.
-From the same example, Figure \ref{fig:local_modes} (b) and (c) show the non-linear relationship between execution time and sample tries in different generations.
-Apart from 1st and 2nd generation, generation 3 to 20 show a general linear relationship i.e.
+From the same example, the sample size for each generation of Run A (Figure \ref{fig:local_modes} (a)) is not proportional to its execution time (Figure \ref{fig:local_modes} (b)), especially for the first two generations and generation 11.
+Figure \ref{fig:local_modes} (c) shows a scatter plot of the required samples and execution time for each generation, where two obvious outliers (generation 1 and generation 11) are observed.
+
+% Figure \ref{fig:local_modes} (b) and (c) show the non-linear relationship between execution time and sample tries in different generations.
+
+% Apart from 1st and 2nd generation, generation 3 to 20 show a general linear relationship i.e.
 $T_t\propto n_t$, where $t$ is the population index, $T_t$ is the execution time of generation $t$ and $n_t$ is the required samples in generation $t$.
-However, this relationship does not hold for the first two generations.
-Generation 1 is to draw samples from prior, but it is done with high efficiency as only a small amount of time are taken; generation 2 takes an unusually large amount of the, but it does not require much more samples.
-The abnormal execution time of generation 1 and two are observed in all other runs, where for $p=24$ generation 1 usually takes c.
+
+% However, this relationship does not hold for the first two generations.
+
+Generation 1 (red pont in Figure \ref{fig:local_modes} (c)) drew samples from prior, but it was done with high efficiency as only a small amount of time are taken; generation 2 (purple pont in Figure \ref{fig:local_modes} (c)) took an unusually large amount of the while it did not require much more samples.
+The abnormal execution time of generation 1 and 2 were observed in all other runs, where for $p=24$ generation 1 usually takes c.
 16s and generation 2 takes c.
@@
 This is believed to be related to the initialisation of the algorithm.
+For local modes, e.g.
+generation 11 (green pont in Figure \ref{fig:local_modes} (c)), high efficiency were achieved.
+
+By looking into the code, some reasons can be found.
+There exist some `cheap' particles.
+If at the first time points, a variable reached unacceptable high value (regarded as \verb|Inf| in \verb|Python|), the ODE solver cannot perform further integration but result in a data full of \verb|NaN|s.
+Such particles cost less time than an usual particle (where values at 120 data points are explicitly calculated and then compared with observed data).
+Local mode, e.g.
+generation 11, might take large amount of these `cheap' particles.
+Also, generation 1 draws samples directly from prior without fitting and using kernels, so it is much faster.
+The huge amount of time observed in generation 2 is believed to be a result of some initialisation steps in the ABC SMC algorithm.
+
+To conclude, the performance experiments revealed reasonable scaling-up behaviour when the \verb|Python| processes did not exceed number of physical cores.
+Local modes lead to high variance in total required number of samples, and the relationship between required sample numbers and required execution time are not ideally linear; they all made further detailed analysis difficult.
 
@@
 The database I/O is also serial.
+It is noticed that the program took no advantage of graphics card, as \verb|multiprocessing| along can only exploit the resources inside CPU.
 

### 2020-08-20 appendix/appendix.tex
Proofread

@@
 
-\section{ABC SMC implementation}
-
-\subsection{Local machine}
+\section{Local machine}
 
@@
 
+Some figures were aligned and labelled in Adobe Photoshop. \verb|R| and Microsoft Excel were used for data processing and analysis.
+
 \begin{table}
@@
 
-\subsection{Remote machine}
+\section{Remote machine}
 
-Most experiments were performaed on Cirrus, its hardware can be found in Table \ref{table:remote_hardware}.
-The program can run in sigle compute-node which contains two processors\footnote{\url{https://www.cirrus.ac.uk/about/hardware.html}}.
+Most experiments were performed on Cirrus, whose hardware can be found in Table \ref{table:remote_hardware}.
+The program can run in single compute-node which contains two processors\footnote{\url{https://www.cirrus.ac.uk/about/hardware.html}}.
 The software environment used in the development and experiments are listed in Table \ref{table:cirrus_soft}.
@@
 
-\begin{table}
+\begin{table}[ht]
     \centering
@@
 
-\begin{table}[H]
+\begin{table}[h]
     \centering
@@
     \end{tabular}
-    \caption{Environment on local machine}
+    \caption{Environment on Cirrus}
     \label{table:cirrus_soft}

### 2020-08-20 background/background.tex
Proofread

@@
 \cite{ref:Tsarouchas} studied the complex interactions that happen at the lesion site.
-The inflammation affects the regeneration: if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration\cite{ref:Tsarouchas}.
-Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, are related the inflammation, and they are dynamically controlled by immunes cells, i.e.
-immune cells produce and control the cytokines-mediated inflammation \cite{ref:Tsarouchas}.
+The inflammation affects the regeneration: if it is promoted, it accelerates the axonal regeneration; if it is inhibited, it reduces the regeneration \cite{ref:Tsarouchas}.
+Further experiments showed that two kinds of cytokines, il-1$\beta$ and tnf-$\alpha$, are related the inflammation, and they are dynamically controlled by immunes cells.
+In other words, immune cells produce and control the cytokines-mediated inflammation \cite{ref:Tsarouchas}.
 Cytokines are small proteins, work as messengers to pass signal across different cells, which are similar to hormones and neurotransmitter.
 
-An interaction map (Figure \ref{fig:map}) can be drawn according to the evidences found in experiments.\cite{ref:Tsarouchas}.
+An interaction map (Figure \ref{fig:map}) can be drawn according to the evidences found in experiments \cite{ref:Tsarouchas}.
 After the injury, a massive immune response was observed, with the increasing numbers of several kinds of cells.
 This dynamical system consists of complex interactions and dynamical feedback between cells and cytokines.
-For example, il-1$\beta$ promotes macrophages' increase.
-As a primary source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
+For example, il-1$\beta$ promotes macrophages' increase while as a primary source, more macrophages leads to more expression of tnf-$\alpha$ mRNA.
 
@@
 \cite{ref:Tsarouchas}.
-Lines ended with arrows represent promoting effect, lines ended with T-connectors represent inhibition}
+Lines ended with arrows represent promoting effects, lines ended with T-connectors represent inhibition}
     \label{fig:map}
@@
 
-This dynamic system is suitable for mathematical modelling, where a set of equations can be written to represent the dynamics of different objects, e.g.
+This dynamic system is suitable for mathematical modelling, where a set of equations can be written to represent the dynamics of different objects, i.e.
 each type of cells and cytokines.
-Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions inside the system, based on the proposed hypothesis \cite{ref:Tsarouchas},  furthermore based our hypothesis which may include different interactions.
+Our interest lies in the modelling of the regeneration period of the dynamic system, and we wish to find that if we can quantitatively study the interactions inside the system, based on the proposed hypotheses \cite{ref:Tsarouchas},  furthermore based our hypotheses which may include different interactions.
 
-The biological background of works in this dissertation is mostly related to the experimental evidence, quantitative data and proposed hypothesis in \cite{ref:Tsarouchas}.
+The biological background of works in this dissertation is mostly related to the experimental evidence, quantitative data and proposed hypotheses in \cite{ref:Tsarouchas}.
 The published data on the quantitative measurement of cells and cytokines make the modelling and further analysis possible.
@@
 They can be used for prediction, classification, testing our understanding or hypothesis and many other tasks; thus, they are widely used to solve practical problems.
-However, for a model that describes the zebrafish spinal cord regeneration process, we were less interested in the model's predictive functionality; the main focus is on the exploration of uncertainties in parameter estimates and comparison with alternative models.
-In this case, we wish to use modelling techniques and implement inference methods, test our hypothesis and hopefully find and compare well-fitted models.
+However, for a model that describes the zebrafish spinal cord regeneration process, we are less interested in the model's predictive functionality; the main focus is on the exploration of uncertainties in parameter estimates and comparison with alternative models.
+In this case, we wish to use modelling techniques, implement inference methods, test our hypothesis and hopefully find and compare well-fitted models.
 
@@
 steady states in the ecosystem).
-If a more precise model is desired, usually more influences or interactions are considered and added as terms of equations to a re-constructed model, e.g.
+If a more precise model is desired, usually more influences or interactions are considered and added as terms of equations, e.g.
 considering environmental factors, a changing environmental capacity and predation from other species in L-V model.
@@
 In the areas where solutions are not analytically feasible, computational techniques are usually applied for approximate solutions.
-Solvers for different models had been proposed to simulate the system or find roots for some special points, such as Explicit Runge-Kutta \cite{dormand1980family} and LSODA.
+Solvers for different models had been proposed to simulate the system or find roots for some special points, such as Explicit Runge-Kutta \cite{dormand1980family} and LSODA algorithm.
 
@@
 Non-linear ordinary differential equations are used in this dissertation, where the only independent variable is time $t$, denoting the post-lesion time (hours).
-Further models can also consider other models in other forms, e.g.
+Further models can also consider models in other forms, e.g.
 stochastic models and partial differential equations.
@@
 These solvers help to simulate the trajectories of the dependent variables, given initial values and a set of fixed parameters.
-The decency of fit can be measured by compared the simulated data and observed real data.
+The decency of fit can be measured by comparing the simulated data and observed real data.
 
@@
 For a popular topic and widely used models, parameters' values can be read from previous literature.
-However, in many other cases, we need to find values explicitly for our model, where numerical values for parameters are mostly preferred as they are suitable for practical analysis and comparison.
+However, in many other cases, we need to find these values explicitly for our model, where numerical values are mostly preferred as they are suitable for practical analysis and comparison.
 
-To find the best parameters' values given observed data, either optimisation-based approaches (e.g.
+To find the best parameters' values given the observed data, either optimisation-based approaches (e.g.
 least-squares fit (LS), maximal likelihood estimation (MLE)) or inference-based approaches can be applied.
 Optimisation-based approaches may face two major drawbacks: over-fitting and local optima \cite{ref:abcsysbio}.
-Bayesian inference approaches gain popularity in recent years, as it can relieve the concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
+Bayesian inference approaches gain popularity in recent years, as it can relieve concerns in over-fitting and provide a measurement for the uncertainty in the inferred values.
 Local optima can still be a concern, and many efforts have been proposed for a more thorough exploration of the parameter space to avoid local modes.
@@
 Some optimisation-based methods attempt to find estimates of parameter values $\theta$ s.t.
-maximise the likelihood.
+maximise likelihood.
 However, such methods will struggle when facing a problem where it is hard to specify an analytical expression for likelihood function, let alone calculating and comparing values of it.
@@
 Markov Chain Monte Carlo (MCMC) \cite{ref:MCMC}.
-As an alternative solution, Approximate Bayesian computation (ABC) approaches do not focus on the evaluation of likelihood function but focus on simulating the data from the given model and observed data, then use the simulated data to approximate the true posterior.
+As an alternative solution, Approximate Bayesian computation (ABC) approaches do not try to evaluate likelihood function but focus on simulating the data from the given model and observed data, then use the simulated data to approximate the true posterior.
 ABC-rejection \cite{ABC_rejection} is a simple example of the ABC process, and further algorithms e.g.
-sequential Monte-Carlo (ABC SMC)\cite{Toni} were proposed to improve the efficiency.
+sequential Monte-Carlo (ABC SMC) \cite{Toni} were proposed to improve the efficiency.
 
-General steps for ABC algorithm includes the following (ABC-rejection):
+General steps for ABC algorithms includes the following (ABC-rejection):
 
@@
     \item Draw a sample $\theta^*$ from parameters' prior distribution. $\theta^*$ is a vector that consists of values for each parameter
-    \item Plug $\theta^*$ into the model, and use the model to generate simulated data $D^*$, where $D^*$ is of the same form as the observed data $D$, e.g. includes same summary statistics
+    \item Plug $\theta^*$ into the model, and use the model to generate simulated data $D^*$, where $D^*$ is of the same form as the observed data $D$, e.g. includes the same summary statistics
     \item The discrepancy between the simulated data and observed data is measured by a distance function $d(D^*, D)$.
 For a pre-fixed threshold value $\epsilon$, if $d(D^*, D)<\epsilon$ then the sample $\theta^*$ is accepted
-    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution then can be approximated by these samples
+    \item Repeat 1, 2 and 3 until enough numbers of samples are accepted. The posterior distribution then can be approximated by these accepted samples
 \end{enumerate}
@@
 The mean values at each time point for each dependent variable are calculated to form $D$.
-For each sample, the ODE is accumulatively solved at the same time points and generate the simulated trajectories $D^*$.
+For each sample, the ODE is accumulatively solved at the same time points and generates the simulated trajectories $D^*$.
 Using other statistics to summarise the raw data are also possible, but care must be taken as the choice of summary statistics can largely affect the resultant posterior \cite{summaryD, summaryD2}.
@@
     \item Set a threshold schedule $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ where $T$ is the total number of populations and $\{\epsilon_1, \epsilon_2, ..., \epsilon_T\}$ is a descending sequence; set the number of particles in each population as $N$, set the population index $t=1$
-    \item if $t=1$, draw samples from prior distribution until $N$ particles are accepted under threshold $\epsilon_1$, i.e. each of $N$ particles generates data $D^*$ s.t. $d(D,D^*)<\epsilon_1$
+    \item if $t=1$, draw a samples from prior distribution until $N$ particles are accepted under threshold $\epsilon_1$, i.e. each of the $N$ particles generates data $D^*$ s.t. $d(D,D^*)<\epsilon_1$
 
-    If $t>1$, draw samples from previous population $\{\theta^*_{t-1}\}$ with weights $\{w_{t-1}\}$, perturb the particles using a perturbation kernel $K(\theta^*|\theta)$ and obtain $\theta^{**}\sim K(\theta^*|\theta)$
-    \item Repeat step 2 until $N$ particles are accept under threshold $\epsilon_t$
+    If $t>1$, draw samples from previous population $\{\theta^*_{t-1}\}$ with weights $\{w_{t-1}\}$, perturb the particles using a perturbation kernel $K(\theta^*|\theta)$ and obtain $\theta^{**}\sim K(\theta^*|\theta)$.
+Repeat drawing samples until $N$ particles are accept under threshold $\epsilon_t$
     \item Calculate the weights of the accepted particles. For particle $i$, the weight is 
@@
 
-    \item Normalise $\{w^i_t\}$, increase population index $t$ by 1
+    \item Normalise $\{w^i_t\}$; increase population index $t$ by 1
     \item Repeat step 2, 3 and 4 until $t>N$
-    \item Output the accepted particles $\{\theta^*_{t}\}$ and their weights $\{w^i_t\}$ in each generation $t$
+    \item Output the accepted particles $\{\theta^*_{t}\}$ and their weights $\{w^i_t\}$ for each generation $t$
 
@@
 In implementations, many factors can affect the efficiency and goodness of fit; some of them are discussed in Chapter 4.
-Rather than a pre-fixed threshold schedule, median or quantile based epsilon schedule can save the total sampling size and thus improve acceptance rates of particles \cite{threshold}.
+Rather than a pre-fixed threshold sequence, quantile-based epsilon schedule (e.g.
+median) can save the total sampling size and thus improve acceptance rates of particles \cite{threshold}.
 Some other improvements include using adaptive population size for each generation \cite{population}, using adaptive distance function \cite{ref:adpt_dis} and switching kernels \cite{ref:kernel}, etc.
@@
 $d$ is a distance measurement, $d(D^*,D)$ represents the discrepancy between observed data and simulated data.
-for each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
+For each population $t$, $\epsilon_t$ is a threshold criterion to determine whether to accept that drawn sample}
     \label{fig:smc}
@@
 
-In addition, ABC SMC is also capable of model selection tasks \cite{model_compare}.
-Instead of draw samples from prior of a single model, we can draw samples from multiple models $m_i$ and approximate $p(m_i, \theta|D)$, which is the joint probability distribution of model $m_i$ and its accepted particles $\theta$.
+In addition, ABC SMC is also capable of model selection task \cite{model_compare}.
+Instead of drawing samples from prior of a single model, we can draw samples from multiple models $m_i$ and approximate $p(m_i, \theta|D)$, which is the joint probability distribution of model $m_i$ and its accepted particles $\theta$.
 In each generation, the marginal distribution $p(m_i|D)$ is calculated as the accepted ${\theta}$ is known; thus models can be ranked according to their probabilities.
 
-The feature that ABC algorithm can perform likelihood-free inference makes it widely used in many dynamic modelling problems, especially in the area of biology \cite{ref:abcsysbio, ref:disease, ref:compare}.
+The feature that ABC algorithms can perform likelihood-free inference makes it widely used in many dynamic modelling problems, especially in the area of biology \cite{ref:abcsysbio, ref:disease, ref:compare}.
 However, ABC is not restricted to these areas, as it can be generally used in inference tasks to give parameter estimates or model rankings.
@@
 
-As we decided to apply ABC SMC on our regeneration models for its high-dimensional parameter estimation and comparison of different models, a code development and experiments workflow was expected to be built.
+As we decided to apply ABC SMC on our regeneration models for high-dimensional parameter estimation and comparison of different models, a code development and experiments workflow was expected to be built.
 The ABC SMC algorithm needs a whole set of mathematical environments (e.g.

### 2020-08-20 conclusion/conclusion.tex
Proofread

@@
 
- Also, ABC SMC successfully helped in the model comparison and thus can be used for suggestions of the best model, although there were uncertainties remained in the process and a large number of population size needed.
+ Also, ABC SMC successfully helped in the model comparison and thus can be used for suggestions of the best model, although there were uncertainties remained in the early generations.
 In our practice, model 5 won with absolute advantages when the threshold is converged to a small value.
@@
  The inference program was mostly run on multi-core machines, and additional scaling-up performance test illustrated the scaling behaviour.
-
- When the number of running \verb|Python| processes is less than the number of physical cores, a considerable scaling performance was observed.
-Studies on the variant required sample numbers revealed that some threshold paths could lead to local optimum and significantly affect the efficiency.
-A non-linear relationship between the required sample numbers and required execution time was also observed and made the detailed analysis complicated.
+When the number of running \verb|Python| processes is less than the number of physical cores (36), a considerable scaling performance was observed.
+Studies on the variant required numbers of sample revealed that some threshold paths could lead to local optimum and significantly affect the efficiency.
+A non-linear relationship between the required number of sample and required execution time was also observed, which all made the detailed analysis complicated.
 
@@
 
- From this application, some suggestions on ABC SMC implementation could be drawn.The prior distribution and interval of each parameter can affect the inference results significantly, and improper prior cold lead to poor fit.
+ From this application, some suggestions on ABC SMC implementation could be drawn.
+The prior distribution and interval of each parameter can affect the inference results significantly, and improper prior cold lead to poor fit.
 In our cases, log-uniform prior gave better fits than uniform prior.
-The threshold path and plot of required samples in each generation can be used to identify possible local modes; if stuck, it takes much more samples than usual to move out.
-To avoid this, large population size and a proper epsilon schedule may help.
-For the uncertainties observed in model selection and some duplicated runs with high variance, a longer run with more generation could help to identify whether the convergency is consistent such that a reproducible inferred results can be obtained.
+The threshold path and plot of required number of samples in each generation can be used to identify possible local modes; if stuck, it takes much more samples than usual to move out.
+To avoid this, large population size and a proper epsilon schedule may help \cite{threshold}.
+For the uncertainties observed in model selection and some duplicated runs with high variance in posteriors, a longer run with more generation could help to identify whether the convergency is consistent such that a reproducible inferred results can be obtained.
 
@@
 
-In terms of the project progress regarding original project proposal, all the proposed subjects were explicitly studied as planned.
+In terms of the project progress regarding the original project proposal, all the proposed subjects were explicitly studied as planned.
 We found that code development cost less time than expected, however, implementations and debugging cost more efforts than expected because of both uncertainties in both the inference ability of the algorithm for our problem, and the implementation options and algorithm hyperparameters.
-Further on our results, we studied some in-depth topics and derived more conclusions.
+Further on our results, we studied some in-depth topics and derived more conclusions, e.g.
+the relationships between time and sample numbers (Figure \ref{fig:local_modes}) and the movement of posterior concentration (Figure \ref{fig:local_para}).
 Some interesting further topics were found in the studies, e.g.
-comparison with other algorithms and implementation optimisations, however, they were not conducted due the insufficient time.
+comparison with other algorithms and implementation optimisations; however, they were not conducted due the insufficient time.
 
@@
 By comparing to these methods, differences in the inference results, algorithm robustness and required computational resources could be revealed and possibly help us in choosing an inference algorism that can balance the trade-off between efficiency and a reasonable good `fit'.
-Also there other implementation options of ABC SMC worth trying, e.g.
+Also, there are some other implementation methods of ABC SMC worth trying, e.g.
 CUDA acceleration and \verb|Julia| implementations\footnote[1]{\url{https://github.com/tanhevg/GpABC.jl}} \footnote[2]{\url{https://github.com/marcjwilliams1/ApproxBayes.jl}}.

### 2020-08-20 implementation/implementation.tex
Proofread

@@
 The scaling-up experiment was performed on Cirrus.
-The profiling and its analysis were performed on a local machine.
+%The profiling and its analysis were performed on a local machine. 
 Some necessary development and debug were performed on the remote machines to ensure that experiments run correctly on the compute node.
@@
 It included the code for the ODE equations, ODE solver, data structures and data format transform functions, etc.
-By doing this in separate code files, we could change implementation options and activate ABC SMC easily in another `trigger' file, without creating definition and duplicated code for the models, observed data and parameter priors.
+By doing this modularisation in separate code files, we could change implementation options and activate ABC SMC easily in another `trigger' file, without creating definition and duplicated code for the models, observed data and parameter priors.
 
-Some missing features were found when using \verb|pyABC|; therefore some modifications were made to the source code of \verb|pyABC|.
-
-For example, the laboratory-measured data were not considered `complete' or `tidy' for modelling and analysis purpose: the available time points for cells and cytokines were different, e.g.
+Some missing features were found when using \verb|pyABC|; therefore some modifications were made to the source code of \verb|pyABC|.For example, the laboratory-measured data were not considered `complete' or `tidy' for modelling and analysis purpose: the available time points for cells and cytokines were different, e.g.
 cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl.
 These missing values were treated as `ignored' when calculating the distance between observed data and simulated data, and represented by \verb|NaN|s.
-However, in some distance calculations, \verb|NaN| was not acceptable.
+However, in the built-in distance calculations, \verb|NaN| was not acceptable.
 To cope with this, the built-in distance function was modified.
@@
 
-Analysis of the program output was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which includes multiple visualisations and summary statistic of the results.
+Analysis of the program output was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which include multiple visualisations and out of summary statistics of the results.
 For different tasks, there are also separated analysis code file.
@@
 To firstly test the ability of inference and observe the results of different options, synthetic trajectories with known true parameter values were used as target data.
-By doing this, we could directly compare the inferred parameters to the true values and compare the simulated trajectories to the true trajectories, then observe the result to see whether the inference is successful and how well the model fit the data.
+By doing this, we could directly compare the inferred parameters to the true values and compare the simulated trajectories to the true trajectories, then observe the result to see whether the inference is successful and how well the model fits the data.
 Also, by setting the same final threshold value, the efficiency under different options can be compared and help us to choose proper options when applying ABC SMC.
@@
     {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot).
-Known parameter values are obtained from a least square fitting of the observed data}
+Known parameter values are obtained from a least square fitting of the observed data.
+Error bar indicates SEM}
     \label{fig:infer_back_data}
@@
 The perturbation kernel is called in the sampling of every particle and determines how the new perturbed particle is chosen; thus it is influential to both computation complexity and the resulting posterior distribution.
-Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optimal), but it may need less computational operations or could generate a population with a higher acceptance rate if the successive epsilon values are close.
+Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optima), but it may need less computational operations or could generate a population with a higher acceptance rate.
 A kernel with a wide variance, or spreading out in a large space could help in resolving the local optima problem by a more thorough exploration of the parameter space.
-However, it can be more computation-intensive and result in a lower acceptance rate.
-A desired optimal kernel should balance the trade-offs; their properties and criteria were discussed in \cite{ref:kernel}.
+However, it can be more computation-intensive and result in lower acceptance rates.
+A desired kernel should balance the trade-offs; their properties and criteria were discussed in \cite{ref:kernel}.
 
@@
 It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks.
-The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `thinner-distributed' perturbation result.
+The $scaling\in(0,1]$ option in \verb|pyABC| will be multiplied to the covariance to produce a `thinner-distributed' perturbation result.
 
@@
 
-The kernel experiment is designed to explore the efficiency of SMC on our dynamical systems.
-Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rates would be our preference.
-The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters.
-The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference.
-As the threshold schedule was fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit, i.e.
-recover of target data trends/features and errors of the inferred parameters compared to true values were not discussed.
+The kernel experiment was designed to explore the efficiency of SMC on our dynamical systems.
+Given the same fixed threshold schedule, kernels that need less number of samples and have higher acceptance rates would be our preference.
+The experiment compared multivariate normal kernel with local MNN kernels with different options, using synthetic data to infer back the parameters.
+The acceptance rate among each generation and total required number of sample were compared to give suggestions on the kernels.
+As the threshold schedule was fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit (i.e.
+recover of target data trends/features and errors of the inferred parameters compared to true values) were not discussed.
 
@@
 
-To compare kernels, a fixed schedule was used with the final epsilon $\epsilon_t=10.0$.
-From the total required samples plot (Figure \ref{fig:kernel1}) the efficiency can be compared.
+To compare kernels, a fixed schedule of thresholds was used with the final epsilon $\epsilon_t=10.0$.
+From the total required number of samples plot (Figure \ref{fig:kernel1}) the efficiency can be compared.
 Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
@@
 As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve.
-Consequently, if greater M was tested, then it would have even lower acceptance rates; the maximum M=2000 (the whole population is considered) would have the lowest acceptance rates.
+Consequently, if a greater M was tested, then it would have even lower acceptance rates; the maximum M=2000 (the whole population is considered) would have the lowest acceptance rates.
 
@@
 
-    \caption[Total required number of sample for different kernels]%
-    {Total required samples of different kernels after 20 populations (2000 particles in each population).
+    \caption[Total required number of samples for different kernels]%
+    {Total required number of samples of different kernels after 20 populations (2000 particles in each population).
 Different color represents different generations (bottom to top: population 1 to population 20)}
@@
     \caption[Acceptance rates of different kernels]%
-    {Acceptance rates of different kernels in 20 populations. Each population has 2000 particles}
+    {Acceptance rates of different kernels in among populations. Each population has 2000 particles}
     \label{fig:acceptance1}
@@
 Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of model 1.
-Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ was also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}, Appendix B2).
+Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ was also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}).
 
-A more `local' kernel is more likely to give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution.
+A more `local' kernel is more likely to give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution more quickly.
 This holds in most cases under a given target final threshold value $\epsilon_t$.
-However, it may not produce the proper result want: a more local kernel, e.g.
-local M-NN with small M and multivariate normal with `scaling' is more likely to be stuck in local optimality, as the kernel is less likely to sample particles that are far from local concentrations, although these local optimal are still accepted before even smaller threshold values is demanded (see Section 5.2 for more discussions).
-One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimal: the last generation takes much more samples to meet the required threshold.
+However, it may not produce the proper result as  we want: a more local kernel, e.g.
+local M-NN with small M or multivariate normal with `scaling' is more likely to be stuck in local optima, as the kernel is less likely to sample particles that are far from local concentrations, although these local optima are still acceptable before even smaller threshold values is demanded (see Section 5.2 for more discussions).
+One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimum: the last generation takes much more samples to meet the required threshold.
 
@@
 To conclude, Local kernels can be fast but unstable in some case.
-For a general parameter inference of our models, multivariate kernels were preferred, as a good fit is our prior target.
+For a general parameter inference for our models, multivariate kernels are preferred, as a good fit is our prior target.
 Some local kernels are also worth trying after the multivariate normal kernel, e.g.
 local M-NN with M$\leq 250$ (for a population size of 2000, i.e.
-12.5\% of the population) and multivariate normal with scaling, but extra cares must be taken: although they are fast, moving in and moving out of local optimum make them more unstable in efficiency.
+12.5\% of the population) and multivariate normal with scaling, but extra cares must be taken: although they are fast, moving in and moving out of local optimum make them unstable in efficiency.
 
@@
 $\Delta x_i = x_{i, simulated}-x_{i, observed}$.
-In this case, data points are 12 time points of four variables, i.e.
+In this case, data points are 12 time points for four variables, i.e.
 48 data points in total.
@@
 Weights can be either pre-set according to prior knowledge of the problem or updated using an adaptive method.
-This method is known as adaptive distance.
+The later method is known as adaptive distance.
 
 Adaptive distance is changing the weights of all data points after each generation iteration, trying to assign informative data points higher weights.
-Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}
+Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}.
 
@@
 Factor is helpful as an option for efficiency.
-When some data points are equally informative, the manually pre-defined can make the distance more focused on certain data points, or behave like a normalisation that can balance the scales of different summarised statistics (here are the 48 mean values).
+When some data points are equally informative, the manually pre-defined factor can make the distance more focuse on certain data points, or behave like a normalisation that can balance the scales of different summarised statistics.
 There are two appliance case studied: (1) factors used as a normalisation option and (2) factors used to give more focus on some features of the observed data.
@@
 rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points, i.e.
-0 - 30 hpl, factors were tried in later experiments.
+0 - 30 hpl.
 Such factor scheme assigns data in the first half of the time point with higher factors (75) and data in the second half with lower factors (25).
@@
 From Figure \ref{fig:factor}, different types of factor gave different performance.
-75:25 factor required much more samples in the first generation (orange bar in the left of Figure \ref{fig:factor}) and also did not improve much in later generations; thus, it is the slowest one.
+75:25 factor required much more samples in the first generation (orange bar in the bottom of 25:75 factor plot, left of Figure \ref{fig:factor}) and also did not improve much in later generations; thus, it is the slowest one.
 The first half of observed data contained more features and information, assigning it with higher factor would make the algorithm think that it is more important to fit the first half and thus focus on the hard part, and consequently more particles were needed in the inference process.
@@
 To conclude, because the factors and weights are directly multiplied to $\Delta x$, and thus the metrics for distance are changed, the direct comparison of the efficiency under the same threshold schedule was unfeasible; we could only compare the results after the same number of generations.
-Factors gave nearly the same inferred results but required a different number of samples, hence cares must be taken when choosing factor schemes.
+Factors gave nearly the same inferred results but required different numbers of samples, hence cares must be taken when choosing factor schemes.
 Adaptive distance did not provide an accurate result as others.
@@
 
-The prior range and distribution can largely affect the inference progress (Figure \ref{fig:size}.
-(a)).
+The prior range and distribution can largely affect the inference progress (Figure \ref{fig:size}).
 Samples were drawn from a high dimensional parameter space in each sampling process.
@@
 Right plot in Figure \ref{fig:size} also indicates that the convergency is slower for wider prior ranges.
-As a result, a proper prior interval based on our experience and belief with seeing the data should be set as narrow as possible, as the inference is largely affected by the prior interval and its length.
+As a result, a proper prior interval based on our experience and belief without seeing the data should be set as narrow as possible, as the inference is largely affected by the prior interval and its length.
 Improper prior could even make the algorithm miss the true posterior \cite{ref:abcsysbio}.
@@
 For our task in the parameter inference of regeneration model, biologically motivated priors were used, which were a general reflection of biological constraints.
-Also, the units of the parameters helped us in identifying the possible scale of the parameters.
+Also, the units of the parameters helped us in identifying the possible scales of the parameters.
 
@@
 the number of particles per generation) were not addressed in this subsection.
-In the real inference, population size was set to be greater than 2000 to avoid insufficient exploration of the previous population; in some case than high variance or uncertainties in the results were observed, population size as large as 20,000 were used.
+In the real inference, population size was set to be greater than 2000 to avoid insufficient exploration of the previous population; in some case than high variance or uncertainties in the results were observed, as large as 20,000 were used as population size.
 If the epsilon did not show a convergency trend, the run could be extended to more generations until a stable result was observed.
@@
 Figure \ref{fig:result123} plots the simulated trajectories of the three models.
-One thousand particles (each particle is a parameter set) were sampled, and the mean (blue) and 25th to 75th percentile interval (grey) were calculated from the data simulated by these 1000 particles.
+One thousand particles (each particle is a parameter set) were sampled for the last generation, and the mean (blue) and 25th to 75th percentile interval (grey) were calculated from the data simulated by these 1000 particles.
 
@@
 All the model probability results are of great discrepancy (e.g.
-100\% model 3 and 0\% for model 1 and 2), so Bayesian factor is not calculated.
+100\% for model 3 and 0\% for model 1 and 2), so Bayesian factor is not calculated.
 
@@
 Together with the conclusions above, model 3 is considered to be the best model to represent the target data so far.
+Considering the motivations in proposing model 2 and 3, the result shows that a decaying $\lambda$ term is reasonable in both theory and practice; simplification of $i_{\beta\Phi}$ term gives an even better model (i.e.
+model 3), which means that the inhibition effect from macrophage can indeed be ignored.
 
@@
 An additional term $f_{\beta\alpha}\beta$ is introduced to production rate, with $f_{\beta\alpha}$ being a new model parameter (positive constant).
-This model meets the biological context, considering the source of tnf-$\alpha$.
+This model meets the biological context considering the source of tnf-$\alpha$.
 The proposed model is written as Equation \ref{eq:model5}.
@@
 
-\begin{figure}[H]
+\begin{figure}[ht!]
     \begin{center}
@@
 
-    \caption[Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison]{Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison.
-Model 5 also wins in further populations (population 25 to 30)}
+    \caption[Model probabilities of model 3, 4 and 5]{Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison. Model 5 also wins in further populations (population 25 to 30)}
     \label{fig:model345cmp}
@@
 Based on the model comparison results, we could conclude that model 5 was the best model among the tested models. 
+It proves that the extra term $f_{\beta\alpha}$ not only makes sense regarding the source of tnf-$\alpha$, but also gives a better model than phenomenological term $d_{\beta\alpha}\beta$.
 Uncertainties were observed in early generations, where similar patterns were also observed in the model comparison of model 1, 2 and 3.
@@
 A thin and concentrated shape indicates a well-inferred parameter, e.g.
-$\mu{\Phi}$ and $\kappa_{\Phi\beta}$.
+$\mu_{\Phi}$ and $\kappa_{\Phi\beta}$.
 Some distributions, however, are skewed to zero, e.g.
 $\mu_N$ and $s_{\alpha\Phi}$.
-Concentrations in the posteriors were presented, and using inter-quartiles (shaded interval in Figure \ref{fig:model5_para}) or mean values to summarise the posteriors is feasible in practice uses, e.g.
+Concentrations in all the posteriors were presented, and using inter-quartiles (shaded interval in Figure \ref{fig:model5_para}) or mean values to summarise the posteriors is feasible in practice uses, e.g.
 simulating data (Figure \ref{fig:overfit}).
@@
 As models 5 won in the model selection, further interest was to explore the parameter-to-model sensitivity of the resultant model.
-PCA was performed on the last generation of inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC.
+PCA was performed on the last generation of the inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC.
 The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less than 5\% of the variance each.
 PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs.
-Thus, they were considered to extend much narrower regions on the posterior distributions and could be used to identify the parameters that the model is sensitive to.
+Thus, they were considered to extend much narrower regions on the posterior distributions and could be used to identify the parameters that the model is most sensitive to.
 
@@
 If we conclude the last 8 PCs, it can be found that the biggest portions ($v_{N\Phi}$, $a$, $\kappa_{N\beta}$, $\kappa_{\Phi\beta}$, $\mu_N$, $\mu_\Phi$, $\nu_{N\Phi}$) mostly come from the equation of neutrophil and macrophage i.e.
-equations of the cells (the un-exploded portion in Figure \ref{fig:pc_pie} a); if we look at the last 8 PCs individually, the same pattern can be observed.
+equations of the cells (the un-exploded wedges in Figure \ref{fig:pc_pie} a); if we look at the last 8 PCs individually, the similar pattern can be observed.
 This might suggest that the model is more sensitive to cells' dynamic rather than cytokines' dynamics.

### 2020-08-20 introduction/introduction.tex
Proofread

@@
 
-Zebrafish can regenerate its spinal cord after injury; tissue regeneration observed at the injury site is a complex biological process with interacting parts and dynamical feedback between different types of cells and molecules.
+Zebrafish can regenerate its spinal cord after injury; tissue regeneration observed at the injury site is a complex biological process with many interacting parts and dynamical feedback between different types of cells and molecules.
 These cells and molecules are involved in the inflammation and repair dynamically, and a complex interacting network can be drawn according to it.
@@
 According to available researches \cite{ref:Tsarouchas}, a mathematical model was first proposed, but the addressed interactions remain to be precisely defined, and quantitatively examined in experiments.
-We want to find the best parameter sets of the model using the published data, and ranking models under specific metrics and possibly find the best-fit model.
-In these steps, models are expected to retain mathematical and biological significance regarding the complex interactions, and differential equations are suitable for this.
+We want to explore the uncertainties in parameter estimates using the published data, ranking models under specific metrics and possibly find the best-fit model.
+In these steps, models are expected to retain mathematical and biological significance regarding the complex interactions, where differential equations are suitable for this.
 
@@
 
-For the parameter estimation tasks, optimisation-based methods usually have trouble in determining global optima from local optimal, and noise in the observed data may lead to over-fitting.
-Compared to optimisation-based methods, Bayesian inference techniques can help to relieve these two concerns\cite{ref:abcsysbio}.
-Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our task as writing down likelihood as a function expression for the dynamic system can be hard.
+For the parameter estimation task, optimisation-based methods usually have trouble in determining the global optima, and noise in the observed data may lead to over-fitting.
+Compared to optimisation-based methods, Bayesian inference techniques can help to relieve these two concerns \cite{ref:abcsysbio}.
+Moreover, approximate inference methods (Approximate Bayesian Computation, ABC) make likelihood-free inference possible, which is suitable for our dynamic system as writing down likelihood as a function expression can be hard.
 
 Such methods introduce computation-intensive sampling-and-test patterns, where a massive number of samples are drawn and followed by complex computations.
-As the sampling is independent of each other in a certain period, the parallel implementation of the algorithm becomes feasible and is expected to accelerate the parameter inference on a large scale when using HPC resources.
+As the sampling is independent of each other in a certain period, the parallel implementation of the algorithm becomes feasible.
+It is expected to accelerate the parameter inference on a large scale when using HPC resources.
 
 This dissertation mainly focuses on the implementations and experiments of parameter estimation using ABC SMC (sequential Monte-Carlo) method, and discussions on the results.
-Additionally, performance experiments illustrated the strong-scaling performance of the algorithm.
-Chapter 2 talks about the background information and theories, Chapter 3 shows how the mathematical models are derived from hypothesis, and Chapter 4 describes the ABC SMC implementations and experiments on the models in detail.
-A brief scaling-up performance experiment is included in Chapter 5, and Chapter 6 summarises the conclusions and states the future works on this topic.
+Additionally, performance experiments illustrates the strong-scaling performance of the algorithm.
+Chapter 2 talks about the background information and theories, Chapter 3 shows how the mathematical models were derived from hypotheses, and Chapter 4 describes the ABC SMC implementations and experiments in detail.
+A brief scaling-up performance experiment is included in Chapter 5, and Chapter 6 summarises the conclusions and states the future works regarding this topic.

### 2020-08-20 modelling/modelling.tex
Proofread

@@
 The only independent variable is post-lesion time $t$.
-The equations consist of four derivatives of dependent variables (number of cells, mRNA expression of cytokines) with respect to $t$.
+The equations consist of four derivatives of dependent variables (numbers of two kinds of cells, mRNA expression of two types of cytokines) with respect to $t$.
 The equations were mostly an interpretation of the proposed interaction maps, with terms correspond to interaction paths.
@@
 Our initial models focused on the changes of four dependent variables $N$, $\Phi $, $\beta$ (for il-1$\beta$) and $\alpha$ (for tnf-$\alpha$) that are most important and determinative in the regeneration.
-$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is recorded as `relative mRNA expression' (compared to the value at the initial time point, i.e.
-0 hpl) and considered unitless.
+$N$ and $\Phi$ is of the unit `number of cells', $\beta$ and $\alpha$ is recorded as `relative mRNA expression' (compared to the value at the initial time point 0 hpl) and considered unitless.
 
-It was noted that the variance of the measured data is relatively high.
+It is noted that the variance of the measured data is relatively high.
 The summary statistic was the mean of measurement data at each time point, assuming that measured data was Gaussian-like distributed.
-To validate this, the distribution of the measured data points was plotted.
-The result was that at most time points the measurement values are Gaussian-distributed, although some distributions were skewed.
-One abnormal distribution was observed at time point 120 hpl for macrophage where there are two concentrations.
+To validate this, the distributions of the measured values was plotted.
+The result was that at most time points the measurement values were Gaussian-distributed, although some distributions were skewed.
+One abnormal distribution was observed at time point 120 hpl for macrophage where there were two concentrations.
 Despite this, the mean value could summarise most data and thus was still used as the target summary data.
-The mean value at each time point is shown in Figure \ref{fig:obs_data}, which is our target trajectories for the models to fit.
+The mean value at each time point is shown in Figure \ref{fig:obs_data}.
+It is our target trajectories for the models to fit.
 
@@
 
-Five models in total are proposed according to different hypotheses.
+Five models in total were proposed according to different hypotheses.
 At first, our tests and implementations of ABC SMC for parameter estimations used only the basic model for developing propose.
@@
 All these models assumed that the involved interactions are within two kinds of cells (neutrophil and macrophage) and two kinds of cytokines (il-1$\beta$ and tnf-$\alpha$), and used the data presented in Figure \ref{fig:obs_data} for the inference task.
-Interactions or influences from other cells or cytokines are not considered as there might not be corresponding data available.
+Interactions or influences from other cells or cytokines were not considered as there might not be corresponding data available.
 If such data or experimental findings are available in the future, we can consider them by then as long as these interactions are indeed necessary for our model.
@@
 A preliminary model (Eqn.
-\ref{eq:model1}) was proposed according to \cite{ref:Tsarouchas} and then mainly used to build and test the code of inference framework.
+\ref{eq:model1}) was proposed according to \cite{ref:Tsarouchas} and then mainly used to build and test the code for the inference framework.
 An interaction map of the model is shown in Figure \ref{fig:m1}.
 This model is a simplification of the process described in \cite{ref:Tsarouchas} (Figure \ref{fig:map}) with some minor interactions ignored.
-This model included 12 parameters, all of which should be positive real numbers.
+This model includes 12 parameters, all of which should be positive real numbers.
 A description of these parameters can be found in Table \ref{table:m1}.
@@
 \end{align}
-It was assumed that there are negative feedbacks the cells and cytokines, i.e.
-overcrowding inhibits the increasing rates.
-As sources, macrophage produces tnf-$\alpha$ and neutrophil produces il-1$\beta$; however, macrophage was assumed to have a negative effect on the il-1$\beta$, term $1+i_{\beta\Phi}\Phi$ in Eqn.
+It was assumed that there are negative feedbacks for all the cells and cytokines, i.e.
+overcrowding inhibits the increasing rate.
+As sources, macrophage produces tnf-$\alpha$ and neutrophil produces il-1$\beta$; however, macrophage was assumed to have a negative effect on the il-1$\beta$; term $1+i_{\beta\Phi}\Phi$ in Eqn.
 \ref{eq:model1} was used to represent this.
 Despite the self-driven increase of immune cells, il-1$\beta$ was assumed to promote the increase of both cells.
-Neutrophil was also assumed to be inhibited by the total immune cells presented in the lesion site, the term $\nu_{N\Phi}N\Phi$ was used to denote this.
+Neutrophil was also assumed to be inhibited by the numbers of immune cells presented in the lesion site; term $\nu_{N\Phi}N\Phi$ was used to denote this.
 
@@
     \end{tabular}
-    \caption{Parameters introduced in the basic model (model 1) [REMOVE unit]}
+    \caption{Parameters introduced in the basic model (model 1)}
     \label{table:m1}
@@
 
-As the observed data indicated, this dynamic system has a steady-state where after a long time ($\geq 120$ hpl), the inflammation is resolved, and immune cells should not be present at the injury site any more.
-Regarding this, the parameter for self-increase rate for the macrophage and neutrophil, i.e.
+As the observed data indicates, this dynamic system has a steady-state where after a long time ($\geq 120$ hpl), the inflammation is resolved, and immune cells should not be present at the injury site any more.
+Regarding this, the parameter for self-increase rate of macrophage and neutrophil, i.e.
 $\lambda_\Phi$ and $\lambda_N$  cannot be constant.
@@
 Some hypothesised interactions may differ from a real case, and there may exist undiscovered mechanisms.
-These could lead to model misspecification, where the model could fit the data but with wrong interaction paths, or the model can hardly fit the data and recover some local features observed in the data.
+These could lead to model misspecification, where the model could fit the data but with wrong interaction paths, or the model can hardly fit the data or recover some local features observed in the data.
 

### 2020-08-20 performance/performance.tex
Proofread

@@
 By default, if the number of cores is not specified, it will automatically read the number of available cores and use them all.
-Cirrus has a 36-core CPU which support hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
+Cirrus has a 36-core CPU which supports hyperthreading, such that the maximal number of cores available to \verb|multiprocessing| is 72.
 
 The experiment was executed on Cirrus, using 8, 16, 24, 36, 54 and 72 cores respectively.
-Each run was repeated ten times and the average execution time, required sampling numbers are recorded.
+Each run was repeated ten times.
+The average execution time and required number of samples are recorded.
 Hyperthreading was enabled when using 54 and 72 cores.
@@
 
-The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling, which is the most time-consuming part.
+The implementation of ABC SMC in \verb|pyabc| enables the parallelisation of sampling process, which is the most time-consuming part.
 The rest part of the program is mostly not parallelised, e.g.
@@
 
-A high variance of total required samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations, where they were supposed to have relatively similar average required samples such that we can compare the performance directly via speedup or efficiency.
+A high variance of total required number of samples was observed in the scaling-up experiments: some single runs required much more samples to finish 20 populations, where they were supposed to have relatively similar average required samples such that we can compare the performance directly via speedup or efficiency.
 
-Due to the variant total required samples drawn our attention to the per-second performance.
+The variant total required number of samples drawn our attention to the per-second performance.
 Although the execution time was not improved by using 54 or 72 cores, the required number of samples were higher than that of $p=36$, which indicates slight improvements.
@@
 
-The variant total required samples for different runs under the same setting was most likely the result of different threshold evolve path.
+The variant total required number of samples for different runs under the same setting was most likely the result of different threshold evolve path.
 Median epsilon schedule and randomness in the sampling process resulted in different sequences of thresholds for different runs, which means different repeated runs may have different convergency path \cite{threshold}.
@@
 To illustrate this, Two independent run, namely Run A and Run B, were chosen from the experiment of $p=24$ and visualised in Figure \ref{fig:local_modes} and Figure \ref{fig:local_para}.
-These two runs used identical options of the algorithm.
-Figure \ref{fig:local_modes} (a) shows the comparison of the required samples, where Run A converged to a local mode, and in population 11 took much more sample tires to move away from the local optimum (Figure \ref{fig:local_modes} (b)).
-Figure \ref{fig:local_para} compares the posterior density distribution before (t=9, cyan) and after (t=13, red) the jumping away: significant movement of concentrations in posterior density distributions are observed, especially in parameters from ODE 1, 2 and 4.
-An interning movement is that there are two concentrations in $\kappa_{\Phi\beta}$ and $\mu_\Phi$ when t=9, and the majority of particles are concentrated in the second concentration with high parameter values; when t=13, the second concentration distribution disappears and all particles are concentrated near zero.
+These two runs used identical options for the algorithm and implementation.
+Figure \ref{fig:local_modes} (a) shows the comparison of the required number of samples, where Run A converged to a local mode.
+In population 11 Run A took much more sample tires to move away from the local optimum.
+Figure \ref{fig:local_para} compares the posterior density distribution before (t=9, cyan) and after (t=13, red) the concentration jumping: significant movement of concentrations in posterior density distributions are observed, especially in parameters from ODE 1, 2 and 4.
+An interning movement is that there are two concentrations in $\kappa_{\Phi\beta}$ and $\mu_\Phi$ when t=9, and the majority of particles are concentrated in the second concentration with higher values; when t=13, the second concentration distribution disappears and all particles are concentrated near zero.
 
@@
 There exist some `cheap' particles.
-If at the first time points, a variable reached unacceptable high value (regarded as \verb|Inf| in \verb|Python|), the ODE solver cannot perform further integration but result in a data full of \verb|NaN|s.
-Such particles cost less time than an usual particle (where values at 120 data points are explicitly calculated and then compared with observed data).
+If at the first time points, a variable reached unacceptably high value (regarded as \verb|Inf| in \verb|Python|), the ODE solver cannot perform further integration but result in the simulated data full of \verb|NaN|s.
+Such particles cost less time than a usual particle (where values at 120 data points are explicitly calculated and then compared with observed data).
 Local mode, e.g.
-generation 11, might take large amount of these `cheap' particles.
+generation 11, might take a large amount of these `cheap' particles, thus it is `efficient' regarding the per-second performance.
 Also, generation 1 draws samples directly from prior without fitting and using kernels, so it is much faster.
@@
 
-To conclude, the performance experiments revealed reasonable scaling-up behaviour when the \verb|Python| processes did not exceed number of physical cores.
-Local modes lead to high variance in total required number of samples, and the relationship between required sample numbers and required execution time are not ideally linear; they all made further detailed analysis difficult.
+To conclude, the performance experiments revealed reasonable scaling-up behaviour when the \verb|Python| processes did not exceed the number of physical cores.
+Local modes lead to high variance in total required number of samples, and the relationship between required number of sample numbers and required execution time are not ideally linear; they all made further detailed analysis difficult.
 
@@
 The database I/O is also serial.
-It is noticed that the program took no advantage of graphics card, as \verb|multiprocessing| along can only exploit the resources inside CPU.
+It is noticed that the program takes no advantage of graphics card, as \verb|multiprocessing| along can only exploit the resources inside CPU.
 
